---
title: "xv20230808_pathway_balance_LAD_calls"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go through the mutation accumulation in BRCA2 mutant tumors

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}

```

## Libraries:

```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
```

## Import balance data

```{r pressure, echo=FALSE}
#Balance data
pathway_balance_RPE1_pools <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/pools_balance/xv20230808_DSB_TRIP_pools_balance.rds")

#Inhibitor effect
inhibitor_balance_effect <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/pools_balance/xv20230809_inhibitors_balance_change.rds")

#IPR position data
RPE1_position <- read.delim2(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/DSB_TRIP_mapping/xv20230519_RPE_pools_mapping.tsv") %>%
  bind_rows(read.delim2(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/DSB_TRIP_mapping/xv20230519_U2OS_pools_mapping.tsv")) %>%
  mutate(end = start + 1)

```

#Import LADs data
```{r}
setwd(in.dir)
# LAD domains
LAD_atlas <- as.data.frame(read.table("import/xv20220329_LAD_atlas_OSF.bed.gz",header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(LAD_atlas) <- c("chr","start","end","length","strand","LAD_class")

#LAD_length plot control
LAD_length_tibble <- LAD_atlas %>% mutate(LAD_length = end - start)

```


#Map indel data for BRCA2 and sporadic tumors
```{r}
#Create function to call unique LADs
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T) #Make LADs into GR ranges

#Function to map IPRs in LADs or iLADs
LADs_status_call_IPR <- function(x) {
  print(dim(x))
  mutations_GR <- makeGRangesFromDataFrame(x, keep.extra.columns = T) #Make a GR object with the coordinated of the deletion
  LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutations_GR) #Find overlaps between deletions and LAD states
  LAD_overlaps_dt <- tibble(x[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6]) #Extract the LAD_status of the mapped deletion
}

#Calls
RPE1_IPR_calls <- LADs_status_call_indel(RPE1_position)

LAD_iLAD_calling <- RPE1_IPR_calls %>%
  separate(name, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity)

#Balance in LADs and iLADs
RPE_balance_LAD <- pathway_balance_RPE1_pools %>% left_join(LAD_iLAD_calling)

#Inhibitor effect
inh_effect_LAD <- inhibitor_balance_effect %>% left_join(LAD_iLAD_calling)

```

#Calculate statistic test
```{r}
test_LAD_iLAD <- map_dfr(c("RPEPro","RPEDef","U2OS"), function(x) {
  #filter data
  cLAD_data <- RPE_balance_LAD %>% filter(cell_line == x & LAD_status == "cLAD")
  print(dim(cLAD_data))
  #test
  ciLAD_data <- RPE_balance_LAD %>% filter(cell_line == x & LAD_status == "ciLAD")
  print(dim(ciLAD_data))
  t.test(cLAD_data$mean_bal, ciLAD_data$mean_bal) %>% broom::tidy() %>% mutate(cell_line = x)
})

#Plot cLAD & ciLADs
ggplot(RPE_balance_LAD %>% filter(grepl("c", LAD_status))) +
  geom_quasirandom(aes(LAD_status, mean_bal)) + 
  facet_wrap(~ cell_line) +
  ggpubr::stat_compare_means(method = "wilcox.test", aes(LAD_status, mean_bal), label.y = 0.5) +
  theme_bw() +
  ylim(c(0,0.6))
```
#Conclusion: Pathway balance differences are also very small between LAD and iLADs

###Test if ∆log2MMEJ::NHEJ are different
```{r}
test_LAD_iLAD <- map_dfr(c("RPEPro","RPEDef","U2OS"), function(x) {
  #filter data
  cLAD_data <- RPE_balance_LAD %>% filter(cell_line == x & LAD_status == "cLAD")
  print(dim(cLAD_data))
  #test
  ciLAD_data <- RPE_balance_LAD %>% filter(cell_line == x & LAD_status == "ciLAD")
  print(dim(ciLAD_data))
  t.test(cLAD_data$mean_bal, ciLAD_data$mean_bal) %>% broom::tidy() %>% mutate(cell_line = x)
})

#Plot cLAD & ciLADs
ggplot(inh_effect_LAD %>% filter(grepl("c", LAD_status))) +
  geom_quasirandom(aes(LAD_status, mean_diff)) + 
  facet_grid(condition ~ cell_line) +
  ggpubr::stat_compare_means(method = "wilcox.test", aes(LAD_status, mean_diff), label.y = 0.5) +
  theme_bw() 
```
#CONCLUSION: This also doesn't help with it. None of inhibitor has strond synergies with LADs.

GENERAL CONCLUSION: THIS METHOD IS NOT BETTER THAN THE PREVIOUS ONE!