---
title: "xv20230329_RPE1_pool_pathway assay"
author: "Xabier Vergara"
date: "2023-07-26"
output:
  html_document: default
  pdf_document: default
---
#Date: 26th July 2023
#Author: Xabier Vergara
#Aim: The aim of this file is to wrap up the filtering strategy and measure if pathway balance in control samples has chromatin biases.

#Load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
```


## Import data
```{r, warning=F}
# List files in mcclintock
pathway_assay_files <- list.files("/DATA/projects/DSBrepair/data/xv20230714_E2286_pool_pathway_assay/indelPCR_counts",pattern = '*[.]co', full.names = T) 

#Read and parse files
clone_pools_pathway_assay <- map_dfr(pathway_assay_files, function(x){
  read_delim(x, show_col_types = F) %>% 
  mutate(filename = gsub("/DATA/projects/DSBrepair/data/xv20230714_E2286_pool_pathway_assay/indelPCR_counts/","",x),
         outcome = paste(call, indel, sep = "_")) %>%
  separate(filename, into = c("exp_n","sample_ID","cell_line","condition","bio_rep","t_rep"))
})
```



#Add chromatin info for RPE1 cells
#Load chromatin info of the IPRs in RPE cells (2000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "2000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "2000_", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-2000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_2000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity, - dam_LMNNB2) %>%
  mutate(binsize = "2000")
```

#Load chromatin info of the IPRs in RPE cells (5000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "5000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "5000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-5000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=5000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_5000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity, -dam_LMNNB2)  %>%
  mutate(binsize = "5000")
```


#Load chromatin info of the IPRs in RPE cells (10000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "10000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "10000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-10000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=10000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_10000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity)  %>%
  mutate(binsize = "10000")
```

#Load chromatin info of the IPRs in RPE cells (20000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "20000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "20000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-20000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=20000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_20000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "20000")
```

#Merge all binsizes in one data table
```{r}
chromatin_pools_binsize <- bind_rows(chromatin_features_pools_20000,chromatin_features_pools_10000,chromatin_features_pools_5000,chromatin_features_pools_2000)
```


#Join with balances and calculate the correlation (do this for all marks) = Spearman's
```{r, warning=F}
#RPE data
values_chromatin_RPE <- left_join(filtered_balance_75, chromatin_pools_binsize)

#Calculate spearman correlation for all chromatin features
chromatin_analysis_pair <- analysis_pair %>% filter(cell_line != "U2OS")
binsizes <- c("2000","5000","10000","20000")

spearman_coefficients <- map_dfr(chromatin_analysis_pair$cell_line, function(x){
  filter_data <- values_chromatin_RPE %>% filter(cell_line == x)
  binsize_filter <- map_dfr(binsizes, function(y){
    bin_filter <- filter_data %>% filter(binsize == y)
    rho_table <- map_dfr(colnames(values_chromatin_RPE[9:19]), function(j){
      balance <- bin_filter$mean_bal
      feature <- as.numeric(as.matrix(bin_filter[j]))
        cor.test(x = balance, y = feature, method = "spearman", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })
})

#chromatin feature order
chromatin_order <- c("dam_H3K9me3","dam_H3K9me2","late_replicating","dam_LMNB2","dam_LMNB1","dam_H3K27me3","chip_H3K4me3","chip_H3K4me2","chip_H3K4me1","chip_H3K27ac","chip_H3K36me3")

#Plot correlation coefficients
ggplot(spearman_coefficients %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Spearman's correlation coefficient of MMEJfreq.") + 
  facet_grid(fct_relevel(binsize,binsizes)~ fct_relevel(cell_line, c("RPE1","RPEPro","RPEDef")))
```

#Join with balances and calculate the correlation (do this for all marks) = Pearson's
```{r, warning=F}
#RPE data
values_chromatin_RPE <- left_join(filtered_balance_75, chromatin_pools_binsize)

#Calculate spearman correlation for all chromatin features
chromatin_analysis_pair <- analysis_pair %>% filter(cell_line != "U2OS")
binsizes <- c("2000","5000","10000","20000")

pearson_coefficients <- map_dfr(chromatin_analysis_pair$cell_line, function(x){
  filter_data <- values_chromatin_RPE %>% filter(cell_line == x)
  binsize_filter <- map_dfr(binsizes, function(y){
    bin_filter <- filter_data %>% filter(binsize == y)
    rho_table <- map_dfr(colnames(values_chromatin_RPE[9:19]), function(j){
      balance <- bin_filter$mean_bal
      feature <- as.numeric(as.matrix(bin_filter[j]))
        cor.test(x = balance, y = feature, method = "pearson", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })
})

#chromatin feature order
chromatin_order <- c("dam_H3K9me3","dam_H3K9me2","late_replicating","dam_LMNB2","dam_LMNB1","dam_H3K27me3","chip_H3K4me3","chip_H3K4me2","chip_H3K4me1","chip_H3K27ac","chip_H3K36me3")

#Plot correlation coefficients
ggplot(pearson_coefficients %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Pearson's correlation coefficient of MMEJfreq.") + 
  facet_grid(fct_relevel(binsize,binsizes)~ fct_relevel(cell_line, c("RPE1","RPEPro","RPEDef")))
```


#Conclusion: Increasing binsizes and using pearson correlation really help

#Plot individual correlations: To check how they look
```{r, fig.height=3, fig.width=10, message=F}
#Plot correlations (These are significant correlations)
ggplot(values_chromatin_RPE, aes(dam_H3K9me3, mean_bal)) + 
  geom_point(aes(size =sd_bal, color = as.character(reps))) + 
  facet_grid(~ cell_line) + 
  geom_smooth(method = "lm",se = F)+ 
  ggpubr::stat_cor(method = "pearson") +
  theme_bw()

#Plot correlations (These are significant correlations)
ggplot(values_chromatin_RPE, aes(chip_H3K36me3, mean_bal)) + 
  geom_point(aes(size =sd_bal, color = as.character(reps))) + 
  facet_grid(~ cell_line) + 
  geom_smooth(method = "lm",se = F)+ 
  ggpubr::stat_cor(method = "spearman") +
  theme_bw()

```


