---
title: "xv20230329_RPE1_mapping_clone"
author: "Xabier Vergara"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
library(broom)
library(scales)
```

# Patwhay balance in clone 32: After mapping
## Import data: Balance data and some other details
```{r, warning=F}
# Import data for high confidence reporters and calculate mean
high_confidence_IPR <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/inhibitors/xv20230919_inhibitors_PRO_DEF_freq.rds")

#Import control data
control_data_DEF <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/inhibitors/xv20230919_inhibitors_control_data.rds")

#Pools used in the experiment
selected_pools <- c("RPE1_Low_1000","RPE1Deff_Low_1000","RPE1Proff_Low_250","U2OS_High_100")

#Import chromatin data
#Export data frame
chromatin_data_RPE1 <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230919_chromatin_data_RPE1.rds")

```

#Calculate differentials per replicate
```{r}
dmso_balance <- high_confidence_IPR %>% filter(gRNA == "LBR2") %>% ungroup() %>% select(barcode, cell_line, bio_rep, dmso_bal = log2_bal)
```

#Calculate diff for these the high_confidence barcodes
```{R}
#Calculate differential
diff_gRNA_RPE1 <- high_confidence_IPR %>% 
  left_join(dmso_balance) %>%
  mutate(gRNA_diff = log2_bal - dmso_bal) %>%
  filter(gRNA != "LBR2")

#Average perturbation
mean_gRNA_effect <- diff_gRNA_RPE1 %>%
  dplyr::group_by(barcode, gRNA,cell_line) %>%
  dplyr::summarise(mean_diff = mean(gRNA_diff),
                   reps = n())

```


#Plot pathway balance changes
```{R}
ggplot(mean_gRNA_effect) +
  ggbeeswarm::geom_quasirandom(aes(cell_line, mean_diff)) +
  facet_wrap(~ gRNA) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()
```

#Join chromatin data
```{r}
#Create RPE1 dataframe
inhibitor_effect_chromatin_RPE1 <- mean_gRNA_effect %>%
  left_join(chromatin_data_RPE1) %>%
  na.omit()
```


#Plot some control plots
```{r, fig.width=3, fig.height=4}
#Import K562 data 
K562_data <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_inhibitors/xv20220929_drug_differentials_log2_MMEJ_NHEJ_balance.rds") %>% dplyr::select(gRNA = drug, barcode,mean_diff =mean.log2foldchange) %>% mutate(cell_line = "K562")

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230919_ATMi_global_effect.pdf", width=3, height=4)
ggplot(mean_gRNA_effect  %>% bind_rows(K562_data) %>% filter(gRNA == "ATMi")) +
  geom_quasirandom(aes(fct_relevel(cell_line,c("K562","DEF","PRO","RPE1")), mean_diff, color = cell_line)) + 
  geom_hline(yintercept = 0, linetype =2) +
  scale_color_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  theme_bw() +
  theme(legend.position = "none") + ylab("∆log2 MMEJ:NHEJ balance") + xlab("Cell line")
dev.off()


```


#Calculate global differences
```{r}
#Calculate mean effect
mean_effects <- mean_inhibitor_effect %>% 
  dplyr::group_by(cell_line, condition) %>%
  dplyr::summarise(global_diff = mean(mean_diff, na.rm = T))

#Calculate p.value calculation
p_val_global <- mean_inhibitor_effect %>% 
  dplyr::group_by(cell_line, condition) %>%
  rstatix::t_test(mean_diff ~ 0)

#Adjust p-value
adj_pvalue <- map_dfr(cells, function(x) {
  p_val_global %>% filter(cell_line == x) %>% mutate(global_p_adj = p.adjust(p, method = "BH"))
}) 

#Create data frame
global_effects <- mean_effects %>% 
  left_join(adj_pvalue) %>%
  select(cell_line, drug = condition, global_diff, global_p_adj)



```


#Calculate slopes for RPE
```{r}
#Create empty dataframe to calculate synergy scores
drug_CCD_slopes_RPE <- tibble(gRNA = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA, cell_line = NA, binsize = NA)

#Loop to run linear models on the values
for (h in unique(inhibitor_effect_chromatin_RPE1$gRNA)) {
  for(i in unique(inhibitor_effect_chromatin_RPE1$cell_line)) {
  for (j in colnames(inhibitor_effect_chromatin_RPE1)[6:16]) { #Run this function for each of the 25 high quality chromatin features
    model.dt <- inhibitor_effect_chromatin_RPE1 %>% filter(gRNA == h & cell_line == i & binsize == 2000) # And For each gene
   if (nrow(model.dt) == 0) {
next
}
    model.epistasis.log2 <- lm(formula = mean_diff ~ unlist(model.dt[j]), data = model.dt) %>% tidy() #Correlation analysis
   drug_CCD_slopes_RPE <- drug_CCD_slopes_RPE %>% tibble::add_row(gRNA = h,binsize =2000, feature = j,cell_line = i, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value)) #Select valuable parameters and save them in data frame
  }
  }
}

#Retain slopes that 
drug_CCD_all_values  <- drug_CCD_slopes_RPE %>% 
  reshape2::dcast(gRNA + feature + cell_line ~ term, value.var = "slope.log2") %>%  #dcast table
  dplyr::select(gRNA, cell_line,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% #Extract information for slopes only
  mutate(CCD_value = case_when(gRNA == "ATMi" & slope < 0 ~ slope,
                               gRNA == "ATMi"& slope > 0 ~ 0,
                               gRNA == "DNAPKi" & slope > 0 ~ slope,
                               gRNA == "DNAPKi" & slope < 0 ~ 0, T ~ slope)) 
#Call M-synergies, N-synergies or no synergies based on the slope and MMEJ:NHEJ differentials
```

#Plot heterochromatin features
```{r}
#K562 inhibitor data
K562_CCD <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220816_drug_CCD_results.rds") %>% dplyr::select(gRNA = drug, CCD_value =CCD_synergy_score,chrom_feature) %>% mutate(cell_line = "K562")

RPE_data_CCD <- drug_CCD_all_values %>%  mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature)) %>% select(gRNA, chrom_feature, CCD_value, cell_line)

drug_CCD_all_values_comb <- bind_rows(K562_CCD,RPE_data_CCD)  %>%
  mutate(brca1_status = case_when(cell_line %in% c("K562","DEF") ~ "BRCA1null",
                                  cell_line %in% c("PRO","RPE1") ~ "BRCA1pro"))

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230919_ATMi_CCD_effects.pdf", width=3, height=4)
ggplot(drug_CCD_all_values_comb %>% filter(chrom_feature %in% c("H3K9me2","H3K9me3","LMNB1","late_replicating","H3K27me3") & gRNA == "ATMi")) +
  geom_quasirandom(aes(fct_relevel(cell_line,c("K562","DEF","PRO","RPE1")), CCD_value, color = cell_line)) + 
  geom_hline(yintercept = 0, linetype =2) +
  scale_color_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  theme_bw() +
  theme(legend.position = "none") + ylab("CCD score") + xlab("Cell line")
dev.off()

#Test
ATMi_CCD_wilcox <- drug_CCD_all_values_comb %>% 
  filter(chrom_feature %in% c("H3K9me2","H3K9me3","LMNB1","late_replicating","H3K27me3") & CCD_value != 0 & gRNA == "ATMi") %>% 
  dplyr::group_by(gRNA) %>%
  rstatix::wilcox_test(CCD_value ~ brca1_status)

```

```{r}
K562_data_chromatin <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_inhibitors/xv20220929_drug_log2_MMEJNHEJ_differentials_chromatin_KO.rds") %>% dplyr::select(gRNA = drug, barcode,mean_diff =mean.log2foldchange, H3K27me3, H3K9me2, H3K9me3) %>% mutate(cell_line = "K562")


#Interaction with triple heterochromatin

CCD_summary <- drug_CCD_all_values_comb %>% filter(chrom_feature %in% c("H3K9me2","H3K9me3","LMNB1","late_replicating","H3K27me3") & gRNA == "ATMi") %>% dplyr::group_by(cell_line) %>% slice_min(CCD_value) %>% filter(CCD_value != 0)

vector_selection <- CCD_summary %>% select(cell_line, gRNA, max_het = chrom_feature) %>% distinct()

#Filter these values
trip_het_table <- inhibitor_effect_chromatin_RPE1 %>% filter(binsize == 2000)
colnames(trip_het_table) <- gsub("dam_|chip_","",colnames(trip_het_table))
all_trip_het_table <- trip_het_table %>% bind_rows(K562_data_chromatin) %>% right_join(vector_selection) 

max_het_correlation <- map2_dfr(vector_selection$cell_line, vector_selection$gRNA, function(x,y){
  filter_dt <- all_trip_het_table %>% filter(gRNA == y & cell_line == x)
  het_feature <- unique(filter_dt$max_het)
  print(het_feature)
  filter_dt %>% select(barcode, gRNA, mean_diff, sel_f = het_feature, max_het, cell_line)
})

#Plot
pdf(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230919_correlation_max_heterochromatin_inhibitor.pdf", width = 3, height = 3)
ggplot(max_het_correlation ,aes(sel_f, mean_diff, color = cell_line)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw() +
  ylab("∆log2 MMEJ::NHEJ") +
  theme(legend.position = "top") +
  scale_color_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  facet_wrap(~ gRNA, ncol = 1, scales = "free_y") + xlab("Triple heterochromatin feature")
dev.off()

```

#CONCLUSION: There are chromatin context dependencies and some of them seem to be conserved. Others dependent on the cell type. 
ATM inhibitor has H3K27me3 CCD in most cell lines and DNAPK inhibitor H3K36me3.

#Plot correlation plots for H3K27me3 & LMNB1 with ATMi and DNAPKi with H3K36me3
```{r, message=F, warning=F}
#With H3K27me3 and ATMi
ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(gRNA == "ATMi" & binsize == 2000), aes(dam_LMNB2,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)


```

#Test balances with large bins and only PRO & DEF
```{r}
ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 20000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K27me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 20000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K9me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 20000 & cell_line %in% c("RPEPro","RPEDef")), aes(chip_H3K27ac,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

```

#Test balances with large bins and only PRO & DEF
```{r}
ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "DNAPKi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K27me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "DNAPKi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K9me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "DNAPKi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(chip_H3K36me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

```
#ATM inhibitor
```{r}
ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K27me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(dam_H3K9me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

ggplot(inhibitor_effect_chromatin_RPE1 %>% filter(condition == "ATMi" & binsize == 2000 & cell_line %in% c("RPEPro","RPEDef")), aes(chip_H3K36me3,mean_diff)) +
  geom_point() +
  geom_smooth(se = F, method = "lm") + 
  ggpubr::stat_cor(method = "pearson") +
  facet_wrap(~ fct_relevel(cell_line, cells)) +
  theme_bw() + geom_hline(yintercept = 0, linetype = 2)

```

#PCR modeling: RPE1 data (only reliable dataset)
```{r}
#Create an empty dt with CCDs of DDR proteins
drug_CCDs_dt_RPE1_test <- tibble(cell_line = NA, drug = NA,binsize = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)
for (j in c("RPEPro","RPEDef")){
  for (i in unique(inhibitor_effect_chromatin_RPE1$condition)){
    gene.library.dt <- filter(inhibitor_effect_chromatin_RPE1, condition == i & cell_line == j & binsize == 2000)
    set.seed(1)
    PCR_model_DDR_test <- pls::pcr(mean_diff~ dam_H3K9me3+dam_H3K27me3+chip_H3K36me3, data=gene.library.dt , validation="CV") #Run principal component regression
    pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = 3) #Run model with three PC
    combined.dt <- tibble(measured = gene.library.dt$mean_diff, predicted = as.numeric(pcr_pred)) #Create table with real and predicted differentials
    pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% broom::glance() #Predicted vs. measured correlation plot
    drug_CCDs_dt_RPE1_test <- drug_CCDs_dt_RPE1_test %>% add_row(cell_line = j, drug = i,binsize = 2000, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value)) #Extract valuable metrics
  }
  }

#Adjust per cell line
adj_drug_dt_RPE1_adjs_pro <- map_dfr(cells, function(x) {
  drug_CCDs_dt_RPE1_test %>% filter(cell_line == x) %>% mutate(p.adj = p.adjust(p.value, method = "BH"))
})
```



#Export data table
```{r}
write_rds(CCD_dt, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/pools_balance/xv20230814_CCD_analysis_summary.rds")

```