---
title: "xv20220826_2_CCD_model"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Script to generate figure 2

Fig. 2. CCDs of 89 DNA repair proteins. 

```{r functions, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
    #write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    readRDS(correct_file)
    #write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  readRDS(correct_file)
  #write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
}
```

# Libraries
```{r libraries, warning=FALSE, message=FALSE}
# libraries:
library(tidyverse)
library(dendextend)
library(ggdendro)
library(grDevices)
library(pheatmap)
library(reshape2)
library(tibble)
library(ggpubr)
library(scales)
library(ggrepel)
library(ggdendro)
```

# Import data tables
```{r import , warning=FALSE, message=FALSE}
#setwd
setwd(in.dir)
# Individual data point (intermediate step)
step2.epistasis <- readRDS_proof("data/processed_data/CCD_analysis","log2_MMEJNHEJ_differentials_chromatin_KO")

#DNA repair CCDs
table_S3_CCDs  <- readRDS_proof("data/processed_data","Table_S7")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

```

# Figure 2A
```{r Fig_2_A , warning=FALSE, message=FALSE}
#Create mock examples
mock.slopes <- tibble(z_score = seq(-3,3,0.25), A = z_score*0 + 0.6,B = z_score*0.1 +0.6, C = z_score*-0.1 - 0.6) %>% reshape2::melt(id.vars = "z_score")

#Plot mock examples
ggplot(mock.slopes, aes(z_score,value,color = variable)) + theme_bw() + 
  geom_smooth() +
  geom_vline(xintercept = 0, linetype = 2) + 
  geom_hline(yintercept = 0, linetype = 2) + 
  ylab("log2(fold change NHEJ:MMEJ)") + xlab("Feature signal (z-score)") + 
  theme(legend.position = "none") + coord_cartesian(ylim= c(-1.5,1.5)) + scale_color_manual(values = c("black","#01665e","#8c510a"))
```
(A) Illustration of M- and N-synergy concepts. For chromatin feature – protein combinations with N-synergy (I in green) ∆log2MMEJ:NHEJ scores increase as the chromatin feature levels increase. For combinations with M-synergy (II in brown) ∆log2MMEJ:NHEJ scores decrease as the chromatin feature levels increase. For combinations with no synergy (III in black) ∆log2MMEJ:NHEJ scores do not correlate with the individual chromatin features.

# Figure 2B:
```{r Fig_2_B, warning=FALSE, message=FALSE}
# Plot RAD50 with LMNB1
ggplot(step2.epistasis %>% 
         dplyr::filter(gene %in% c("RAD50")),
       aes(LMNB1,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = -0.114),method = "lm") +
  stat_cor(label.x = -1.5, label.y = -0.7) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(RAD50 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")

# Source data table
fig2b <- step2.epistasis %>% 
  dplyr::filter(gene %in% c("RAD50")) %>%
  select(x_LMNB1 = LMNB1, y_deltalog2bal = mean.log2foldchange) %>%
  mutate(figure = "Figure_2B")

write.table(fig2b, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig2b.txt", row.names = F)

```
(B) M-synergy example: linear fit (brown line) of RAD50 ∆log2MMEJ:NHEJ scores with LMNB1 interaction levels. Slope (synergy score) and p-value of the correlation are shown in the graph. 

# Figure 2C
```{r Fig_2_C , warning=FALSE, message=FALSE}
# Plot MDC1 and H2AFZ
ggplot(step2.epistasis %>% 
         filter(gene == "MDC1"),
       aes(H2AFZ,mean.log2foldchange)) + 
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "black") +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  stat_cor()+
  theme_bw() + ylab("log2(MDC1 MMEJ:NHEJ balance)") 

# Source data table
fig2c <- step2.epistasis %>% 
  dplyr::filter(gene %in% c("MDC1")) %>%
  select(x_H2AFZ = H2AFZ, y_deltalog2bal = mean.log2foldchange) %>%
  mutate(figure = "Figure_2C")

write.table(fig2b, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig2c.txt", row.names = F)
```
(C) No synergy example: linear fit (black line) of MDC1 ∆log2MMEJ:NHEJ scores with H2AFZ levels. Slope (synergy score) and two-sided p-value of the correlation are shown in the graph. 

# Figure 2D
```{r Fig_2_D, warning=FALSE, message=FALSE}
# Separate data based on the significant IPRs that they have (based on z-score)
significant_proteins <- table_S3_CCDs %>% filter(CCD_model_p_adj < 0.05)

# dcaast with selected
slope.prot.features.dcast.hits <- significant_proteins %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap <- pheatmap(t(slope.prot.features.dcast.hits), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 4)
signif_prots.gene.order <- rev(rownames(slope.prot.features.dcast.hits[signif_prots_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_proteins) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order),fct_relevel(chrom_feature, heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

# Dendrogram
ggdendrogram(rev(signif_prots_heatmap$tree_col))

# Source data table
fig2d <- significant_proteins %>% 
  select(x_gene = gene, y_chromatin = chrom_feature, fill_synergyscore = CCD_synergy_score) %>%
  mutate(figure = "Figure_2D")

write.table(fig2d, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig2d.txt", row.names = F)
```
(D) Heatmap of synergy scores of all 89 proteins with significant CCDs. Proteins (columns) mentioned in the text are highlighted in bold. Chromatin features (rows) are colored and ordered as in Fig. 1C.


# Output chunk
```{r, eval = FALSE, echo = FALSE}
setwd(in.dir)
#Extract 4 clusters
cluster_data <- cutree(signif_prots_heatmap$tree_col, k=4) 
cluster_tibble <- tibble(gene = names(cluster_data), cluster = cluster_data) %>% mutate(cluster_name = case_when(cluster == "1" ~ "M-synergy_2", cluster == "2" ~ "M-synergy_1", cluster == "3" ~ "N-synergy", cluster == "4" ~"M-synergy_3"))

#cluster summary
cluster_summary <- cluster_tibble %>% dplyr::group_by(cluster_name) %>% dplyr::summarise(c = n())

#Export cluster data to use it in file 4
saveRDS_proof(cluster_tibble, "data/processed_data/xv20220927_DR_CCD_cluster_data.rds")
```
