---
title: "xv20220623_4_interactions_data"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Script to generate figure 4 in the paper. 
Fig. 4: Proteins that physically interact tend to have similar CCD patterns. 

```{r function, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
    #write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    base::readRDS(file = correct_file)
    #write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  base::readRDS(file = correct_file)
  #write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
}
```

# Libraries
```{r libraries, warning=FALSE, message=FALSE}
# setwd
setwd(in.dir)

# libraries:
library(tidyverse)
library(reshape2)
library(broom)
library(ggrepel)
library(dendextend)
library(umap)
library(igraph)
library(lsa)
library(tibble)
library(utils)
library(base)
library(scales)
library(openxlsx)
library(ggbeeswarm)
```

# Import data tables
```{r input files, warning=FALSE, message=FALSE}
# setwd
setwd(in.dir)

# Proteins used in the validation experiment
proteins_gRNA <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen","gRNA_gene") #gRNA gene conversion

#DNA repair protein KO CCD analysis
table_S3_CCDs <- readRDS_proof("data/processed_data/","S7")

#Filter previous data table only for significant CCDs: Created here
significant_proteins <- table_S3_CCDs %>% filter(CCD_model_p_adj < 0.05) %>% na.omit() #Melt
ddr_significant_CCDs_dcast <- significant_proteins %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene") #dcast

#Import validation experiment data
#Processed CCD data (4 cell lines)
PRO_CCD <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data","PRO_CCD")
DEF_CCD <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data","DEF_CCD")
K562_CCD_all <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data","Table_S7") %>%
  mutate(cell_line = "K562") %>% 
  select(-screen_position,-DR_effect_type) %>% 
  dplyr::group_by(gene) %>%
  dplyr::mutate(pathway_signif = case_when(sum(CCD_synergy_score < 0) == 0 & sum(CCD_synergy_score > 0) != 0 ~ "NHEJ",
                                    sum(CCD_synergy_score > 0) == 0 & sum(CCD_synergy_score < 0) != 0 ~ "MMEJ",
                                    sum(CCD_synergy_score > 0) != 0 & sum(CCD_synergy_score < 0) != 0 ~ "both",
                                    T ~ "none"),
                n = 19,
                toxicity = "non_toxic")

K562_CCD_validation <- K562_CCD_all %>%
  right_join(proteins_gRNA) %>%
  select(-gRNA)

#Bind all table together
cell_lines_CCD <- bind_rows(PRO_CCD, DEF_CCD,K562_CCD_validation)

#Cluster data (named as cluster tibble)
cluster_tibble <- readRDS_proof("data/processed_data","DR_CCD_cluster_data")

#Inhibitor data
CCD_inhibitors_data <- readRDS_proof("data/processed_data/","drug_CCD")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

#Viability data
viability_data <- readRDS(file= "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230908_dna_quantification_viability.rds")

#log2 chromatin values
#All log2_values
PRO_log2_chrom <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_validation_RPE1/","PRO_log2")
DEF_log2_chrom <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_validation_RPE1/","DEF_log2")

#Bind both data frames
DEF_PRO_log2_chrom <- bind_rows(PRO_log2_chrom, DEF_log2_chrom)

#Chromatin concordance
chromatin_concordance <- DEF_PRO_log2_chrom %>% select(-gene,-gRNA,-mean.log2foldchange, -n_rep, -binsize, - dam_LMNB2) %>% distinct() %>% melt() %>% mutate(chrom_feature = str_extract(variable, "(H3|l|L).*"))

#Filter out chromatin 
filter_out_chromatin <- chromatin_concordance %>% filter(value > 0.5) %>% dplyr::group_by(cell_line, chrom_feature) %>% dplyr::summarise(IPR_n = n()) %>% filter(IPR_n < 3)

```

# Import BIOgrid database
```{r database, warning=FALSE, message=FALSE}
# setwd
setwd(in.dir)

#Import human database
human.biogrid.database <- read.table(file = "import/BIOGRID_database/xv20220530_BIOGRID-ALL-4.4.209.tab3.txt", sep = "\t", fill = T)[c(1,4,5,8,9,12,13,14,36,37)] %>% filter(V36 == "Homo sapiens" & V37 == "Homo sapiens")

#Human physical biogrid database
h.physical.biogrid.db <- human.biogrid.database %>% filter(V13 == "physical" & !V12 %in% c("Biochemical Activity","Far Western","Protein-RNA","Protein-peptide","Reconstituted Complex"))

#filter significant proteins
significant_protein.vector <- c(significant_proteins$gene %>% unique(), "BRE","C19orf40","NDNL2","C20orf196","OBFC1","UFD1L") #Vector with significant proteins
ddr.ccd.physical.biogrid.db <- h.physical.biogrid.db %>% filter(V8 %in% significant_protein.vector & V9 %in% significant_protein.vector) #Filter interaction database for interaction happening between proteins with CCDs

#Change colnames
colnames(ddr.ccd.physical.biogrid.db) <- c("interaction_ID","gene1_ID", "gene2_ID", "gene1","gene2","int_method","int_type","publication","sp_gene1", "sp_gene2")

#Repair names to match screening data
ddr.ccd.physical.biogrid.rep.db <- ddr.ccd.physical.biogrid.db %>% mutate(gene1 = case_when(gene1 == "BRE" ~ "BABAM2",
                                                                                               gene1 == "C19orf40" ~"FAAP24",
                                                                                               gene1 == "NDNL2" ~ "NSMCE3",
                                                                                               gene1 == "C20orf196" ~ "SHILD1",
                                                                                               gene1 == "OBFC1" ~ "STN1",
                                                                                               gene1 == "UFD1L" ~ "UFD1",
                                                                                               T ~ gene1),
                                                                          gene2 = case_when(gene2 == "BRE" ~ "BABAM2",
                                                                                               gene2 == "C19orf40" ~"FAAP24",
                                                                                               gene2 == "NDNL2" ~ "NSMCE3",
                                                                                               gene2 == "C20orf196" ~ "SHILD1",
                                                                                               gene2 == "OBFC1" ~ "STN1",
                                                                                               gene2 == "UFD1L" ~ "UFD1",
                                                                                               T ~ gene2))

#Unique interactions
interacting_pairs <- ddr.ccd.physical.biogrid.db %>% dplyr::select(gene1,gene2) %>% distinct() %>% dim()
```
Conclusion: There are 138 physically interacting protein pairs between proteins with significant CCDs

# Figure 4A
## Data processing
```{r prep_Fig_3AB, warning=FALSE, message=FALSE}
# setwd
setwd(in.dir)

#Create a dataframe with all possible protein pairs
signif.protein.vector.ddr <- significant_proteins %>% pull(gene) %>% unique() #Get gene vector with CCDs
comb_signif.ddr <- combn(signif.protein.vector.ddr, 2, simplify = F) #Create all possible protein pairs between DR prots with CCDs
comb_signif_dt.ddr <- do.call("rbind", comb_signif.ddr) %>% as.data.frame() #Create data frame
colnames(comb_signif_dt.ddr) <- c("gene1","gene2") #Change column names

#Calculate cosine distances between each gene
cos.distance <- cosine(t(as.matrix(ddr_significant_CCDs_dcast))) #as.matrix
melt.dist.matrix <- reshape2::melt(cos.distance) #as.tibble
colnames(melt.dist.matrix) <- c("gene1", "gene2", "cos.distance") #Change column names

#Filter out redundant pairs (A-B & B-A): There is no directionality in these interactions
dist.matrix.int.biogrid <-comb_signif_dt.ddr %>% left_join(melt.dist.matrix)

#Add interaction data to all possible pairs
dist.int.partners.biogrid <- dist.matrix.int.biogrid %>% 
  left_join(ddr.ccd.physical.biogrid.rep.db) %>% #Add interaction data
  mutate(interaction = case_when(is.na(int_type) ~ FALSE, T ~ TRUE)) %>% #Does each pair interact or not?
  dplyr::group_by(gene1, gene2, cos.distance) %>% #Clean possible redundant pairs in BioGRID
  dplyr::summarise(interaction = case_when(sum(interaction) == 0 ~ F, T ~ T)) #Summarize phyisical interactions

#Summary numbers of interaction
summary_interacting_pairs <- dist.int.partners.biogrid %>% dplyr::group_by(interaction) %>% dplyr::summarise(c = n())

#Sample random pairs
sample_correlations.ddr.biogrid <- tibble() #Empty tibble for sampling
n.interacting.biogrid <- dist.int.partners.biogrid %>% filter(interaction == T) %>% nrow() #Number of physical interactions in our data

#Sampling loop to select random samples of equal size as interacting pairs
   for (i in c(1:1000)){ #n=1000
  set.seed(i)
  t <- dist.int.partners.biogrid %>% ungroup() %>% sample_n(n.interacting.biogrid) %>% mutate(seed =i)
  sample_correlations.ddr.biogrid <- sample_correlations.ddr.biogrid %>% bind_rows(t)
   }

#Create density matrices for plotting later
density.matrix.datapoints <- tibble()
for (i in c(1:1000)){
  seed.unique <- filter(sample_correlations.ddr.biogrid, seed == i) %>% pull(cos.distance)%>% as.numeric()
  dens.seed.unique <- density(seed.unique, from = -1, to = 1) %>% tidy() %>% mutate(seed = i) %>% mutate(round_x = round(x, digits = 2)) %>% dplyr::group_by(round_x) %>% dplyr::summarise(mean_y = mean(y), seed = i)
  density.matrix.datapoints <- density.matrix.datapoints %>% bind_rows(dens.seed.unique)
}

#Calculate mean and sd to plot in density plots for random effects
summarise.mean.sd_sample <- density.matrix.datapoints %>% dplyr::group_by(round_x) %>% dplyr::summarise(avg_y = mean(mean_y), sd_y = sd(mean_y), counts = n())

#Calculate real interaction density data
interacting.prots <- filter(dist.int.partners.biogrid, interaction == T)
real_density_plot <- density(interacting.prots %>% pull(cos.distance), from = -1, to = 1) %>% tidy()  %>% mutate(round_x = round(x, digits = 2)) %>% dplyr::group_by(round_x) %>% dplyr::summarise(mean_y = mean(y))
```

## Plot cosine similarities between proteins that interact.
```{r Fig_4A, warning=FALSE, message=FALSE,fig.width=2.75,fig.height=2.5}
ggplot(summarise.mean.sd_sample) + 
  geom_ribbon(aes(round_x, ymax = avg_y + sd_y, ymin = avg_y - sd_y), alpha = 0.2, fill = "grey40") +
  geom_line(aes(round_x,avg_y), color = "grey40", linetype = 2) + 
  geom_line(data = real_density_plot, aes(round_x,mean_y), color = "red") +
    theme_bw() + theme(legend.position = "top",
        panel.grid = element_blank()) + 
  ylab("Density") + xlab("Cosine distance")

#Mean value
mean_per_iteration <- sample_correlations.ddr.biogrid %>% 
  dplyr::group_by(seed) %>% 
  dplyr::summarise(mean_cosdist = mean(cos.distance))

pval <- t.test(mean_per_iteration$mean_cosdist, mu = 0.3641)

z_score <- (0.3641 - mean(mean_per_iteration$mean_cosdist)) /sd(mean_per_iteration$mean_cosdist)

p_value = 2*pnorm(abs(z_score), lower.tail=F)
  
#99CI
CI_99 <-  quantile(mean_per_iteration$mean_cosdist, c(0.025,0.975))

print(CI_99)

#Real_data mean
mean(interacting.prots %>% pull(cos.distance))

# Source data table
fig4a <- summarise.mean.sd_sample %>%
  left_join(real_density_plot) %>%
  select(x_similarity = round_x, y_realdata = mean_y, y_meanrandom = avg_y, y_sdrandom = sd_y) %>%
  mutate(figure = "Figure_4A")

write.table(fig4a, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4a_bottom.txt", row.names = F)
```
(A) Distributions of pairwise similarity scores for CCD patterns across the 25 chromatin features, between interacting proteins (red; 118 pairs) and between randomly picked protein pairs (mean ± s.d. of 1,000 draws of 118 random pairs). 

# Figure 4B
```{r Fig_4B, warning=FALSE, message=FALSE, fig.width = 3, fig.height=2}
#Figure 3B: Random mean distributions
sample_correlation_ddr_mean_biogrid <- sample_correlations.ddr.biogrid %>% dplyr::group_by(seed) %>% dplyr::summarise(m = mean(cos.distance))
normal_fit_data <- 
real_correlation_mean_biogrid <- dist.int.partners.biogrid %>% filter(interaction == T) %>% pull(cos.distance) %>% mean()


#pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/figure_4_Interaction_data/xv20230927_Fig4B_mean_distribution_changes.pdf", width = 3, height = 2)
ggplot(sample_correlation_ddr_mean_biogrid) + 
  geom_histogram(aes(m), bins = 100) + 
  geom_linerange(aes(real_correlation_mean_biogrid,ymin = 0, ymax = 25),color = "red") + 
  coord_cartesian(expand = T, xlim = c(0,0.4)) + 
  theme_bw() +
  theme(panel.grid = element_blank())
#dev.off()

# Source data table
fig4b <- sample_correlation_ddr_mean_biogrid %>%
  select(x_meansimilarity = m, sampling_seed = seed) %>%
  mutate(figure = "Figure_4B")

write.table(fig4b, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4b.txt", row.names = F)
```
(B) Mean similarity score of 118 interacting protein pairs (red line) compared to the distribution of mean similarity scores of 1,000 random draws as in (A) (grey histogram), indicating that high similarities of CCD patterns of interacting protein pairs cannot be explained by random chance (two-sided p-value = 0.0006).

# Figure 4C
## Data processing: Plot showing number of cliques
```{r prep_fig_4C, warning=FALSE, message=FALSE}
#CCD data visualization as UMAP
set.seed(4) #Get consistend umap
umap.hits <- umap(ddr_significant_CCDs_dcast) #umap algorithm {umap package}
colnames(umap.hits$layout) <- c("A","B") #Change colnames
cluster_umap <- cluster_tibble %>% separate(cluster_name, into = c("pathway", "cluster_b"), sep = "_") #separate cluster data into pathway and interaction type
umap.hits.plot <- umap.hits$layout %>% as.data.frame() %>% rownames_to_column(var = "gene") %>% left_join(cluster_umap) #Import data from previous line
umap.hits.plot[which(umap.hits.plot$pathway == "N-synergy"),]$cluster_b <- 3 #Manually change nadd N-synergy clusters
umap.hits.plot[which(umap.hits.plot$gene == "CHAF1A"),]$cluster_b <- 1

#Find highest order cliques in the data {igraph package}
igraph_interaction_biogrid <- graph_from_data_frame(interacting.prots) #Create igraph object with interaction data

#Find maximum cliques in the data
maximum_cliques_data <- largest_cliques(igraph_interaction_biogrid) #Find largest cliques in the network (In this case five 4-node cliques)

#Find plot summary of all cliques in the data
all_max_cliques <- max_cliques(igraph_interaction_biogrid)

#Extract element numbers in the dataset
summary_max_cliques <- tibble(clique = NA, elements = NA)
for (i in 1: length(all_max_cliques)) {
  summary_max_cliques <- summary_max_cliques %>% add_row(clique = i, elements = length(names(all_max_cliques[[i]])))  
}
#Summary plot
ggplot(summary_max_cliques) +
  geom_bar(aes(elements)) +
  theme_bw() + xlab("Clique complexity")

```

## Plotting umap visualization (This figure was further edited in Illustrator)
```{r Fig_4C_plot, warning=FALSE, message=FALSE}
# Clique data table: Summarizing maximum_cliques_data output

all_cliques <- tibble(gene = c("ATM","FANCD2","MDC1","TOPBP1","FANCM","FANCF","FANCG","FANCD2","BLM","RMI2","FANCD2","FANCM"), clique = c(rep(1,4), rep(2,4),rep(3,4)), type = c(rep("ATM",4), rep("FA",4), rep("mixed",4))) %>% left_join(umap.hits.plot)
#Plot with umap with 4-node cliques
#pdf(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/figure_4_Interaction_data/xv20240424_Fig3C_umap_visualization.pdf", height=4, width = 6)
ggplot() +
     geom_polygon(data = all_cliques, aes(x = A, y = B, group = clique, fill = type), alpha = 0.4) +
     stat_density2d(data = umap.hits.plot, geom = "polygon", aes(A,B, fill = cluster_b, alpha = ..level..), bins = 8) + xlim(-4,5) + ylim(-4,5) +
     theme_bw() + geom_point(data = umap.hits.plot, aes(A,B, color = gene %in% all_cliques$gene, shape = pathway), size = 3) + 
     theme(panel.grid = element_blank()) + 
     scale_fill_manual(values = c("#9E75B2","#F088B6","#FAC075","#1D71B8","#E94E1B","#B2B2B2")) + scale_alpha(range = c(0.15,0.3)) +
     geom_text_repel(data = subset(umap.hits.plot, gene %in% all_cliques$gene), aes(A,B,label = gene), force = 40) + 
     ylab("UMAP2") + xlab("UMAP1")
#dev.off()
# Source data table
fig4c <- umap.hits.plot %>% 
  left_join(all_cliques) %>%
  select(id_gene = gene, x_umap1 = A, y_umap2 = B, symbol_pathway = pathway, background_cluster = cluster_b, highlight_clique = clique, description_clique = type) %>%
  mutate(figure = "Figure_4C")

write.table(fig4c, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4c.txt", row.names = F)
```
(C) Uniform Manifold Approximation and Projection (UMAP) visualization of proteins with CCDs. Each dot represents a protein, with the shape indicating the type of synergy. Color clouds show the major chromatin state that explains each CCD. Three ‘cliques’ of four interacting proteins are shown as colored quadrangles. Proteins shared between multiple cliques are marked by concentric circles with the color of each clique.

# Data processing for figures 4D-H
```{r prep_figD_H, warning=FALSE, message=FALSE}
#Select genes in cliques
cosine.cliques.dcast <- significant_proteins %>% filter(gene %in% all_cliques$gene) %>% reshape2::dcast(chrom_feature ~ gene, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "chrom_feature")

#Generate cosine distance matrix
cosine.d.cliques <- cosine(as.matrix(cosine.cliques.dcast)) %>% reshape2::melt()
```

# Figure 4D
```{r Fig_3_D, warning=FALSE, message=FALSE}
#Clique ATMi
ggplot(cosine.d.cliques %>% filter(Var1 %in% c("ATM","FANCD2","TOPBP1","MDC1") & Var2 %in% c("ATM","FANCD2","TOPBP1","MDC1"))) + 
     geom_tile(aes(fct_relevel(Var1,c("ATM","FANCD2","TOPBP1","MDC1")), fct_relevel(Var2,c("ATM","FANCD2","TOPBP1","MDC1")), fill = value))+ 
     theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + 
     coord_fixed(expand = F,ratio = 1) + scale_fill_gradient2(low = "#af8dc3", mid = "#f7f7f7", high = "#7fbf7b", limits = c(-1,1)) + 
     geom_text(aes(fct_relevel(Var1,c("ATM","FANCD2","TOPBP1","MDC1")), fct_relevel(Var2,c("ATM","FANCD2","TOPBP1","MDC1")),label = round(value, digits = 2))) + theme(axis.title = element_blank())

# Source data table
fig4d <- cosine.d.cliques %>% 
  filter(Var1 %in% c("ATM","FANCD2","TOPBP1","MDC1") & 
           Var2 %in% c("ATM","FANCD2","TOPBP1","MDC1")) %>% 
  select(x_gene = Var1, y_gene = Var2, fill_similarity = value) %>%
  mutate(figure = "Figure_4D")

write.table(fig4d, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4d.txt", row.names = F)
```

# Figure 4E
```{r Fig_3_E, warning=FALSE, message=FALSE}
#Clique FA
ggplot(cosine.d.cliques %>% filter(Var1 %in% c("FANCD2","FANCM","FANCG","FANCF") & Var2 %in% c("FANCD2","FANCM","FANCG","FANCF"))) + 
     geom_tile(aes(fct_relevel(Var1,c("FANCD2","FANCM","FANCG","FANCF")), fct_relevel(Var2,c("FANCD2","FANCM","FANCG","FANCF")), fill = value))+ 
     theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + 
     coord_fixed(expand = F,ratio = 1) + scale_fill_gradient2(low = "#af8dc3", mid = "#f7f7f7", high = "#7fbf7b", limits = c(-1,1)) + 
     geom_text(aes(fct_relevel(Var1,c("FANCD2","FANCM","FANCG","FANCF")), fct_relevel(Var2,c("FANCD2","FANCM","FANCG","FANCF")),label = round(value, digits = 2))) + theme(axis.title = element_blank())

# Source data table
fig4e <- cosine.d.cliques %>% 
  filter(Var1 %in% c("FANCD2","FANCM","FANCG","FANCF") & 
           Var2 %in% c("FANCD2","FANCM","FANCG","FANCF")) %>% 
  select(x_gene = Var1, y_gene = Var2, fill_similarity = value) %>%
  mutate(figure = "Figure_4E")

write.table(fig4e, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4e.txt", row.names = F)
```

# Figure 4F
```{r Fig_4F, warning=FALSE, message=FALSE}
#Clique Mixed
ggplot(cosine.d.cliques %>% filter(Var1 %in% c("FANCD2","FANCM","BLM","RMI2") & Var2 %in% c("FANCD2","FANCM","BLM","RMI2"))) + 
     geom_tile(aes(fct_relevel(Var1,c("FANCD2","FANCM","BLM","RMI2")), fct_relevel(Var2,c("FANCD2","FANCM","BLM","RMI2")), fill = value))+ 
     theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + 
     coord_fixed(expand = F,ratio = 1) + scale_fill_gradient2(low = "#af8dc3", mid = "#f7f7f7", high = "#7fbf7b", limits = c(-1,1)) + 
     geom_text(aes(fct_relevel(Var1,c("FANCD2","FANCM","BLM","RMI2")), fct_relevel(Var2,c("FANCD2","FANCM","BLM","RMI2")),label = round(value, digits = 2))) + theme(axis.title = element_blank())

# Source data table
fig4f <- cosine.d.cliques %>% 
  filter(Var1 %in% c("FANCD2","FANCM","BLM","RMI2") & Var2 %in% c("FANCD2","FANCM","BLM","RMI2")) %>% 
  select(x_gene = Var1, y_gene = Var2, fill_similarity = value) %>%
  mutate(figure = "Figure_4F")

write.table(fig4f, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4f.txt", row.names = F)
```
(D-F) CCD similarity score matrix of proteins in (D) ATM clique, (E) FA clique and (F) mixed clique. 

# Data processing for Figures 4G-I
```{r prep_figG_i, warning=FALSE, message=FALSE}

cell_lines <- c("PRO","DEF","K562")
#
cosine_similarity_cell_line <- map_dfr(cell_lines, function(x) {
  one_cell_line <- cell_lines_CCD %>% 
    anti_join(filter_out_chromatin) %>%
    na.omit() %>%
    filter(cell_line == x)
  #Create dcast for DEF CCD table
  one_cell_line_CCD_dcast <- one_cell_line %>% 
    reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% na.omit() %>%
    column_to_rownames(var = "gene")
  #Calculate cosine distances between each gene
  one_cell_cos.distance <- cosine(t(as.matrix(one_cell_line_CCD_dcast))) #as.matrix
  one_cell_melt.dist.matrix <- reshape2::melt(one_cell_cos.distance) %>% mutate(cell_line = x)#as.tibble
})

#Clique summary table
all_cliques <- tibble(gene = c("ATM","FANCD2","MDC1","TOPBP1","FANCM","FANCF","FANCG","FANCD2","BLM","RMI2","FANCD2","FANCM"), clique = c(rep(1,4), rep(2,4),rep(3,4)), type = c(rep("ATM",4), rep("FA",4), rep("mixed",4)))

```

# Figure 4G 
```{r Fig_4G, warning=FALSE, message=FALSE, fig.width=2, fig.height=3}
#Plot diferent
clique1_data <- cosine_similarity_cell_line %>% 
  filter(Var1 %in% c("ATM","FANCD2","TOPBP1","MDC1") &
         Var2 %in% c("ATM","FANCD2","TOPBP1","MDC1") &
         Var1 != Var2)

#Similarity scores across cell lines
#pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/figure_4_Interaction_data/xv20230927_Fig4E_clique_1_conservation_other_cell_lines.pdf", width = 2, height = 3)
ggplot(clique1_data %>% select(cell_line, value) %>% distinct(), aes(fct_relevel(cell_line,c("K562","PRO","DEF")), value)) +
  stat_summary(geom = "col", aes(fill = cell_line)) +
  geom_quasirandom() + 
  theme_bw() +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482")) +
  coord_cartesian(ylim = c(-1,1)) +
  theme(axis.title.x = element_blank(), legend.position = "none", panel.grid = element_blank()) + ylab("Cosine similarity score")
#dev.off()

# Source data
fig4d <- umap.hits.plot %>% 
  left_join(all_cliques) %>%
  select(id_gene = gene, x_umap1 = A, y_umap2 = B, symbol_pathway = pathway, background_cluster = cluster_b, highlight_clique = clique, description_clique = type) %>%
  mutate(figure = "Figure_4D")

write.table(fig4c, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/source_data/xv20240417_sourcedata_fig4c.txt", row.names = F)

```

# Figure 4H
```{r Fig_4H, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=2.5}
#Plot diferent
clique3_data <- cosine_similarity_cell_line %>% 
  filter(Var1 %in% c("FANCD2","FANCM","FANCG","FANCF") &
         Var2 %in% c("FANCD2","FANCM","FANCG","FANCF") &
         Var1 != Var2)

#pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/figure_4_Interaction_data/xv20231027_Fig4F_clique_FA1_conservation_other_cell_lines.pdf", width = 2, height = 3)
ggplot(clique3_data %>% select(cell_line, value) %>% distinct(), aes(fct_relevel(cell_line,c("K562","PRO","DEF")), value)) +
  stat_summary(geom = "col", aes(fill = cell_line)) +
  geom_quasirandom() + 
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482")) +
  coord_cartesian(ylim = c(-1,1)) +
  geom_hline(yintercept = 0, linetype =2) +
  theme(axis.title.x = element_blank(), legend.position = "none",
        panel.grid = element_blank()) + ylab("Cosine similarity score")
#dev.off()

```

# Figure 4I
```{r Fig_4I, warning=FALSE, message=FALSE,}
clique4_data <- cosine_similarity_cell_line %>% 
  filter(Var1 %in% c("FANCD2","FANCM","BLM","RMI2") &
         Var2 %in% c("FANCD2","FANCM","BLM","RMI2") &
         Var1 != Var2)

#pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/figure_4_Interaction_data/xv20231027_Fig4Fb_mixed_clique_conservation_other_cell_lines.pdf", width = 2, height = 3)
ggplot(clique4_data %>% select(cell_line, value) %>% distinct(), aes(fct_relevel(cell_line,c("K562","PRO","DEF")), value)) +
  stat_summary(geom = "col", aes(fill = cell_line)) +
  geom_quasirandom() + 
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482")) +
  coord_cartesian(ylim = c(-1,1)) +
  geom_hline(yintercept = 0, linetype =2) +
  theme(axis.title.x = element_blank(), legend.position = "none",
        panel.grid = element_blank()) + ylab("Cosine similarity score")
#dev.off()

```
(G-I) Similarity scores of proteins in (G) ATM clique, (H) FA clique and (I) mixed clique in K562, RPE-1 p53KO and p53/BRCA1dKO cells.

# Data processing for Figures 4J-N
```{r, warning=FALSE, message=FALSE}
#Import interesting genes
cosine.highlights.genes <- table_S3_CCDs %>% filter(gene %in% c("ATM", "CHEK2","PRKDC")) %>% na.omit()
#Import inhibitor data
cosine.highlights.drugs <- CCD_inhibitors_data %>% filter(drug %in% c("ATMi", "DNAPKi")) %>% dplyr::select(gene = drug, chrom_feature,CCD_synergy_score)
#Bind all data
cosine.highlights.all <- bind_rows(cosine.highlights.genes, cosine.highlights.drugs) 

#dcast
cosine.highlights.dcast <- cosine.highlights.all %>% reshape2::dcast(chrom_feature ~ gene, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "chrom_feature")

#Cosine distance
cosine.d.highlingts <- cosine(as.matrix(cosine.highlights.dcast)) %>% melt()
```

# Figure 4J
```{r, warning=FALSE, message=FALSE}
#dcast epistatic interactions (PARP proteins)
ATM.proteins.CCD <- table_S3_CCDs %>% filter(gene %in% c("ATM", "CHEK2"))

#dcast epistatic interactions (PARP inhibitors)
ATM_inhibitors <- CCD_inhibitors_data %>% filter(drug %in% c("ATMi"))

#Cosine distance between all of them
#Prepare for binding
ATM.inhibitor.sel <- ATM_inhibitors %>% dplyr::select(gene = drug, chrom_feature, CCD_synergy_score)
ATM.combined <- ATM.proteins.CCD %>% dplyr::select(gene, chrom_feature, CCD_synergy_score) %>% bind_rows(ATM.inhibitor.sel) 
ATM.dcast <- ATM.combined %>% reshape2::dcast(chrom_feature ~ gene, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "chrom_feature")
cosine.d.ATM<- cosine(as.matrix(ATM.dcast))

#PARP combined heatmap
ggplot(ATM.combined) + 
  geom_tile(aes(fct_relevel(gene, c("ATM","ATMi","CHEK2")), fct_relevel(chrom_feature,heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.1,0, 0.1), limits = c(-0.15,0.15), oob = squish)  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")
```

#Figure 4K
```{r, warning=FALSE, message=FALSE}
#FA proteins order
FA_prots <- c("FANCM","FAAP24","FANCG","FANCA","FANCF","FANCC","FANCB","FANCL","FANCD2","FANCI")

# Filter FANC proteins
FA.proteins.CCD <- table_S3_CCDs %>% filter(gene %in% FA_prots) 

# dCast FANC proteins
FA.proteins.dcast <- FA.proteins.CCD %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Cosine distance between all of them
cosine.d.FANC <- cosine(as.matrix(t(FA.proteins.dcast)))

#Plot heatmap
ggplot(FA.proteins.CCD) + 
  geom_tile(aes(fct_relevel(gene,FA_prots), fct_relevel(chrom_feature,heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e",  breaks = c(-0.1,0, 0.1), limits = c(-0.15,0.15))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

```


#Figure 4L
```{r, warning=FALSE, message=FALSE}
#dcast epistatic interactions (PARP proteins)
SMC5_compl <- table_S3_CCDs %>% filter(gene %in% c("SMC5","NSMCE1","NSMCE3","NSMCE2","NSMCE4","SMC6")) %>% na.omit()

SMC5.dcast <- SMC5_compl %>% reshape2::dcast(chrom_feature ~ gene, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "chrom_feature")
cosine.d.SMC5<- cosine(as.matrix(SMC5.dcast))

#PARP combined heatmap
ggplot(SMC5_compl) + 
  geom_tile(aes(fct_relevel(gene,c("NSMCE1","NSMCE3","SMC5","SMC6")), fct_relevel(chrom_feature,heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.1,0, 0.1), limits = c(-0.15,0.15), oob = squish)  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")
```

# Figure 4M
```{r, warning=FALSE, message=FALSE}
#dcast epistatic interactions (PARP proteins)
DNAPK.proteins.CCD <- table_S3_CCDs %>% filter(gene == "PRKDC") %>% dplyr::select(gene, chrom_feature, CCD_synergy_score)
DNAPK.inhibitors <- CCD_inhibitors_data %>% filter(drug == "DNAPKi") %>% dplyr::select(gene = drug, chrom_feature, CCD_synergy_score)

# Bind rows
DNAPK.perturbations <- bind_rows(DNAPK.proteins.CCD, DNAPK.inhibitors)

# DNAPK inhibition
ggplot(DNAPK.perturbations) + 
    geom_tile(aes(fct_relevel(gene, c("PRKDC","DNAPKi")), fct_relevel(chrom_feature,heatmap.chromatin.order), fill = CCD_synergy_score)) +
    scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), breaks = c(-0.1,0, 0.1), oob = squish)  + 
    coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

```

# Figure 4N
```{r, warning=FALSE, message=FALSE}
#BRC1A complex proteins
BRCA1A_complex <- c("BRCC3","UIMC1","ABRAXAS1","BABAM2")

# Separate data based on the significant IPRs that they have (based on z-score)
UIMC1_BRCC3<- table_S3_CCDs %>% filter(gene %in% BRCA1A_complex)  %>% na.omit()

# heatmap: All heatmaps together
ggplot(UIMC1_BRCC3) + 
  geom_tile(aes(fct_relevel(gene, BRCA1A_complex),fct_relevel(chrom_feature, heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

```

# Export Table S8
This table includes the protein pairs that interact with each other.
```{r, warning=FALSE, message=FALSE}
setwd(in.dir)
#Prepare the file to export
export_protein_protein_table <- ddr.ccd.physical.biogrid.rep.db %>% left_join(melt.dist.matrix)

#Table S4
write.xlsx(export_protein_protein_table, file = "data/supplementary_tables/xv20220913_Table_S8_BioGRID_interaction_CCD.xlsx")
```
