---
title: "xv20211230_4_epistasis_model"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Script to generate figure #2 in the paper. This figure illustrates DDR chromatin context dependencies (CCDs). Figure outline:

A- Mock data to describe M-, N- and no-synergies
B- 


This script generates the plots for figure 4 and S5

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
    #write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    readRDS(correct_file)
    #write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  readRDS(correct_file)
  #write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
}
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(dendextend)
library(ggdendro)
library(grDevices)
library(pheatmap)
library(reshape2)
library(tibble)
library(ggpubr)
library(scales)
```

# Import data tables
```{r}
#setwd
setwd(in.dir)
# Individual data point (intermediate step)
step2.epistasis <- readRDS_proof("data/processed_data/CCD_analysis","log2_MMEJNHEJ_differentials_chromatin_KO")

#DNA repair CCDs
table_S3_CCDs  <- readRDS_proof("data/processed_data","Table_S3")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

```

# Figure 2A: Mock examples
```{r}
#Create mock examples
mock.slopes <- tibble(z_score = seq(-3,3,0.25), A = z_score*0 + 0.6,B = z_score*0.1 +0.6, C = z_score*-0.1 - 0.6) %>% reshape2::melt(id.vars = "z_score")

#Plot mock examples
ggplot(mock.slopes, aes(z_score,value,color = variable)) + theme_bw() + 
  geom_smooth() +
  geom_vline(xintercept = 0, linetype = 2) + 
  geom_hline(yintercept = 0, linetype = 2) + 
  ylab("log2(fold change NHEJ:MMEJ)") + xlab("Feature signal (z-score)") + 
  theme(legend.position = "none") + coord_cartesian(ylim= c(-1.5,1.5)) + scale_color_manual(values = c("black","#01665e","#8c510a"))
```

# Figure 2B-D: Examples
  #2B: Favors MMEJ - RAD50 & LMNB1
```{r}
# Plot RAD50 with LMNB1
ggplot(step2.epistasis %>% 
         dplyr::filter(gene %in% c("RAD50")),
       aes(LMNB1,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = -0.114),method = "lm") +
  stat_cor(label.x = -1.5, label.y = -0.7) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(RAD50 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#2C: No interaction - MDC1 and H2AFZ
```{r}
# Plot MDC1 and H2AFZ
ggplot(step2.epistasis %>% 
         filter(gene == "MDC1"),
       aes(H2AFZ,mean.log2foldchange)) + 
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "black") +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  stat_cor()+
  theme_bw() + ylab("log2(MDC1 MMEJ:NHEJ balance)") 
```

# Figure 4D: Matrix for hits
```{r}
# Separate data based on the significant IPRs that they have (based on z-score)
significant_proteins <- table_S3_CCDs %>% filter(CCD_model_p_adj < 0.05)

# dcaast with selected
slope.prot.features.dcast.hits <- significant_proteins %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap <- pheatmap(t(slope.prot.features.dcast.hits), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 4)
signif_prots.gene.order <- rev(rownames(slope.prot.features.dcast.hits[signif_prots_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_proteins) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order),fct_relevel(chrom_feature, heatmap.chromatin.order), fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

# Dendrogram
ggdendrogram(rev(signif_prots_heatmap$tree_col))
```

#Output chunk
```{r}
setwd(in.dir)
#Extract 4 clusters
cluster_data <- cutree(signif_prots_heatmap$tree_col, k=4) 
cluster_tibble <- tibble(gene = names(cluster_data), cluster = cluster_data) %>% mutate(cluster_name = case_when(cluster == "1" ~ "M-synergy_2", cluster == "2" ~ "M-synergy_1", cluster == "3" ~ "N-synergy", cluster == "4" ~"M-synergy_3"))

#cluster summary
cluster_summary <- cluster_tibble %>% dplyr::group_by(cluster_name) %>% dplyr::summarise(c = n())

#Export cluster data to use it in file 4
saveRDS_proof(cluster_tibble, "data/processed_data/xv20220802_DR_CCD_cluster_data.rds")
```
