---
title: "xv20230329_RPE1_mapping_clone"
author: "Xabier Vergara"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(rtracklayer)
library(caTools)
```


#Load chromatin info of the IPRs in RPE cells
```{r}
setwd("/DATA/projects/DSBrepair/data/xv20230608_K562_PMA_Ayoub/IPR_location/")
# Load chip data
dam_files <- list.files(path = "dam", pattern = "2000", full.names = T)

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_clone5_megaK <- chromatin_features_pools_dam %>% 
  reshape2::melt() %>% 
  mutate(treatment = case_when(grepl("DMSO", variable) ~ "DMSO",
                               grepl("PMA", variable) ~ "PMA"),
         antibody = case_when(grepl("H3K9me3", variable) ~ "H3K9me3",
                              grepl("H3K9me2", variable) ~ "H3K9me2",
                              grepl("LMNB2", variable) ~ "LMNB2",
                              grepl("H3K27me3", variable) ~ "H3K27me3")) %>%
  dplyr::select("barcode" = name,treatment, antibody, "IPR_score" = value)


#Load mean values as well
#Make a dataframe with dam files
chromatin_features_pools_dam_mean <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_clone5_megaK_mean <- chromatin_features_pools_dam_mean %>% 
  reshape2::melt() %>% 
  mutate(treatment = case_when(grepl("DMSO", variable) ~ "DMSO",
                               grepl("PMA", variable) ~ "PMA"),
         antibody = case_when(grepl("H3K9me3", variable) ~ "H3K9me3",
                              grepl("H3K9me2", variable) ~ "H3K9me2",
                              grepl("LMNB2", variable) ~ "LMNB2",
                              grepl("H3K27me3", variable) ~ "H3K27me3")) %>%
  dplyr::select("barcode" = name,treatment, antibody, "IPR_score" = value)
```
#Load chromatin info of the IPRs in K562 pools
```{r}
setwd("/DATA/projects/DSBrepair/data/xv20230608_K562_PMA_Ayoub/Pool_IPR_location/")
# Load chip data
dam_files <- list.files(path = "dam", pattern = "2000", full.names = T)

#Make a dataframe with dam files
chromatin_features_RSTP2_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_RSTP2_pools_megaK <- chromatin_features_RSTP2_pools_dam %>% 
  reshape2::melt() %>% 
  mutate(treatment = case_when(grepl("DMSO", variable) ~ "DMSO",
                               grepl("PMA", variable) ~ "PMA"),
         antibody = case_when(grepl("H3K9me3", variable) ~ "H3K9me3",
                              grepl("H3K9me2", variable) ~ "H3K9me2",
                              grepl("LMNB2", variable) ~ "LMNB2",
                              grepl("H3K27me3", variable) ~ "H3K27me3")) %>%
  dplyr::select("barcode" = name,treatment, antibody, "IPR_score" = value)


#Load mean values as well
#Make a dataframe with dam files
chromatin_features_RSTP2_pools_dam_mean <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_RSTP2_pools_megaK_mean <- chromatin_features_RSTP2_pools_dam_mean %>% 
  reshape2::melt() %>% 
  mutate(treatment = case_when(grepl("DMSO", variable) ~ "DMSO",
                               grepl("PMA", variable) ~ "PMA"),
         antibody = case_when(grepl("H3K9me3", variable) ~ "H3K9me3",
                              grepl("H3K9me2", variable) ~ "H3K9me2",
                              grepl("LMNB2", variable) ~ "LMNB2",
                              grepl("H3K27me3", variable) ~ "H3K27me3")) %>%
  dplyr::select("barcode" = name,treatment, antibody, "IPR_score" = value)
```
# Check replicate correlations
```{r}
#list_files
setwd("/DATA/projects/DSBrepair/data/xv20230608_K562_PMA_Ayoub/results/tracks/normalized/bin-20kb/")
file_list_rep <- list.files(pattern = "K562.*20kb.bw")

#load files
PMA_replicate_tracks <- map_dfr(file_list_rep, function(x){
  import(x) %>% as_tibble() %>% 
    mutate(file = x, smooth_score = runmean(score, k = 5)) %>% separate(file, into = c("rep","cell","treatment","antibody","extensionA","extensionB"))
})

PMA_rep_tracks_dcast <- PMA_replicate_tracks %>% reshape2::dcast(seqnames+start+end+antibody+treatment~rep, value.var = "score")

#Plot correlation (20kb bins)
ggplot(PMA_rep_tracks_dcast, aes(E2260,E2267)) +
  geom_point(alpha = 0.05, size = 0.1) + 
  ggpubr::stat_cor() + 
  geom_smooth(method = "lm") + 
  facet_grid(treatment~antibody) +
  theme_bw() + coord_fixed()

PMA_rep_tracks_dcast_cor <- PMA_replicate_tracks %>% reshape2::dcast(seqnames+start+end~rep+treatment+antibody, value.var = "score") %>% dplyr::select(-seqnames, -start, -end) 


correlation_samples <- cor(PMA_rep_tracks_dcast_cor,use = "complete.obs", method = "spearman")
pheatmap(correlation_samples)
```

# Load Tracks genomewide
```{r}
#list_files
setwd("/DATA/projects/DSBrepair/data/xv20230608_K562_PMA_Ayoub/results/tracks/normalized/bin-20kb/")
file_list <- list.files(pattern = "K562.*-combined")

#load files
PMA_diff_tracks <- map_dfr(file_list, function(x){
  import(x) %>% as_tibble() %>% 
    mutate(file = x, smooth_score = runmean(score, k = 3)) %>% separate(file, into = c("cell","treatment","antibody","extension"), sep = "_")
})

PMA_diff_tracks_dcast <- PMA_diff_tracks %>% reshape2::dcast(seqnames+start+end+antibody~treatment, value.var = "score")

#Plot correlation (20kb bins)
ggplot(PMA_diff_tracks_dcast %>% filter(antibody != "H3K9me2"), aes(DMSO,PMA)) +
  geom_point(alpha = 0.05, size = 0.1) + 
  ggpubr::stat_cor() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~antibody, ncol = 1) +
  theme_bw() + coord_fixed()

#Plot density plot
ggplot(PMA_diff_tracks %>% filter(antibody != "H3K9me2")) +
  geom_density(aes(score, fill = treatment, color = treatment), alpha = 0.25) +
  facet_wrap(~antibody) +
  theme_bw()

```

#Plot genomic coordinates

#LMNB2
```{r}
  ggplot(PMA_diff_tracks %>% 
         filter(antibody %in% c("LMNB2","H3K9me3") & seqnames == "chr3" & start < 90000000)) +
  geom_col(aes(start,smooth_score, color = antibody)) + 
  facet_wrap(~antibody+treatment,ncol = 1) + 
  theme_bw() + 
  geom_hline(yintercept =0, linetype =2)+
  theme(legend.position = "none", panel.grid.minor = element_blank()) + 
  xlab("chr3")
```



#H3K27me3
```{r}
  ggplot(PMA_diff_tracks %>% 
         filter(antibody == "H3K27me3" & seqnames == "chr3" & start < 9000000)) +
  geom_col(aes(start,smooth_score), color = "grey30") + 
  facet_wrap(~treatment,ncol = 1) + 
  theme_bw() + 
  theme(legend.position = "top") + 
  xlab("chr3")
```


#Chromatin features changes in clone #5 IPRs
```{r}
#Load screening data
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS') %>%
  filter(binsize == 2000) %>% select("barcode" = ID, H3K9me3, "LMNB2" = LMNB1, H3K27me3) %>% reshape2::melt() %>% filter()
colnames(clone5_z.score_chrom_tib) <- c("barcode","antibody","reference")
#Prepare for plotting
chrom_features_clone5 <- chromatin_features_clone5_megaK %>% reshape2::dcast(barcode + antibody ~ treatment) %>% left_join(clone5_z.score_chrom_tib)

ggplot(chrom_features_clone5 %>% filter(antibody != "H3K9me2"), aes(DMSO, reference)) + 
  geom_point() + 
  ggpubr::stat_cor() + 
  geom_smooth(method = "lm") +
  facet_wrap(~ antibody) +
  coord_fixed() +
  theme_bw()
```
#CONCLUSION: Once more we show that reporter take the expected chromatin type. This time, even if the technique used to profile chromatin features and the presence of DMSO didn't alter the chromatin landscape.

#Chromatin features changes in RSTP2 pools
```{r}
#Load screening data
RSTP2_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS') %>%
  filter(binsize == 2000) %>% select("barcode" = ID, H3K9me3, "LMNB2" = LMNB1, H3K27me3) %>% reshape2::melt() %>% filter()
colnames(RSTP2_z.score_chrom_tib) <- c("barcode","antibody","reference")
#Prepare for plotting
chrom_features_RSTP2_pools <- chromatin_features_RSTP2_pools_megaK %>% reshape2::dcast(barcode + antibody ~ treatment) %>% left_join(RSTP2_z.score_chrom_tib)

ggplot(chrom_features_RSTP2_pools %>% filter(antibody != "H3K9me2"), aes(DMSO, reference)) + 
  geom_point() + 
  ggpubr::stat_cor() + 
  geom_smooth(method = "lm") +
  facet_wrap(~ antibody) +
  coord_fixed() +
  theme_bw()

#Prepare for plotting (with mean)
chrom_features_RSTP2_pools_mean <- chromatin_features_RSTP2_pools_megaK_mean %>% reshape2::dcast(barcode + antibody ~ treatment) %>% left_join(RSTP2_z.score_chrom_tib)

ggplot(chrom_features_RSTP2_pools_mean %>% filter(antibody != "H3K9me2"), aes(DMSO, reference)) + 
  geom_point() + 
  ggpubr::stat_cor() + 
  geom_smooth(method = "lm") +
  facet_wrap(~ antibody) +
  coord_fixed() +
  theme_bw()
```


#Change between treated and un-treated
```{r}
ggplot(chrom_features_clone5 %>% filter(antibody %in% c("H3K9me3","LMNB2")), aes(DMSO, PMA)) + 
  geom_point() + 
  ggpubr::stat_cor(label.x = -1, label.y = -1.75) + 
  geom_smooth(method = "lm",se = F) +
  facet_wrap(~ antibody, ncol = 1) +
  coord_fixed() +
  theme_bw() +
  geom_abline(linetype = 2)

#Plot with mean values before z_score normalization
chrom_values_mean <-  chromatin_features_clone5_megaK_mean %>% reshape2::dcast(barcode + antibody ~ treatment)

ggplot(chrom_values_mean %>% filter(antibody %in% c("H3K9me3","LMNB2")), aes(DMSO, PMA)) + 
  geom_point() + 
  ggpubr::stat_cor(label.x = -1, label.y = -1.75) + 
  geom_smooth(method = "lm",se = F) +
  facet_wrap(~ antibody, ncol = 1) +
  coord_fixed() +
  theme_bw() +
  geom_abline(linetype = 2)


```
# CCONCLUSION: Correlation between PMA and DMSO samples are there. LMNB2 signal changes the most globally, we observe the decrease of dynamic range. Still, there are some reporters that deviate from the global trend, suggesting that these change their chromatin state above the general effects. H3K9me3 does not change that much and general. H3K27me3 has some sites that are changed quite some, but the trend stays close to the diagonal.


#Change between treated and un-treated in RSTP2_pools
```{r}
ggplot(chrom_features_RSTP2_pools %>% filter(antibody %in% c("H3K9me3","LMNB2")), aes(DMSO, PMA)) + 
  geom_point() + 
  ggpubr::stat_cor(label.x = -1, label.y = -1.75) + 
  geom_smooth(method = "lm",se = F) +
  facet_wrap(~ antibody, ncol = 1) +
  coord_fixed() +
  theme_bw() +
  geom_abline(linetype = 2)

#Plot with mean values before z_score normalization
chrom_values_mean_RSTP2 <-  chromatin_features_RSTP2_pools_megaK_mean %>% reshape2::dcast(barcode + antibody ~ treatment)

ggplot(chrom_values_mean_RSTP2 %>% filter(antibody %in% c("H3K9me3","LMNB2","H3K27me3")), aes(DMSO, PMA)) + 
  geom_point() + 
  ggpubr::stat_cor(label.x = -1, label.y = -1.75) + 
  geom_smooth(method = "lm",se = F) +
  facet_wrap(~ antibody, ncol = 1) +
  coord_fixed() +
  theme_bw() +
  geom_abline(linetype = 2)


```


#Quantify changes in the reporters: Residual (deviation from the genomewide trend)
```{r}
antibodies_unique <- unique(chrom_features_RSTP2_pools$antibody)
linear_model_pool <- map_dfr(antibodies_unique, function(x) {
  tib <- chrom_features_RSTP2_pools %>% filter(antibody == x)
  res <- resid(lm(PMA ~ DMSO, tib))
  tib %>% mutate(residual = res)
})


ggplot(linear_model_pool %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,residual, barcode = barcode)) + 
  facet_wrap(~antibody) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  theme_bw() + coord_fixed()


```
#In clone 5
```{r}
antibodies_unique <- unique(chrom_features_clone5$antibody)
linear_model_clone <- map_dfr(antibodies_unique, function(x) {
  tib <- chrom_features_clone5 %>% filter(antibody == x)
  res <- resid(lm(PMA ~ DMSO, tib))
  tib %>% mutate(residual = res)
})


ggplot(linear_model_clone %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,residual, barcode = barcode)) + 
  facet_wrap(~antibody) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  theme_bw() + coord_fixed()

#Combined plot
ggplot(linear_model_pool %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,residual)) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  geom_point(data= linear_model_clone %>% filter(antibody != "H3K9me2")
             ,aes(DMSO,residual, barcode = barcode), color = "red") + 
  facet_wrap(~antibody) +
  theme_bw() + coord_fixed()


```

#Quantify changes in the reporters: log2 foldchange (general deviation)
```{r}
log2FC_feature_pools <- chrom_features_RSTP2_pools %>% mutate(deltaZ = PMA - DMSO)

ggplot(log2FC_feature_pools %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,deltaZ, barcode = barcode)) + 
  facet_wrap(~antibody) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  theme_bw() + coord_fixed()

log2FC_feature_clone <- chrom_features_clone5 %>% mutate(deltaZ = PMA - DMSO)

ggplot(log2FC_feature_clone %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,deltaZ, barcode = barcode)) + 
  facet_wrap(~antibody) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  theme_bw() + coord_fixed()


#Combined plot
ggplot(log2FC_feature_pools %>% filter(antibody != "H3K9me2")) + 
  geom_point(aes(DMSO,deltaZ, barcode = barcode)) + 
  geom_hline(yintercept = 0, linetype= 2) + 
  geom_point(data= log2FC_feature_clone %>% filter(antibody != "H3K9me2")
             ,aes(DMSO,deltaZ, barcode = barcode), color = "red") + 
  facet_wrap(~antibody) +
  theme_bw() + coord_fixed()


```

#Correlate changes between them
```{r}
# 
deltaZ_correlation <- log2FC_feature %>% reshape2::dcast(barcode ~ antibody, value.var = "deltaZ")

linear_model_cor <- linear_model %>% reshape2::dcast(barcode ~ antibody, value.var = "residual")

ggplot(linear_model_cor,aes(H3K9me3,LMNB2)) + geom_point() + geom_smooth(method = "lm") + ggpubr::stat_cor(method = "pearson") 
ggplot(linear_model_cor,aes(H3K27me3,LMNB2)) + geom_point() + geom_smooth(method = "lm") + ggpubr::stat_cor() 
```

#Check tracks per replicate
```{r}
#Load bed file
coordinates <- read.table(file = "/DATA/projects/DSBrepair/data/xv20230608_K562_PMA_Ayoub/IPR_location/xv20230608_coordinates_clone5.txt", header=T)

plot_changes <- map(coordinates$IPR_barcode,function(x){
  position <- filter(coordinates, IPR_barcode == x) %>% pull(start)
  chrom <- filter(coordinates, IPR_barcode == x) %>% pull(chr)
  
  ggplot(PMA_replicate_tracks %>% 
         filter(antibody %in% c("H3K9me3","LMNB2") & seqnames == chrom & start > position - 5000000 & start < position + 5000000)) +
  geom_line(aes(start,smooth_score,color = treatment), position = "dodge") + 
  facet_grid(rep~antibody) + 
  theme_bw() + 
  geom_vline(xintercept = position, linetype = 2) +
  geom_hline(yintercept = 0)+
  ggtitle(x)
})

#Make pdf
print(plot_changes)



```

