---
title: "xv20230831_CCD_RPE1_validation"
author: "Xabier Vergara"
date: "2023-08-31"
output: html_document
---

Aim: In this file, I will explore some of the possible plots, where I get general conclusions from K562 and RPE1wt data.

```{r setup, include=FALSE}
library(tidyverse)
```

# Import data

```{r}
#Proteins_validation
proteins_gRNA <- tibble(gRNA = c("36","38","40","42","44","46","48","50","52","54","56","58","60","62","64","66","68","70","72","74"),
                        gene = c("ATM","ATR","BLM","BOD1L1","BRCA2","BRCC3","CHAF1A","CHEK2","FAAP24","FANCD2","FANCG","FANCM","MDC1","PARP1","POLL","RAD50","RBBP8","RMI2","SMC5","TOPBP1"))

#Processed data
PRO_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230908_CCD_analysis_summary_PRO_3reps.rds") %>% left_join(proteins_gRNA) %>% mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature))
K562_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220819_Table_S7_DR_screen_results.rds") %>% left_join(proteins_gRNA)
RPE1_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230831_CCD_summary_table_RPE1wt.rds") %>% left_join(proteins_gRNA) %>% mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature))

#All log2_values
RPE1wt_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230831_delta_log2_balance_RPE1wt.rds") %>% left_join(proteins_gRNA)
K562_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_log2_MMEJNHEJ_differentials_chromatin_KO.rds") %>% left_join(proteins_gRNA)

#K562 data on proteins in the validation screen
K562_validation_prots <- K562_CCD %>% na.omit() %>% filter(chrom_feature %in% c("H3K4me1","H3K36me3","LMNB1","H3K27me3","H3K27ac","H3K4me2","H3K4me3","H3K9me3","H3K9me2","late_replicating"))

```

#Plot correlation of global effects
```{r pressure, echo=FALSE}
#Global effect correlation data
global_effect_correlation <- PRO_CCD %>% 
  select(gene,gRNA,PRO_global = global_diff) %>% 
  left_join(K562_validation_prots %>% 
              select(gene,gRNA, K562_global = global_diff)) %>%
  left_join(RPE1_CCD %>% 
              select(gene,gRNA, RPE1_global = global_effect)) %>%
  distinct()

#Plot this global effect
ggplot(global_effect_correlation,aes(PRO_global,K562_global)) + 
  geom_smooth(method = "lm", se = F) +
  geom_point() +
  ggpubr::stat_cor(method = "spearman") +
  theme_bw() + 
  geom_abline(linetype = 2) +
  coord_fixed()

#Plot this global effect
ggplot(global_effect_correlation,aes(RPE1_global,PRO_global)) + 
  geom_smooth(method = "lm", se = F) +
  geom_point() +
  ggpubr::stat_cor(method = "spearman") +
  theme_bw() + 
  geom_abline(linetype = 2) +
  coord_fixed()



```

#Calculate cosine distance for all proteins
```{r}
all_validation <- PRO_CCD %>%
  select(gene,gRNA,chrom_feature,PRO_CCD = CCD_score) %>% 
  left_join(RPE1_CCD %>% 
              select(gene,gRNA,chrom_feature, RPE1_CCD = CCD_value)) %>%
  na.omit()

#Calculate cosine distance per gene
K562_RPE1_distance_all <- map_dfr(unique(all_validation$gene), function(x){
  CCD_table <- all_validation %>%
    filter(gene == x) %>%
    select(PRO_CCD, RPE1_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Join CCD_scores
K562_RPE_distances <- K562_RPE1_distance_all %>% dplyr::left_join(PRO_CCD %>% select(gene, gRNA,fdr)) 

ggplot(K562_RPE_distances %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist, fill = fdr < 0.1)) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("Protein") + 
  theme_bw() + 
  theme(axis.text.x =  element_text(angle = 90,
                                    hjust = 1,
                                    vjust = 0.5))
```

#Calculate cosine distance for all proteins
```{r}
all_validation <- DEF_CCD %>%
  select(gene,gRNA,chrom_feature,DEF_CCD = CCD_score) %>% 
  left_join(RPE1_CCD %>% 
              select(gene,gRNA,chrom_feature, RPE1_CCD = CCD_value)) %>%
  na.omit()

#Calculate cosine distance per gene
DEF_RPE1_distance_all <- map_dfr(unique(all_validation$gene), function(x){
  CCD_table <- all_validation %>%
    filter(gene == x) %>%
    select(DEF_CCD, RPE1_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Join CCD_scores
DEF_RPE_distances <- DEF_RPE1_distance_all %>% dplyr::left_join(DEF_CCD %>% select(gene, gRNA,fdr)) 

ggplot(DEF_RPE_distances %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist, fill = fdr < 0.1)) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("Protein") + 
  theme_bw() + 
  theme(axis.text.x =  element_text(angle = 90,
                                    hjust = 1,
                                    vjust = 0.5))
```
#Conclusion: DEF data is good quality, I should check the same filtering strategy for RPE1.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
