---
title: "xv20211230_3_Function_conserved"
output: html_document
---

# Script to generate figure #3 in the paper. This figure shows how protein function is conserved in most of cases, with few exceptions. Figure outline:
A- Heatmap & dendrograam with hits

Data in text:
A-...
B-...

Supplementary figure 4:
A- All proteins where one IPR has significant difference
B- Previously described genes = MRE11,Rad50,NBS1 or PARP1
C- Some proteins have perturb in both ways


This script generates the plots for figure 3 and S4

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(gprofiler2)
library(dendextend)
library(Hmisc)
library(scales)
library(ggdendro)
```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_epistasis.rds")
```

# Figure 3A: 19 IPR
```{r}
# Select hits 
hit.genes <- hits.both.screens  %>% filter(pathway != "NA") %>% separate(ID_gene, into = "gene", remove = T)

# Per 19 IPR heatmap
balance.IPR.hits <-per_19IPR_data %>% filter(gene %in% hit.genes$gene & !(gene == "KAT5" & library == "DNA_repair")) %>% dplyr::group_by(gene, barcode) %>% dplyr::summarise(IPR.z.score = mean(IPR.z.score, na.rm = T))

#dcast for heatmap
dcast.balance.IPR.hits <- balance.IPR.hits  %>% reshape2::dcast(gene ~ barcode, value.var = "IPR.z.score") %>% column_to_rownames(var = "gene")

#plot heatmap silently
heatmap.balance.IPR.hits <- pheatmap(dcast.balance.IPR.hits, silent = T)

#Extract gene and IPR orders
gene.o.balance.IPR.hits <- rownames(dcast.balance.IPR.hits[heatmap.balance.IPR.hits$tree_row[["order"]],])
IPR.o.balance.IPR.hits <- colnames(dcast.balance.IPR.hits[,heatmap.balance.IPR.hits$tree_col[["order"]]])

# heatmap
ggplot(balance.IPR.hits) + 
  geom_tile(aes(fct_relevel(gene,gene.o.balance.IPR.hits),fct_relevel(barcode, IPR.o.balance.IPR.hits), fill = IPR.z.score)) +
  scale_fill_gradient2( low = "#EB2030",mid = "grey90", high = "#2E358F", limits = c(-15,15), oob = squish)  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top", axis.text.y = element_blank()) + coord_fixed(expand = F,ratio = 1)

# Dendrogram
ggdendrogram(heatmap.balance.IPR.hits$tree_row)

```
# Number in text

#NOTE: This will change, but every number that I refer in the text will available in this chunk
```{r}
```

#Supplementary data

# Figure S4A, heatmap with all genees that have at least 1 IPR significant
```{r}
# Plot those that have at least one barcode significan
per_19IPR_data_hits <- filter(per_19IPR_data, abs(IPR.z.score) > 2.65 & count == 3) %>% pull(gene) %>% unique()


balance.IPR.hits.all <-per_19IPR_data %>% filter(gene %in% per_19IPR_data_hits & !(gene == "KAT5" & library == "DNA_repair")) %>% dplyr::group_by(gene, barcode) %>% dplyr::summarise(IPR.z.score = mean(IPR.z.score, na.rm = T))

#dcast for heatmap
dcast.balance.IPR.hits.all <- balance.IPR.hits.all  %>% reshape2::dcast(gene ~ barcode, value.var = "IPR.z.score") %>% column_to_rownames(var = "gene")

#plot heatmap silently
heatmap.balance.IPR.hits.all <- pheatmap(dcast.balance.IPR.hits.all, silent = T)

#Extract gene and IPR orders
gene.o.balance.IPR.hits.all <- rownames(dcast.balance.IPR.hits.all[heatmap.balance.IPR.hits.all$tree_row[["order"]],])
IPR.o.balance.IPR.hits.all <- colnames(dcast.balance.IPR.hits.all[,heatmap.balance.IPR.hits.all$tree_col[["order"]]])

# Plot
ggplot(balance.IPR.hits.all) + 
  geom_tile(aes(fct_relevel(gene,gene.o.balance.IPR.hits.all),fct_relevel(barcode, IPR.o.balance.IPR.hits.all), fill = IPR.z.score)) +
  scale_fill_gradient2( low = "#EB2030",mid = "grey90", high = "#2E358F", limits = c(-15,15), oob = squish)  + theme_bw() + theme(axis.text.x = element_blank(), legend.position = "top", axis.text.y = element_blank()) + coord_fixed(ratio = 5)

# Dendrogram
ggdendrogram(heatmap.balance.IPR.hits.all$tree_row)
```

#Fig S4B: Examples of genes involved in different processes
```{r}
# Filter proteins highlighted 
ggplot(per_19IPR_data %>% filter(gene %in% c("MRE11","RAD50","NBN","PARP1"))) +
  geom_quasirandom(aes(fct_relevel(gene, c("RAD50","NBN","MRE11","PARP1")),IPR.z.score)) +
  ylim(c(-15,5)) +
  geom_hline(yintercept = 0, linetype =2) +
  theme_bw()
```
# Fig S4C: Genes that have effects in both diirections
```{r}
# Is there any that has significant perturbation in both situation
genes.both.dir <- balance.IPR.hits.all %>% mutate(pos = IPR.z.score > 2.65, neg = IPR.z.score < -2.65) %>% dplyr::group_by(gene) %>% dplyr::summarise(pos = sum(pos), neg = sum(neg))

# Genes effect in both directions
genes.both.dir.list <- filter(genes.both.dir, pos > 0 & neg > 0) %>% pull(gene)

#dcast for heatmap
dcast.balance.IPR.hits.both <- balance.IPR.hits.all %>% filter(gene %in% genes.both.dir.list) %>% reshape2::dcast(gene ~ barcode, value.var = "IPR.z.score") %>% column_to_rownames(var = "gene")

#plot heatmap silently
heatmap.balance.IPR.hits.both <- pheatmap(dcast.balance.IPR.hits.both, silent = T)

#Extract gene and IPR orders
gene.o.balance.IPR.hits.both <- rownames(dcast.balance.IPR.hits.both[heatmap.balance.IPR.hits.both$tree_row[["order"]],])
IPR.o.balance.IPR.hits.both <- colnames(dcast.balance.IPR.hits.both[,heatmap.balance.IPR.hits.both$tree_col[["order"]]])

# Plot
ggplot(balance.IPR.hits.all %>% filter(gene %in% genes.both.dir.list)) + 
  geom_tile(aes(fct_relevel(gene,gene.o.balance.IPR.hits.both),fct_relevel(barcode, IPR.o.balance.IPR.hits.both), fill = IPR.z.score)) +
  geom_point(data = balance.IPR.hits.all %>% filter(gene %in% genes.both.dir.list & abs(IPR.z.score) > 2.65),aes(fct_relevel(gene,gene.o.balance.IPR.hits.both),fct_relevel(barcode, IPR.o.balance.IPR.hits.both), fill = IPR.z.score)) +
  scale_fill_gradient2( low = "#EB2030",mid = "grey90", high = "#2E358F", limits = c(-5,5), oob = squish)  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top", axis.text.y = element_blank()) + coord_fixed(ratio = 1)

# Dendrogram
ggdendrogram(heatmap.balance.IPR.hits.both$tree_row)

# Add library stip
hits.genese.both.dir <- hits.both.screens %>% separate(ID_gene, into = "gene") %>% filter(gene %in% genes.both.dir.list)

ggplot(hits.genese.both.dir %>% dplyr::group_by(gene,library) %>% dplyr::summarise(count = n())) +
  geom_bar(aes(library,fill = library)) +
  theme_bw() +
  scale_fill_manual(values = rev(levels(as.factor(library.colors$color))))

```

