---
title: "xv20211230_5_epistasis_inhibitors"
output: html_document
---
# Script to generate figure #5 in the paper. This figure shows that epistatic interactions can be bigger if drugs are used instead of KO. Figure outline:
A- Effect of drug treatment in MMEJ:NHEJ balance
B- Matrix for all treatments
C- Slopes are bigger in inhibition than KO
D- PARP1 and olaparib correlation
E- ATM and ATMi correlation

Data in text:
A-...
B-...

Supplementary figure 6:
A- PARP1 vs. veliparib
B- PARP1 vs. talazoparib
C- ATRi vs. VE-821
D- ATRi vs. VE-822

This script generates the plots for figure 5 and S6

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(ggdendro)
```


# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_kinase_inhib_epistasis.rds")

```

#Figure 5A: Effect in pathway balance
```{r}
# Mean effect over replicates per experiment
mean.combined.inhibitor.data <- combined.inhibitor.data.filt %>% dplyr::group_by(drug,barcode,exp) %>% dplyr::summarise(m.ratio = mean(NHEJMMEJratio), na.rm = T)

#DMSO control data
DMSO.control <- mean.combined.inhibitor.data %>% filter(drug == "DMSO") %>% ungroup() %>%  select(barcode,DMSO.ratio = m.ratio, exp)

# Calculate log2 differences
log2.fc.drugs <- mean.combined.inhibitor.data %>% left_join(DMSO.control, by = c("barcode", "exp")) %>% mutate(log2.fc = log2(m.ratio/DMSO.ratio))

# Data plot
plot.dt.log2.fc.drugs <- log2.fc.drugs %>% filter(!drug %in% c("DMSO","GFP")) %>% dplyr::group_by(barcode,drug) %>% dplyr::summarise(log2.fc = mean(log2.fc, na.rm = T))

# Plot
ggplot(plot.dt.log2.fc.drugs, aes(drug,log2.fc)) + 
geom_quasirandom() + 
stat_mean(geom = "point", color = "red") +
geom_hline(yintercept = 0, linetype = 2) + 
theme_bw() +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

test.perturbation.drug <- plot.dt.log2.fc.drugs %>%
  dplyr::group_by(drug) %>%
  wilcox_test(log2.fc ~ 0)

```


# Figure 6B: Heatmap
```{r}
#Prepare for dcast
mean.slope.protein.featurese.inh <- slope.protein.features.inhibitors.filt %>% filter(drug != "DMSO") %>% dplyr::group_by(drug,feature) %>% dplyr::summarise(epistasis = mean(epistasis, na.rm = T))

#dcast epistatic interactions
slope.prot.features.scale.dcast.inhibitors <- mean.slope.protein.featurese.inh %>% reshape2::dcast(drug ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "drug")

#Heatmap for slope differences
heatmap.slope.diff.inhibitors <- pheatmap(slope.prot.features.scale.dcast.inhibitors, silent = T)
heatmap.gene.order.slope.diff.inhibitors <- rownames(slope.prot.features.scale.dcast.inhibitors[dendextend::rotate(heatmap.slope.diff.inhibitors$tree_row, order = c("DNAPKi","ATMi","ATRi_VE821","ATRi_VE822","veliparib","talazoparib","olaparib"))[["order"]],])
heatmap.chromatin.order.slope.diff.inhibitors <- colnames(slope.prot.features.scale.dcast.inhibitors[,heatmap.slope.diff.inhibitors$tree_col[["order"]]])

#Plot correlations (As a heatmap with corrected pvalues)
ggplot(mean.slope.protein.featurese.inh) + 
  geom_tile(aes(fct_relevel(feature, heatmap.chromatin.order.slope.diff.inhibitors),fct_relevel(drug,heatmap.gene.order.slope.diff.inhibitors), fill = epistasis)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# dendogram
ggdendrogram(dendextend::rotate(heatmap.slope.diff.inhibitors$tree_row, order = c("DNAPKi","ATMi","ATRi_VE821","ATRi_VE822","veliparib","talazoparib","olaparib")), rotate =T)

```
# Figure 6C: Inhibitors have bigger epistatic interactions than KOs
```{r}
#Prepare data tables
selected.genes.drugs <- slope.protein.features.all %>% filter(gene %in% c("PARP1","ATM","ATR","PRKDC")) %>% select("gene", "feature","epistasis") %>% mutate(class = "protein", perturbation = paste(gene, "KO", sep = "_"))

drug.comparison <- mean.slope.protein.featurese.inh %>% select("perturbation" = "drug", "feature","epistasis") %>% mutate(class = "drug", gene = case_when(grepl("ATM", perturbation) ~ "ATM", grepl("ATR",perturbation) ~ "ATR", gene = grepl("DNAPK", perturbation) ~ "PRKDC", T ~ "PARP1"))

# dt.for.plotting
comparison.plot <- bind_rows(drug.comparison,selected.genes.drugs)

ggplot(comparison.plot) + geom_quasirandom(aes(fct_reorder(perturbation, desc(class)),epistasis, color = class)) + facet_wrap(~ fct_relevel(gene,c("ATM","PARP1","ATR","PRKDC")), scales = "free_x") +
theme_bw() +
theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5,hjust = 1), legend.position = "top")

#Test kolmogorov-Smirnof (read more about this test)
ks.test.inhibitors <- tibble(statistic = NA, p.value = NA, method = NA, alternative = NA)

# Create loop 
for (i in unique(comparison.plot$gene)){
  gene.dt <- filter(comparison.plot, gene == i)
  KO.values <- gene.dt %>% filter(class == "protein") %>% pull(slope.log2) 
    if (length(unique(gene.dt$perturbation)) == 2){
      drug.dt <- gene.dt %>% filter(class == "drug") 
      drug.value <- drug.dt %>% pull(slope.log2)
      ks_test <- ks.test(KO.values,drug.value) %>% broom::tidy() %>% mutate(gene = i, perturbation = unique(drug.dt$perturbation))
      ks.test.inhibitors <- bind_rows(ks.test.inhibitors, ks_test)
    }
  else{
      drug.dt <- gene.dt %>% filter(class == "drug")
    for (j in unique(drug.dt$perturbation)) {
      drug.value <- drug.dt %>% filter(perturbation == j) %>% pull(slope.log2)
      ks_test <- ks.test(KO.values,drug.value) %>% broom::tidy() %>% mutate(gene = i, perturbation = j)
      ks.test.inhibitors <- bind_rows(ks.test.inhibitors, ks_test)
    }
  }
  }

```

#Figure 6D: PARP1 olaparib
```{r}
# olaparib vs PARP1
genes <- slope.protein.features.all %>% filter(gene %in% c("PARP1","ATM","ATR","PRKDC")) %>% reshape2::dcast(feature ~ gene, value.var = "epistasis")

#dcast for plotting
inhibitors.dcast <- as.data.frame(t(slope.prot.features.scale.dcast.inhibitors)) %>% rownames_to_column(var = "feature")

gene.drug.perturbation <- left_join(inhibitors.dcast , genes)

#Plots (main figure)
ggplot(gene.drug.perturbation, aes(olaparib, PARP1)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```

#Figure 6E: ATMi vs. ATMKO
```{r}
ggplot(gene.drug.perturbation, aes(ATMi, ATM)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```
# Number in text

#NOTE: This will change, but every number that I refer in the text will available in this chunk
```{r}
```

#Supplementary figures

#Figure S6A: Veliparib vs PARP1
```{r}
ggplot(gene.drug.perturbation, aes(veliparib, PARP1)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```

#Figure S6B: Talazoparib vs PARP1
```{r}
ggplot(gene.drug.perturbation, aes(talazoparib, PARP1)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```

#Figure S6C: VE821 vs ATRKO
```{r}
ggplot(gene.drug.perturbation, aes(ATRi_VE821, ATR)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```

#Figure S6D: VE822 vs ATRKO
```{r}
ggplot(gene.drug.perturbation, aes(ATRi_VE822, ATR)) + geom_point() + geom_smooth(method = "lm") + stat_cor() + theme_bw()
```
