---
title: "xv20220623_CCD_preprocessing"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  log_path <- paste0(in.dir,"log")
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    #print(mode(log_path))
    #base::write(c("test", as.character(sys_time),"Output", object), file = log_path, ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```

# Libraries
```{r libraries, warning = FALSE}
# libraries:
library(tidyverse)
library(reshape2)
library(broom)
library(dendextend)
library(pls)
library(broom)
library(stats)
library(MASS)
library(rstatix)
library(parallel)
library(data.table)
library(readxl)
```



# Data import
```{r, message=FALSE, warnings=FALSE}
trip_tib_mut_2000 = readRDS(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RDS"))
trip_tib_2000 = readRDS(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RDS"))

chromsizes<-read.table("/home/r.schep/mydata/data/genomes/GRCh38/hg38_chromsizes2.txt",header=F, stringsAsFactors = F)
names(chromsizes)<-c("chr_r","pos_r")

load(paste0(in.dir, "Analyis_Mapping_RSTP2_2000.RData"))

clone5barcodes <- c("AGGGCGTAAAATATTT.B",
                    "TATGGCTGTCGGGTAG.B",
                    "TGTCCCTTAGTACTTT.B",
                    "AGAAAATAATATGACG.B",
                    "CGGCCTGAAGGTCAGG.B",
                    "TTGAACGCGGGCTCGG.B",
                    "GCTAACATCACGAATC.B",
                    "GCGCACCCTTTAATTG.B",
                    "ACTGTCGAGTTGTCCG.B",
                    "CCGGGGACGTATGCAC.B",
                    "TCTTTTGAGGAGCTGA.B",
                    "ATATCGTTGCTGGAGA.B",
                    "CATCCACCACACTTCA.B",
                    "ACCCCTAAAGGCGCTG.B",
                    "ATACTATATTTAACGG.B",
                    "CATTTCTGATCAATAA.B",
                    "GAGCGCGTCACCGGGT.B",
                    "GTACCTCTCGATAGTG.B",
                    "TGGCCAATATTTGTCT.B")

clone5_domains <- readRDS("/DATA/projects/DSBrepair/data/R/cl20200310_domains_clone5.RDS")
```

## Data processing
```{r preprocess experiments}
trip_tib_mut_all <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp_pool, .keep_all = TRUE)
trip_lbr2_DMSO_tib <- trip_tib_2000 %>%
  filter(exp_pool=="mean_RSTP2_2000_LBR2_DMSO_64")

trip_tib <- trip_tib_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_64"))
trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_64", "mean_RSTP2_2000_LBR2_NU7441_64"))

trip_tib_mut_all_selection <- trip_tib_mut_all %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_64", "mean_RSTP2_2000_LBR2_NU7441_64"))

mean_trip_tib_2000 <- trip_tib_2000 %>%
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp_pool, pathway, .keep_all = TRUE)

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]


mutation_barcodes <- trip_tib_2000 %>% 
  filter(exp_pool=="mean_RSTP2_2000_LBR2_64", nexp > 1) %>% 
  distinct(barcode, exp_pool, .keep_all = TRUE)


mutationintegrations <-na.omit(analysis.mapped.integrations.df[c("barcode", "seqname","start")])
colnames(mutationintegrations) <- c("barcode", "chr_r","pos_r")
dim(mutationintegrations)
mutationintegrations$mapping <- "mapped"
mutationintegrations[mutationintegrations$barcode %in% c(mutation_barcodes$barcode, clone5barcodes) , ]$mapping <- "filtered"

mutationintegrations <- changeLevels(mutationintegrations)
chromsizes <- changeLevels(chromsizes)
chromsizes <- chromsizes[!is.na(chromsizes$chr_r), ]
mutationintegrations <- mutationintegrations[!is.na(mutationintegrations$chr_r), ]
chromsizes$pos_r <- as.numeric(chromsizes$pos_r)


chrom.mods.cols <- seq(grep("binsize", names(trip_tib))+1, grep("H3K27me3_domain", names(trip_tib))-1)
```

## Panel S2G
```{r FigS2G pathways with HR siRNA, fig.height=5, fig.width=7}
siRNA_HR_exps <- trip_tib_mut_all %>% 
  filter(replicate %in% c('mean'), 
         drug == "Shield", 
         siRNA %in% c("siNT", "siPolQ", 
                      "siRad51", "siBRCA1", 
                      "siBRCA2"), 
         cell_line == "RSTP2_clone5", 
         plasmid == "LBR2", 
         time == 72) %>% 
  pull(exp_pool) %>% 
  unique()

trip_lbr2_siRNA_HR <- trip_tib_2000 %>%
  filter(exp_pool %in% c(siRNA_HR_exps))

trip_mut_lbr2_all_siRNA_HR <- trip_tib_mut_all %>%
  filter(exp_pool %in% c(siRNA_HR_exps))

#Select columns for analysis
HR_columns_select_mean <- trip_mut_lbr2_all_siRNA_HR %>% select(siRNA, barcode, reads_MMEJ, reads_NHEJ) %>% distinct()

#Calculate ∆log2MMEJNHEJ
HR_columns_mean_computed <- HR_columns_select_mean %>%
  mutate(log2MMEJNHEJ = log2(reads_MMEJ/reads_NHEJ))


```



### Calculation of differentials of log2 MMEJ:NHEJ (Formula #3)
```{r}
#Extract non-targetting samples from this dataset
wt.set.ratio <- filter(HR_columns_mean_computed , siRNA == "siNT") %>% select(barcode, wt.log2ratio = log2MMEJNHEJ)

# Compute log2 fold changes in different set ups for all the KO samples
log2.distance.ratio <- HR_columns_mean_computed %>% 
  na.omit() %>% #Remove NA
  left_join(wt.set.ratio, by = c("barcode")) %>% #Join mock KOs
  mutate(log2.foldchange = log2MMEJNHEJ - wt.log2ratio) %>% #Calculate differential of log2MMEJ:NHEJ 
  dplyr::group_by(siRNA,barcode) %>% #Group by sample
  dplyr::summarise(mean.log2foldchange = mean(log2.foldchange,na.rm = T), n_rep = n()) %>% #Compute differential of log2MMEJ:NHEJ
  ungroup()
```
  
  
#Step 4: Identification of DR proteins with global effects on MMEJ:NHEJ balance
```{r}
#Compute mean MMEJ:NHEJ differential over all IPRs
mean.diff.balance <- log2.distance.ratio %>% 
  filter(siRNA != "siNT") %>%
  group_by(siRNA) %>% 
  dplyr::summarise(mean.fc = mean(mean.log2foldchange, na.rm = T))

#Run t.test with all the data
test.mean.diff.balance <- log2.distance.ratio %>% 
  filter(siRNA != "KO") %>%
  dplyr::group_by(siRNA) %>% 
  t_test(mean.log2foldchange ~ 0) %>% 
  mutate(p.adj = p.adjust(p, method = "BH")) %>% 
  left_join(mean.diff.balance, by = "siRNA")

#Prepare data frame for export
mean.diff.balance.export <- test.mean.diff.balance %>%
  dplyr::select(siRNA,p,p.adj,mean.balance.diff = mean.fc)

#Significant MMEJ
mean_significant_MMEJ <- filter(mean.diff.balance.export, p.adj < 0.001 & mean.balance.diff < 0) %>% nrow() #149 proteins favor MMEJ
mean_significant_NHEJ <- filter(mean.diff.balance.export, p.adj < 0.001 & mean.balance.diff > 0) %>% nrow() #16 proteins favor NHEJ

#Print
print(paste("A total of", mean_significant_MMEJ, "DR proteins favor MMEJ globally with an FDR < 0.001"))
print(paste("A total of", mean_significant_NHEJ, "DR proteins favor NHEJ globally with an FDR < 0.001"))
```

# Step5: Identification of repair proteisn with CCDs: three-step linear modeling

## Sub-step A: Selection of DR proteins with Effects on MMEJ:NHEJ balance (I will consider the global balance means)

# Substep B: Principal component regression to identify repair proteins with CCDs
```{r}
#Filter genes that do not significantly perturb balance and add chromatin info
DDR.KO.CCD.dt <- log2.distance.ratio %>%
  filter(siRNA != "siNT") %>%
  left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID")) #Join chromatin info

#Create an empty dt with CCDs of DDR proteins
DDR_KO_CCDs_dt <- tibble(siRNA = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)

for (i in unique(DDR.KO.CCD.dt$siRNA)){
gene.library.dt <- filter(DDR.KO.CCD.dt, siRNA == i)
set.seed(1)
PCR_model_DDR_test <- pcr(mean.log2foldchange~CTCF+EZH2+H2AFZ+H3K27ac+H3K27me3+H3K36me3+H3K4me1+H3K4me2+H3K4me3+H3K79me2+H3K9me2+H3K9me3+H4K5acK8ac+HDAC1+HDAC2+HDAC3+POL2+POL2AS2+SMC3+LMNB1+Dam+DNAse+m5C+late_replicating+TTseq, data=gene.library.dt , validation="CV") #Run principal component regression

pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = 3) #Run model with three PC
combined.dt <- tibble(measured = gene.library.dt$mean.log2foldchange, predicted = as.numeric(pcr_pred)) #Create table with real and predicted differentials
pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% glance() #Predicted vs. measured correlation plot
DDR_KO_CCDs_dt <- DDR_KO_CCDs_dt %>% add_row(siRNA = i, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value)) #Extract valuable metrics
}

#Calculate
adj_p.value_model <- DDR_KO_CCDs_dt %>%
  mutate(p.adj = p.adjust(p.value, method = "BH")) %>%
  select(-r.squared) %>%
  mutate(pathway_signif = case_when(siRNA %in% c("siRad51","siBRCA1","siPolQ") ~ "MMEJ",
                                    T ~ "NHEJ"))

```


#Sub-step C: Linear modeling to identify individual DR proteins - chromatin feature links
```{r}
#Create empty dataframe to calculate synergy scores
DR_KO_CCD_slopes <- tibble(siRNA = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

#Loop to run linear models on the values
for (h in unique(DDR.KO.CCD.dt$siRNA)) {
  for (j in colnames(DDR.KO.CCD.dt)[7:31]) { #Run this function for each of the 25 high quality chromatin features
    model.dt <- DDR.KO.CCD.dt %>% filter(siRNA == h) # And For each gene
   if (nrow(model.dt) == 0) {
next
}
    model.epistasis.log2 <- lm(formula = mean.log2foldchange ~ unlist(model.dt[j]), data = model.dt) %>% tidy() #Correlation analysis
   DR_KO_CCD_slopes <- DR_KO_CCD_slopes %>% add_row(siRNA = h, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value)) #Select valuable parameters and save them in data frame
  }
}

#Retain slopes that 
DR_KO_CCD_all_values <- DR_KO_CCD_slopes %>% 
  reshape2::dcast(siRNA + feature ~ term, value.var = "slope.log2") %>%  #dcast table
  dplyr::select(siRNA ,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% #Extract information for slopes only
  left_join(adj_p.value_model %>% 
              dplyr::select(siRNA, model_p.value = p.value, model_p.adj = p.adj, pathway_signif), by = "siRNA") %>% #Join with CCDs p_value data table and change some names
  mutate(CCD_value = case_when(pathway_signif == "MMEJ" & slope < 0 ~ slope, pathway_signif == "MMEJ" & slope > 0 ~ 0, pathway_signif == "NHEJ" & slope > 0 ~ slope, pathway_signif == "NHEJ" & slope < 0 ~ 0, T ~ slope)) %>% #Call M-synergies, N-synergies or no synergies based on the slope and MMEJ:NHEJ differentials
  na.omit()

#How many M-, N- and no synergies
synergy_per_gene <- DR_KO_CCD_all_values %>% 
  dplyr::group_by(siRNA) %>% 
  dplyr::summarise(M_synergy = sum(CCD_value < 0),
                   N_synergy = sum(CCD_value > 0)) %>%
  distinct() %>% 
  mutate(CCD_effect = case_when(M_synergy > 0 & N_synergy == 0 ~ "M_synergy",
                                N_synergy > 0 & M_synergy == 0 ~ "N_synergy",
                                N_synergy > 0 & M_synergy > 0 ~ "both", T ~ "none"))

synergy_summaries <- synergy_per_gene %>% dplyr::group_by(CCD_effect) %>% dplyr::summarise(c = n())
```

#Step 6: Compute estimated ∆log2MMEJ:NHEJ scores for CCD effects and compared to global ∆log2MMEJ:NHEJ
##Pre-processing chunk: load all ChIP track z-scores
```{r}
#
chromatin_features_plot <-  clone5_z.score_chrom_tib %>% reshape2::melt(id.vars = c("pool","ID","binsize"))

##99% CI for each feature
CI99_chromatin_features <- chromatin_features_plot %>% dplyr::group_by(variable) %>% dplyr::summarise(CI0.5 = quantile(value, 0.005), CI99.5 = quantile(value, 0.995), range = CI99.5 - CI0.5, mean = mean(value))

##DNA repair protein class (CCD, global, both)
global_effect_proteins <- mean.diff.balance.export %>% filter(p.adj < 0.001) %>% pull(siRNA) %>% unique() #Filter genes with significant global ∆log2MMEJ:NHEJ scores
CCD_proteins <- DR_KO_CCD_all_values %>% filter(model_p.adj < 0.05) %>% pull(siRNA) %>% unique() #Filter genes with significand CCD ∆log2MMEJ:NHEJ scores
#All classes
both_CCD_global_proteins <- intersect(global_effect_proteins, CCD_proteins) #Check for overlap
only_global <- setdiff(global_effect_proteins, both_CCD_global_proteins) #Only global effects
only_CCD <- setdiff(CCD_proteins, both_CCD_global_proteins) #Only CCDs

# Create a summary table with three protein classes
classes_summary <- tibble(siRNA = c(both_CCD_global_proteins,only_global,only_CCD), 
                          class = c(rep("both", length(both_CCD_global_proteins)),rep("global",length(only_global)), rep("CCD", length(only_CCD))))

##Compute estimated CCD ∆log2MMEJ:NHEJ score
estimated_FC_CCD <- DR_KO_CCD_all_values %>% 
  left_join(CI99_chromatin_features, by = c("feature" ="variable")) %>% #Join 99% CI for chromatin
  mutate(estim_fc_CCD = range*CCD_value) %>% #Calculate range
  full_join(classes_summary, by = "siRNA") #Join class they belong

##Select the maximum estimated effect for each gene
max_estimated_FC_CCD_gene <- estimated_FC_CCD %>% 
  dplyr::group_by(siRNA) %>% 
  dplyr::summarise(m_CCD_FC_nhej = max(estim_fc_CCD), 
                   m_CCD_FC_mmej = min(estim_fc_CCD)) %>% 
  reshape2::melt(value.name = "m_CCD_FC", id.vars = "siRNA") %>% 
  filter(m_CCD_FC != 0)

## Filter max(CCDs)
max_estimated_CCD_FC <- estimated_FC_CCD %>% 
  mutate(m_CCD_FC = estim_fc_CCD) %>% 
  right_join(max_estimated_FC_CCD_gene) %>% 
  dplyr::select(-m_CCD_FC) %>%
  full_join(classes_summary)

#Left_join mean.fc
estimated_CCD_global <- max_estimated_CCD_FC %>% 
  full_join(mean.diff.balance.export, by = "siRNA")

#Summary table class of DR proteins
summary_DR_prot_class <- estimated_CCD_global %>% dplyr::select(class,siRNA) %>% distinct() %>% dplyr::group_by(class) %>% dplyr::summarise(count = n())

#Print output of the chunk
print(paste("We find",summary_DR_prot_class$count[1] + summary_DR_prot_class$count[2], "DR proteins with CCD. Among these,", summary_DR_prot_class$count[1],"have global effects as well and", summary_DR_prot_class$count[2], "have exclusively CCD effect. Additionally, we find", summary_DR_prot_class$count[3], "DR proteins that have global effect only"))

```

#Step 7: Data visualization
This step will be performed with the output files in separate scripts 

#Create table S4: Mean ∆log2MMEJ:NHEJ balance of each well.
```{r}
export_screen_result_TS4 <- log2.distance.ratio %>% 
  dplyr::select(siRNA, IPR_barcode = "barcode", delta_log2_MMEJ_NHEJ = "mean.log2foldchange")
```


#Create table S7: Data to export all major features
```{r}
export_screen_result_TS7 <- estimated_FC_CCD %>% 
  full_join(mean.diff.balance.export) %>%
  dplyr::select(siRNA, CCD_model_p_adj = "model_p.adj",chrom_feature = "feature", CCD_synergy_score = "CCD_value",CCD_estim_diff = "estim_fc_CCD", global_diff = "mean.balance.diff", global_p_adj = "p.adj", DR_effect_type = "class") %>%
  mutate(global_effect = case_when(global_p_adj < 0.001 & global_diff < 0 ~ "favorMMEJ",
                                   global_p_adj < 0.001 & global_diff > 0 ~ "favorNHEJ",
                                   T ~ "none"),
         cell_line = "K562_clone5") %>%
  left_join(synergy_per_gene %>% 
              select(siRNA, CCD_effect))

export_screen_result_TS7[is.na(export_screen_result_TS7$DR_effect_type), "DR_effect_type"] <- "none"
export_screen_result_TS7[is.na(export_screen_result_TS7$CCD_effect), "CCD_effect"] <- "none"

```


#Output files: Export all files needed for making figures
```{r, eval = FALSE, echo = FALSE}
setwd("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/")

#Step 2: dcast data table for 
saveRDS_proof(log2.distance.ratio, file = "data/processed_data/CCD_siRNA/xv20231123_siRNA_differentials_log2_MMEJ_NHEJ_balance.rds")

#Step 4: diff log2 MMEJ:NHEJ values for DR proteins with chromatin scores
saveRDS_proof(DDR.KO.CCD.dt, file = "data/processed_data/CCD_siRNA/xv20231123_siRNA_log2_MMEJNHEJ_differentials_chromatin_KO.rds")

#Step 5: save global changes for all proteins
saveRDS_proof(mean.diff.balance.export, file = "data/processed_data/CCD_siRNA/xv20231123_siRNA_global_diff_MMEJNHEJ.rds")

#Save final results as rds
saveRDS_proof(export_screen_result_TS7, file = "data/processed_data/xv20231123_Table_S7_siRNA_Mol_Cell_results.rds")

```

#Export supplementary tables in the paper
```{r, eval = FALSE, echo = FALSE}
setwd(in.dir)
#Export for supplementary files (.xlsx file)

#Table S4: ∆log2MMEJ:NHEJ scores for 519 proteins and 19 IPRs.
#write.xlsx(export_screen_result_TS4, file = "data/supplementary_tables/xv20220819_Table_S4a_delta_log2_MMEJ_NHEJ.xlsx")

#Table S6: Genomic coordinates, chromatin feature scores and barcodes of 19 IPRs in K562 clone 5
#write.xlsx(export_screen_result_TS6, file = "data/supplementary_tables/xv20220819_Table_S6_clone_5_chromatin_features.xlsx")

#Table S7: Chromatin context dependent effects of proteins in the screen.
write.xlsx(export_screen_result_TS7, file = "data/supplementary_tables/xv20231117_Table_S7f_siRNA_global_CCD_MMEJ_NHEJ_results.xlsx", keepNA = T)

```
