---
title: "xv20220712_S7_editing_efficiency"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S7 in the short version of the paper.

Fig. S7: Editing efficiency estimation. A) Transcribed (n = 8) and non-transcribed (n = 11) IPRs in clone 5. We classified IPRs based on the transcription-related feature signals (light orange). B) Editing efficiencies in transcribed IPRs for each replicate in mock transfected samples (n = 33). Error bars show mean ± sd.

```{r functions, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries, warning=FALSE, message=FALSE}
# libraries:  
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
```



# Import data tables
```{r import, warning=FALSE, message=FALSE}

setwd(in.dir)
#Import MMEJ:NHEJ balance ratio measurements per replicate
ddr.screen.all <- readRDS_proof( "data/processed_data/CCD_analysis","raw_data")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")

```

#Supplementary Figures

#Figure S7A
Modified from Fig. 1C
A) Transcribed (n = 8) and non-transcribed (n = 11) IPRs in clone 5. We classified IPRs based on the transcription-related feature signals (light orange). 

#Figure S7B: 
B) Editing efficiencies in transcribed IPRs for each replicate in mock transfected samples (n = 33). Error bars show mean ± sd.
```{r}
editing_efficiency_WT <-  ddr.screen.all %>% filter(sample == "WT") %>% dplyr::select(barcode,well,plate, rep, freqCut)

#Reporters in transcribed regions 
transcr_reporters <- clone5_bc_cluster[c(12:19)]
editing_eff_transcr_reporters <- editing_efficiency_WT %>% filter(barcode %in% transcr_reporters) 

#
ggplot(editing_eff_transcr_reporters) + stat_summary(aes(fct_relevel(barcode, transcr_reporters),freqCut, fill = rep), geom = "col", fun = "mean", position = "dodge") + stat_summary(aes(fct_relevel(barcode, transcr_reporters),freqCut, group = rep), geom = "errorbar",fun.data = "mean_sd", position = "dodge") + theme_bw() + theme(axis.text.x = element_blank())

```

