---
title: "xv20220712_S7_editing_efficiency"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S7 in the short version of the paper.

Fig. S7: Editing efficiency estimation. A) Transcribed (n = 8) and non-transcribed (n = 11) IPRs in clone 5. We classified IPRs based on the transcription-related feature signals (light orange). B) Editing efficiencies in transcribed IPRs for each replicate in mock transfected samples (n = 33). Error bars show mean ± sd.

```{r functions, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries, warning=FALSE, message=FALSE}
# libraries:  
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
```



# Import data tables
```{r import, warning=FALSE, message=FALSE}

setwd(in.dir)
#Import MMEJ:NHEJ balance ratio measurements per replicate
ddr.screen.all <- readRDS_proof( "data/processed_data/CCD_analysis","raw_data")

inhibtor.screen.all <- readRDS_proof("data/processed_data/CCD_inhibitors","repair_metrics")

revision_data_RPE1 <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/","RPE1_control")
revision_data_PRO <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/","PRO_control")
revision_data_DEF <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/","DEF_control")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")

#Import delta log2 balance effects
PRO_log2 <-  readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_PRO_all_mean_freq_0.0025.rds") %>% filter(is.finite(log2_bal))
K562_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_screening_raw_data_repair_metrics.rds")
DEF_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230911_DEF_all_mean_freq_0.0075_.rds") %>%
  filter(is.finite(log2_bal))
RPE1_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_RPE1_min_read_5.rds") %>% 
  filter(!gRNA %in% c("LBR2","halfLBR2","empty") & cell_line == "RPE1")

#Import CCD effects
CCD_effects_all_cells <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230917_CCD_summary_cell_lines.rds")

```

#Data arrangement in K562 cells
```{r}
#Merge transfection and nucleofection examples
editing_efficiency_WT <-  ddr.screen.all %>% filter(sample == "WT") %>% dplyr::select(barcode,well,plate, rep, freqCut) %>% mutate(delivery = "lipofection")
editing_efficiency_inhibitor <-  inhibtor.screen.all %>% filter(drug == "DMSO") %>% dplyr::select(barcode, rep = replicate, freqCut) %>% mutate(delivery = "nucleofection")

mergedt_K562_data <- editing_efficiency_WT %>% bind_rows(editing_efficiency_inhibitor)

#Calculate value per replicate
mean_per_replicate_TIF <- mergedt_K562_data %>%
  dplyr::group_by(barcode, delivery) %>%
  dplyr::summarise(mean_TIF = mean(freqCut))
 
#Reporters in transcribed regions 
transcr_reporters <- clone5_bc_cluster[c(12:19)]
editing_eff_transcr_reporters_inhibitor <- editing_efficiency_inhibitor %>% filter(barcode %in% transcr_reporters) 
```

#Supplementary Figures

#Figure S7A
Modified from Fig. 1C
A) Transcribed (n = 8) and non-transcribed (n = 11) IPRs in clone 5. We classified IPRs based on the transcription-related feature signals (light orange). 

#Figure S7B: 
B) Editing efficiencies in transcribed IPRs for each replicate in mock transfected samples (n = 33). Error bars show mean ± sd.
```{r}
#Plot as beeswarm plot
ggplot(mean_per_replicate_TIF %>% filter(barcode %in% transcr_reporters)) +
  geom_quasirandom(aes(delivery, mean_TIF)) + theme_bw() + ylim(c(0,1))

 #Correlation of both measurements
t_test_IPR <- mean_per_replicate_TIF %>% 
  filter(barcode %in% transcr_reporters) %>%
  ungroup() %>%
  t_test(mean_TIF ~ delivery, paired = T)

```
#CONCLUSION: In all IPRs nucleofection efficiency is higher than before.

#Data arrangement in RPE1 cells
```{r, fig.width=3, fig.height=4}
#Merge transfection and nucleofection examples
editing_efficiency_RPE1 <-  revision_data_RPE1 %>% mutate(TIF = 1 - (wt_0/total_IPR_reads)) %>% filter(gRNA == "LBR2") %>% select(barcode, cell_line,rep = bio_rep, TIF)
editing_efficiency_PRO <-  revision_data_PRO %>% mutate(TIF = 1 - (wt_0/total_IPR_reads)) %>% filter(gRNA == "halfLBR2") %>% select(barcode, cell_line,rep = bio_rep, TIF)
editing_efficiency_DEF <-  revision_data_DEF %>% mutate(TIF = 1 - (wt_0/total_IPR_reads)) %>% filter(gRNA == "LBR2") %>% select(barcode, cell_line,rep = bio_rep, TIF)

#K562 data
editing_efficiency_K562 <- mergedt_K562_data %>% 
  filter(delivery == "lipofection" & barcode %in% transcr_reporters) %>%
  ungroup() %>%
  dplyr::group_by(rep) %>%
  dplyr::summarise(mean_TIF = mean(freqCut)) %>%
  mutate(cell_line = "K562")



#Merge RPE1 data
merge_RPE1_data_TIF <- bind_rows(editing_efficiency_RPE1, editing_efficiency_PRO, editing_efficiency_DEF) %>% 
  dplyr::group_by(rep,cell_line) %>% 
  dplyr::summarise(mean_TIF = mean(TIF)) %>%
  bind_rows(editing_efficiency_K562)

#Calcualte TIF
t_test_TIF <- merge_RPE1_data_TIF %>%
  ungroup() %>%
  t_test(mean_TIF ~ cell_line, ref.group = "K562", alternative = "less")

#Plot as beeswarm plot
pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S6_editing_efficiency/xv20231006_TIF_validation_experiment.pdf", width = 3, height= 4)
ggplot(merge_RPE1_data_TIF,aes(fct_relevel(cell_line, c("K562","DEF","PRO","RPE1")), mean_TIF)) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom() +
  ylab("Total indel frequency") +
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())
dev.off()
```

#How to plot this effects?
```{r, fig.width=3, fig.height=4}
global_effects_plot <- CCD_effects_all_cells %>% select(gene,global_diff, cell_line) %>% distinct() %>% na.omit()
#Overall ∆log2 balance global changes
pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S6_editing_efficiency/xv20231006_global_effect.pdf", width = 3, height= 4)
ggplot(global_effects_plot, aes(fct_relevel(cell_line, c("K562","DEF","PRO","RPE1")), abs(global_diff))) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom() +
  ylab("abs(Global ∆log2 MMEJ::NHEJ)") +
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())
dev.off()

#Perform test (differences between cell lines?)

global_effect_test <- map_dfr(c("DEF","PRO","RPE1"), function(x) {
  cell_line1 <- CCD_effects_all_cells %>% filter(cell_line == "K562" & toxic == F & gRNA != "NTC") %>% select(cell_line, gRNA, global_diff) 
  cell_line2 <- CCD_effects_all_cells %>% filter(cell_line == x & toxic == F & gRNA != "NTC") %>% select(cell_line, gRNA, global_diff)
  proteins <- intersect(cell_line1$gRNA, cell_line2$gRNA)
  cell_lines <- bind_rows(cell_line1, cell_line2) %>% filter(gRNA %in% proteins) %>% distinct() %>% mutate(abs_global = abs(global_diff))
  cell_lines %>% ungroup() %>% wilcox_test(abs_global ~ cell_line, paired = T, ref.group = "K562", alternative = "less")
  }) %>% mutate(fdr = p.adjust(p, method = "BH"))

```
#Conclusion: Effect sizes are also larger RPE1 cells compared to K562

#Is this true for CCDs?
#How to plot this effects?
```{r, fig.width=2, fig.height=3, warning=F}
#Create data frame with only matching pairs
CCD_effect_matched <- map(c("DEF","PRO","RPE1"), function(x) {
  cell_line1 <- CCD_effects_all_cells %>% filter(cell_line == "K562" & toxic == F & gRNA != "NTC" & CCD_score != 0) %>% select(cell_line, gRNA, CCD_score, chrom_feature) %>% mutate(assay = paste(gRNA, chrom_feature, sign(CCD_score), sep = "_"))
  cell_line2 <- CCD_effects_all_cells %>% filter(cell_line == x & toxic == F & gRNA != "NTC" & CCD_score != 0) %>% select(cell_line, gRNA, CCD_score, chrom_feature) %>% mutate(assay = paste(gRNA, chrom_feature, sign(CCD_score), sep = "_"))
  proteins <- intersect(cell_line1$assay, cell_line2$assay)
  assay <- tibble(assay = proteins) %>% separate(assay, into = c("gRNA","chrom_feature"))
  cell_lines <- bind_rows(cell_line1, cell_line2) %>% right_join(assay) %>% distinct() %>% mutate(effect_CCD = abs(CCD_score)) %>% na.omit()
  })

#Overall ∆log2 balance global changes (K562 vs. DEF)
ggplot(CCD_effect_matched[[1]] %>% na.omit() %>% filter(CCD_score != 0), aes(fct_relevel(cell_line, c("K562","DEF")), effect_CCD)) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom(alpha = 0.5) +
  ylab("abs(CCD score)") +
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())

ggplot(CCD_effect_matched[[2]] %>% na.omit() %>% filter(CCD_score != 0), aes(fct_relevel(cell_line, c("K562","PRO")), abs(CCD_score))) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom(alpha = 0.5) +
  ylab("abs(CCD score)") +
  theme_bw()  +
  scale_fill_manual(values = c("grey70","#95c482")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())

ggplot(CCD_effect_matched[[3]] %>% na.omit() %>% filter(CCD_score != 0), aes(fct_relevel(cell_line, c("K562","RPE1")), abs(CCD_score))) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom(alpha = 0.5) +
  ylab("abs(CCD score)") +
  theme_bw()  +
  scale_fill_manual(values = c("grey70","#98bad5")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())

#Perform test (differences between cell lines?)
cell_lines_pairs <- t(combn(c("K562","DEF","PRO","RPE1"), 2)) %>% as_tibble()

CCD_effect_test <- map_dfr(c("DEF","PRO","RPE1"), function(x) {
  cell_line1 <- CCD_effects_all_cells %>% filter(cell_line == "K562" & toxic == F & gRNA != "NTC" & CCD_score != 0) %>% select(cell_line, gRNA, CCD_score, chrom_feature) %>% mutate(assay = paste(gRNA, chrom_feature, sign(CCD_score), sep = "_"))
  cell_line2 <- CCD_effects_all_cells %>% filter(cell_line == x & toxic == F & gRNA != "NTC" & CCD_score != 0) %>% select(cell_line, gRNA, CCD_score, chrom_feature) %>% mutate(assay = paste(gRNA, chrom_feature, sign(CCD_score), sep = "_"))
  proteins <- intersect(cell_line1$assay, cell_line2$assay)
  assay <- tibble(assay = proteins) %>% separate(assay, into = c("gRNA","chrom_feature"))
  cell_lines <- bind_rows(cell_line1, cell_line2) %>% right_join(assay) %>% distinct() %>% mutate(effect_CCD = abs(CCD_score))
  cell_lines %>% ungroup() %>% wilcox_test(effect_CCD ~ cell_line, ref.group = "K562", alternative = "less")
  }) %>% mutate(fdr = p.adjust(p, method = "BH"))
```

#Is this true for CCDs? - Take the strongest per gene
#How to plot this effects?
```{r, fig.width=3, fig.height=4, warning=F}
chrom_features_RPE <- CCD_effects_all_cells %>% filter(cell_line == "DEF") %>% pull(chrom_feature) %>% unique()


#Create data frame with only matching pairs
strongest_CCD_proteins <- CCD_effects_all_cells %>%
  filter(toxic == F & CCD_score != 0 & chrom_feature %in% chrom_features_RPE & fdr < 0.25) %>%
  mutate(effect_CCD = abs(CCD_score)) %>%
  dplyr::group_by(gene,cell_line) %>%
  dplyr::slice_max(effect_CCD)

pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S6_editing_efficiency/xv20231006_CCD_effect.pdf", width = 3, height= 4)
ggplot(strongest_CCD_proteins, aes(fct_relevel(cell_line, c("K562","DEF","PRO","RPE1")), effect_CCD)) +
  stat_summary(geom = "col",fun.y = "mean", aes(fill = cell_line)) +
  geom_quasirandom() +
  ylab("abs(Max. CCD ∆log2MMEJ:NHEJ)") +
  theme_bw()  +
  scale_fill_manual(values = c("#9f86bf","grey70","#95c482","#98bad5")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())
dev.off()


CCD_greater_test <- map_dfr(c("DEF","PRO","RPE1"), function(x) {
  cell_line1 <- strongest_CCD_proteins %>% filter(cell_line == "K562" & toxic == F & gRNA != "NTC" & fdr < 0.25) %>% select(cell_line, gRNA, effect_CCD) 
  cell_line2 <- strongest_CCD_proteins %>% filter(cell_line == x & toxic == F & gRNA != "NTC" & fdr < 0.25) %>% select(cell_line, gRNA, effect_CCD)
  proteins <- intersect(cell_line1$gRNA, cell_line2$gRNA)
  cell_lines <- bind_rows(cell_line1, cell_line2) %>% filter(gRNA %in% proteins) %>% distinct()
  cell_lines %>% ungroup() %>% t_test(effect_CCD ~ cell_line, paired = T, ref.group = "K562", alternative = "less")
  }) %>% mutate(fdr = p.adjust(p, method = "fdr"))

```


#Are effect sizes bigger if editing frequency is bigger: Arrange data
```{r}
#Create table with mean ∆log2balance and TIF
merge_RPE1_data_bal <- bind_rows(DEF_log2, PRO_log2 , RPE1_log2) %>%
  filter(!gRNA %in% c("halfLBR2","LBR2","empty"))

#Extract controls
control_RPE1_cells <- filter(merge_RPE1_data_bal, gRNA == "NTC") %>%
  select(barcode, cell_line, dmso_log2_bal = log2_bal)

#Left_join
RPE1_cells <- merge_RPE1_data_bal %>% left_join(control_RPE1_cells) %>%
  mutate(delta_log2_bal = log2_bal- dmso_log2_bal) %>%
  dplyr::group_by(cell_line, bio_rep, gRNA) %>%
  dplyr::summarise(mean_delta = abs(mean(delta_log2_bal, na.rm = T))) %>%
  ungroup() %>%
  dplyr::group_by(bio_rep, cell_line) %>%
  dplyr::summarise(overall_delta = mean(mean_delta))  %>%
  select(rep = bio_rep, overall_delta, cell_line)


#K562 log2 plot
control_K562_log2 <- K562_log2 %>% 
  filter(sample == "WT") %>%
  dplyr::group_by(barcode, rep) %>%
  dplyr::summarise(dmso_control = mean(log2MMEJNHEJratio))

K562_log2_data <- K562_log2 %>%
  left_join(proteins_gRNA) %>%
  na.omit() %>%
  left_join(control_K562_log2) %>%
  mutate(delta_log2_bal = log2MMEJNHEJratio- dmso_control) %>%
  dplyr::group_by(rep, gRNA) %>%
  dplyr::summarise(mean_delta = abs(mean(delta_log2_bal, na.rm = T))) %>%
  ungroup() %>%
  dplyr::group_by(rep) %>%
  dplyr::summarise(overall_delta = mean(mean_delta)) %>%
  mutate(cell_line = "K562")


```

#Is there correlation between editing frequency and effect sizes?
```{r}
#Bind editing and ∆log2 balance
correlation_effect_size <- K562_log2_data %>% 
  bind_rows(RPE1_cells) %>%
  left_join(merge_RPE1_data_TIF)

#Are these two variable correlated
correlation_test <- cor.test(correlation_effect_size$overall_delta, correlation_effect_size$mean_TIF, method = "spearman") %>% tidy()

#plot scatter plot
ggplot(correlation_effect_size, aes(overall_delta, mean_TIF)) +
  geom_point(aes(color = cell_line)) +
  stat_cor(method = "spearman",label.x = 1.5, label.y.npc = "bottom") + 
  theme_bw() + ylab("Total indel frequency per replicate") + xlab("Average ∆log2 MMEJ::NHEJ per replicate")  +
  scale_color_manual(values = c("#9f86bf","grey70","#95c482","#98bad5"))
```

```

