---
title: "xv20230831_CCD_RPE1_validation"
author: "Xabier Vergara"
date: "2023-08-31"
output: html_document
---

Aim: In this file, I will explore some of the possible plots, where I get general conclusions from K562 and RPE1wt data.

```{r setup, include=FALSE}
library(tidyverse)
```

# Import data

```{r}
#Proteins_validation
proteins_gRNA <- tibble(gRNA = c("36","38","40","42","44","46","48","50","52","54","56","58","60","62","64","66","68","70","72","74"),
                        gene = c("ATM","ATR","BLM","BOD1L1","BRCA2","BRCC3","CHAF1A","CHEK2","FAAP24","FANCD2","FANCG","FANCM","MDC1","PARP1","POLL","RAD50","RBBP8","RMI2","SMC5","TOPBP1"))

#Processed data
RPE1wt_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230831_CCD_summary_table_RPE1wt.rds") %>% left_join(proteins_gRNA) %>% mutate(chrom_feature = str_extract(pattern = "(?<=_).*", feature))
K562_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220819_Table_S7_DR_screen_results.rds") %>% left_join(proteins_gRNA)

#All log2_values
RPE1wt_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230831_delta_log2_balance_RPE1wt.rds") %>% left_join(proteins_gRNA)
K562_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_log2_MMEJNHEJ_differentials_chromatin_KO.rds") %>% left_join(proteins_gRNA)

#K562 data on proteins in the validation screen
K562_validation_prots <- K562_CCD %>% na.omit() %>% filter(chrom_feature %in% c("H3K4me1","H3K36me3","LMNB1","H3K27me3"))

```

#Plot correlation of global effects
```{r pressure, echo=FALSE}
#Global effect correlation data
global_effect_correlation <- RPE1wt_CCD %>% 
  select(gene,gRNA,RPE_global = global_effect) %>% 
  left_join(K562_validation_prots %>% 
              select(gene,gRNA, K562_global = global_diff)) %>%
  distinct()

#Plot this global effect
ggplot(global_effect_correlation,aes(RPE_global,K562_global, gene = gene)) + 
  geom_smooth(method = "lm", se = F) +
  geom_point() +
  ggpubr::stat_cor() +
  theme_bw() + 
  geom_abline(linetype = 2) +
  coord_fixed()

```

#Plot correlation of CCD effects (How similar are they?)
```{r pressure, echo=FALSE}
#Global effect correlation data
CCD_effect_correlation <- RPE1wt_CCD %>%
  filter(CCD_p.adj < 0.1) %>%
  select(gene,gRNA,chrom_feature,RPE_CCD = CCD_value) %>% 
  left_join(K562_validation_prots %>% 
              select(gene,gRNA,chrom_feature, K562_CCD = CCD_synergy_score))

#Calculate cosine distance per gene
K562_RPE1_distance <- map_dfr(unique(CCD_effect_correlation$gene), function(x){
  CCD_table <- CCD_effect_correlation %>%
    filter(gene == x) %>%
    select(RPE_CCD, K562_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Bind to initial data table
CCD_correlation_cosine <- CCD_effect_correlation %>% left_join(K562_RPE1_distance)

```
#Conclusion: Cosine distances look quite good overall

#Conclusion: Most of the proteins we highlight in the screen, have similar synergies

#Calculate cosine distance for all proteins
```{r}
all_validation <- RPE1wt_CCD %>%
  select(gene,gRNA,chrom_feature,RPE_CCD = CCD_value,CCD) %>% 
  left_join(K562_validation_prots %>% 
              select(gene,gRNA,chrom_feature, K562_CCD = CCD_synergy_score))

#Calculate cosine distance per gene
K562_RPE1_distance_all <- map_dfr(unique(all_validation$gene), function(x){
  CCD_table <- all_validation %>%
    filter(gene == x) %>%
    select(RPE_CCD, K562_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

ggplot(K562_RPE1_distance_all %>% select(gene,cosine_dist) %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist), fill = ) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("Protein") 

```

#Plot some of them as data points
```{R}
#Modify cell lines
RPE1_modified <- RPE1wt_log2 %>% 
  select(barcode,gRNA, gene, mean_diff, H3K4me1 = chip_H3K4me1, H3K36me3 = chip_H3K36me3, H3K27me3 = dam_H3K27me3, LMNB1 = dam_LMNB1) %>% 
  mutate(cell_line = "RPE1")

#K562_modified_data
K562_modified <- K562_log2 %>% 
  select(barcode,gRNA, gene, mean_diff=mean.log2foldchange, H3K4me1, H3K36me3, H3K27me3, LMNB1) %>% 
  mutate(cell_line = "K562")

#Generate a data frame for both
combined <- bind_rows(RPE1_modified, K562_modified)

#ATMko and LMNB1
ggplot(combined %>% filter(gene == "ATM"),aes(LMNB1, mean_diff, color = cell_line)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

#PARP1ko and H3K27me3
ggplot(combined %>% filter(gene == "PARP1"),aes(H3K4me1, mean_diff, color = cell_line)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

#SMC5ko and H3K27me3
ggplot(combined %>% filter(gene == "SMC5"),aes(LMNB1, mean_diff, color = cell_line)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

#FANCD2ko and H3K27me3
ggplot(combined %>% filter(gene == "FANCD2"),aes(LMNB1, mean_diff, color = cell_line)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

#BRCA2ko and H3K4me1
ggplot(combined %>% filter(gene == "BRCA2"),aes(H3K4me1, mean_diff, color = cell_line)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw()

```
#Conclusion: With very noisy data, we validated several phenotypes! I am not sure this will get better even if we sequence deeper. 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
