---
title: "xv20230831_CCD_RPE1_validation"
author: "Xabier Vergara"
date: "2023-08-31"
output: html_document
---

Aim: In this file, I will explore some of the possible plots, where I get general conclusions from K562 and RPE1wt data.

```{r setup, include=FALSE}
library(tidyverse)
```

# Import data

```{r}
#Proteins_validation
proteins_gRNA <- tibble(gRNA = c("36","38","40","42","44","46","48","50","52","54","56","58","60","62","64","66","68","70","72","74"),
                        gene = c("ATM","ATR","BLM","BOD1L1","BRCA2","BRCC3","CHAF1A","CHEK2","FAAP24","FANCD2","FANCG","FANCM","MDC1","PARP1","POLL","RAD50","RBBP8","RMI2","SMC5","TOPBP1"))

#Processed data
PRO_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230908_CCD_analysis_summary_PRO_3reps.rds") %>% left_join(proteins_gRNA) %>% mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature))
K562_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220819_Table_S7_DR_screen_results.rds") %>% left_join(proteins_gRNA)
RPE1_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_CCD_analysis_summary_RPE1_freq_0.002_n3.rds") %>% left_join(proteins_gRNA) %>% mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature))

#All log2_values
RPE1wt_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_log2_balance_chromatin_RPE1.rds") %>% left_join(proteins_gRNA)
K562_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_log2_MMEJNHEJ_differentials_chromatin_KO.rds") %>% left_join(proteins_gRNA)

#K562 data on proteins in the validation screen
K562_validation_prots <- K562_CCD %>% na.omit() %>% filter(chrom_feature %in% c("H3K4me1","H3K36me3","LMNB1","H3K27me3","H3K27ac","H3K4me2","H3K4me3","H3K9me3","H3K9me2","late_replicating"))

```

#Plot correlation of global effects
```{r pressure, echo=FALSE}
#Global effect correlation data
global_effect_correlation <- RPE1_CCD %>% 
  select(gene,gRNA,RPE1_global = global_diff) %>% 
  left_join(K562_validation_prots %>% 
              select(gene,gRNA, K562_global = global_diff)) %>%
  distinct()

#Plot this global effect
ggplot(global_effect_correlation,aes(RPE1_global,K562_global)) + 
  geom_smooth(method = "lm", se = F) +
  geom_point() +
  ggpubr::stat_cor(method = "spearman") +
  theme_bw() + 
  geom_abline(linetype = 2) +
  coord_fixed()

```

#Plot M- / N- synergies
```{r, fig.width=2.5,fig.height=2.5}
#Create a table to plot total M- & N-synergies
M_N_synergy_summary <- RPE1_CCD %>% 
  na.omit() %>%
  select(gRNA, cell_line, fdr, pathway, gene) %>%
  distinct() %>%
  dplyr::group_by(pathway, cell_line) %>%
  dplyr::summarise(fdr_0.1 = sum(fdr < 0.1),
                   ns = sum(fdr > 0.1)) %>%
  reshape2::melt()

#Which proteins have M- N-synergies & both synergies
ggplot(M_N_synergy_summary) +
  geom_col(aes(fct_relevel(pathway,c("MMEJ","NHEJ","both")),value, fill = variable)) +
  facet_wrap(~ variable) + theme_bw() +
  theme(legend.position = "none") + 
  ylab("Numbef of proteins") +
  xlab("Type of CCD") +
  scale_fill_manual(values = c("coral", "grey70"))

```


#Calculate cosine distance for all proteins
```{r}
all_validation <- RPE1_CCD %>%
  select(gene,gRNA,chrom_feature,RPE1_CCD = CCD_score) %>% 
  left_join(K562_CCD %>% 
              select(gene,gRNA,chrom_feature, K562_CCD = CCD_synergy_score)) %>%
  na.omit()

#Calculate cosine distance per gene
K562_RPE1_distance_all <- map_dfr(unique(all_validation$gene), function(x){
  CCD_table <- all_validation %>%
    filter(gene == x) %>%
    select(RPE1_CCD, K562_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Join CCD_scores
K562_RPE_distances <- K562_RPE1_distance_all %>% dplyr::left_join(RPE1_CCD %>% select(gene, gRNA,fdr)) 

ggplot(K562_RPE_distances %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist, fill = fdr < 0.1)) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("Protein") + 
  theme_bw() + 
  theme(axis.text.x =  element_text(angle = 90,
                                    hjust = 1,
                                    vjust = 0.5)) +
  ggtitle("CCD pattern similarity", 
          subtitle = "Comparison between RPE1-hTERT p53 KO & K562") +
  ylab("Cosine similarity score") + 
  scale_fill_manual(values = c("grey70", "coral"))
```

#Plot some examples
```{r, fig.width=4,fig.height=4}

RPE1_modified <- RPE1wt_log2 %>% 
  filter(binsize == 2000) %>%
  select(barcode,gRNA, gene, mean_diff, H3K4me1 = chip_H3K4me1, H3K36me3 = chip_H3K36me3, H3K27me3 = dam_H3K27me3, LMNB1 = dam_LMNB1, H3K9me3 = dam_H3K9me3, H3K4me3 = chip_H3K4me3, late_replicating) %>% 
  mutate(cell_line = "RPE1wt")

#K562_modified_data
K562_modified <- K562_log2 %>% 
  select(barcode,gRNA, gene, mean_diff=mean.log2foldchange, H3K4me1, H3K36me3, H3K27me3, LMNB1, H3K9me3, H3K4me3, late_replicating) %>% 
  mutate(cell_line = "K562")

#Generate a data frame for both
combined <- bind_rows(RPE1_modified, K562_modified)

#Plot ATM-KO vs. LMNB1
#PARP1 and H3K4me3
ggplot(combined %>% filter(gene == "PARP1"),aes(H3K4me3, mean_diff, color = fct_relevel(cell_line,c("PRO","K562")))) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw() +
  ylab("∆log2 MMEJ::NHEJ PARP1-KO") +
  theme(legend.position = "top")  +
  labs(color = "Cell line")


#ATR and H3K9me3
ggplot(combined %>% filter(gene == "ATM" & cell_line != "RPE1"),aes(H3K9me3, mean_diff, color = fct_relevel(cell_line,c("PRO","K562")))) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw() +
  ylab("∆log2 MMEJ::NHEJ ATR-KO") +
  theme(legend.position = "top") +
  labs(color = "Cell line")

#FANCD2 and late_replicating
ggplot(combined %>% filter(gene == "FANCD2" & cell_line != "RPE1"),aes(LMNB1, mean_diff, color = fct_relevel(cell_line,c("PRO","K562")))) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = F) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw() +
  ylab("∆log2 MMEJ::NHEJ ATR-KO") +
  theme(legend.position = "top") +
  labs(color = "Cell line")

```



#Global effects
```{r, fig.width=6, fig.height = 3}
#Gene order based on K562 data
combined_order <- combined %>% filter(cell_line == "RPE1") %>% na.omit() %>% dplyr::group_by(gene,gRNA) %>% dplyr::summarise(mean = mean(mean_diff)) %>% arrange(mean) %>% pull(gene) 

ggplot(combined %>% na.omit() %>% filter(cell_line != "RPE1")) +
  stat_summary(aes(fct_relevel(gene,combined_order),mean_diff,fill = fct_relevel(cell_line, c("PRO","K562"))),  position = position_dodge2(preserve = "single"), geom = "bar", fun.data = "mean_sd") + 
  stat_summary(aes(fct_relevel(gene,combined_order),mean_diff, group = fct_relevel(cell_line, c("PRO","K562"))), position = position_dodge2(preserve = "single"), geom = "errorbar", fun.data = "mean_sd",width = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5),
        legend.position = "top",
        axis.title.x = element_blank()) +
  geom_hline(yintercept = 0, linetype = 2) +
  ylab("∆log2 MMEJ::NHEJ") +
  labs(fill = "Cell line")

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
