---
title: "xv20230831_CCD_RPE1_validation"
author: "Xabier Vergara"
date: "2023-08-31"
output: html_document
---

Aim: In this file, I will explore some of the possible plots, where I get general conclusions from K562 and RPE1wt data.

```{r setup, include=FALSE}
library(tidyverse)
```

# Import data

```{r}
#Proteins_validation
proteins_gRNA <- tibble(gRNA = c("36","38","40","42","44","46","48","50","52","54","56","58","60","62","64","66","68","70","72","74"),
                        gene = c("ATM","ATR","BLM","BOD1L1","BRCA2","BRCC3","CHAF1A","CHEK2","FAAP24","FANCD2","FANCG","FANCM","MDC1","PARP1","POLL","RAD50","RBBP8","RMI2","SMC5","TOPBP1"))

#Processed data
PRO_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_CCD_analysis_summary_PRO_freq_0.0025_n3.rds") %>% 
  left_join(proteins_gRNA) %>% 
  mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature)) %>%
  select(gene, fdr, chrom_feature, CCD_score,global_diff, global_p_adj, gRNA, cell_line, pathway)


K562_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220819_Table_S7_DR_screen_results.rds") %>% 
  left_join(proteins_gRNA) %>%
  mutate(cell_line = "K562") %>%
  select(gene, fdr = CCD_model_p_adj, chrom_feature, CCD_score =CCD_synergy_score,global_diff, global_p_adj, gRNA, cell_line) %>% na.omit() %>%
  dplyr::group_by(gene) %>%
  dplyr::mutate(pathway = case_when(sum(CCD_score < 0) == 0 & sum(CCD_score > 0) != 0 ~ "NHEJ",
                                    sum(CCD_score > 0) == 0 & sum(CCD_score < 0) != 0 ~ "MMEJ",
                                    sum(CCD_score > 0) != 0 & sum(CCD_score < 0) != 0 ~ "both",
                                    T ~ "none"))

RPE1_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_CCD_analysis_summary_RPE1_5reads_n3.rds") %>% 
  left_join(proteins_gRNA) %>% 
  mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature)) %>%
  select(gene, fdr, chrom_feature, CCD_score,global_diff, global_p_adj, gRNA, cell_line, pathway)

DEF_CCD <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_CCD_analysis_summary_DEF_all_viable_3reps.rds") %>% 
  left_join(proteins_gRNA) %>% 
  mutate(chrom_feature = str_extract(pattern = "(H3|l|L).*", feature)) %>%
  select(gene, fdr, chrom_feature, CCD_score,global_diff, global_p_adj,gRNA, cell_line, pathway)

#All combined CCD summary table
all_cell_lines_summary_CCD <- bind_rows(PRO_CCD, K562_CCD, DEF_CCD, RPE1_CCD)

viability_data <- readRDS(file= "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230908_dna_quantification_viability.rds")

#All log2_values
PRO_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_log2_balance_chromatin_PRO.rds") %>% left_join(proteins_gRNA)
K562_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_log2_MMEJNHEJ_differentials_chromatin_KO.rds") %>% left_join(proteins_gRNA) %>% na.omit()
DEF_log2 <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230911_log2_balance_chromatin_DEF.rds") %>% left_join(proteins_gRNA) %>% filter(reps == 3)

#K562 data on proteins in the validation screen
K562_validation_prots <- K562_CCD %>% na.omit() %>% filter(chrom_feature %in% c("H3K4me1","H3K36me3","LMNB1","H3K27me3","H3K27ac","H3K4me2","H3K4me3","H3K9me3","H3K9me2","late_replicating"))

#Export data frame
chromatin_data_RPE1 <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230806_DSB_TRIP_pool_chromatin_data.rds")

```

#Load control data
```{r}
#Clone #5 (screen)
k562_control <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_analysis/xv20220819_screening_raw_data_repair_metrics.rds")

#RPE1 BRCA1/p53 dKO (LBR2)
def_control <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_DEF_control_data.rds")

#RPE1 p53 KO (halfLBR2)
pro_control <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_PRO_control_data.rds")

#RPE1 parental (halfLBR2)
rpe_control <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230912_RPE1_control_data.rds")

```


#Create a data-frame with LBR2 samples
```{r}
#For K562
K562_wt <- k562_control %>% filter(sample == "WT") %>%
  dplyr::group_by(barcode) %>%
  dplyr::summarise(mean_balance = mean(log2MMEJNHEJratio)) %>%
  mutate(cell_line = "K562")

#RPE1 BRCA1/p53 dKO (LBR2)
DEF_wt <- def_control %>% filter(gRNA == "LBR2") %>%
  dplyr::group_by(barcode) %>%
  dplyr::summarise(mean_balance = mean(log2_bal)) %>%
  mutate(cell_line = "DEF")

#RPE1 p53 KO (halfLBR2)
PRO_wt <- pro_control %>% filter(gRNA == "halfLBR2") %>%
  dplyr::group_by(barcode) %>%
  dplyr::summarise(mean_balance = mean(log2_bal)) %>%
  mutate(cell_line = "PRO")

#Bind all in the same dataframe
wt_all_cell_lines <- bind_rows(K562_wt, DEF_wt, PRO_wt)

#Wilcox test difference
wilcox_test <- wt_all_cell_lines %>%
  rstatix::wilcox_test(mean_balance ~ cell_line)

```

#Plot balance in all cell lines
```{R, fig.width = 2, fig.height= 3}

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230914_PRO_DEF_mean_effect_balance.pdf",width = 2, height =  3)
ggplot(wt_all_cell_lines %>% filter(cell_line != "K562"), aes(fct_relevel(cell_line,c("PRO","DEF")), mean_balance, color = cell_line)) + 
  geom_quasirandom() +
  theme_bw() + 
  xlab("Cell line") + 
  ylab("log2 MMEJ::NHEJ balance") +
  ylim(c(-3,0)) +
  stat_summary(geom = "hpline", color = "grey20") +
  scale_color_manual(values = c("#95c482","#9f86bf")) +
  theme(legend.position =  "none")
dev.off()


```



#Global effects: I need to import all the log2 values from K562, PRO and DEF
```{r, fig.width=8, fig.height = 4}
#Combine data files
K562_global_effects <- K562_log2 %>% select(gene, barcode, mean_log2_bal = mean.log2foldchange) %>% mutate(cell_line = "K562")
DEF_global_effects <- DEF_log2 %>% select(gene, barcode, mean_log2_bal = mean_diff) %>% mutate(cell_line = "DEF")
PRO_global_effects <- PRO_log2 %>% select(gene, barcode, mean_log2_bal = mean_diff) %>% mutate(cell_line = "PRO")

three_cell_lines_log2 <- bind_rows(K562_global_effects, DEF_global_effects, PRO_global_effects)


#Gene order based on all data
combined_order <- three_cell_lines_log2 %>% filter(cell_line == "K562") %>% dplyr::group_by(gene) %>% dplyr::summarise(avg_balance = mean(mean_log2_bal)) %>% arrange(avg_balance) %>% pull(gene)

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230914_global_effect_all_cells.pdf",width = 8, height =  4)
ggplot(three_cell_lines_log2 %>% na.omit()) +
  stat_summary(aes(fct_relevel(gene,combined_order),mean_log2_bal,fill = fct_relevel(cell_line, c("PRO","DEF","K562"))),  position = position_dodge2(preserve = "single"), geom = "bar", fun.data = "mean_sd") + 
  stat_summary(aes(fct_relevel(gene,combined_order),mean_log2_bal, group = fct_relevel(cell_line, c("PRO","DEF","K562"))), position = position_dodge2(preserve = "single"), geom = "errorbar", fun.data = "mean_sd",width = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5),
        legend.position = "top",
        axis.title.x = element_blank()) +
  geom_hline(yintercept = 0, linetype = 2) +
  ylab("∆log2 MMEJ::NHEJ") +
  labs(fill = "Cell line") +
  scale_fill_manual(values = c("#9f86bf","#95c482","grey90"))
dev.off()

```

#Plot correlation between PRO & DEF

```{r}
PRO_DEF_dcast <- three_cell_lines_log2 %>% 
  distinct() %>%
  filter(cell_line != "K562") %>% 
  dplyr::group_by(gene, cell_line) %>%
  dplyr::summarise(global_diff = mean(mean_log2_bal)) %>%
  reshape2::dcast(gene ~ cell_line) %>%
  na.omit()


#Calculate linear moodel for cell_line specific differences
PRO_DEF_dt<- map_dfr(PRO_DEF_dcast$gene, function(x){
  gene_dt <- filter(three_cell_lines_log2 %>%
                      filter(gene == x & cell_line != "K562")) %>% distinct()
  lm_dt <- lm(mean_log2_bal ~ cell_line, data = gene_dt) %>% tidy() %>% mutate(gene = x)
}) %>% filter(term != "(Intercept)") %>% mutate(fdr = p.adjust(p.value, method = "BH"))
```



```{r, fig.width=3, fig.height=4}
#Calculate correlation with two gene with differential effects
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230914_correlation_PRO_DEF.pdf",width = 3, height =  3)
ggplot(PRO_DEF_dcast, aes(DEF, PRO)) + 
  geom_point(aes(color = gene %in% c("ATM","MDC1"))) +
  ggpubr::stat_cor(method = "spearman") + 
  geom_abline(linetype = 2) +
  theme_bw() + 
  coord_fixed(ylim = c(-1.5,2), xlim = c(-1.5,2), expand = F) + xlab("∆log2 MMEJ::NHEJ") +
  ylab("∆log2 MMEJ::NHEJ") +
  theme(legend.position = "none") +
  geom_text_repel(data = PRO_DEF_dcast %>% filter(gene %in% c("ATM","MDC1")),
                  aes(label = gene)) +
  scale_color_manual(values = c("grey50","#9f86bf"))
dev.off()
```

#Plot differencial effects of these genes
```{r, fig.width=3.5, fig.height=2.5}
tmp <- three_cell_lines_log2 %>%
  filter(cell_line != "K562") %>%
  distinct() %>% na.omit()

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230917_MDC1_ATM_PRO_DEF.pdf",width = 3.5, height =  2.5)
ggplot(tmp %>% filter(gene %in% c("ATM","MDC1") & cell_line != "RPE1")) +
  geom_quasirandom(aes(cell_line, mean_log2_bal, color = cell_line)) + 
  facet_wrap(~ gene) +
  geom_hline(yintercept = 0, linetype =2) +
  scale_color_manual(values = c("#9f86bf","#95c482")) +
  theme_bw() + ylab("∆log2 MMEJ:NHEJ") + xlab("Cell line")
dev.off()

```


#Plot M- and N- synergies
```{r, fig.width=7, fig.hetigh = 10}
toxic_values <- viability_data %>%
  filter(dna_score < 0.33 & cell_line != "RPE1") %>%
  ungroup() %>%
  dplyr::select(cell_line, gRNA) %>%
  distinct() %>%
  left_join(proteins_gRNA) %>% mutate(plot = "toxicity")

synergy_values <- all_cell_lines_summary_CCD %>%
  select(gRNA, cell_line, pathway,global_diff,gene,fdr) %>%
  distinct() %>%
  full_join(toxic_values) %>%
  mutate(pathway = case_when(plot == "toxicity" ~ "toxicity",
                             fdr > 0.25 ~ "none",
                             fdr <= 0.25 ~ pathway)) %>%
  filter(!is.na(gene))

#Manual order
gene_list_order <- c("TOPBP1","RBBP8","ATR","SMC5","RAD50","ATM","MDC1","CHEK2","FANCM","FANCD2","FAAP24","FANCG","PARP1","RMI2","BLM","BOD1L1","CHAF1A","BRCC3","BRCA2","POLL")

pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/RPE_PRO_DEF/xv20230915_CCD_effects_validation.pdf", width = 10, height = 5)
ggplot() + 
  geom_tile(data = synergy_values %>% filter(cell_line %in% c("K562","DEF","PRO")), aes(fct_relevel(cell_line,c("K562","DEF","PRO")), fct_relevel(gene, gene_list_order), fill =pathway)) +
  theme_bw() + coord_fixed() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(fill = "CCD type (FDR ~ 0.25)") + 
  scale_fill_manual(values = c("#849223", "#8C510A","#01665E","grey90","grey70")) +
  xlab("Cell line") + ylab("Proteins with CCDs")
dev.off()

```

#Correlation of fdr (test)
```{R}
fdr_dcast <- all_cell_lines_summary_CCD %>%
  select(gene, cell_line, CCD_score, chrom_feature) %>%
  distinct() %>%
  reshape2::dcast(gene + chrom_feature ~ cell_line, value.var = "CCD_score")

```
#Cosine distance calculations
```{r}
#Filter out values that do not reach high enough values
filter_out_chromatin <- tibble(cell_line = "PRO",
                               chrom_feature = c("H3K9me3","LMNB1","H3K27me3"))

#Filter chromatin features that are not good
filtered_CCDs <- all_cell_lines_summary_CCD %>%
  filter(!is.na(gene) & cell_line != "RPE1") %>%
  anti_join(filter_out_chromatin) %>%
  reshape2::dcast(gene + chrom_feature ~ cell_line, value.var = "CCD_score")

#Calculate cosine distance for PRO
PRO_K562_distance_all <- map_dfr(unique(filtered_CCDs$gene), function(x){
  CCD_table <- filtered_CCDs %>%
    filter(gene == x) %>%
    select(PRO, K562) %>%
    na.omit()
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Join CCD_scores
PRO_DEF_distances_all <- PRO_K562_distance_all %>% dplyr::left_join(PRO_CCD %>% select(gene, gRNA,fdr)) 

ggplot(PRO_DEF_distances_all %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist, fill = fdr < 0.25)) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("DNA repair proteins KO in validation experiment") + 
  theme_bw() + 
  theme(axis.text.x =  element_text(angle = 90,
                                    hjust = 1,
                                    vjust = 0.5)) + 
  ggtitle("CCD pattern similarity", 
           subtitle = "Comparison between RPE1-hTERT p53 KO") +
  ylab("Cosine similarity score") + 
  scale_fill_manual(values = c("grey70", "#9f86bf"))




```

#Calculate cosine distance for all proteins
```{r}
all_validation <- PRO_CCD %>%
  select(gene,gRNA,chrom_feature,PRO_CCD = CCD_score) %>% 
  left_join(DEF_CCD %>% 
                select(gene,gRNA,chrom_feature,DEF_CCD = CCD_score)) %>% 
  na.omit()

#Calculate cosine distance per gene
PRO_DEF_distance_all <- map_dfr(unique(all_validation$gene), function(x){
  CCD_table <- all_validation %>%
    filter(gene == x) %>%
    select(PRO_CCD, DEF_CCD) 
  
  #calculate concordance
  cosine_value <- cosine(as.matrix(CCD_table))[1,2]
  tibble(gene = x, cosine_dist = cosine_value)
})

#Join CCD_scores
PRO_DEF_distances_all <- PRO_DEF_distance_all %>% dplyr::left_join(DEF_CCD %>% select(gene, gRNA,fdr)) 

ggplot(PRO_DEF_distances_all %>% distinct()) +
  geom_col(aes(reorder(gene, dplyr::desc(cosine_dist)),cosine_dist, fill = fdr < 0.1)) +
  theme_bw() + 
  geom_hline(yintercept = 0, linetype =2) +
  xlab("DNA repair proteins KO in validation experiment") + 
  theme_bw() + 
  theme(axis.text.x =  element_text(angle = 90,
                                    hjust = 1,
                                    vjust = 0.5)) + 
  ggtitle("CCD pattern similarity", 
           subtitle = "Comparison between RPE1-hTERT p53 KO & p53/BRCA1 dKO") +
  ylab("Cosine similarity score") + 
  scale_fill_manual(values = c("grey70", "coral"))
```
