---
title: "xv20230123_RPE_clone_IPR"
author: "Xabier Vergara"
date: "2023-01-23"
output: html_document
---

#Info: In this document, I will analyse the mapping and pathway assay of RPE1 clones.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Script to analyse the IPRs in RPE clones

```{r library}
library(tidyverse)
library(rtracklayer)
```

#Load all clones data
```{r}
#List files
dir="/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20230123_revision_scripts/RPE_DSB_TRIP/clone_mapping/"
setwd(dir)
clone_files <- list.files(pattern= "XV", full.names = T)

clone_integrations <- map_dfr(clone_files, function(x){
  setwd(dir)
  read_delim(x, show_col_types = F) %>%
    mutate(filename = x) %>%
    separate(filename, into = c("a","owner","clone_number", "filetype"))
})
```
#Filter integrations with more than 5 reads
```{r pressure, echo=FALSE}
#Filtered locations
filt_clone_IPR <- clone_integrations %>% filter(read_count_1 > 1 & read_count_2 > 1)

#Are there clones that are repeated?
repeated_positions <- filt_clone_IPR %>% 
  dplyr::group_by(chr,start,end) %>%
  dplyr::summarise(count = n())

#plot IPRs
ggplot(filt_clone_IPR) + 
  geom_point(aes(clone_number, log10(read_count), color = chr)) +
  theme_bw()

ggplot(filt_clone_IPR) + 
  geom_bar(aes(clone_number)) +
  theme_bw()

```
#CONCLUSION: Clone #32 has multiple integrations, but the other clones seem to be the same clone with 2 to 3 IPRs.

#Quick control: check if they are in heterochromatin vs euchromatin. Very rough estimation taking 20kb bin surrounding the integration site
```{r}
# pA-DamID tracks
pA_DamID_dir <- "/DATA/projects/DSBrepair/data/xv20221110_E2061_pA_DamID_het/results/tracks/normalized/bin-20kb/"

#tracks
control_pA_DamID <- list.files(path = pA_DamID_dir, pattern = "tracr.*comb", full.names = T)[2:5]

#IPR_coordinates (in repeated_positions)

chromatin_values_pA_DamID <- map_dfr(control_pA_DamID, function(x){
  track <- import(x) %>% as_tibble()
  tmp <- map2_dfr(repeated_positions$chr, repeated_positions$start, function(x,y) {
    track %>% 
      filter(seqnames == x & start < y & end > y) %>%
      select(chr = seqnames, width, score) %>%
      mutate(start = y)
  })
  tmp %>% mutate(antibody = str_extract(x,"(?<=tracr_).*(?=_comb)"))
})

#Combine with the clone and position they are
position_chromatin_feat <- filt_clone_IPR %>% left_join(chromatin_values_pA_DamID)


```

#Check pathway assay
#Import files
```{r, warning=F,message=F}
# List files in mcclintock
pathway_assay_files <- list.files("/DATA/projects/DSBrepair/data/xv20230123_RPE_clones_targetted_pA_DamID/indelPCR_counts",pattern = '*[.]co', full.names = T) 

#Read and parse files

RPE1_clones_pathway_assay <- map_dfr(pathway_assay_files, function(x){
  read_delim(x, show_col_types = F) %>% 
  mutate(filename = gsub("/DATA/projects/DSBrepair/data/xv20230123_RPE_clones_targetted_pA_DamID/indelPCR_counts/","",x),
         outcome = paste(call, indel, sep = "_")) %>%
  separate(filename, into = c("seq_run","sample_ID","exp","ID","NA","clone","gRNA","replicate"))
})

#Table with curated barcodes (these are the 29 IPRs of clone 12 and three on other clones)
clone_32_bc <- tibble(clone = "32", barcode = c("TGGCCTGAGGAGGCGA","AGCTCAGACTAGATGG","GGACGTGCCAACTTGC","TGCATGTAGACCTCGG","CATATCTACTAAAGT","TCCATCTGCTATTGGT","AGCCTGATCGGATAAT","AGAGACATTTATGTAC","GAATAAGTCTTATCCG","CAATATGCCCCTTAGG","ACCCATAGCCAGCCAT","GGTTGTTAACCCGAGG","AACTTTGAACACCAAG","AGAAACTGGTGCCATG","TACTACTAGTTCGCAT","ATTCGCCTTACACAGA","TACGCATTTTTACGAT","GTATTGATAACAATCC","TCTCAATAACAGCAGG","TTGCTGGACTGTCTTT","ACGTGCTACCGTCAAC","GTACAAGCTACCTGTG","ACGTGATTGTTACGA","TTGTATTCCTAACGCC","GGCGAATTCAAGAAGT","CCTCCGTCTCACCTAC","CCGGCGAGCTTCAGCG","ACCCATCGACCACGGT","TAAAGTTATATACAAC"))

other_clones_bc <- tibble(clone = rep(c("9","11","12","14","16","19","23","27","30"), 3), barcode = c(rep("AGGATACACCACTTAT",9),rep("CCCCCAGCTGCACGCT",9),rep("CTGTCCGCATTTTTCC",9)))

#All correct barcodes
barcodes_RPE <- bind_rows(clone_32_bc,other_clones_bc)
```

#Perform some quality checks with the samples
```{r, message=F}
#How many reads I have per parcode and clone
#total counts per barcode in each clone
read_per_barcode <- RPE1_clones_pathway_assay %>%
  dplyr::group_by(barcode, clone,replicate, gRNA) %>%
  dplyr::summarise(total = sum(count))

#Plot number of total reads per barcode
ggplot(read_per_barcode) + 
  geom_point(aes(barcode,log10(total))) + 
  facet_wrap(~clone+gRNA) +
  geom_hline(yintercept = 2.5, color = "red", linetype = 2) +
  theme_bw() +
  theme(axis.text.x = element_blank())

#Filter barcode total read > 250
barcodes_RPE_clones <- read_per_barcode %>%
  filter(total > 250)

#Tile of barcodes per clone
ggplot(barcodes_RPE_clones) +
  geom_tile(aes(barcode,interaction(clone,gRNA), fill = total)) +
  facet_wrap(~replicate) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```
#CONCLUSION: The threshold between backgound integrations and real integrations is quite clear. I will filter them based on the total amount of reads. It confirms that all clones have the same three IPRs
#NOTE #2: There is something weirg, but there seem to be some cross talk between samples or clone contamination. But this is not consistent across replicates. I need to double check this.

#Calculate frequencies and create dcast tables
```{r}
# Check editing efficiency
pathway_assay_clones <- RPE1_clones_pathway_assay %>% 
  right_join(barcodes_RPE) %>%
  left_join(barcodes_RPE_clones) %>%
  na.omit() %>%
  mutate(freq = count/total,
         mut = paste(call, gsub("-","",indel), sep = "_"))

#dcast total counts
dcast_counts_clones <- pathway_assay_clones %>% 
  reshape2::dcast(barcode + exp + clone + gRNA + replicate ~ mut, value.var = "count", fill=0, fun.aggregate = mean)

#dcast total counts
dcast_freq_clones <- pathway_assay_clones %>% 
  reshape2::dcast(barcode + exp + clone + gRNA + replicate ~ mut, value.var = "freq", fill= 0, fun.aggregate = mean)

colnames(dcast_freq_clones)[6:length(colnames(dcast_freq_clones))] <- paste0("freq_", colnames(dcast_counts_clones)[6:length(dcast_counts_clones)])

#combine both
dcast_counts_freq_clones <- dcast_counts_clones %>%
  left_join(dcast_freq_clones)
```
#CONCLUSION: the samples look decent, cutting efficiency is nearly 100% at 72h post transfection

#Data processing
# I will process the data as I did previously
```{r}
# Process kinase inhibitor data
pathway_assay_clones.processed <- dcast_counts_freq_clones %>% 
  filter(del_7 + ins_1 > 30) %>% #Apply the read number filtering step
  mutate(freqCut = 1 - freq_wt_0,
         log2MMEJNHEJratio = log2(del_7 / ins_1),
         MMEJ_freq = del_7/(del_7+ins_1)) %>%
  dplyr::select(barcode, exp,clone,gRNA, replicate,log2MMEJNHEJratio,MMEJ_freq,freqCut, del_7, ins_1, wt_0)  %>% distinct()

#Filter wells with mean cutting efficiency lower than 25%
filter.out.wells <- pathway_assay_clones.processed %>% 
  dplyr::group_by(exp, replicate,gRNA,clone) %>%
  dplyr::summarise(mean.cut = mean(freqCut, na.rm = T)) %>%
  filter(mean.cut < 0.90) %>% 
  dplyr::select(exp,replicate,gRNA,clone)

#Filter based on cutting efficiency
filtered.RPE.clones.tib <- pathway_assay_clones.processed %>% 
  anti_join(filter.out.wells, by = c("exp", "replicate","gRNA","clone"))

```
#CONCLUSION: values are normalized and filtered out the samples with low editing efficiency

#Control plots
```{r}
# Reproducibility across cell replicates
dcast_replicates_log2 <- filtered.RPE.clones.tib %>% reshape2::dcast(barcode + clone + gRNA ~ replicate, value.var = "MMEJ_freq") %>% na.omit()

dcast_replicates_freq <- filtered.RPE.clones.tib %>% reshape2::dcast(barcode + clone + gRNA ~ replicate, value.var = "freqCut") %>% na.omit()

```
#CONCLUSION: Editing efficiency is nearly 100% in all of them. I only get lower editing efficiency if I transfect x10 less gRNA. Could I use that to check for differences in editing eff???
#CONCLUSION #2: Data looks noisier than in K562 cells, measurements are not so reproducible. At least in clone #32. I had problems with plating and replating, this might have affected read out.

#average values for replicates
```{r}
#Plot averages with sd
average_values <- filtered.RPE.clones.tib %>%
  dplyr::group_by(barcode,clone,gRNA) %>%
  dplyr::summarise(mean_ratio = mean(MMEJ_freq),
                   sd_ratio = sd(MMEJ_freq),
                   mean_cut = mean(freqCut),
                   sd_cut = sd(freqCut))
```

#Plot clone #32 data
```{r}
#Plot with error bars
ggplot(filter(average_values),
       aes(reorder(barcode,mean_ratio),mean_ratio)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("9","11","12","14","16","19","23","27","30")), ncol = 5) + 
  geom_errorbar(aes(ymax = mean_ratio+sd_ratio, 
                    ymin = mean_ratio-sd_ratio)) +
  ylab("MMEJ relative frequency") +
  ggtitle("MMEJ frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + xlab("IPRs in clones")

#Plot for power point
#Plot with error bars
p1 = ggplot(filter(average_values),
       aes(reorder(barcode,mean_ratio),mean_ratio)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("9","11","12","14","16","19","23","27","30")), ncol = 5) + 
  geom_errorbar(aes(ymax = mean_ratio+sd_ratio, 
                    ymin = mean_ratio-sd_ratio)) +
  ylab("MMEJ relative frequency") +
  ggtitle("MMEJ frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")

ggplotly(p1)

```


#Plot clone #32 data
```{r}
#Plot with error bars
ggplot(filter(average_values),
       aes(reorder(barcode,mean_ratio),mean_cut)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("9","11","12","14","16","19","23","27","30")), ncol = 5) + 
  geom_errorbar(aes(ymax = mean_cut+sd_cut, 
                    ymin = mean_cut-sd_cut)) +
  ylab("MMEJ relative frequency") +
  ggtitle("MMEJ frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + xlab("IPRs in clones")

#Plot for power point
#Plot with error bars
ggplot(filter(average_values),
       aes(reorder(barcode,mean_ratio),mean_cut)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("9","11","12","14","16","19","23","27","30")), ncol = 5) + 
  geom_errorbar(aes(ymax = mean_cut+sd_cut, 
                    ymin = mean_cut-sd_cut)) +
  ylab("Editing relative frequency") +
  ggtitle("Editing frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")+ ylim(c(0,1))

```

## Plot DamID values for different values
```{r}
#plot as heatmap
heatmap_dt <- position_chromatin_feat %>% 
  mutate(coordinate = paste(chr, start, sep = ":")) %>% 
  reshape2::dcast(coordinate+clone_number ~ antibody, value.var = "score")

#Plot clone 32 as heatmap
clone_32_dt <- heatmap_dt %>%
  filter(clone_number == 32) %>%
  select(-clone_number) %>%
  column_to_rownames(var = "coordinate")

#plot
rg <- max(abs(clone_32_dt))
pheatmap(t(as.matrix(clone_32_dt)),
         breaks = seq(-rg, rg, length.out = 100),
         main = "pA_DamID_clone_32_RPE1")


#For other clones
others_dt <- heatmap_dt %>%
  filter(clone_number == 11) %>%
  select(-clone_number) %>%
  column_to_rownames(var = "coordinate")

#for clone 32
pheatmap(t(as.matrix(others_dt)),
         breaks = seq(-rg, rg, length.out = 100),
         main = "pA_DamID_other_clones_RPE1")
```
#CONCLUSION: Only some integrations in clone 32 are in LADs (one or two maybe)

## Plot zoom in of two types of clones
```{r}
# Editing efficiency in clone 32 & clone 11 (as example)
ggplot(filter(average_values, clone %in% c("32","11")),
       aes(reorder(barcode,mean_ratio),mean_cut)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("32","1")), ncol = 1) + 
  geom_errorbar(aes(ymax = mean_cut+sd_cut, 
                    ymin = mean_cut-sd_cut)) +
  ylab("Editing relative frequency") +
  ggtitle("Editing frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")+ ylim(c(0,1))

# Editing efficiency in clone 32 & clone 11 (as example)
ggplot(filter(average_values, clone %in% c("32","11")),
       aes(reorder(barcode,mean_ratio),mean_ratio)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("32","1")), ncol = 1) + 
  geom_errorbar(aes(ymax = mean_ratio+sd_ratio, 
                    ymin = mean_ratio-sd_ratio)) +
  ylab("MMEJ frequency") +
  ggtitle("MMEJ in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")

```

#Can we use R1 (with 10x less gRNA) to check changes in balance with different transfection eff.?
```{r}
R1_values <- pathway_assay_clones.processed %>%
  filter(replicate == "R1") %>%
  dplyr::group_by(barcode,clone,gRNA) %>%
  dplyr::summarise(R1_ratio = mean(MMEJ_freq),
                   R1_cut= mean(freqCut))

#Plot correlations

correlation_tibble <- average_values %>% left_join(R1_values)


#Individual values
# Editing efficiency in clone 32 & clone 11 (as example)
ggplot(filter(correlation_tibble, clone %in% c("32","11"))) + 
  geom_point(aes(reorder(barcode,mean_ratio),R1_cut)) + 
  facet_wrap(~ fct_relevel(clone, c("32","1")), ncol = 1) + 
  ylab("Editing relative frequency") +
  ggtitle("Editing frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")+ ylim(c(0,1))

# Editing efficiency in clone 32 & clone 11 (as example)
ggplot(filter(correlation_tibble, clone %in% c("32","11"))) + 
  geom_point(aes(reorder(barcode,mean_ratio),R1_ratio)) + 
  facet_wrap(~ fct_relevel(clone, c("32","1")), ncol = 1) + 
  ylab("MMEJ relative frequency") +
  ggtitle("MMEJ frequency in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")

# Editing efficiency in clone 32 & clone 11 (as example)
ggplot(filter(average_values, clone %in% c("32","11")),
       aes(reorder(barcode,mean_ratio),mean_ratio)) + 
  geom_point() + 
  facet_wrap(~ fct_relevel(clone, c("32","1")), ncol = 1) + 
  geom_errorbar(aes(ymax = mean_ratio+sd_ratio, 
                    ymin = mean_ratio-sd_ratio)) +
  ylab("MMEJ frequency") +
  ggtitle("MMEJ in clone 32 (29 IPRs) and other clones (3 IPRs)") + 
  theme_bw() +
  theme(axis.text.x = element_blank()) + xlab("IPRs in clones")

#MMEJ frequency
ggplot(correlation_tibble %>% filter(clone == "32"), aes(mean_ratio,R1_ratio)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(color = "blue") +
  stat_cor(method = "spearman", color = "red",label.y = 0.1) + 
  ylab("MMEJ freq in R1 (n=1, x10 times less gRNA)") +
  xlab("MMEJ freq (n =3)") +
  theme_bw()

#editing frequency
ggplot(correlation_tibble %>% filter(clone == "32"), aes(mean_cut,R1_cut)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(color = "blue") +
  stat_cor(method = "spearman", color = "red",label.y = 0.67) + 
  ylab("Editing freq in R1 (n=1, x10 times less gRNA)") +
  xlab("Editing freq (n = 3)") +
  theme_bw()

```
