---
title: "xv20230329_RPE1_mapping_clone"
author: "Xabier Vergara"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
library(broom)
library(scales)
```

## Import data: Balance data and some other details
```{r, warning=F}
# Import data for high confidence reporters and calculate mean
high_confidence_IPR <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230911_DEF_all_mean_freq_0.0075_.rds") %>%
  filter(is.finite(log2_bal))

#Import control data
control_data_DEF <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230909_DEF_control_data.rds")

#Pools used in the experiment
selected_pools <- c("RPE1_Low_1000","RPE1Deff_Low_1000","RPE1Proff_Low_250","U2OS_High_100")

#Import chromatin data
#Export data frame
chromatin_data_RPE1 <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230919_DSB_TRIP_pool_chromatin_data.rds")

```

#Calculate correlation with chromatin features in controls
```{r, fig.width=6, fig.height=5}
#Calculate average balances
average_log2_balance <- control_data_DEF %>%
  filter(gRNA == "LBR2") %>%
  dplyr::group_by(gRNA, barcode, cell_line) %>%
  dplyr::summarise(mean_log2_bal = mean(log2_bal),
                   reps = n()) %>%
  left_join(chromatin_data_RPE1 %>% select(-dam_LMNB2))


#Calculate spearman correlation coefficient per cell line and chromatin feature
spearman_coefficients <- map_dfr(unique(average_log2_balance$gRNA), function(x){
  filter_data <- average_log2_balance %>% filter(gRNA == x)
  binsize_filter <- map_dfr(binsizes, function(y){
    bin_filter <- filter_data %>% filter(binsize == y)
    rho_table <- map_dfr(colnames(bin_filter[6:15]), function(j){
      balance <- bin_filter$mean_log2_bal
      feature <- bin_filter %>% pull(j)
      cor.test(balance, feature, method = "spearman", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })
})

#Chromatin feature order for plotting
chromatin_order <- c("dam_H3K9me3","dam_H3K9me2","late_replicating","dam_LMNB1","dam_H3K27me3","chip_H3K4me3","chip_H3K4me2","chip_H3K4me1","chip_H3K27ac","chip_H3K36me3")

#Plot correlation coefficients
ggplot(spearman_coefficients %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Spearman's correlation coefficient of MMEJfreq.") + 
  facet_wrap(~ fct_relevel(binsize,binsizes)) +
  theme(legend.position = "right") +
  ggtitle("Chromatin feature correlation at different binsizes", subtitle = "RPE1-hTERT p53-BRCA1 dKO DSB-TRIP pool")
```
#Plot reproducibility of the data: Only IPRs present in three replicates
```{r, warning= T}
#Data reproducibility
plot_reproducibility_dt <- high_confidence_IPR %>%
  reshape2::dcast(barcode+cell_line + gRNA ~ bio_rep, value.var = "log2_bal") %>% na.omit()

#Data reproducibility in the screen (RPE1 data complete)

#R1 vs. R2
ggplot(plot_reproducibility_dt, aes(R1,R2)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  ggpubr::stat_cor() +
  theme_bw() +
  coord_fixed() +
  xlab("log2 MMEJ::NHEJ balance R1") +
  ylab("log2 MMEJ::NHEJ balance R2")

#R2 vs. R3
ggplot(plot_reproducibility_dt, aes(R2,R3)) + 
  geom_point() +
  geom_smooth(method = "lm") + 
  ggpubr::stat_cor() +
  theme_bw() +
  coord_fixed() +
  xlab("log2 MMEJ::NHEJ balance R2") +
  ylab("log2 MMEJ::NHEJ balance R3")

#R1 vs. R3
ggplot(plot_reproducibility_dt, aes(R1,R3)) + 
  geom_point() +
  geom_smooth(method = "lm", se = F) + 
  ggpubr::stat_cor() +
  theme_bw() +
  coord_fixed() +
  xlab("log2 MMEJ::NHEJ balance R1") +
  ylab("log2 MMEJ::NHEJ balance R3")

```

#Control MMEJ frequency: Only for barcodes with 5 reads on both MMEJ and NHEJ
```{r}
MMEJ_balance_control <- high_confidence_IPR %>%
  filter(gRNA %in% c("halfLBR2","NTC","36","64")) %>%
  mutate(gene = case_when(gRNA == "halfLBR2" ~ "LBR2",
                          gRNA == "NTC" ~ "NTC",
                          gRNA == "36" ~ "ATM",
                          gRNA == "64" ~ "POLL"))

#Plot MMEJ frequency
ggplot(MMEJ_balance_control) + 
  geom_quasirandom(aes(fct_relevel(gene,c("LBR2","NTC","ATM","POLL")),log2_bal)) + 
  facet_grid(cell_line~ bio_rep) +
  theme_bw() +
  xlab("Sample") + 
  ylab("log2 MMEJ:NHEJ balance")

```

#Calculate differentials per replicate
```{r}
dmso_balance <- high_confidence_IPR %>% filter(gRNA == "NTC") %>% ungroup() %>% select(barcode, cell_line, bio_rep, dmso_bal = log2_bal)
```

#Calculate diff for these the high_confidence barcodes
```{R}
#Calculate differential
diff_gRNA_RPE1 <- high_confidence_IPR %>% 
  left_join(dmso_balance) %>%
  mutate(gRNA_diff = log2_bal - dmso_bal) %>%
  filter(!gRNA %in% c("empty","halfLBR2","LBR2")) %>%
  na.omit()

#Average perturbation
mean_gRNA_effect <- diff_gRNA_RPE1 %>%
  dplyr::group_by(barcode, gRNA) %>%
  dplyr::summarise(mean_diff = mean(gRNA_diff),
                   reps = n())

  
```

#Run test for each individual barcode
```{r}
#calculate per barcode
IPR_direction <- diff_gRNA_RPE1 %>% 
  filter(gRNA != "NTC") %>%
  dplyr::group_by(barcode,gRNA) %>%
  mutate(reps = n()) %>%
  filter(reps == 3) %>%
  t_test(gRNA_diff ~ 1, mu = 0) %>%
  ungroup() %>%
  dplyr::group_by(gRNA)

#summarise how many MMEJ or NHEJ
summary_MMEJ <- IPR_direction %>%
  filter(p < 0.05) %>%
  dplyr::group_by(gRNA) %>%
  dplyr::summarise(IPR_M = sum(statistic < 0),
                   IPR_N = sum(statistic > 0)) %>%
  mutate(pathway = case_when(IPR_M > 0 & IPR_N == 0 ~ "MMEJ",
                   IPR_N > 0 & IPR_M == 0 ~ "NHEJ",
                   T ~ "both"))
```

#Join chromatin data
```{r}
#Create RPE1 dataframe
inhibitor_effect_chromatin_RPE1 <- mean_gRNA_effect %>%
  left_join(chromatin_data_RPE1 %>% filter(cell_line %in% c("DEF")) %>% distinct()) %>%
  filter(reps == 3)
```

#Calculate balance vs. read frequency on selected values
```{R}
#Average phenotype
control_filter_quality <- diff_gRNA_RPE1 %>%
  dplyr::group_by(barcode, cell_line, gRNA) %>%
  dplyr::summarise(mean_freq = mean(IPR_read_freq),
                   mean_bal = mean(log2_bal))

ggplot(control_filter_quality) +
  geom_point(aes(mean_freq, mean_bal)) +
  theme_bw()
```

#Check the amount of measurements per KO and filter everything below 10
```{r, fig.width=4,fig.height=3, warning =F, message=F}
#Summary
summary_IPR_KO <- inhibitor_effect_chromatin_RPE1 %>%
  filter(binsize == "2000") %>%
  dplyr::group_by(gRNA, cell_line) %>%
  dplyr::summarise(IPR_n = n())

#Plot 
ggplot(summary_IPR_KO) + 
  geom_col(aes(gRNA,IPR_n))+ 
  theme_bw() +
  ggtitle("Number of IPR per gRNA", subtitle = "RPE1-hTERT p53-BRCA1 dKO DSB-TRIP pool") +
  ylab("Number of IPR with reliable data") +
  xlab("DNA repair protein KO")

#Filter only proteins with more than 8 IPRs
prots_with_8_IPR <- summary_IPR_KO %>%
  filter(IPR_n > 5) %>%
  pull(gRNA) %>%
  unique() 

```

#Calculate global differences
```{r}
#Calculate mean effect
mean_effects <- mean_gRNA_effect %>% 
  dplyr::group_by(gRNA) %>%
  dplyr::summarise(global_diff = mean(mean_diff, na.rm = T))

#Calculate p.value calculation
p_val_global <- mean_gRNA_effect %>% 
  dplyr::group_by(gRNA) %>%
  rstatix::t_test(mean_diff ~ 0) %>%
  mutate(global_p_adj = p.adjust(p, method = "BH")) %>%
  select(gRNA,global_p = p, global_p_adj)

#Table for joining
global_summary <- left_join(mean_effects,p_val_global)

```

#PCR modeling: RPE1 data
```{r}
#PCA model
pca_model_dt <- inhibitor_effect_chromatin_RPE1 %>% filter(gRNA %in% prots_with_8_IPR & binsize == 2000) %>% left_join(proteins_gRNA) %>% na.omit()


#Create an empty dt with CCDs of DDR proteins
drug_CCDs_dt_RPE1 <- tibble(cell_line = NA, gRNA = NA,binsize = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)



  for (i in unique(pca_model_dt$gene)){
    gene.library.dt <- filter(pca_model_dt, gene == i)
    set.seed(1)
    PCR_model_DDR_test <- pls::pcr(mean_diff~ dam_H3K27me3+chip_H3K36me3+dam_LMNB1+chip_H3K4me1+chip_H3K4me2+chip_H3K4me3+dam_H3K9me2+chip_H3K27ac+dam_H3K9me2+late_replicating, data=gene.library.dt, validation = "CV") #Run principal component regression
    pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = 3) #Run model with three PC
    combined.dt <- tibble(measured = gene.library.dt$mean_diff, predicted = as.numeric(pcr_pred)) #Create table with real and predicted differentials
    pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% broom::glance() #Predicted vs. measured correlation plot
    drug_CCDs_dt_RPE1 <- drug_CCDs_dt_RPE1 %>% add_row(cell_line = j, gRNA = i, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value)) #Extract valuable metrics
  }

#Adjust per cell line
adj_drug_dt_RPE1_adjs <- map_dfr(cells, function(x) {
  drug_CCDs_dt_RPE1 %>% filter(cell_line == x) %>% mutate(fdr = p.adjust(p.value, method = "BH"))
})

```

#Calculate slopes for RPE
```{r}
#Create empty dataframe to calculate synergy scores
drug_CCD_slopes_RPE <- tibble(gRNA = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA, cell_line = NA, binsize = NA)

#Loop to run linear models on the values
for (h in unique(pca_model_dt$gRNA)) {
  for(i in unique(pca_model_dt$cell_line)) {
  for (j in colnames(pca_model_dt)[c(5:15)]) { #Run this function for each of the 25 high quality chromatin features
    model.dt <- inhibitor_effect_chromatin_RPE1 %>% filter(gRNA == h & cell_line == i & binsize == 2000) # And For each gene
   if (nrow(model.dt) == 0) {
next
}
    model.epistasis.log2 <- lm(formula = mean_diff ~ unlist(model.dt[j]), data = model.dt) %>% tidy() #Correlation analysis
   drug_CCD_slopes_RPE <- drug_CCD_slopes_RPE %>% add_row(gRNA = h,binsize =2000, feature = j,cell_line = i, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value)) #Select valuable parameters and save them in data frame
  }
  }
}

#Retain slopes that 
drug_CCD_all_values <- drug_CCD_slopes_RPE %>% 
  reshape2::dcast(gRNA + feature + cell_line ~ term, value.var = "slope.log2") %>%  #dcast table
  dplyr::select(gRNA, cell_line,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])')  %>% #Extract information for slopes only
  left_join(adj_drug_dt_RPE1_adjs) %>% 
  filter(!feature %in%  c("dam_LMNB2")) %>%
  left_join(summary_MMEJ) %>%
  left_join(global_summary) %>%
  mutate(pathway = case_when(is.na(pathway) & global_diff < 0 ~ "MMEJ",
                             is.na(pathway) & global_diff > 0 ~ "NHEJ",
                             T ~ pathway),
         CCD_score = case_when(pathway == "MMEJ" & slope < 0 ~ slope,
                               pathway == "NHEJ" & slope > 0 ~ slope,
                             T ~ 0)) %>%
  select(-binsize) %>% na.omit()

```


#Export data table
```{r}
write_rds(drug_CCD_all_values, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_CCD_analysis_summary_DEF_all_viable_3reps.rds") #In this data, all toxic KO are kept as well

#Export log2 balances
write_rds(inhibitor_effect_chromatin_RPE1, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230915_log2_balance_chromatin_DEF_all.rds") #In this data, all toxic KO are kept as well

```
