---
title: "xv20230830_validation_screen_chromatin_bias_controls"
author: "Xabier Vergara"
date: "2023-07-26"
output:
  html_document: default
  pdf_document: default
---
#Date: 30th August 2023
#Author: Xabier Vergara
#Aim: The aim of this file is to wrap up the filtering strategy and measure if pathway balance in control samples has chromatin biases in the control samples

#Load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
```


## Import data: Balance data and some other details
```{r, warning=F}
# Import data for high confidence reporters and calculate mean
high_confidence_IPR <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230830_pathway_balance_5reads_per_indel.rds") %>%
  dplyr::group_by(barcode, cell_line, gRNA) %>%
  dplyr::summarise(mean_bal = mean(log2_bal, na.rm = T),
                   reps = n())


#Pools used in the experiment
selected_pools <- c("RPE1_Low_1000","RPE1Deff_Low_1000","RPE1Proff_Low_250","U2OS_High_100")

#Pair of 
```

#Add chromatin info for RPE1 cells
#Load chromatin info of the IPRs in RPE cells (2000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_U2OS_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "2000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "2000_", full.names = T)

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, log2_mean) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "(?<=2018_).*(?=_E-MTAB)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "log2_mean")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_pools_2000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "2000")
```


#Load chromatin info of the IPRs in RPE cells (5000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_U2OS_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "5000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "5000_", full.names = T)

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, log2_mean) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "(?<=2018_).*(?=_E-MTAB)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "log2_mean")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=5000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_pools_5000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "5000")
```

#Load chromatin info of the IPRs in RPE cells (10000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_U2OS_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "10000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "10000_", full.names = T)

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, log2_mean) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "(?<=2018_).*(?=_E-MTAB)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "log2_mean")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=10000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_pools_10000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "10000")
```

#Load chromatin info of the IPRs in RPE cells (20000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_U2OS_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "20000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "20000_", full.names = T)

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, log2_mean) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "(?<=2018_).*(?=_E-MTAB)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "log2_mean")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=20000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_pools_20000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "20000")
```

#Load chromatin info of the IPRs in RPE cells (25000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_U2OS_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "250000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "250000_", full.names = T)

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, log2_mean) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "(?<=2018_).*(?=_E-MTAB)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "log2_mean")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, mean) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=250000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "mean")
}) %>% purrr::reduce(left_join, by = "name")

#Bind both
chromatin_features_pools_250000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "RPEDef",
                               iPCR_cell == "RPE1Proff" ~ "RPEPro",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "250000")
```


#Merge all binsizes in one data table
```{r}
chromatin_pools_binsize <- bind_rows(chromatin_features_pools_2000,
                                     chromatin_features_pools_5000,
                                     chromatin_features_pools_10000,
                                     chromatin_features_pools_20000,
                                     chromatin_features_pools_250000)
```


#Join with balances and calculate the correlation (do this for all marks) = Spearman's
```{r, warning=F}
#Merge RPE data with chromatin data with different bins
values_chromatin_RPE_controls <- left_join(high_confidence_IPR, chromatin_pools_binsize) %>%
  filter(gRNA %in% c("LBR2") & reps > 1) %>% na.omit()

#Calculate spearman correlation for all chromatin features
cell_lines <- c("U2OS")
binsizes <- c("2000","5000","10000","20000")

#Calculate spearman correlation coefficient per cell line and chromatin feature
spearman_coefficients <- map_dfr(cell_lines, function(x){
  filter_data <- values_chromatin_RPE_controls %>% filter(cell_line == x)
  binsize_filter <- map_dfr(binsizes, function(y){
    bin_filter <- filter_data %>% filter(binsize == y)
    rho_table <- map_dfr(colnames(bin_filter[6:14]), function(j){
      balance <- bin_filter$mean_bal
      feature <- bin_filter %>% pull(j)
      cor.test(balance, feature, method = "spearman", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })
})

#Chromatin feature order for plotting
chromatin_order <- c("dam_H3K9me3","dam_H3K9me2","late_replicating","dam_LMNB1","dam_LMNB2","dam_H3K27me3","chip_H3K4me3","chip_H3K4me2","chip_H3K4me1","chip_H3K27ac","chip_H3K36me3")

#Plot correlation coefficients
ggplot(spearman_coefficients %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Spearman's correlation coefficient of MMEJfreq.") + 
  facet_grid(fct_relevel(binsize,binsizes)~ fct_relevel(cell_line, c("RPE1","RPEPro","RPEDef")))
```
#Conclusion: There is nothing on the U2OS data. As seen before! I will not use this data in the screen.