---
title: "xv20230329_RPE1_pool_pathway assay"
author: "Xabier Vergara"
date: "2023-07-26"
output:
  html_document: default
  pdf_document: default
---
#Date: 28th August 2023
#Author: Xabier Vergara

#Aim: The aim of this file is to start with the first checks on the screening validation data. Answer questions like, do we need more sequencing depth, did all cell lines work, did all replicates work nicely?

#Conclusions: Overall, it looks better than last experiment. But, the conclusions remain very similar to the previous experiment. With the weird result on U2OS cells, that show positive correlation between balance and TIF.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
```

## Import data
```{r, warning=F, message=F}
# List files in mcclintock
pathway_assay_files <- list.files("/DATA/projects/DSBrepair/data/xv20230825_DSB_TRIP_revision/indelPCR_counts",pattern = '*[.]co', full.names = T) 

#Load sample info for the experiment
experiment_info <- readRDS("/DATA/projects/DSBrepair/data/xv20230825_DSB_TRIP_revision/xv20230825_sample_info.rds")

#Read and parse files
clone_pools_pathway_assay <- map_dfr(pathway_assay_files, function(x){
  read_delim(x, show_col_types = F) %>% 
    mutate(ID = str_extract(x,"7444_[:digit:].*(?=.count.table)"),
           outcome = paste(call, indel, sep = "_")) %>%
    left_join(experiment_info)
})
```


#Number of mapped IPRs with indel data (most optimistic)
```{r}
selected_pools <- c("RPE1_Low_1000","RPE1Deff_Low_1000","RPE1Proff_Low_250","U2OS_High_100")

#Import barcode sequences
#RPE cells
RPE_IPR <- read_delim("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/DSB_TRIP_mapping/xv20230519_RPE_pools_mapping.tsv") %>%
  select(name) %>%
  separate(name, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity"))

U2OS_IPR <- read_delim("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/DSB_TRIP_mapping/xv20230519_U2OS_pools_mapping.tsv") %>%
  select(name) %>%
  separate(name, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity"))

#Bind both
all_IPRs <- bind_rows(RPE_IPR, U2OS_IPR) %>% mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% filter(pool %in% selected_pools)
```
#CONCLUSION: We are recovering the whole complexity for U2OS & PRO (input 400 ng). With DEF & RPE we are not capturing full complexity (input 200 ng).


```{r}
#Samples with low viability
toxic_conditions <- tibble(cell_line = c(rep("PRO",3), rep("DEF",5)), gRNA = c("38","68","74","38","58","68","72","74"))

#Filter by replicate
indel_data <- clone_pools_pathway_assay %>% 
  filter(cell_line %in% c("PRO","DEF")) %>%
  anti_join(toxic_conditions) %>%
  dplyr::group_by(barcode,outcome, well, cell_line, gRNA, bio_rep) %>%
  dplyr::summarise(counts = sum(count, na.rm = F)) %>% left_join(all_IPRs %>% select(barcode, pool)) %>% na.omit()
```


#Calculate total reads per sample and per barcode
```{r}
#Reads per sample
total_sample_reads <- indel_data %>%
  dplyr::group_by(well, cell_line, gRNA,bio_rep) %>%
  dplyr::summarise(total_sample_reads = sum(counts))

#Reads per barcode 
total_IPR_reads <- indel_data %>%
  dplyr::group_by(well,cell_line, gRNA, bio_rep, barcode) %>%
  dplyr::summarise(total_IPR_reads = sum(counts))

#Left_join
relative_barcode_reads <- left_join(total_sample_reads, total_IPR_reads) %>%
  mutate(IPR_read_freq = total_IPR_reads/total_sample_reads)

```


#Compute pathway balance in controls
```{r}
#Filter signature indels in control samples
control_balance_bio <- indel_data %>% filter( outcome %in% c("ins_1","del_-7","wt_0"))

#Look at read numbers
control_balance_bio_dcast <- control_balance_bio %>% reshape2::dcast(barcode + cell_line + gRNA + bio_rep ~ outcome, value.var = "counts", fill = 0, fun.aggregate = mean)

#Control_measurements (bio_rep)
pathway_metric_control_bio <- control_balance_bio_dcast %>%
  mutate(balance = `del_-7`/(ins_1 + `del_-7`),
         log2_bal = log2(`del_-7`/ins_1),
         TIF = 1 - (wt_0/(`del_-7`+ins_1+wt_0))) %>%
  na.omit() %>% left_join(relative_barcode_reads)
```


#Plot relative read frequency
```{r}
plot_example <- pathway_metric_control_bio %>% filter(gRNA %in% c("halfLBR2","NTC","36","empty"))

#Make plots gRNA
ggplot(plot_example) +
  geom_quasirandom(aes(bio_rep, IPR_read_freq)) + 
  facet_grid(cell_line~gRNA) +
  theme_bw() 
```

#Compute data by minimum read number
```{r}
#Filter by read number and plot noise level
read_count <- c(0,1,2,5,10,15,20,50,100,200,300)

filter_read_count_test <- map_dfr(read_count , function(x) {
  pathway_metric_control_bio %>%
    filter(`del_-7` >= x & ins_1 >= x) %>%
    mutate(read_filter = x)
})

#Number of IPRs in each
IPR_summary <- filter_read_count_test %>% 
  dplyr::group_by(cell_line, gRNA, read_filter, bio_rep) %>% 
  dplyr::summarize(IPR_n = n()) 

IPR_summary_dcast <- IPR_summary %>% 
  reshape2::dcast(cell_line + gRNA + bio_rep ~ read_filter, value.var  = "IPR_n")

```
#Conclusion: we always get some barcodes, but these are extremely low in RPE1 cells


#Check reproducibility among replicates
```{r}
#Calculate co-ocurrence per replicate
replicates_IPR <- filter_read_count_test %>%
  dplyr::group_by(barcode,cell_line, read_filter,gRNA) %>%
  dplyr::summarise(rep_n = n()) %>%
  ungroup() %>%
  dplyr::group_by(cell_line, read_filter,rep_n,gRNA) %>%
  dplyr::summarise(capture_IPR = n())

total_IPR_rep <- filter_read_count_test %>%
  select(cell_line, barcode, read_filter,gRNA) %>% 
  distinct() %>%
  dplyr::group_by(cell_line,read_filter,gRNA) %>%
  dplyr::summarise(total_IPR = n())

#Calculate frequency
freq_IPR_capture <- replicates_IPR %>% 
  left_join(total_IPR_rep) %>%
  mutate(freq_capture = capture_IPR/total_IPR)

#Absolute_numbers
IPR_reps <- replicates_IPR %>% reshape2::dcast(cell_line + rep_n ~ read_filter)

```

#Calculate frequency in absolute number of IPRs
```{r}
#Total mapped IPR per pool
summary_mapped <- all_IPRs %>% 
  dplyr::group_by(iPCR_cell) %>% 
  dplyr::summarise(total_IPR = n()) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "DEF",
                               iPCR_cell == "RPE1Proff" ~ "PRO",
                               T ~ iPCR_cell)) %>%
  select(cell_line, total_IPR)

#Calculate frequency out of all IPRs mapped
freq_IPR_capture_total <- replicates_IPR %>% 
  left_join(summary_mapped) %>%
  mutate(freq_capture = capture_IPR/total_IPR)

#Frequency
ggplot(freq_IPR_capture_total) + 
  geom_col(aes(x = fct_relevel(as.character(read_filter),as.character(read_count)),
               y = freq_capture,
               fill = as.character(rep_n)),
           position = "dodge") +
  labs(fill = "IPR_reps") +
  facet_wrap(~ cell_line) + 
  theme_bw() + labs(x = "read_filter")

#How many IPRs have at least n=2
n2_IPRs <- freq_IPR_capture_total %>% 
  filter(rep_n > 1) %>%
  dplyr::group_by(cell_line, read_filter,gRNA) %>%
  dplyr::summarise(IPR_n = sum(capture_IPR)) %>%
    left_join(summary_mapped) %>%
  mutate(freq_capture_n2 = IPR_n/total_IPR)
```

#Plot summary IPR number
```{r}
#Frequency
ggplot(n2_IPRs) + 
  geom_col(aes(x = gRNA,
               y = freq_capture_n2,
               fill = fct_relevel(as.character(read_filter),as.character(read_count))),
           position = "dodge") +
  labs(fill = "read_filter") +
  facet_wrap(~ cell_line) +
  theme_bw() +
  theme(legend.position = "top")
```

# CONCLUSION: IPRs captured per sample, look better than before! 


#Is pathway balance reproducible among replicates?
```{r, message=F, warning=F}
read_counts_filter_reprod <- filter_read_count_test %>%
  reshape2::dcast(barcode + cell_line + gRNA + read_filter ~ bio_rep, value.var = "log2_bal")

#Calculate mean & sd
read_counts_filter_reprod_summary <- read_counts_filter_reprod %>% 
  rowwise() %>% 
  mutate(mean_bal = mean(c(R1,R2,R3),na.rm = T),
         sd_bal = sd(c(R1,R2,R3), na.rm = T),
         reps = 3 - sum(is.na(c(R1,R2,R3)))) %>%
  distinct()


#R1 vs. R2
ggplot(read_counts_filter_reprod_summary %>% filter(reps == 3), aes(R1,R2)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor() +
  facet_grid(cell_line ~ read_filter, scales = "free") +
  theme_bw()

#R1 vs. R2
ggplot(read_counts_filter_reprod_summary %>% filter(reps == 3), aes(R2,R3)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor() +
  facet_grid(cell_line ~ read_filter, scales = "free") +
  theme_bw()

#R1 vs. R2
ggplot(read_counts_filter_reprod_summary %>% filter(reps == 3), aes(R1,R3)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor() +
  facet_grid(cell_line ~ read_filter, scales = "free") +
  theme_bw()

```

#Conclusion: To get high reproducibility in the results, we really need to be very stringent with the cut-off. I will try to repeat a similar filtering with total number of read cut-off and see how it looks.


#Plot balances and TIF
```{r}
#Plot balance
ggplot(filter_read_count_test %>% filter(gRNA %in% c("36","64","NTC","LBR2","halfLBR2") & read_filter == 5)) + 
  geom_quasirandom(aes(fct_relevel(gRNA,c("LBR2","halfLBR2","NTC","36","64")),log2_bal)) +
  facet_grid(cell_line ~ bio_rep) + theme_bw()
```

#Calculate delta log2 MMEJ:NHEJ (+1 > 10 reads & -7 > 10 reads)
```{r}
#Filter barcodes with reads
filtered_indel_data <- pathway_metric_control_bio %>%
  filter(`del_-7` > 10 & ins_1 > 10)

#extract control data
control_indel_data <- filtered_indel_data %>%
  filter(gRNA == "NTC") %>%
  select(barcode, cell_line, bio_rep, ntc_log2_bal = log2_bal)

#Prepare data frame
diff_data <- filtered_indel_data %>%
  left_join(control_indel_data) %>%
  na.omit() %>% 
  mutate(diff_balance = log2_bal - ntc_log2_bal)

#combine replicates
combined_diff_data <- diff_data %>%
  dplyr::group_by(barcode, cell_line, gRNA) %>%
  dplyr::summarise(mean_diff_balance = mean(diff_balance),
                   reps = n()) %>%
  filter(reps == 3)
```

#Plot mean values per cell line
```{r}
mean_per_cell_line <- combined_diff_data %>%
  filter(reps == 3 & !gRNA %in% c("NTC","LBR2","halfLBR2","empty")) %>%
  dplyr::group_by(cell_line, gRNA) %>%
  dplyr::summarise(global_effect = mean(mean_diff_balance, na.rm = T),
                   sd_global = sd(mean_diff_balance, na.rm = T))

mean_per_cell_line_dcast <- mean_per_cell_line %>% 
  reshape2::dcast(gRNA ~ cell_line, value.var = "global_effect")

```

#Plot global effect similarity in RPE1 and K562 cells
```{r, fig.width=6, fig.height = 3}
#Gene order based on K562 data
ggplot(combined_diff_data %>% filter(reps == 3 & !gRNA %in% c("NTC","LBR2","halfLBR2","empty") & cell_line == "PRO")) +
  stat_summary(aes(gRNA,mean_diff_balance,fill = fct_relevel(cell_line, c("PRO","DEF"))),  position = position_dodge2(preserve = "single"), geom = "bar", fun.data = "mean_sd") + 
  stat_summary(aes(gRNA,mean_diff_balance, group = fct_relevel(cell_line, c("PRO","DEF"))), position = position_dodge2(preserve = "single"), geom = "errorbar", fun.data = "mean_sd",width = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5),
        legend.position = "top",
        axis.title.x = element_blank()) +
  geom_hline(yintercept = 0, linetype = 2) +
  ylab("∆log2 MMEJ::NHEJ") +
  labs(fill = "Cell line")

#Test ∆log2 different than 0
wilcox_test <- combined_diff_data %>% 
  filter(reps == 3) %>%
  dplyr::group_by(cell_line, gRNA) %>%
  t_test(mean_diff_balance ~ 0) %>%
  mutate(p.adj = p.adjust(p, method = "fdr"))


```




#Correlation by cell type
```{r} 
#DEF vs. PRO
ggplot(mean_per_cell_line_dcast, aes(DEF, PRO)) + 
  geom_point(aes(color= gRNA)) +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(method = "spearman") +
  geom_abline(linetype = 2) +
  theme_bw()
```
#CONCLUSION: Global pathway balance effect are conserved across cell lines, in the validation set up.

#Calculate significance of globat values for volcano plot
```{r}
#Calculate wilcoxon test per gRNA
global_analysis_test_data <- combined_diff_data %>%
  filter(reps >1 & !gRNA %in% c("NTC","LBR2","halfLBR2","empty")) %>%
  dplyr::group_by(cell_line,gRNA) %>%
  wilcox_test(mean_diff_balance ~ 0) %>%
  ungroup() %>%
  dplyr::mutate(adj.p = p.adjust(p, method = "BH")) %>%
  left_join(mean_per_cell_line)

#Plot volcano plot
tmp <- ggplot(global_analysis_test_data, aes(global_effect, -log10(adj.p), gRNA = gRNA)) + 
  geom_point() +
  facet_wrap(~ cell_line) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = -log10(0.001), linetype = 2)

ggplotly(tmp)

```
#CONCLUSION: I need higher quality data in RPE1. I will need this to say something about pathway balance. I need to retrieve higher number of IPRs.

#Are there differential global effects on pathway balance between PRO and DEFs.
```{r}
wilcox_test <- combined_diff_data %>%
  filter(reps == 3) %>%
  filter(!gRNA %in% c("58","72","empty","halfLBR2","LBR2","NTC")) %>%
  dplyr::group_by(gRNA) %>%
  wilcox_test(mean_diff_balance ~ cell_line) %>%
  mutate(fdr = p.adjust(p, method = "BH"))

```
#NOTE: ATM and MDC1 have a smaller effect on BRCA1 KO cell lines. For the rest, they have very similar effect between both cell lines.

#Export data
```{r}
write_rds(filtered_indel_data, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230905_PRO_DEF_pathway_balance_10reads_per_indel.rds")

```
