---
title: "xv20230830_validation_screen_chromatin_bias_controls"
author: "Xabier Vergara"
date: "2023-07-26"
output:
  html_document: default
  pdf_document: default
---
#Date: 30th August 2023
#Author: Xabier Vergara
#Aim: The aim of this file is to wrap up the filtering strategy and measure if pathway balance in control samples has chromatin biases in the control samples

#Load libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(ggbeeswarm)
library(pheatmap)
```


## Import data: Balance data and some other details
```{r, warning=F}
# Import data for high confidence reporters and calculate mean
high_confidence_IPR <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen/xv20230830_pathway_balance_freq_per_indel.rds") %>%
  dplyr::group_by(barcode, cell_line, gRNA) %>%
  dplyr::summarise(mean_bal = mean(log2_bal, na.rm = T),
                   reps = n())


#Pools used in the experiment
selected_pools <- c("RPE1_Low_1000","RPE1Deff_Low_1000","RPE1Proff_Low_250","U2OS_High_100")

```

#Add chromatin info for RPE1 cells
#Load chromatin info of the IPRs in RPE cells (2000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "2000_", full.names = T)
dam_files <- list.files(path = "dam", pattern = "2000_", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-2000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=2000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_2000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "DEF",
                               iPCR_cell == "RPE1Proff" ~ "PRO",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity, - dam_LMNNB2) %>%
  mutate(binsize = "2000")
```


#Load chromatin info of the IPRs in RPE cells (5000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "5000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "5000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-5000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=5000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_5000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "DEF",
                               iPCR_cell == "RPE1Proff" ~ "PRO",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity, -dam_LMNNB2)  %>%
  mutate(binsize = "5000")
```


#Load chromatin info of the IPRs in RPE cells (10000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "10000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "10000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-10000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=10000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_10000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "DEF",
                               iPCR_cell == "RPE1Proff" ~ "PRO",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity)  %>%
  mutate(binsize = "10000")
```

#Load chromatin info of the IPRs in RPE cells (20000)
```{r, warning=F}
setwd("/DATA/projects/DSBrepair/data/DSB_TRIP_cell_lines_mapping/xv20230519_all_RPE_pools_mapping/")
# Load chip data
chip_files <- list.files(path = "chip/site_means", pattern = "20000", full.names = T)
dam_files <- list.files(path = "dam", pattern = "20000", full.names = T)
repli_files <- "coverage/RPE1_DSB_TRIP_pools-20000_summary.tsv"

#Make a dataframe with all the files
chromatin_features_pools_chip <- map(chip_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(ID, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("chip_",str_extract(x, "H.*(?=_)"))) %>%
    reshape2::dcast(ID ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "ID")

#Make a dataframe with dam files
chromatin_features_pools_dam <- map(dam_files, function(x) {
  read.delim(x, header = T) %>%
    dplyr::select(name, z_score) %>%
    na.omit() %>%
    mutate(file = paste0("dam_",str_extract(x, "(?<=20000_).*(?=.txt)"))) %>%
    reshape2::dcast(name ~ file, value.var = "z_score")
}) %>% purrr::reduce(left_join, by = "name")

#Make a column for repliseq
repliseq_dataframe <- map_dfr(repli_files, function(x) {
  read.delim(x, header = T) %>%
    na.omit() %>%
    mutate(ratio_r1 = repliseq_r1_late - repliseq_r1_early, ratio_r2 = repliseq_r2_late - repliseq_r2_early) %>%
    mutate(late_replicating = scale((ratio_r1 + ratio_r2)/2)) %>%
    dplyr::select("ID" = barcode, late_replicating)
})

#Bind both
chromatin_features_pools_20000 <-  left_join(chromatin_features_pools_chip, chromatin_features_pools_dam, by = c("ID"="name")) %>%
  left_join(repliseq_dataframe) %>%
  separate(ID, into = c("barcode","iPCR_cell","iPCR_transfection","iPCR_complexity")) %>%
  mutate(pool = paste(iPCR_cell, iPCR_transfection,iPCR_complexity, sep = "_")) %>% 
  filter(pool %in% selected_pools) %>%
  mutate(cell_line = case_when(iPCR_cell == "RPE1Deff" ~ "DEF",
                               iPCR_cell == "RPE1Proff" ~ "PRO",
                               T ~ iPCR_cell)) %>%
  select(-pool, -iPCR_cell, -iPCR_transfection, -iPCR_complexity) %>%
  mutate(binsize = "20000")
```

#Merge all binsizes in one data table
```{r}
#Bind the chromatin data for all binsizes
chromatin_pools_binsize <- bind_rows(chromatin_features_pools_20000,
                                     chromatin_features_pools_10000,
                                     chromatin_features_pools_5000,
                                     chromatin_features_pools_2000)
```




#Join with balances and calculate the correlation (do this for all marks) = Spearman's
```{r, warning=F}
#Merge RPE data with chromatin data with different bins
values_chromatin_RPE_controls <- left_join(high_confidence_IPR, chromatin_pools_binsize) %>%
  filter(gRNA %in% c("LBR2") & reps > 1) %>% na.omit()

#Calculate spearman correlation for all chromatin features
cell_lines <- c("RPE1","PRO","DEF")
binsizes <- c("2000","5000","10000","20000")

#Calculate spearman correlation coefficient per cell line and chromatin feature
spearman_coefficients <- map_dfr(cell_lines, function(x){
  filter_data <- values_chromatin_RPE_controls %>% filter(cell_line == x)
  binsize_filter <- map_dfr(binsizes, function(y){
    bin_filter <- filter_data %>% filter(binsize == y)
    rho_table <- map_dfr(colnames(bin_filter[6:16]), function(j){
      balance <- bin_filter$mean_bal
      feature <- bin_filter %>% pull(j)
      cor.test(balance, feature, method = "spearman", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })
})

#Chromatin feature order for plotting
chromatin_order <- c("dam_H3K9me3","dam_H3K9me2","late_replicating","dam_LMNB1","dam_LMNB2","dam_H3K27me3","chip_H3K4me3","chip_H3K4me2","chip_H3K4me1","chip_H3K27ac","chip_H3K36me3")

#Plot correlation coefficients
ggplot(spearman_coefficients %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Spearman's correlation coefficient of MMEJfreq.") + 
  facet_grid(fct_relevel(binsize,binsizes)~ fct_relevel(cell_line, c("RPE1","RPEPro","RPEDef")))
```

#Test data bining
```{r}
#Arrange control data set
indel_data_inh <- high_confidence_IPR %>%
  filter(reps > 1) %>% 
  left_join(chromatin_pools_binsize) %>%
  filter(binsize == "2000" & gRNA == "LBR2")

#Melt chromatin data
melt_chromatin <- indel_data_inh %>%
  select(-mean_bal,-reps) %>%
  reshape2::melt() %>%
  separate(variable, into = c("tech", "antibody")) %>%
  select(-tech) %>%
  left_join(indel_data_inh %>% 
              select(barcode,cell_line, gRNA,mean_bal) %>% 
              distinct())

#Calculate bins per antibody
binned_control_chromatin_data <- map_dfr(unique(melt_chromatin$antibody), function(x) {
  map_dfr(unique(melt_chromatin$cell_line), function(y) {
  melt_chromatin %>%
    filter(antibody == x & cell_line == y) %>%
    mutate(percentile = ntile(value, n = 20)) %>%
    group_by(percentile,antibody,cell_line) %>%
    dplyr::summarise(bin_balance = mean(mean_bal),
                     bin_min = case_when(percentile == 1 ~ -10,
                                         T ~ min(value)),
                     bin_max = case_when(percentile == 20 ~ 10,
                                         T ~ max(value)),
                     bin_mean_chrom = mean(value),
                     IPR_n = n()) %>% distinct()
})
})

```

#Binned data plotting (as before, spearman correlations)
```{r}
#Calculate spearman correlation coefficient per cell line and chromatin feature
spearman_coefficients_binned <- map_dfr(cell_lines, function(x){
    rho_table <- map_dfr(unique(binned_control_chromatin_data$antibody), function(j){
        filter_data <- binned_control_chromatin_data %>% filter(cell_line == x & antibody == j) %>% na.omit()
      balance <- filter_data$bin_balance
      feature <- filter_data$bin_mean_chrom
      cor.test(balance, feature, method = "spearman", use = "pairwise.complete.obs") %>% broom::tidy() %>% mutate(chromatin = j)
    })
  rho_table %>% mutate(cell_line = x, binsize = y) %>% distinct()
  })

#Chromatin feature order for plotting
chromatin_order <- c("H3K9me3","H3K9me2","replicating","LMNB1","LMNB2","H3K27me3","H3K4me3","H3K4me2","H3K4me1","H3K27ac","H3K36me3")

#Plot correlation coefficients
ggplot(spearman_coefficients_binned %>% distinct()) + 
  geom_col(aes(fct_relevel(chromatin,chromatin_order),estimate, fill = p.value < 0.05)) + 
  coord_flip() +
  theme_bw() + 
  xlab("Chromatin features") +
  ylab("Spearman's correlation coefficient of MMEJfreq.") + 
  facet_grid(~ fct_relevel(cell_line, cell_lines))

```


#Conclusion: 
I definitely need more data on RPE1 cells. It is quite sparce! Data is noisy in general mainly PRO and DEF (correlation coefficients are not great). They get much better if we bin the data, it might be a matter of other factores hidded in the pool (# IPR/cell or other things). But, the correlations the bias euchromatin/heterochromatin is consistent across different cell lines.

