---
title: "xv20220712_S1_set_up_screening"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S1 in the short version of the paper.

Fig. S1: Screen replicate reproducibility and distribution of log2MMEJ:NHEJ values. A-C) Pair-wise correlations of log2MMEJ:NHEJ values of individual IPRs between replicate experiments R1, R2 and R3, after application of quality filters as described in step 3 of the data processing. R denotes Pearson correlation coefficient. D) Dynamic range of ∆log2MMEJ:NHEJ balances after averaging of rep-licates. 


```{r functions, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries, message=FALSE, warning=FALSE}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
library(ggpubr)
library(ggbeeswarm)
```



# Import data tables
```{r Import_data, warning=FALSE}
setwd(in.dir)

print("For this script the following data was inputed:")

#Import MMEJ:NHEJ balance ratio measurements per replicate
ddr.screen.all.sel <- readRDS_proof( "data/processed_data/CCD_analysis","raw_data")

#log2 fold change
log2_fold_change_IPR <- readRDS_proof( "data/processed_data/CCD_analysis","differentials_log2_MMEJ")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")
```


#Figure S1A-C

 A-C) Pair-wise correlations of log2MMEJ:NHEJ values of individual IPRs between replicate experiments R1, R2 and R3, after application of quality filters as described in step 3 of the data processing. R denotes Pearson correlation coefficient.

```{r Fig_S1_A_C, message=FALSE, warning=FALSE}

#dcast data table for plotting reproducibility of the ratio
balance.ratio.list <- ddr.screen.all.sel %>% reshape2::dcast(gene + barcode + sample + well + plate ~ rep, value.var = "log2MMEJNHEJratio") %>% arrange(sample)

# Plot reproducibility DNA_repair library
#R1 vs. R2 (Fig. S1B)
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S1_general_controls/xv20231009_log2balance_corplot_R1_R2.pdf",width = 4, height = 4)
ggplot(balance.ratio.list) + 
  geom_point(aes(R1,R2, color = fct_relevel(sample,c("WT","POLQ"))))+ 
  geom_smooth(aes(R1,R2), method = "lm")+ 
  xlab("R1 (log2 MMEJ:NHEJ balance))") + ylab("R2 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R1,R2), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))
dev.off()

#R1 vs. R3 (Fig. S1C)
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S1_general_controls/xv20231010_log2balance_corplot_R1_R3.pdf",width = 4, height = 4)
ggplot(balance.ratio.list) + 
  geom_point(aes(R1,R3, color = fct_relevel(sample,c("WT","POLQ"))))+ 
  geom_smooth(aes(R1,R3), method = "lm")+ 
  xlab("R1 (log2 MMEJ:NHEJ balance))") + ylab("R3 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R1,R3), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))
dev.off()

#R2 vs. R3 (Fig. S1D)
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S1_general_controls/xv20231010_log2balance_corplot_R2_R3.pdf",width = 4, height = 4)
ggplot(balance.ratio.list) + 
  geom_point(aes(R2,R3, color = fct_relevel(sample,c("WT","POLQ","KO"))))+ 
  geom_smooth(aes(R2,R3), method = "lm")+ 
  xlab("R2 (log2 MMEJ:NHEJ balance))") + ylab("R3 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R2,R3), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))
dev.off()
```


#Figure S1D: log2 fold change effects of controls and KO (dynamic range)
D) Dynamic range of ∆log2MMEJ:NHEJ balances after averaging of replicates. 

```{r Fig_S1_E, warning=FALSE, message=FALSE, fig.width=4,fig.height=3}
#Plot log2 fold changes
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/figure_S1_general_controls/xv20231010_delta_log2_measurements.pdf",width = 4, height = 3)
ggplot(log2_fold_change_IPR %>% arrange(sample)) +
  geom_quasirandom(aes(fct_relevel(barcode,clone5_bc_cluster),mean.log2foldchange, color = sample), dodge.width = 0.5, size = 0.8) + 
  theme_bw() + 
  theme(axis.text.x = element_blank(),
        legend.position = "top",panel.grid = element_blank()) + 
  scale_color_manual(values = c("#DADADA","#465584","#949DCF")) +
  geom_hline(yintercept = 0, linetype = 2) +
  ylab("∆log2MMEJ:NHEJ") +
  xlab("Integrated pathway reporter (IPR)")
dev.off()
```
