---
title: "xv20220712_S2_CCDs_z_score"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S2 in the short version of the paper. This figure describes the z-score normalization and CCDs calculations. 
A- Plate design
B- Correlation in log2 pathway balance
C- Positive control effect size


This script generates the plots for figure 1

```{r}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
```



# Import data tables
```{r}
setwd(in.dir)
#Clone 5 chromatin Version cl20201026
clone5_z.score_chrom_tib <- readRDS_proof("/DATA/projects/DSBrepair/data/R","ChIP_zscore_selection")

#Import MMEJ:NHEJ balance ratio measurements per replicate
ddr.screen.all <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data","screen_ddr_all_computations")
ddr.screen.all.sel <- ddr.screen.all %>% dplyr::select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep = bio.rep,t.rep = FinRep, freqCut) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#log2 fold change
log2_fold_change_IPR_KO <- readRDS_proof( "data/processed_data","fold_change_ratio_KO")
log2_fold_change_IPR_WT <- readRDS_proof( "data/processed_data/CCD_analysis","fold_change_ratio_WT")
log2_fold_change_IPR_POLQ <- readRDS_proof( "data/processed_data/CCD_analysis","fold_change_ratio_POLQ")


# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")

# Individual data point (intermediate step)
step2.epistasis <- readRDS_proof("data/processed_data/CCD_analysis","fold_change_ratio_KO")
 
```

#Supplementary Figures

#Figure S5A: Examples RAD50 vs LMNB1
```{r}
# Plot RAD50 with LMNB1
ggplot(step2.epistasis %>% 
         filter(gene %in% c("RAD50")),
       aes(LMNB1,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = -0.114),method = "lm") +
  stat_cor(label.x = -1.5, label.y = -0.7) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(RAD50 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#Figure S5A: Examples RAD50 vs LMNB1
```{r}
# Plot RAD50 with H3K4me3
ggplot(step2.epistasis %>% 
         filter(gene %in% c("RAD50")),
       aes(H3K4me3,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = -0.114),method = "lm") +
  stat_cor(label.x = -0.5, label.y = -0.7) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(RAD50 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#Figure S5A: Examples BRCA2 vs. H3K4me1
```{r}
# Plot BRCA2 with H3K4me1
ggplot(step2.epistasis %>% 
         filter(gene %in% c("BRCA2")),
       aes(H3K4me1,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = 0.14),method = "lm") +
  stat_cor(label.x = -1.5, label.y = 0.4) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(BRCA2 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#Figure S5D: Example BRCA2 vs. LMNB1
```{r}
# Plot BRCA2 with H3K36me3
ggplot(step2.epistasis %>% 
         filter(gene %in% c("BRCA2")),
       aes(H3K9me2,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = 0.14),method = "lm") +
  stat_cor(label.x = -1.5, label.y = 0.4) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(BRCA2 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#Figure S5D: Example BOD1L1 vs. H3K4me1
```{r}
# Plot BRCA2 with H3K36me3
ggplot(step2.epistasis %>% 
         filter(gene %in% c("BOD1L1")),
       aes(H3K4me1,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = 0.14),method = "lm") +
  stat_cor(label.x = -1.5, label.y = 0.4) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(BOD1L1 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```

#Figure S5E: Example BOD1L1 vs. late_replicating
```{r}
# Plot BRCA2 with H3K36me3
ggplot(step2.epistasis %>% 
         filter(gene %in% c("BOD1L1")),
       aes(late_replicating,mean.log2foldchange)) + 
  geom_point() +
  geom_smooth(aes(color = -0.14),method = "lm") +
  stat_cor(label.x = -1.5, label.y = 0.4) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(BOD1L1 FC MMEJ:NHEJ balance)") + theme(legend.position = "none")
```


 