---
title: "xv20220928_S3_CCDs_pcr"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S3 in the short version of the paper.

Fig. S3: Principal component regression analysis. A) Principal Component Regression workflow. First, a principal component analysis (PCA) was run with 25 chromatin feature values for each IPR. Second, exploration of the PCA revealed that the first three PCs recapitulate most of the chromatin feature variance. Third, a linear model with three principal components was ran for each protein and the performance was assessed by predicted vs. measured comparison. B) Percentage of variance ex-plained by the first five PCs. C) Bar graph showing the weight of each chromatin feature for PC1, PC2 and PC3. Bars are colourded according to the chromatin context they represent: Triple heterochroma-tin (purple), H3K27me3 heterochromatin (pink), transcription (light orange) and enhancers/promoters (orange).

```{r functions, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries, warning=FALSE,message=FALSE}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
library(pls)
library(ggpubr)
```

# Import data tables
```{r import, warning=FALSE, message=FALSE}
setwd(in.dir)

print("For this script the following data was inputed:")

#
diff_KO_chromatin <- readRDS_proof("/data/processed_data/CCD_analysis","differentials_chromatin_KO")


# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")
```

#Calculate PCR for a single protein RAD50
```{r}
# Filter RAD50 data
RAD50_data_diff <- filter(diff_KO_chromatin, gene == "RAD50")

#Run PCR and calculate all the components
PCR_example <- pcr(mean.log2foldchange~CTCF+EZH2+H2AFZ+H3K27ac+H3K27me3+H3K36me3+H3K4me1+H3K4me2+H3K4me3+H3K79me2+H3K9me2+H3K9me3+H4K5acK8ac+HDAC1+HDAC2+HDAC3+POL2+POL2AS2+SMC3+LMNB1+Dam+DNAse+m5C+late_replicating+TTseq, data=RAD50_data_diff , validation="CV") #Run principal component regression

#Predict the values with the model with 3PCs
RAD50_pred <- predict(PCR_example, RAD50_data_diff, ncomp = 3) 
```

#Supplementary Figures

#Figure S3A: Cartoon

A) Principal Component Regression workflow. First, a principal component analysis (PCA) was run with 25 chromatin feature values for each IPR. Second, exploration of the PCA revealed that the first three PCs recapitulate most of the chromatin feature variance. Third, a linear model with three principal com-ponents was ran for each protein and the performance was assessed by predicted vs. measured com-parison.

```{r, warning=FALSE, message=FALSE}
# PC1 vs PC2 plot
PC_plot <- PCR_example$scores[,c(1,2)] %>% as.data.frame()

#Plot scatterplot
ggplot(PC_plot) + geom_point(aes(`Comp 1`,`Comp 2`)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
  theme_bw() +
  coord_fixed() +
  xlab("PC1 (37.4%)") +
  ylab("PC2 (25%)")

# Example of predicted vs. measured plot
pred_vs_measured <- tibble(measured = RAD50_data_diff$mean.log2foldchange, predicted = as.numeric(RAD50_pred)) #Set dataframe

#Plot scatter plot and correlations
ggplot(pred_vs_measured, aes(predicted,measured)) + 
  geom_point() +
  stat_cor() +
  geom_smooth(method = "lm") +
  theme_bw() +
  coord_cartesian()

```

#Figure S3B:
B) Percentage of variance explained by the first five PCs. 
```{r, warning=FALSE, message=FALSE}
#Accumulated variance explained
acc_var <- c(0,37.37,62.36,72.95,79.09,83.93)

#Extract variance explained by each component
pc_variance <- sapply(acc_var[-1], FUN = function(x) {
  x - acc_var[which(acc_var == x) - 1]
})

#dt to plot
var_plot_tib <- tibble(pc = c(1:5), var = pc_variance)
#plot
ggplot(var_plot_tib) +
  geom_col(aes(pc, var)) +
  theme_bw()
```
#Figure S3C
C) Bar graph showing the weight of each chromatin feature for PC1, PC2 and PC3. Bars are colourded according to the chromatin context they represent: Triple heterochromatin (purple), H3K27me3 hetero-chromatin (pink), transcription (light orange) and enhancers/promoters (orange).
```{r}
#Axis orientation
PC1_3_projections <- PCR_example$projection[,c(1:3)] %>% as.data.frame() %>% 
  rownames_to_column(var = "feature") %>%
  reshape2::melt()

#Plot
ggplot(PC1_3_projections) +
  geom_col(aes(fct_relevel(feature, heatmap.chromatin.order),value)) +
  facet_wrap(~ variable, ncol = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```


