---
title: "xv20230817_mutations_analysis"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

I will explore the new data that Mathijs sent me, and figure it out if we can use it in the paper.

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20231017_indel_data_mathijs_new/BOOTSTRAP"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}

```

## Libraries:
```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
library(ggbeeswarm)
```

#Import indel data
```{r, warning=F, message=F}
setwd(in.dir)

#Function to import all the data
#List of all driver mutations
driver_mutations_indels <- list.files(path = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20231017_indel_data_mathijs_new/BOOTSTRAP", full.names = T)[c(1:4,6,7)]

control_indel_file_names <- map_dfr(driver_mutations_indels, function(y) {
  control_directory <- paste0(y,"/controls_indel")
  #Import all data for a single driver mutation
tumor_file_names <- list.files(path = control_directory, full.names = T)
  mutation_files <- map_dfr(tumor_file_names,function(x) {
    tibble(file_names = list.files(path = x, full.names = T))
  })
  })

#list of files
control_indel_data <- map_dfr(control_indel_file_names$file_names, function(x) {
  read.delim(x) %>%
    mutate(file = str_extract(x, "(?<=BOOTSTRAP/).*LAD")) %>%
    separate(file, into = c("driver",NA,"project","chromatin"), sep = "/")
}) %>% reshape2::melt(vars.id = c("TYPE","driver","project")) %>%
  reshape2::dcast(TYPE+driver+project~chromatin+variable, fun.aggregate = sum)  %>% 
  mutate(sample = "control")


#Load samples
indel_file_names <- map_dfr(driver_mutations_indels, function(y) {
  #Import all data for a single driver mutation
tumor_file_names <- list.files(path = y, full.names = T)
  mutation_files <- map_dfr(tumor_file_names,function(x) {
    tibble(file_names = list.files(path = x, full.names = T))
  })
  }) %>% filter(grepl(".txt", file_names))

#list of files
mutant_indel_data <- map_dfr(indel_file_names$file_names, function(x) {
  read.delim(`x`) %>%
    mutate(file = str_extract(x, "(?<=BOOTSTRAP/).*LAD")) %>%
    separate(file, into = c("driver","project","chromatin"), sep = "/")
}) %>% reshape2::melt(vars.id = c("TYPE","driver","project")) %>%
  reshape2::dcast(TYPE+driver+project~chromatin+variable, fun.aggregate = sum)  %>% 
  mutate(sample = "mutant")

#Load data from manuscript
#Tumor analysis
indel_data_tumors <- readRDS_proof("data/processed_data/","SCC_indel_MMEJ_NHEJ")

#Process data
distribution_tumor_indel_pathway <- indel_data_tumors %>% 
  filter(ciLAD == 1 | cLAD == 1) %>%
  mutate(LAD_status = case_when(ciLAD == 1 ~ "ciLAD",
                                cLAD == 1 ~ "cLAD")) %>%
  dplyr::group_by(tumor_type, tumor_id, LAD_status, total_mutation, CLASS) %>% dplyr::summarise(mut_count = n(), .groups = "drop") %>% 
  reshape2::dcast(tumor_type + tumor_id ~ CLASS+LAD_status, value.var = "mut_count", fill = 0)

#Tumor analysis
SV_MH_del_data_tumors <- readRDS_proof("data/processed_data/","SCC_SV_MH_deletions")
indel_data_tumors <- readRDS_proof("data/processed_data/","SCC_indel_MMEJ_NHEJ")

#Import bootstrap counts
boot_matched_distribution <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/","bootstrap")

```

#Load tumor info data
```{r}
#Import qualitative data
pcaw_TCGA_donor <- read_delim("/DATA/projects/DSBrepair/data/xv20230818_PCAWG_data_revision/DATA_PCAWG/INFO/pcawg_TCGA_to_donorID.txt")
selection <- read_delim("/DATA/projects/DSBrepair/data/xv20230818_PCAWG_data_revision/DATA_PCAWG/INFO/selection.txt")

#Combine them
tumor_info <- left_join(pcaw_TCGA_donor,selection, c("TCGA_ID" = "sample_id"))

#Multiple driver genes
multiple_drivers_data <- mutant_indel_data %>% 
  select(TYPE, project,driver) %>% 
  distinct() %>%
  dplyr::group_by(TYPE, project) %>%
  dplyr::summarise(counts = n()) %>%
  filter(counts > 1)

#Controls that are not properly assigned
controls_with_driver_mutations <- control_indel_data %>%
  filter(TYPE %in% unique(mutant_indel_data$TYPE))

```

#Filter data from repeated samples
```{r}
filtered_indels_samples <- mutant_indel_data %>% 
  bind_rows(control_indel_data) %>%
  filter(!TYPE %in% multiple_drivers_data$TYPE) %>%
  anti_join(controls_with_driver_mutations)
```

#Number of tumors per tumor type
```{r, fig.width=5, fig.height=4}
tumor_count_summary <- filtered_indels_samples %>%
  select(project,driver, sample, TYPE) %>%
  distinct() %>%
  dplyr::count(project, driver, sample)

#Plot
pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231128_S9A_tumor_counts.pdf", width = 5, height = 4)
ggplot(tumor_count_summary) +
  geom_col(aes(fct_relevel(driver, c("EGFR","BRCA2","ATM","SETD2","MEN1","ATRX","BRCA1","TRRAP")),n,fill = project), color = "grey40", size = 0.25) +
  facet_wrap(~ fct_relevel(sample, c("mutant","control")), scales = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top") + 
  ylab("Number of genomes")
dev.off()
```

# Tumor biases (per project)
```{r}
total_changes_indels_project <- filtered_indels_samples %>% 
  dplyr::group_by(project,driver,sample) %>%
  dplyr::summarise(sum_CILAD_NHEJ = sum(CILAD_NHEJ)+1,
                   sum_CILAD_MMEJ = sum(CILAD_MMEJ)+1,
                   sum_CLAD_NHEJ = sum(CLAD_NHEJ)+1,
                   sum_CLAD_MMEJ = sum(CLAD_MMEJ)+1,
                   sample_n = n()) %>%
  rowwise() %>%
  mutate(LAD_bal = log2(sum_CLAD_MMEJ/sum_CLAD_NHEJ),
                 iLAD_bal = log2(sum_CILAD_MMEJ/sum_CILAD_NHEJ),
         global_bal = log2((sum_CLAD_MMEJ+sum_CILAD_MMEJ)/(sum_CLAD_NHEJ+sum_CILAD_NHEJ)),
                 bias = LAD_bal - iLAD_bal,
         total_indel = sum(sum_CILAD_NHEJ,sum_CILAD_MMEJ, sum_CLAD_NHEJ,sum_CLAD_MMEJ)) %>%
  filter(total_indel > 10)

```

# Biases by tissue type in controls
```{r, fig.width=4,fig.height=3.5}
bias_per_tissue_type <- filtered_indels_samples %>%
  filter(sample == "control") %>%
  select(TYPE,driver, project, CILAD_MMEJ, CILAD_NHEJ,CLAD_NHEJ, CLAD_MMEJ) %>%
  distinct() %>%
  dplyr::group_by(project) %>%
  dplyr::summarize(sum_CILAD_NHEJ = sum(CILAD_NHEJ)+1,
                   sum_CILAD_MMEJ = sum(CILAD_MMEJ)+1,
                   sum_CLAD_NHEJ = sum(CLAD_NHEJ)+1,
                   sum_CLAD_MMEJ = sum(CLAD_MMEJ)+1,
                   sample_n = n()) %>%
  rowwise() %>%
  mutate(LAD_bal = log2(sum_CLAD_MMEJ/sum_CLAD_NHEJ),
                 iLAD_bal = log2(sum_CILAD_MMEJ/sum_CILAD_NHEJ),
         global_bal = log2((sum_CLAD_MMEJ+sum_CILAD_MMEJ)/(sum_CLAD_NHEJ+sum_CILAD_NHEJ)),
                 bias = LAD_bal - iLAD_bal,
         total_indel = sum(sum_CILAD_NHEJ,sum_CILAD_MMEJ, sum_CLAD_NHEJ,sum_CLAD_MMEJ)) %>%
  filter(total_indel > 10)

#Plot differences
pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231128_S9B_control_tissue_type.pdf", width=4,height=3.5)
ggplot(bias_per_tissue_type) +
  geom_col(aes(project, global_bal)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ylab("∆log2MMEJ:NHEJ het/eu")
dev.off()
```


#Bootstrap per tumor-type (I did it per project)
```{r, message=F, warning=F}
#Create simplified table
simplified_table_filter <- simplified_table
#calculate mean and sd
null_distriution <- boot_matched_distribution %>%
  filter(is.finite(bias)) %>%
  dplyr::group_by(driver,project) %>%
  dplyr::summarise(mean_null = mean(bias),
                   sd_null = sd(bias))

#plot densities
mean_project_value <- total_changes_indels_project %>%
  filter(sample == "mutant") %>%
  filter(is.finite(bias)) %>%
  select(project, driver, bias, sample_n) %>%
  left_join(null_distriution) %>%
  mutate(z_score = (bias - mean_null)/sd_null) %>% na.omit()

driver_significance <- mean_project_value %>%
  dplyr::group_by(driver) %>%
  dplyr::summarise(combined = sum(z_score, na.rm = T)/sqrt(n()),
         p_value = 2*pnorm(abs(combined), lower.tail=F)) %>%
  mutate(fdr = p.adjust(p_value, method = "BH"))


#Combinations to plot
combinations_to_plot <- total_changes_indels_project %>%
  filter(sample == "mutant") %>%
  select(project, driver) %>%
  distinct()

```
#Conclusion: This also gives significant effects, on the right direction.

#Plot disttributions per tumor type
```{r, fig.width=6, fig.height=5, warning=F, message = F}
#ggplot difference densities
pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231128_S9C_distribution_per_tissue.pdf", width = 7, height = 6)
ggplot(boot_matched_distribution %>% right_join(combinations_to_plot), aes(project,bias)) + 
  geom_quasirandom(color = "grey80") + 
  stat_summary(geom = "errorbar", fun.min = function(z) { quantile(z,0.05) },
               fun.max = function(z) { quantile(z,0.95) }, color = "red", width = 0.25) +
  theme_bw() + geom_point(data = total_changes_indels_project  %>% right_join(combinations_to_plot), aes(project,bias, color = sample), size = 3) +
  facet_wrap(~ fct_relevel(driver, c("ATM","ATRX","SETD2","MEN1","BRCA2","BRCA1")), scales = "free", ncol = 3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid = element_blank(),
        legend.position = "top")
dev.off()
```

#Make a table that count in how many tissues the pattern is mantained
```{r}
phenotype_conserved <- total_changes_indels_project %>%
  reshape2::dcast(project+driver ~ sample, value.var = "bias") %>%
  na.omit() %>%
  dplyr::group_by(driver) %>%
  dplyr::summarise(tissue_higher = sum(mutant > control),
                   tissue_lower = sum(mutant < control))

```

# Tumor biases (per patient)
```{r}
total_changes_indels_patient <- filtered_indels_samples %>% 
  rowwise() %>%
  mutate(total_indel = sum(CILAD_NHEJ,CILAD_MMEJ, CLAD_NHEJ,CLAD_MMEJ))


#Original dataset
total_changes_indels_HNSCC <- distribution_tumor_indel_pathway %>% 
  rowwise() %>%
  mutate(total_indel = sum(cNHEJ_ciLAD,cNHEJ_cLAD, MMEJ_ciLAD,MMEJ_cLAD))
```


# Figure S9D: Plot mutation accumuation
```{r, fig.width=3,fig.height=4}
mutation_accumulation_table <- filtered_indels_samples %>%
  rowwise() %>%
  mutate(total_indel = sum(CILAD_NHEJ,CILAD_MMEJ,CLAD_NHEJ,CLAD_MMEJ)) %>%
  dplyr::group_by(driver, sample) %>%
  dplyr::summarise(median_mut = median(total_indel),
                   mean_mut = mean(total_indel),
                   n_sample = n())

pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231128_S9C_total_indel_accumulation.pdf", width = 3, height = 4)
ggplot(mutation_accumulation_table) +
  geom_col(aes(driver, median_mut, fill = sample), position = "dodge") +
  theme_bw() +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  ylab("Median number of mutations per genome")
dev.off()
 
```


# Figure S9E: Differences in indel accumulation
```{r, fig.width=2,fig.height=4}
mutation_accumulation_table <- total_changes_indels_patient %>%
  select(TYPE, driver,sample, total_indel) %>%
  distinct() %>%
  dplyr::group_by(driver,sample) %>%
  dplyr::summarise(median_mut = median(total_indel),
                   mean_mut = mean(total_indel),
                   n_sample = n())

#Calculate median mutation number between HNSSC
mutation_accumulation_table_HNSCC <- total_changes_indels_HNSCC %>%
  select(tumor_id,tumor_type, total_indel) %>%
  distinct() %>%
  dplyr::group_by(tumor_type) %>%
  dplyr::summarise(median_mut = median(total_indel),
                   mean_mut = mean(total_indel))

pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231117_S8E_HNSCC_total_indel_accumulation.pdf", width = 2, height = 4)
ggplot(mutation_accumulation_table_HNSCC) +
  geom_col(aes(fct_relevel(tumor_type, c("HHNSC_indel","BRCA2mut_indel")), median_mut, fill = tumor_type), position = "dodge") +
  theme_bw() +
  theme(legend.position = "top",
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank()) + 
  ylab("Median number of mutations per genome")
dev.off()

```



#Analyse SV data from the manuscript in the same way
```{r}
SV_distribution_tumor_indel_pathway <- SV_MH_del_data_tumors %>% 
  filter(ciLAD == 1 | cLAD == 1) %>%
  mutate(LAD_status = case_when(ciLAD == 1 ~ "ciLAD",
                                cLAD == 1 ~ "cLAD")) %>%
  dplyr::group_by(type, tumor_id, LAD_status) %>% dplyr::summarise(mut_count = n(), .groups = "drop") %>% 
  reshape2::dcast(type + tumor_id ~ LAD_status, value.var = "mut_count", fill = 0)

#Export SV counts
openxlsx::write.xlsx(SV_distribution_tumor_indel_pathway, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/supplementary_tables/xv20231128_TableS9c_SV_HNSCC_BRCA2.xlsx")

#Calculate mean values
SV_mean_balances_bias <- SV_distribution_tumor_indel_pathway %>%
  dplyr::group_by(type) %>%
  dplyr::summarise(CILAD_SV = sum(ciLAD),
                   CLAD_SV = sum(cLAD),
                   sample_n = n()) %>%
  rowwise() %>%
  mutate(bias = log2(CLAD_SV/CILAD_SV),
         total_indel = sum(CLAD_SV,CLAD_SV))

```

#Perform bootstrapping among controls samples
```{r}
#Filter controls
SV_controls_HNSCC <- SV_distribution_tumor_indel_pathway %>%
  filter(type == "sporadic_HPVneg")


#bootstrap_general_tumors
boot_HNSCC_SV <-  map_dfr(c(1:1000), function(j) {
    set.seed(j)
      SV_controls_HNSCC %>% 
        ungroup() %>%
        sample_n(nrow(SV_controls_HNSCC),replace = TRUE) %>%
        mutate(iteration = j) %>% 
          dplyr::group_by(iteration) %>%
          dplyr::summarise(CILAD_SV = sum(ciLAD)+1,
                   CLAD_SV = sum(cLAD)+1,
                   sample_n = n()) %>%
  rowwise() %>%
  mutate(bias = log2(CLAD_SV/CILAD_SV),
         total_indel = sum(CLAD_SV,CLAD_SV))
        })

#calculate mean and sd
null_distriution_HNSCC_SV <- boot_HNSCC_SV %>%
  mutate(control = "HNSCC_control") %>%
  filter(is.finite(bias)) %>%
  dplyr::group_by(control) %>%
  dplyr::summarise(mean_null = mean(bias),
                   sd_null = sd(bias))

#plot densities
mean_HNSCC_value_SV <- SV_mean_balances_bias %>%
  filter(is.finite(bias)) %>%
  select(type, bias, sample_n) %>%
  bind_cols(null_distriution_HNSCC_SV %>% select(-control)) %>%
  mutate(z_score = (bias - mean_null)/sd_null) %>% na.omit()

HNSCC_significance_SV <- mean_HNSCC_value_SV %>%
  mutate(p_value = 2*pnorm(abs(z_score), lower.tail=F)) 

```

#Figure S8E
(D) Total number of large deletions with MH at break sites in euchromatin (Eu.) and constitutive lamina-associated heterochromatin (Het.) in BRCA2+/+ and BRCA2-/-.
```{r Fig_S8_D, warning=FALSE, message=FALSE, fig.width=2,fig.height=3}
#Plot SV S8D
plot_SV_data <- SV_mean_balances_bias %>% select(-sample_n,-bias, -total_indel) %>% reshape2::melt()

pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220117_paper_figure_short/supplementary_figures/fiigure_S8_tumor_analysis/xv20231031_S8E_SV_differences.pdf", width = 2, height = 3)
ggplot(plot_SV_data,aes(fct_relevel(variable, c("CILAD_SV","CLAD_SV")),value, fill = variable)) +
  stat_summary(geom = "col", fun.y = "median") +
  facet_wrap(~ fct_relevel(type, c("sporadic_HPVneg","BRCA2mut")), ncol = 1) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title.x = element_blank()) +
  ylab("MH_Del") +
  scale_fill_manual(values = c("#F39200","#5D277F"))
dev.off()


```
