---
title: "xv20220926_ATM_DNAPK_inhibitors"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S6: 

Text in supplementary:

Fig. S6: ATM and DNAPK inhibitor effects. A) ∆log2MMEJ:NHEJ of ATM and DNAPKcs inhibitor of each IPR. B) Adjusted p-values of Student’s t-test comparing ∆log2MMEJ:NHEJ scores in ATM and DNAPKcs inhibited compared to the vehicle control (n = 3) for each reporter. This test was used to con-firm significance of the changes in the log2MMEJ:NHEJ.

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
library(ggpubr)
library(ggbeeswarm)
```

# Import data tables
```{r}
setwd(in.dir)

#Import MMEJ:NHEJ balance ratio measurements per replicate
drug_summary_test <- readRDS_proof( "data/processed_data/CCD_inhibitors/","drug_t_test")

drug_summary_ratio <- readRDS_proof( "data/processed_data/CCD_inhibitors/","drug_log2_MMEJ")

clone5_bc_cluster <- read.table(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/figures/xv20220927_barcode_clustering.txt")

```


#Plot for supplementary figures (Check if this in supps)

A) ∆log2MMEJ:NHEJ of ATM and DNAPKcs inhibitor of each IPR.
```{r}
# General perturbation
ggplot(drug_summary_ratio %>% filter(drug %in% c("ATMi","DNAPKi"))) + geom_quasirandom(aes(drug, mean.log2foldchange)) + geom_hline(aes(yintercept = 0), linetype = 2) + theme_bw()
```

B) Adjusted p-values of Student’s t-test comparing ∆log2MMEJ:NHEJ scores in ATM and DNAPKcs inhibited compared to the vehicle control (n = 3) for each reporter. This test was used to confirm significance of the changes in the log2MMEJ:NHEJ.
```{r}
#p.values
significant.inh.t.test_levels <- drug_summary_test %>% mutate(signif_level = case_when(p.adj > 0.05 ~ "ns", p.adj > 0.01 & p.adj <= 0.05 ~ "*", p.adj > 0.001 & p.adj <= 0.01 ~ "**", p.adj > 0.0001 & p.adj <= 0.001 ~ "***", T ~ "****"))

ggplot(significant.inh.t.test_levels) + geom_tile(aes(fct_relevel(barcode, clone5_bc_cluster$x), drug, fill = -log10(p.adj))) + coord_fixed(ratio = 1) +  scale_fill_gradient2(limits = c(0,6), breaks = c(0,2,4,6)) + geom_text(aes(fct_relevel(barcode, clone5_bc_cluster$x), drug, label = signif_level)) + theme_bw() + theme(axis.title.x  = element_blank(), legend.position = "top",axis.text.x = element_blank())
```

