---
title: "xv20220826_2_CCD_model"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# Script to generate Supplementary Figure 5

Fig. S5: M- and N-synergies of DNA repair proteins per Gene Ontology category. 

```{r functions, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
    #write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    readRDS(correct_file)
    #write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  readRDS(correct_file)
  #write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  }
}
```

# Libraries
```{r libraries, warning=FALSE, message=FALSE}
# libraries:
library(tidyverse)
library(dendextend)
library(ggdendro)
library(grDevices)
library(pheatmap)
library(reshape2)
library(tibble)
library(ggpubr)
library(scales)
library(ggrepel)
library(ggdendro)
library(readxl)
```

# Import data tables
```{r import , warning=FALSE, message=FALSE}
#setwd
setwd(in.dir)

#DNA repair CCDs
table_S3_CCDs  <- readRDS_proof("data/processed_data","Table_S7_DR") %>% filter(CCD_model_p_adj < 0.05) %>% distinct()

#Import all genes targetted
library_composition <- read_xlsx(path = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/Info_well_plate_gene.xlsx") %>% pull(gene)

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

#Import all different GO terms
GO_files <- list.files(path = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/GO_terms", pattern = "xv20231116.*txt", full.names = T)

#GO files per category
GO_files_all <- map_dfr(GO_files, function(x) {
  read.delim(x, header= F) %>% mutate(GO_ID = str_extract(x, "(?=human_GO).*(?<=.txt)"))
})

```

# Figure S5A
```{r Fig_5A, warning=FALSE, message=FALSE}
# Separate data based on the significant IPRs that they have (based on z-score)
GO_0006303_genes <- GO_files_all %>% filter(grepl("0006303", GO_ID)) %>% pull(V1) %>% unique()
GO_0006303 <- table_S3_CCDs %>% filter(gene %in% GO_0006303_genes & !is.na(chrom_feature))

# dcaast with selected
GO_0006303_dcast <- GO_0006303 %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
GO_0006303_heatmap <- pheatmap(t(GO_0006303_dcast), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 4)
GO_0006303_order <- rev(rownames(GO_0006303_dcast[GO_0006303_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/GO_terms/xv20231116_GO0006303_significants.pdf", width = 5, height = 4)
ggplot(GO_0006303) + 
  geom_tile(aes(fct_relevel(gene,GO_0006303_order),
                fct_relevel(chrom_feature, heatmap.chromatin.order),
                fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "right") + 
  coord_fixed(expand = F,ratio = 1) + 
  xlab("DNA repair proteins with significant chromatin dependencies") + 
  ylab("Chromatin features") +
  ggtitle("GO:0006303 = double-strand break repair involved in non-homologous end joining")
#dev.off()

```
(A-E) CCD patterns of DNA repair proteins in (A) end-joining repair (GO:0006303) 


# Figure S5B
```{r Fig_5B, warning=FALSE, message=FALSE, fig.width=5, fig.height=4}
# Separate data based on the significant IPRs that they have (based on z-score)
GO_0000729_genes <- GO_files_all %>% filter(grepl("0000729", GO_ID)) %>% pull(V1) %>% unique()
GO_0000729 <- table_S3_CCDs %>% filter(gene %in% GO_0000729_genes) %>% distinct()

#Percentage of significant
genes_library <- intersect(GO_0000729_genes, library_composition) #18 proteins
genes_with_CCD <- GO_0000729 %>% pull(gene) %>% unique() #4 proteins

#Percentage 4/18

# dcaast with selected
GO_0000729_dcast <- GO_0000729 %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
GO_0000729_heatmap <- pheatmap(t(GO_0000729_dcast), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 4)
GO_0000729_order <- rev(rownames(GO_0000729_dcast[GO_0000729_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/GO_terms/xv20231116_GO0000729_significants.pdf", width = 5, height = 4)
ggplot(GO_0000729) + 
  geom_tile(aes(fct_relevel(gene,GO_0000729_order),
                fct_relevel(chrom_feature, heatmap.chromatin.order),
                fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "right") + 
  coord_fixed(expand = F,ratio = 1) + 
  xlab("DNA repair proteins with significant chromatin dependencies") + 
  ylab("Chromatin features") +
  ggtitle("GO:0000729 DNA double-strand break processing")
#dev.off()
```
(B) DSB processing (GO:0000729)

# Figure S5C
```{r Fig_5C, warning=FALSE, message=FALSE}
# Separate data based on the significant IPRs that they have (based on z-score)
GO_0045002_genes <- GO_files_all %>% filter(grepl("0045002", GO_ID)) %>% pull(V1) %>% unique()
GO_0045002 <- table_S3_CCDs %>% filter(gene %in% GO_0045002_genes)

# dcaast with selected
GO_0045002_dcast <- GO_0045002 %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
GO_0045002_heatmap <- pheatmap(t(GO_0045002_dcast), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 2)
GO_0045002_order <- rev(rownames(GO_0045002_dcast[GO_0045002_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/GO_terms/xv20231116_GO0045002_significants.pdf", width = 5, height = 4)
ggplot(GO_0045002) + 
  geom_tile(aes(fct_relevel(gene,GO_0045002_order),
                fct_relevel(chrom_feature, heatmap.chromatin.order),
                fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "right") + 
  coord_fixed(expand = F,ratio = 1) + 
  xlab("DNA repair proteins with significant chromatin dependencies") + 
  ylab("Chromatin features") +
  ggtitle("GO:0045002 = double-strnad break repair by single-strand annealing")
#dev.off()
```
(C) single-strand annealing repair (GO:0045002)

# Figure S5D
```{r Fig_5D, warning=FALSE, message=FALSE}
# Separate data based on the significant IPRs that they have (based on z-score)
GO_1990918_genes <- GO_files_all %>% filter(grepl("1990918", GO_ID)) %>% pull(V1) %>% unique()
GO_1990918 <- table_S3_CCDs %>% filter(gene %in% GO_1990918_genes) %>% distinct()

# dcaast with selected
GO_1990918_dcast <- GO_1990918 %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
GO_1990918_heatmap <- pheatmap(t(GO_1990918_dcast), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 2)
GO_1990918_order <- rev(rownames(GO_1990918_dcast[GO_1990918_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/GO_terms/xv20231116_GO1990918_significants.pdf", width = 5, height = 4)
ggplot(GO_1990918) + 
  geom_tile(aes(fct_relevel(gene,GO_1990918_order),
                fct_relevel(chrom_feature, heatmap.chromatin.order),
                fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "right") + 
  coord_fixed(expand = F,ratio = 1) + 
  xlab("DNA repair proteins with significant chromatin dependencies") + 
  ylab("Chromatin features") +
  ggtitle("GO:1990918 DNA double-strand break processing")
#dev.off()
```
(D) meiotic recombination (GO:1990918)

# Figure S5E
```{r Fig_5E, warning=FALSE, message=FALSE}
# Separate data based on the significant IPRs that they have (based on z-score)
GO_0000724_genes <- GO_files_all %>% filter(grepl("0000724", GO_ID)) %>% pull(V1) %>% unique()
GO_0000724 <- table_S3_CCDs %>% filter(gene %in% GO_0000724_genes)

# dcaast with selected
GO_0000724_dcast <- GO_0000724 %>% reshape2::dcast(gene ~ chrom_feature, value.var = "CCD_synergy_score") %>% column_to_rownames(var = "gene")

#Heatmap
GO_0000724_heatmap <- pheatmap(t(GO_0000724_dcast), silent = T, clustering_method = "ward.D", cluster_rows = T, cutree_cols = 4)
GO_0000724_order <- rev(rownames(GO_0000724_dcast[GO_0000724_heatmap$tree_col[["order"]],]))

# This are the effects

# heatmap: All heatmaps together
#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20230307_revision_figures/GO_terms/xv20231116_GO0000724_significants.pdf", width = 5, height = 4)
ggplot(GO_0000724) + 
  geom_tile(aes(fct_relevel(gene,GO_0000724_order),
                fct_relevel(chrom_feature, heatmap.chromatin.order),
                fill = CCD_synergy_score)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "right") + 
  coord_fixed(expand = F,ratio = 1) + 
  xlab("DNA repair proteins with significant chromatin dependencies") + 
  ylab("Chromatin features") +
  ggtitle("GO:0000724 = double-strand break repair involved in homologus recombination")
#dev.off()
```
(E) homologous recombination (GO:0000724). Chromatin features are colored as in Fig. 1C and pro-teins arranged by CCD pattern similarity. Proteins highlighted in the text are highlighted in bold.
