---
title: "xv20211121_epistasis_model_bas"
output: html_document
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
```

# Aesthetic legend for the whole paper
```{r}
# Pathway color palette (MMEJ, wt/no_effect, NHEJ)
pathway.colors <- c("#EB2030","grey90","#2E358F")
# Pathway symbol palette ()
pathway.symbol <- c(19,17)

# Chromatin color palette
#chrom_colors = c( c("euchromatin" = "#F7941D", "other" = "#838687",  "H3K27me3" = "#D21F8A", "triple" = "#662D91"))
chromatin.colors <- c("#D21F8A","#838687","#F7941D","#662D91")
```

# Import data tables

```{r, include= FALSE}
#dcast table
#ddr
ddr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_dcast.rds")
#chromatin
chr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_dcast.rds")
  
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

#All variable table
ddr.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20220110_screen_ddr_all_computations.rds")
ddr.screen.all.sel <- ddr.screen.all %>% dplyr::select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep = bio.rep,t.rep = FinRep, freqCut) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#Chromatin all 
chromatin.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20220110_chromatin_library_all.rds")
chromatin.screen.all.sel <- chromatin.screen.all %>% dplyr::select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep,t.rep, freqCut) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "Chromatin")

# bind both dataset
both.screen.all.sel <- bind_rows(ddr.screen.all.sel,chromatin.screen.all.sel) %>% filter(!gene %in% c("DMSO","LBR_ctrl"))

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"

#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

#Chromatin data
clone5_chrom_tib <- readRDS("/DATA/projects/DSBrepair/data/R/rs20200519_clone5_newdoms_chromatin.RDS")
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)

```

#This is a analysis pipeline to study chromatin context dependencies (CCD)
#This is a proposal to get to a final version where we can infer statistical significance from the data

#I am currently working on these steps:
  1- log2 MMEJ:NHEJ ratio calculation
  2- Filter based on z-score of this data: Filter genes that have 1 significant IPR (z-score > 1.95) in at least 2 replicates.
  3- Compute log2 MMEJ:NHEJ fold changes based on WT values: Compare with WT samples (This shows that data is reproducible and R are better for KO than WT)
  4- Compute CCDs with WT wells (random CCDs) and KO samples
    A- Table with mean CCDs
    B- Table with all CCDs per replicate
  5- Test for non-random CCDs (KS test, I could even use the WT from the other screen, it won't change much). These are the interactions that matter and they correlate decently between replicates
    A- non-random genes based on mean CCDs
    B- non-random genes based on all CCDs per replicate
  

#Check reproducibility at every step
#Step 1 = log2 NHEJ:MMEJ ratio
```{r}
#Compute the log2 MMEJ:NHEJ balance: Initial dataset
log2.ratio.screen.detail.all_follow_up <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::select(bio.rep, NHEJMMEJratio, gene, gRNA,barcode,well,plate) %>% mutate(log2ratio = log2(1/NHEJMMEJratio))
### THIS WAS NOT PLOTTED
#dcast data table for plotting reproducibility of the ratio 
balance.ratio.list_fup <- log2.ratio.screen.detail.all_follow_up %>% reshape2::dcast(gene + gRNA + barcode + well + plate ~ bio.rep, value.var = "log2ratio")

# Plot reproducibility Chromatin library
ggplot(balance.ratio.list_fup) + geom_point(aes(B1,B2)) + xlab("R1 (log2 MMEJ:NHEJ balance)") + ylab("R2 (MMEJ:NHEJ balance)") + geom_abline() + theme_bw() 

#Extract non-targetting samples from this dataset
wt.set.ratio <- filter(log2.ratio.screen.detail.all , sample == "WT") %>% dplyr::group_by(barcode, t.rep,library) %>% dplyr::summarise(wt.log2ratio = mean(log2ratio, na.rm = T))

```

#Step 2: Filter based on z-score on log2 MMEJ:NHEJ
```{r}
# Compute z-score with wt log2(MMEJ:NHEJ ratio)
null.distribtution.log2Ratio_fup <- filter(log2.ratio.screen.detail.all_follow_up , gene == "tracr") %>% dplyr::group_by(barcode, bio.rep) %>% dplyr::summarise(wt.log2ratio = mean(log2ratio, na.rm = T), sd.wt.log2ratio = sd(log2ratio, na.rm = T), c = n())

# Calculate z-scores based on this values
z.scores.ratio.IPR_fup <- log2.ratio.screen.detail.all_follow_up %>%  left_join(null.distribtution.log2Ratio_fup) %>% mutate(z.score.log2ratio = (log2ratio - wt.log2ratio)/sd.wt.log2ratio) %>% select(bio.rep, gRNA, gene, barcode,z.score.log2ratio, well,plate)
  
#Combine replicates
z.scores.ratio.mean_fup <- z.scores.ratio.IPR_fup %>% dplyr::group_by(gRNA,gene,barcode,well,plate) %>% dplyr::summarise(mean.z.score.IPR = sum(z.score.log2ratio, na.rm = T)/sqrt(n()), count = n())

# Filter z-scores with two significant IPRs in at least 2 replicates
z.scores.ratio.IPR.filtered_fup <- filter(z.scores.ratio.mean_fup, abs(mean.z.score.IPR) > 1.95 & count > 1) %>% dplyr::group_by(gene,gRNA,well,plate) %>% dplyr::summarise(MMEJ_signif = sum(mean.z.score.IPR < -1.95), NHEJ_signif = sum(mean.z.score.IPR > 1.95)) %>% filter(MMEJ_signif >= 1 | NHEJ_signif >= 1) %>% mutate(pathway_signif = case_when(MMEJ_signif == 0 ~ "NHEJ", NHEJ_signif == 0 ~ "MMEJ", T ~ "both"))

```

#Step 3: Compute log2 MMEJ:NHEJ fold changes
```{r}
# controls
controls <- c("LBR2","tracr","no")
# Compute log2 fold changes in different set ups for all the KO samples
log2.distance.ratio.KO_fup <- log2.ratio.screen.detail.all_follow_up %>% na.omit() %>%  left_join(null.distribtution.log2Ratio_fup, by = c("bio.rep","barcode"))%>% filter(!gene %in% controls) %>% mutate(log2.foldchange = log2ratio - wt.log2ratio) %>% left_join(z.scores.ratio.IPR.filtered_fup) %>% na.omit()

# Compute log2 fold changes in different set ups for all the KO samples
log2.distance.ratio.controls_fup <- log2.ratio.screen.detail.all_follow_up %>%  left_join(null.distribtution.log2Ratio_fup, by = c("bio.rep","barcode")) %>% filter(gene == "tracr") %>% mutate(log2.foldchange = log2ratio - wt.log2ratio) %>% left_join(z.scores.ratio.IPR.filtered_fup) %>% select(-gRNA) %>% na.omit()

```


# Run the model: For this I need average values => I'll include r.squared and adj.r.squared as well
```{r}
#DDR library
# KO samples per replicate
gene.KO.CCD.dt_fup <- log2.distance.ratio.KO_fup %>% dplyr::group_by(gene,gRNA,barcode) %>% dplyr::summarise(mean.log2foldchange = mean(log2.foldchange,na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID")) %>% distinct() %>% na.omit()

model.pcr.fup <- tibble(gene = NA, gRNA = NA, num_comp = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)

for (i in unique(gene.KO.CCD.dt_fup$gene)){
  for (k in unique(gene.KO.CCD.dt_fup$gRNA)) {
  for (j in 1:10){
gene.library.dt <- filter(gene.KO.CCD.dt_fup, gene == i & gRNA == k)
if (nrow(gene.library.dt) < 19) {
  next
}
# Run a model per gene
set.seed(1)
PCR_model_DDR_test <- pls::pcr(mean.log2foldchange~CTCF+EZH2+H2AFZ+H3K27ac+H3K27me3+H3K36me3+H3K4me1+H3K4me2+H3K4me3+H3K79me2+H3K9me2+H3K9me3+H4K5acK8ac+HDAC1+HDAC2+HDAC3+POL2+POL2AS2+SMC3+LMNB1+Dam+DNAse+m5C+late_replicating+TTseq, data=gene.library.dt , validation="CV")

pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = j)
combined.dt <- tibble(measured = gene.library.dt$mean.log2foldchange, predicted = as.numeric(pcr_pred))
pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% glance()
model.pcr.fup <- model.pcr.fup %>% add_row(gene = i, gRNA = k, num_comp = j, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value))
  }
  }
}
# KO samples per replicate
gene.controls.CCD.dt_fup <- log2.distance.ratio.controls_fup %>% dplyr::group_by(well,gene,barcode) %>% dplyr::summarise(mean.log2foldchange = mean(log2.foldchange,na.rm = T)) %>% select(gene, gRNA = well, barcode, mean.log2foldchange) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID")) %>% distinct() %>% na.omit()

nt.model.value_chromatin_fup <- tibble(gene = NA, gRNA = NA, num_comp = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)

for (i in unique(gene.controls.CCD.dt_fup$gene)){
  for (k in unique(gene.controls.CCD.dt_fup$gRNA)) {
  for (j in 1:10){
gene.library.dt <- filter(gene.controls.CCD.dt_fup, gene == i & gRNA == k)
if (nrow(gene.library.dt) < 19) {
  next
}
# Run a model per gene
set.seed(1)
PCR_model_DDR_test <- pls::pcr(mean.log2foldchange~CTCF+EZH2+H2AFZ+H3K27ac+H3K27me3+H3K36me3+H3K4me1+H3K4me2+H3K4me3+H3K79me2+H3K9me2+H3K9me3+H4K5acK8ac+HDAC1+HDAC2+HDAC3+POL2+POL2AS2+SMC3+LMNB1+Dam+DNAse+m5C+late_replicating+TTseq, data=gene.library.dt , validation="CV")

pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = j)
combined.dt <- tibble(measured = gene.library.dt$mean.log2foldchange, predicted = as.numeric(pcr_pred))
pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% glance()
nt.model.value_chromatin_fup <- nt.model.value_chromatin_fup %>% add_row(gene = i, gRNA = k, num_comp = j, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value))
  }
  }
}


#non-targetings
significant.nt_fup <- nt.model.value_chromatin_fup %>% dplyr::select(num_comp, p.value, gene, gRNA) %>% mutate(sample = "nt") 

#significant genes
significant.genes_p.value_fup <- model.pcr.fup %>% dplyr::select(num_comp, p.value, gene, gRNA) %>% mutate(sample = "ko") %>% bind_rows(significant.nt_chromatin_fup) %>% filter(num_comp == 3) %>% mutate(p.adj = p.adjust(p.value, method = "BH")) %>% select(gene,p.value,p.adj, sample, gRNA)

#How maany nt are picked
sgnf.values_fup <- significant.genes_p.value_fup %>% filter(p.adj < 0.05) %>% dplyr::group_by(sample) %>% dplyr::summarise(count = n())
#Total numbers
all.values <- significant.genes_p.value_chromatin %>% dplyr::group_by(sample) %>% dplyr::summarise(all_count = n())
hits.percentage.chromatin <- all.values %>% left_join(sgnf.values) %>% mutate(freq = count/all_count )
```


#Step 4: Calculate CCDs for WT samples: 

# Calculate CCDs for KO samples: 
# A) Calculate this after computing the mean
```{r}
# Compute slopes for all experiments
gene.KO.CCD_fup <- tibble(gRNA = NA, gene = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

for (k in unique(gene.KO.CCD.dt_fup$gRNA)) {
for (h in unique(gene.KO.CCD.dt_fup$gene)) {
  for (j in colnames(gene.KO.CCD.dt_fup)[7:31]) {
    model.dt <- gene.KO.CCD.dt_fup %>% filter(gRNA == k & gene == h)
   if (nrow(model.dt) == 0) {
next
}
    model.epistasis.log2 <- lm(formula = mean.log2foldchange ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
   gene.KO.CCD_fup <- gene.KO.CCD_fup %>% add_row(gRNA = k,gene = h, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
  }
}
}

#Calculate KO proteins (terms)
gene.KO.CCD.mean.fup <- gene.KO.CCD_fup %>% reshape2::dcast(gRNA + gene + feature ~ term, value.var = "slope.log2") %>% dplyr::select(gRNA, gene ,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% left_join(significant.genes_p.value_fup %>% select(gene,gRNA, model_p.value = p.value, model_p.adj = p.adj)) %>% left_join(z.scores.ratio.IPR.filtered_fup ) %>% mutate(cross.x = -intercept/slope, slope_cor = case_when(cross.x < 1 ~ slope, cross.x > 1 ~ 0)) %>% mutate(CCD_value = case_when(pathway_signif == "MMEJ" & slope_cor < 0 ~ slope_cor, pathway_signif == "MMEJ" & slope_cor > 0 ~ 0, pathway_signif == "NHEJ" & slope > 0 ~ slope_cor, pathway_signif == "NHEJ" & slope < 0 ~ 0, T ~ slope_cor))

#Filter significant hits
fup_proteins_selected <- gene.KO.CCD.mean.fup %>% na.omit() %>% filter(model_p.adj < 0.05)

```

#Save every step
```{r}
#Step 2: Save log2 fold change data set: This will be saved separately as KO, WT and POLQ
saveRDS(log2.distance.ratio.KO, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_fold_change_ratio_KO.rds")
saveRDS(log2.distance.ratio.WT, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_fold_change_ratio_WT.rds")
saveRDS(log2.distance.ratio.POLQ, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_fold_change_ratio_POLQ.rds")

#Export epistasis 2 data table
saveRDS(gene.KO.CCD.dt_tmp, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_CCD_analysis_preprocessing.rds")

#Export CCDs from the mean values
saveRDS(DDR.CCD.values.mean, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_DDR_CCD_mean.rds")
saveRDS(chromatin.CCD.values.mean, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_chromatin_CCD_mean.rds")

#Export CCDs from calculation per replicate
saveRDS(DDR.CCD.values.pr.path, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_DDR_CCD_per_rep.rds")
saveRDS(chromatin.CCD.values.pr.path, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220314_Chromatin_CCD_per_rep.rds")

#Export CCD of significant genes based on model
saveRDS(gene.KO.CCD.mean.fup, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220503_follow_up_experiment_mll.rds")

# There not significant mean values

```

