---
title: "AGM20201222_screen_mmej_sep"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

#Report of the Chromatin DSB screening - This will be the last version.
I will start importing data from previous scripts
A - Data frame with different scores to use as input
B - Table with clustering data (Needs to be included once Christ finalises his analysis)

I will process data based on MMEJscore first. I filter data based gene editing % and read counts. Filtering strategy:
  Gene editing %: There are some wells that have 20-30% cutting efficiency - These I'll discard. Modification 20210609, For that I filtered out every well that has an average cutting efficiency lower than 25%".
  Read count: To avoid having low count read-outs, I will filter out things that have less than 30 mut.reads (+1 & -7)

Read count filtering was performed in the previous script. In this report, I will filter wells with low count reads and I will perform z-score normalization for MMEJscore. I also thought to do the same processing for cutting efficiency, but this is not possible because technical noise in transfection efficiency.


```{r}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}
```

```{r setup, include=FALSE}
library(tidyverse) 
library(ggbeeswarm)
library(corrplot)
library(pheatmap)
library(viridis)
library(factoextra)
library(NbClust)
library(ggpubr)
library(readxl)
library(gridExtra)
library(ggrepel)
```
## Functions
```{r, echo=F}
#Plot correlations
replicate.correlation.plot <- function(dt, i, j, k){
  ggplot(dt, aes_string(x = i, y = j,color = k)) +
    geom_point(alpha = 0.5) +
    theme_bw() + 
    coord_cartesian(expand = F) +
    theme(aspect.ratio = 1)
}

#mybreaks
myBreaks <- function(test, p) {c(seq(min(test), 0, length.out=ceiling(p/2) + 1), 
              seq(max(test)/p, max(test), length.out=floor(p/2)))}
```

#Columns in DDRDSB
 [1] "plate"         "replicate"     "bio.rep"       "well"          "barcode"       "MMEJscore"    
 [7] "NHEJscore"     "freqMMEJ"      "freqNHEJ"      "freqCut"       "NHEJMMEJratio" "sample"       
[13] "indelreads"    "FinRep"  

#Columns in new.domains.c5.bis
[1] "barcode"          "H3K27me3"         "H3K9me2"          "LAD"              "late_replicating"
[6] "euchromatin"      "transcription"    "chromatin" 

#Note 1: If both data tables have these columns it should run properly, if there are columns that are named differently. These names have to be modified in the script.

#Note 2: Go through the script and change every path after pdf() function to your home directory.

## load data
```{r}
#File with all parameters
DDRDSB <-readRDS("~/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/AGM20210527_screen_data_agm_ddr_final_genes.rds")

#Chromatin data
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')
```

#How data tables looks
```{r}
#Plot screen data
head(DDRDSB)

#Plot cluster data
head(clone5_z.score_chrom_tib)
```

## Technical noise variability
# Possible Suppl. figure in the paper. Noise evaluation in negative controls and KOs
To calculate z-scores, it is important that the noise levels of each of the measurements (MMEJscore and cutting efficiency). If the noise is too big or measurement inconsistent, I will only pick noise differences and not biological differences. A way for estimating the noise level is to average every meaasurement per well and compare these. Any technical variability should be mainly be picked here and then I will plot correlation between different replicates. 


# Technical noise in negative controls (WT background)
```{r}
# Possible supp technical figure with background noise levels. This is the reason why we exclusively focus on MMEJscore. 

# Calculate average mmej and cut efficiencies
noise.var.dt <- DDRDSB %>% 
  filter(sample == "WT") %>% 
  dplyr::group_by(well, plate, FinRep) %>% 
  dplyr::summarise(mean.cut = mean(freqCut, na.rm = T), mean.mmej = mean(MMEJscore, na.rm = T))

# Plot gene editing efficiency noise level
cut.var.dt <- noise.var.dt %>% 
  reshape2::dcast(well + plate ~ FinRep, value.var = "mean.cut") %>%
  na.omit()

p.c.var.1 <- replicate.correlation.plot(cut.var.dt, "R1","R4","plate") + theme(legend.position = "top") + ylim(c(0.2,0.8)) + xlim(c(0.2,0.8))
p.c.var.2 <- replicate.correlation.plot(cut.var.dt, "R1","R5","plate")+ theme(legend.position = "top") + ylim(c(0.2,0.8)) + xlim(c(0.2,0.8))
p.c.var.3 <- replicate.correlation.plot(cut.var.dt, "R4","R5","plate")+ theme(legend.position = "top")+ ylim(c(0.2,0.8)) + xlim(c(0.2,0.8))
g.c.var <- arrangeGrob(p.c.var.1,p.c.var.2,p.c.var.3, nrow = 1)

# Plot MMEJscore gene editing efficiency noise level
mmej.var.dt <- noise.var.dt %>% 
  reshape2::dcast(well + plate ~ FinRep, value.var = "mean.mmej") %>%
  na.omit()

p.m.var.1 <- replicate.correlation.plot(mmej.var.dt, "R1","R4","plate") + theme(legend.position = "top") + ylim(c(0,0.5)) + xlim(c(0,0.5))
p.m.var.2 <- replicate.correlation.plot(mmej.var.dt, "R1","R5","plate")+ theme(legend.position = "top")+ ylim(c(0,0.5)) + xlim(c(0,0.5))
p.m.var.3 <- replicate.correlation.plot(mmej.var.dt, "R4","R6","plate")+ theme(legend.position = "top")+ ylim(c(0,0.5)) + xlim(c(0,0.5))
g.m.var <- arrangeGrob(p.m.var.1,p.m.var.2,p.m.var.3, nrow = 1)

#Combine both plots together with tittles
t.c <- text_grob("WT background average cutting efficiency - Color code: Plate #")
t.m <- text_grob("WT background average MMEJscore - Color code: Plate #")

#These plots were already exported to pdf, so I wont run them.

#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20210608_KO_noise_bis.pdf", height = 5)
grid.arrange(t.c,g.c.var,t.m,g.m.var, ncol = 1, heights = c(0.5,5,0.5,5))
#dev.off()
```

##Technical noise plot (KO transfected)
```{r}
# Calculate average mmej and cut efficiencies
noise.var.KO.dt <- DDRDSB %>% 
  filter(sample == "KO") %>% 
  dplyr::group_by(well, plate, FinRep) %>% 
  dplyr::summarise(mean.cut = mean(freqCut, na.rm = T), mean.mmej = mean(MMEJscore, na.rm = T))

# Plot gene editing efficiency noise level
cut.var.KO.dt <- noise.var.KO.dt %>% 
  reshape2::dcast(well + plate ~ FinRep, value.var = "mean.cut") %>%
  na.omit()

p.c.var.4 <- replicate.correlation.plot(cut.var.KO.dt, "R1","R4","plate") + theme(legend.position = "none") + ylim(c(0,1)) + xlim(c(0,1))
p.c.var.5 <- replicate.correlation.plot(cut.var.KO.dt, "R1","R5","plate")+ theme(legend.position = "none") + ylim(c(0,1)) + xlim(c(0,1))
p.c.var.6 <- replicate.correlation.plot(cut.var.KO.dt, "R4","R5","plate")+ theme(legend.position = "none") + ylim(c(0,1)) + xlim(c(0,1))
g.c.var.KO <- arrangeGrob(p.c.var.4,p.c.var.5,p.c.var.6, nrow = 1)

# Plot MMEJscore gene editing efficiency noise level
mmej.var.KO.dt <- noise.var.KO.dt %>% 
  reshape2::dcast(well + plate ~ FinRep, value.var = "mean.mmej") %>%
  na.omit()

p.m.var.4 <- replicate.correlation.plot(mmej.var.KO.dt, "R1","R4","plate") + theme(legend.position = "none") + ylim(c(0,1)) + xlim(c(0,1))
p.m.var.5 <- replicate.correlation.plot(mmej.var.KO.dt, "R1","R5","plate")+ theme(legend.position = "none")+ ylim(c(0,1)) + xlim(c(0,1))
p.m.var.6 <- replicate.correlation.plot(mmej.var.KO.dt, "R4","R5","plate")+ theme(legend.position = "none")+ ylim(c(0,1)) + xlim(c(0,1))
g.m.var.KO <- arrangeGrob(p.m.var.4,p.m.var.5,p.m.var.6, nrow = 1)

#Combine both plots together with tittles
t.c.KO <- text_grob("KO average cutting efficiency - Color code: Plate #")
t.m.KO <- text_grob("KO average MMEJscore - Color code: Plate #")
grid.arrange(t.c.KO,g.c.var.KO,t.m.KO,g.m.var.KO, ncol = 1, heights = c(0.5,5,0.5,5))

#These plots were already exported to pdf, so I wont run them.

#pdf(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20210608_KO_noise_bis.pdf", height = 5)
grid.arrange(t.c.KO,g.c.var.KO,t.m.KO,g.m.var.KO, ncol = 1, heights = c(0.5,5,0.5,5))
#dev.off()

```

## Apply a filter based on gene editing efficiency per well (At this step, I'll filter out the data)
```{r}
#Filter wells with mean cutting efficiency lower than 25%
filter.out.wells <- noise.var.KO.dt %>%
  filter(mean.cut < 0.25) %>% 
  select(well,plate,FinRep)

#final filter
DDRDSB <- DDRDSB %>% 
  anti_join(filter.out.wells, by = c("plate", "well", "FinRep"))
```

## MMEJscore data processing
MMEJ score processing contains three different steps:
  1. Selection of WT background breaks (negative controls) and calculation of z-scores based on mean & sd (per barcode and replicate)
  2. Combination of IPRs by chromatin type as dependent variables
  3. Combination of replicates as independent variables
  
```{r}
#STEP 1 : Calculate z-scores

# Filter data for only wt DSBs
wt.filter <- DDRDSB %>% 
  filter(FinRep %in% c("R1","R4","R5"), sample == "WT") %>% 
  dplyr::group_by(FinRep)

wt.DDRDSB <- wt.filter %>% 
  dplyr::group_by(barcode,FinRep) %>% 
  dplyr::summarise(WT.mean = mean(MMEJscore, na.rm = T), WT.sd = sd(MMEJscore, na.rm = T)) %>% 
  right_join(wt.filter, by = c("barcode","FinRep"))

#Export data for future plots
wt.mmej.score <- wt.DDRDSB %>% 
  dplyr::select(barcode,FinRep,MMEJscore = WT.mean) %>% 
  distinct() %>% 
  left_join(new.domains.c5.bis %>% 
              dplyr::select(barcode,chromatin)) %>% 
  mutate(gene = "WT")

# Calculate z-score (per Fin Rep)
z.scores.calc.mmej <- DDRDSB %>% 
  filter (FinRep %in% c("R1","R4","R5")) %>% 
  left_join(wt.DDRDSB %>% 
              dplyr::select(barcode,FinRep, WT.mean,WT.sd), by = c("barcode","FinRep")) %>% 
  mutate(z.score = (MMEJscore - WT.mean)/WT.sd) %>% distinct()

#STEP 2 : Combine z-scores per chromatin

# Calculate z-score (per Fin REep)
ch.z.scores.calc.mmej <- z.scores.calc.mmej %>% 
  left_join(new.domains.c5.bis %>% 
              dplyr::select(barcode,chromatin), by = c("barcode")) %>%
  dplyr::group_by(FinRep, well,plate,chromatin,sample, gene) %>%
  dplyr::mutate(ch.z.score = sum(z.score, na.rm = T)/n()) %>% 
  distinct()

# STEP 3: Combine three replicates together and calculate p.value conversion of Z score (t population with 1 or two freedom degrees)
cchz.scores.calc.mmej <- ch.z.scores.calc.mmej %>% 
  dplyr::select(FinRep, well,plate,chromatin,sample, gene, ch.z.score) %>% 
  distinct() %>% 
  dplyr::group_by(well,plate,chromatin,sample, gene) %>% 
  dplyr::mutate(mmej.SCscore = sum(ch.z.score, na.rm = T)/sqrt(n())) %>% 
  ungroup() %>% 
  select(well,plate,chromatin,sample,gene,mmej.SCscore) %>% 
  distinct() %>% 
  dplyr::group_by(chromatin) %>% 
  distinct() %>% 
  right_join(ch.z.scores.calc.mmej, by = c("well", "plate","sample", "gene","chromatin")) %>% ungroup() 


# STEP 4: Remove variables and export data (filter data that doesn't have a p.value - single value)
chz.scores.mmej <- cchz.scores.calc.mmej %>% 
  filter(!is.na(gene)) %>% 
  dplyr::select(gene,barcode,chromatin,mmej.SCscore, mmej.ch.score = ch.z.score,MMEJscore,freqCut,mmej.z.score = z.score,bio.rep,FinRep,plate,well,sample) 


#STEP 5: dcast genes per chromatin value and replicate
chz.mmej.dcast <- reshape2::dcast(chz.scores.mmej %>% 
                                    ungroup() %>%
                                    dplyr::select(gene,chromatin,mmej.SCscore) %>%
                                    distinct(), gene ~ chromatin, value.var = "mmej.SCscore", fun.aggregate = mean)

## The output of this chunk is a melted data table that allows multiple grouping (chz.scores.mmej) and a casted dt that get each average chromatin score for each gene (chz.mmej.dcast)

```

#Correlation plots for each different step in controls and KO (Supplementary figure 2)

##Correlation plots MMEJscore 
```{r}
# Correlation at STEP 0
controls.DDRDSB <- DDRDSB %>% 
  filter(FinRep %in% c("R1","R4","R5"),sample != "KO") %>%
  reshape2::dcast(plate + well + barcode + sample ~ FinRep, value.var = "MMEJscore") %>%
  na.omit()

pa <- replicate.correlation.plot(controls.DDRDSB,"R1","R4","sample") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")
pb <- replicate.correlation.plot(controls.DDRDSB,"R1","R5","sample") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")
pc <- replicate.correlation.plot(controls.DDRDSB,"R4","R5","sample") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")
rcp.l.up <- arrangeGrob(pa,pb,pc, nrow = 1)

# Correlation at STEP 1
control.z.mmej.screen.full <- z.scores.calc.mmej %>% 
  filter(sample %in% c("WT","POLQ")) %>%
  reshape2::dcast(well + plate + sample + barcode ~ FinRep, value.var = "z.score", fun.aggregate = mean) %>%
  na.omit()

pd <- replicate.correlation.plot(control.z.mmej.screen.full,"R1","R4","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
pe <- replicate.correlation.plot(control.z.mmej.screen.full,"R1","R5","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
pf <- replicate.correlation.plot(control.z.mmej.screen.full,"R4","R5","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
rcp.l.mid <- arrangeGrob(pd,pe,pf, nrow = 1)

# Correlation at STEP 2
control.chz.mmej.screen.full <- ch.z.scores.calc.mmej %>% 
  filter(sample %in% c("WT","POLQ")) %>%
  reshape2::dcast(well + plate + sample + chromatin ~ FinRep, value.var = "ch.z.score", fun.aggregate = mean) %>%
  na.omit()

pg <- replicate.correlation.plot(control.chz.mmej.screen.full,"R1","R4","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
ph <- replicate.correlation.plot(control.chz.mmej.screen.full,"R1","R5","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
pi <- replicate.correlation.plot(control.chz.mmej.screen.full,"R4","R5","sample") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
rcp.l.down <- arrangeGrob(pg,ph,pi, nrow = 1)

#Density plot after combination
pj <- ggplot(cchz.scores.calc.mmej %>% dplyr::select(chromatin, well,plate,mmej.SCscore,sample) %>% distinct() %>% filter(sample %in% c("WT","POLQ"))) + geom_jitter(aes(mmej.SCscore,"Controls", group = sample, color = sample)) + theme_bw() + coord_cartesian(expand = T) + theme(axis.title.y = element_blank(), axis.text.y = element_text(angle = 90,hjust = 0.5))

# legends
p.leg.control <- cowplot::get_legend(replicate.correlation.plot(control.chz.mmej.screen.full,"R4","R5","sample") + theme(legend.position = "bottom"))
title.main <- text_grob("Controls sample correlation per replicate over steps")
title.n.up <- text_grob("STEP 0 :  MMEJscore per barcode")
title.n.mid <- text_grob("STEP 1 : Z scores per barcode")
title.n.down <- text_grob("STEP 2 : Z scores per chromatin type")
title.n.final <- text_grob("STEP 3 : Z scores difference combination")

#Final plot saved
cor.plot.mmej.c <- arrangeGrob(title.n.up,rcp.l.up,title.n.mid,rcp.l.mid,title.n.down,rcp.l.down,title.n.final,pj, ncol = 1, heights = c(0.1,3,0.1,3,0.1,3,0.2,2))
```

##Control plots
```{r}
#CONTROL PLOT: Correlation plot between biological replicates
control.correlation <- chz.scores.mmej %>% 
  filter(sample %in% c("POLQ","WT")) %>%
  select(bio.rep,mmej.ch.score,chromatin,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(chromatin + sample + well + plate ~ FinRep, value.var = "mmej.ch.score") %>%
  rowwise() %>%
  mutate(mmej.SCscore = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(control.correlation) + 
  geom_point(aes(R1,R5, color = mmej.SCscore)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 2") + 
  ggtitle("Control correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") +
  facet_wrap(~ chromatin)

#CONTROL PLOT2: KO correlation
KO.correlation <- chz.scores.mmej %>% 
  filter(sample == "KO") %>%
  select(bio.rep,mmej.ch.score,chromatin,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(chromatin + sample + well + plate ~ FinRep, value.var = "mmej.ch.score") %>%
  rowwise() %>%
  mutate(mmej.SCscore = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(KO.correlation) + 
  geom_point(aes(R1,R5, color = mmej.SCscore)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 3") + 
  ggtitle("KO correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") + 
  facet_wrap(~ chromatin) 
```  


##Same plot for KO transfected wells
```{r}
# Correlation at STEP 0
KO.DDRDSB <- DDRDSB %>% 
  filter(FinRep %in% c("R1","R4","R5"),sample == "KO") %>%
  reshape2::dcast(plate + well + barcode ~ FinRep, value.var = "MMEJscore") %>%
  na.omit()

#Plot_control correlation between replicates (MMEJscore)
p1 <- replicate.correlation.plot(KO.DDRDSB,"R1","R4","barcode") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")
p2 <- replicate.correlation.plot(KO.DDRDSB,"R1","R5","barcode") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")
p3 <- replicate.correlation.plot(KO.DDRDSB,"R4","R5","barcode") + xlim(c(0,1)) + ylim(c(0,1)) + theme(legend.position = "none")


#Grid three plots to one panel
rcp.n.up <- arrangeGrob(p1,p2,p3, nrow = 1)

# Correlation at STEP 1
KO.z.mmej.screen.full <- z.scores.calc.mmej %>% 
  filter(FinRep %in% c("R1","R4","R5"),sample == "KO") %>%
  reshape2::dcast(well + plate + sample + barcode ~ FinRep, value.var = "z.score", fun.aggregate = mean) %>%
  na.omit()

p4 <- replicate.correlation.plot(KO.z.mmej.screen.full,"R1","R4","barcode") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
p5 <- replicate.correlation.plot(KO.z.mmej.screen.full,"R1","R5","barcode") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
p6 <- replicate.correlation.plot(KO.z.mmej.screen.full,"R4","R5","barcode") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")

#Grid three plots to one panel
rcp.n.mid <- arrangeGrob(p4,p5,p6, nrow = 1)

# Correlation at STEP 2
KO.chz.mmej.screen.full <- ch.z.scores.calc.mmej %>% 
  filter(sample == "KO") %>%
  reshape2::dcast(well + plate + sample + chromatin ~ FinRep, value.var = "ch.z.score", fun.aggregate = mean) %>%
  na.omit()

p7 <- replicate.correlation.plot(KO.chz.mmej.screen.full,"R1","R4","chromatin") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
p8 <- replicate.correlation.plot(KO.chz.mmej.screen.full,"R1","R5","chromatin") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")
p9 <- replicate.correlation.plot(KO.chz.mmej.screen.full,"R4","R5","chromatin") + xlim(c(-15,10)) + ylim(c(-15,10)) + theme(legend.position = "none")

#Grid three plots to one panel
rcp.n.down <- arrangeGrob(p7,p8,p9, nrow = 1)

#Density plot after combination
p10 <- ggplot(cchz.scores.calc.mmej %>% dplyr::select(chromatin, well,plate,mmej.SCscore,sample) %>% distinct() %>% filter(sample == "KO")) + geom_jitter(aes(mmej.SCscore,chromatin, group = sample, color =chromatin)) + theme_bw() + coord_cartesian(expand = T) + theme(axis.title.y = element_blank())

#Tittles
title.main <- text_grob("Controls sample correlation per replicate over steps")
title.n.up <- text_grob("STEP 0 :  MMEJscore per barcode")
title.n.mid <- text_grob("STEP 1 : Z scores per barcode")
title.n.down <- text_grob("STEP 2 : Z scores per chromatin type")
title.n.final <- text_grob("STEP 3 : Z scores difference combination")
                                     
# grid arrange
grid.arrange(title.n.up,rcp.n.up,title.n.mid,rcp.n.mid,title.n.down,rcp.n.down,title.n.final,p10, ncol = 1, heights = c(0.1,2.5,0.1,2.5,0.1,2.5,0.1,2.5))

#This plot was already exported, I will block this lines

#pdf(file = "~/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/export/xv20210608_KO_z_scores.pdf", height = 5)
#grid.arrange(title.n.up,rcp.n.up,title.n.mid,rcp.n.mid,title.n.down,rcp.n.down,title.n.final,p10, ncol = 1, heights = c(0.1,2.5,0.1,2.5,0.1,2.5,0.1,2.5))
#dev.off()

```

#Other clasification
I will clasify the data in other categories:
1. Euchromatin & Heterochromatin
2. LAD vs. iLADs
3. Early replication vs. late replication
4. Triple heterochromatin vs. rest

```{r}
#I will start arranging the IPR-chromatin datatable
c5.extra.classification <- new.domains.c5 %>%
  mutate(eu_het = case_when(euchromatin == 1 | transcription == 1 ~ "euchromatin",
                   TRUE ~ "heterochromatin"),
         is_triple = factor(if_else(chromatin == 'late_replicating-LAD-H3K9me2',1,0)))

# Add this data table to the z.score table
other.chr.z.scores <- z.scores.calc.mmej %>% 
  left_join(c5.extra.classification %>% 
              dplyr::select(barcode,LAD,late_replicating,eu_het,is_triple), by = c("barcode"))
```

# Calculate z-scores LAD vs. iLADs
```{r}
# calculate LAD.z.score
LAD.z.score <- other.chr.z.scores %>% 
  dplyr::group_by(FinRep, well,plate,LAD,sample, gene) %>%
  dplyr::mutate(LAD.z.score = sum(z.score, na.rm = T)/n()) %>% 
  distinct()

# calculate over different replicates
LAD.final.z.score <- LAD.z.score %>% 
  dplyr::select(FinRep, well,plate,LAD,sample, gene, LAD.z.score) %>% 
  distinct() %>% 
  dplyr::group_by(well,plate,LAD,sample, gene) %>% 
  dplyr::mutate(mmej.LADscore = sum(LAD.z.score, na.rm = T)/sqrt(n())) %>% 
  ungroup() %>% 
  select(well,plate,LAD,sample,gene,mmej.LADscore) %>% 
  distinct() %>% 
  dplyr::group_by(LAD) %>% 
  distinct() %>% 
  right_join(LAD.z.score, by = c("well", "plate","sample", "gene","LAD")) %>% ungroup() 

#Control plots (reproducibility & differences)

LAD.correlation <- LAD.z.score %>% 
  filter(sample == "KO") %>%
  select(bio.rep,LAD.z.score,LAD,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(LAD + sample + well + plate ~ FinRep, value.var = "LAD.z.score") %>%
  rowwise() %>%
  mutate(mmej.LAD.score = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(LAD.correlation) + 
  geom_point(aes(R1,R5, color = mmej.LAD.score)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 3") + 
  ggtitle("KO correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") + 
  facet_wrap(~ LAD) 

# 2nd plot differences between LAD and iLAD
LAD_vs_iLAD <- LAD.final.z.score %>% 
  filter(sample == "KO") %>%
  select(mmej.LADscore,LAD,gene) %>%
  distinct() %>%
  reshape2::dcast(gene ~ LAD, value.var = "mmej.LADscore", fun.aggregate = mean) %>%
  mutate(is_ATR = gene == "ATR") %>%
  arrange(is_ATR) %>%
  na.omit()


lad_vs_ilad_plot <- ggplot(LAD_vs_iLAD) + 
  geom_point(aes(`0`,`1`, color = gene == "ATR" )) +
  theme_bw() +
  xlab("z-score iLAD") + 
  ylab("z-score LAD") + 
  ggtitle("LADs and iLADs correlation") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_hline(yintercept = c(-2.5,2.5), linetype= 3) +
  geom_vline(xintercept = c(-2.5,2.5), linetype = 3) + 
  coord_fixed(xlim = c(-12,35), ylim = c(-12,35))

```
# Do the same for late replicating vs. early replicating
```{r}
#calculate repli.z.score
repli.z.score <- other.chr.z.scores %>% 
  dplyr::group_by(FinRep, well,plate,late_replicating,sample, gene) %>%
  dplyr::mutate(repli.z.score = sum(z.score, na.rm = T)/n()) %>% 
  distinct()

# calculate over different replicates
repli.final.z.score <- repli.z.score %>% 
  dplyr::select(FinRep, well,plate,late_replicating,sample, gene, repli.z.score) %>% 
  distinct() %>% 
  dplyr::group_by(well,plate,late_replicating,sample, gene) %>% 
  dplyr::mutate(mmej.repliscore = sum(repli.z.score, na.rm = T)/sqrt(n())) %>% 
  ungroup() %>% 
  select(well,plate,late_replicating,sample,gene,mmej.repliscore) %>% 
  distinct() %>% 
  dplyr::group_by(late_replicating) %>% 
  distinct() %>% 
  right_join(repli.z.score, by = c("well", "plate","sample", "gene","late_replicating")) %>% ungroup() 

#Control plots (reproducibility & differences)

repli.correlation <- repli.z.score %>% 
  filter(sample == "KO") %>%
  select(bio.rep,repli.z.score,late_replicating,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(late_replicating + sample + well + plate ~ FinRep, value.var = "repli.z.score") %>%
  rowwise() %>%
  mutate(mmej.repliscore = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(repli.correlation) + 
  geom_point(aes(R1,R5, color = mmej.repliscore)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 3") + 
  ggtitle("KO correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") + 
  facet_wrap(~ late_replicating) 

# 2nd plot differences between LAD and iLAD
late_vs_early <- repli.final.z.score %>% 
  filter(sample == "KO") %>%
  select(mmej.repliscore,late_replicating,gene) %>%
  distinct() %>%
  reshape2::dcast(gene ~ late_replicating, value.var = "mmej.repliscore", fun.aggregate = mean) %>%
  mutate(is_ATR = gene == "ATR") %>%
  arrange(is_ATR) %>%
  na.omit()


late_vs_early.plot <- ggplot(late_vs_early) + 
  geom_point(aes(`0`,`1`, color = is_ATR)) +
  theme_bw() +
  xlab("z-score early replicating") + 
  ylab("z-score late replicating") + 
  ggtitle("Early vs. late replication") + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_hline(yintercept = c(-2.5,2.5), linetype= 3) +
  geom_vline(xintercept = c(-2.5,2.5), linetype = 3) +
  coord_fixed(xlim = c(-12,35), ylim = c(-12,35))

```
# Now with euchromatin vs. heterochromatin

```{r}
#calculate repli.z.score
eu.z.score <- other.chr.z.scores %>% 
  dplyr::group_by(FinRep, well,plate,eu_het,sample, gene) %>%
  dplyr::mutate(eu.z.score = sum(z.score, na.rm = T)/n()) %>% 
  distinct()

# calculate over different replicates
eu.final.z.score <- eu.z.score %>% 
  dplyr::select(FinRep, well,plate,late_replicating,sample, gene, eu.z.score) %>% 
  distinct() %>% 
  dplyr::group_by(well,plate,eu_het,sample, gene) %>% 
  dplyr::mutate(mmej.euscore = sum(eu.z.score, na.rm = T)/sqrt(n())) %>% 
  ungroup() %>% 
  select(well,plate,eu_het,sample,gene,mmej.euscore) %>% 
  distinct() %>% 
  dplyr::group_by(eu_het) %>% 
  distinct() %>% 
  right_join(eu.z.score, by = c("well", "plate","sample", "gene","eu_het")) %>% ungroup() 

#Control plots (reproducibility & differences)

eu.correlation <- eu.z.score %>% 
  filter(sample == "KO") %>%
  select(bio.rep,eu.z.score,eu_het,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(eu_het + sample + well + plate ~ FinRep, value.var = "eu.z.score") %>%
  rowwise() %>%
  mutate(mmej.euscore = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(eu.correlation) + 
  geom_point(aes(R1,R5, color = mmej.euscore)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 3") + 
  ggtitle("KO correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") + 
  facet_wrap(~ eu_het) 

# 2nd plot differences between LAD and iLAD
eu_vs_het <- eu.final.z.score %>% 
  filter(sample == "KO") %>%
  select(mmej.euscore,eu_het,gene) %>%
  distinct() %>%
  reshape2::dcast(gene ~ eu_het, value.var = "mmej.euscore", fun.aggregate = mean) %>%
  mutate(is_ATR = gene == "ATR" ) %>%
  arrange(is_ATR) %>%
  na.omit()


eu_vs_het.plot <- ggplot(eu_vs_het) + 
  geom_point(aes(euchromatin,heterochromatin,color = is_ATR)) +
  theme_bw() +
  ggtitle("Euchromatin vs. heterochromatin") +
  xlab("z-score euchromatin") + 
  ylab("z-score heterochromatin") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_hline(yintercept = c(-2.5,2.5), linetype= 3) +
  geom_vline(xintercept = c(-2.5,2.5), linetype = 3) +
  coord_fixed(xlim = c(-15,45), ylim = c(-15,45))

```

# I will add a 4th correlation to this: Triple heterochromatin vs. rest
```{r}
#calculate repli.z.score
th.z.score <- other.chr.z.scores %>% 
  dplyr::group_by(FinRep, well,plate,is_triple,sample, gene) %>%
  dplyr::mutate(th.z.score = sum(z.score, na.rm = T)/n()) %>% 
  distinct()

# calculate over different replicates
th.final.z.score <- th.z.score %>% 
  dplyr::select(FinRep, well,plate,is_triple,sample, gene, th.z.score) %>% 
  distinct() %>% 
  dplyr::group_by(well,plate,is_triple,sample, gene) %>% 
  dplyr::mutate(mmej.thscore = sum(th.z.score, na.rm = T)/sqrt(n())) %>% 
  ungroup() %>% 
  select(well,plate,is_triple,sample,gene,mmej.thscore) %>% 
  distinct() %>% 
  dplyr::group_by(is_triple) %>% 
  distinct() %>% 
  right_join(th.z.score, by = c("well", "plate","sample", "gene","is_triple")) %>% ungroup() 

#Control plots (reproducibility & differences)

th.correlation <- th.z.score %>% 
  filter(sample == "KO") %>%
  select(bio.rep,th.z.score,is_triple,sample,well,plate,FinRep) %>%
  distinct() %>%
  reshape2::dcast(is_triple + sample + well + plate ~ FinRep, value.var = "th.z.score") %>%
  rowwise() %>%
  mutate(mmej.thscore = (R1 + R4 + R5)/sqrt(3)) %>%
  na.omit()

ggplot(th.correlation) + 
  geom_point(aes(R1,R5, color = mmej.thscore)) +
  coord_fixed(xlim = c(-12,25), ylim = c(-12,25)) +
  theme_bw() +
  xlab("z-score replicate 1") + 
  ylab("z-score replicate 3") + 
  ggtitle("KO correlation between two biological replicates") +
  scale_color_gradient2(mid = "grey90") + 
  facet_wrap(~ is_triple) 

# 2nd plot differences between LAD and iLAD
triple.cor <- th.final.z.score %>% 
  filter(sample == "KO") %>%
  select(mmej.thscore,is_triple,gene) %>%
  distinct() %>%
  reshape2::dcast(gene ~ is_triple, value.var = "mmej.thscore", fun.aggregate = mean) %>%
  mutate(is_ATR = gene == "ATR" ) %>%
  arrange(is_ATR) %>%
  na.omit()


tripl.cor.plot <- ggplot(triple.cor) + 
  geom_point(aes(`0`,`1`,color = is_ATR)) +
  theme_bw() +
  ggtitle("Triple heterochromatin vs. others") +
  xlab("z-score Others") + 
  ylab("z-score Triple heterochromatin") +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_hline(yintercept = c(-2.5,2.5), linetype= 3) +
  geom_vline(xintercept = c(-2.5,2.5), linetype = 3) +
  coord_fixed(xlim = c(-10,30), ylim = c(-10,30))

```

# Plot all three plots
```{r}
#Plot all three correlations
lad_vs_ilad_plot
late_vs_early.plot
eu_vs_het.plot
tripl.cor.plot

#Export to pdf
pdf("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/export/xv20210610_correlation_ATR.pdf")
lad_vs_ilad_plot
late_vs_early.plot
eu_vs_het.plot
tripl.cor.plot
dev.off()
```

###Conclusions: This looks nice and process cleans data quite nicely. The outcome of this initial step is a dataframe with multiple categories and a z-score for each of them. We can answer several question with this dataset:
1- Which genes lead to differential effects in chromatin?
2- Is there any difference between LAD and iLADs when it comes to repair?
3- Which genes have a global effect on DDR repair?

After this quality control plots this data tables are fixed and I will export this data sets.


```{r}
# Export both data tables: full detail and dcast data table

#Export initial data table with ratio
saveRDS(DDRDSB %>% filter(FinRep %in% c("R1","R4","R5") & !gene %in% c("DMSO","LBR_ctrl")), "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20220110_screen_ddr_all_computations.rds")

#All data with all the relevant info by chromatin type
saveRDS(chz.scores.mmej, "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")

# dcast data only by chromatin type
saveRDS(chz.mmej.dcast %>%
          filter(gene != "LBR_ctrl"), file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_dcast.rds")

# All data per chromatin type
saveRDS(cchz.scores.calc.mmej %>% 
          filter(sample %in% c("WT","POLQ")) %>% 
          dplyr::select(gene = sample,barcode,chromatin,mmej.SCscore, mmej.ch.score = ch.z.score,MMEJscore,freqCut,mmej.z.score = z.score,bio.rep,FinRep,plate,well,sample), file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")

#Dcast version of all the data for the different correlations by chromatin features
saveRDS(LAD_vs_iLAD, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210610_lad_vs_ilad.rds")
saveRDS(eu_vs_het, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210610_euchr_vs_heteroch.rds")
saveRDS(late_vs_early, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210610_late_vs_early.rds")
saveRDS(triple.cor, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210610_triple_het_corr.rds")


```
