---
title: "xv20211121_epistasis_model_bas"
output: html_document
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
```

# Aesthetic legend for the whole paper
```{r}
# Pathway color palette (MMEJ, wt/no_effect, NHEJ)
pathway.colors <- c("#EB2030","grey90","#2E358F")
# Pathway symbol palette ()
pathway.symbol <- c(19,17)

# Chromatin color palette
#chrom_colors = c( c("euchromatin" = "#F7941D", "other" = "#838687",  "H3K27me3" = "#D21F8A", "triple" = "#662D91"))
chromatin.colors <- c("#D21F8A","#838687","#F7941D","#662D91")
```

# Import data tables

```{r, include= FALSE}
#dcast table
#ddr
ddr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_dcast.rds")
#chromatin
chr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_dcast.rds")
  
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

#All variable table
ddr.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20220110_screen_ddr_all_computations.rds")
ddr.screen.all.sel <- ddr.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep = bio.rep,t.rep = FinRep) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#Chromatin all 
chromatin.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20220110_chromatin_library_all.rds")
chromatin.screen.all.sel <- chromatin.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep,t.rep) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "Chromatin")

# bind both dataset
both.screen.all.sel <- bind_rows(ddr.screen.all.sel,chromatin.screen.all.sel) %>% filter(!gene %in% c("DMSO","LBR_ctrl"))

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"

#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

#Chromatin data
clone5_chrom_tib <- readRDS("/DATA/projects/DSBrepair/data/R/rs20200519_clone5_newdoms_chromatin.RDS")
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)

```

# I will test NHEJ:MMEJ ratio reproducibility first
```{r}
# Select WT population
wt.follow.up <- chromatin.followup %>% filter(gene == "tracr")
wt.follow.up.ratio <- wt.follow.up %>% reshape2::dcast(barcode + well + plate ~ bio.rep, value.var = "NHEJMMEJratio")
ggplot(wt.follow.up) + geom_density(aes(NHEJMMEJratio, color = plate, linetype = bio.rep))

ggplot(wt.ddr.screen.all.dcast.ratio) + geom_point(aes(R1,R4)) + geom_abline()
ggplot(wt.ddr.screen.all.dcast.ratio) + geom_point(aes(R1,R2)) + geom_abline()

wt.ddr.screen.all.dcast.mmejscore <- wt.ddr.screen.all.data %>% reshape2::dcast(barcode + well + plate + library ~ t.rep, value.var = "MMEJscore")

ggplot(wt.ddr.screen.all.dcast.mmejscore) + geom_point(aes(R1,R4)) + geom_abline()
ggplot(wt.ddr.screen.all.dcast.mmejscore) + geom_point(aes(R1,R2)) + geom_abline()

```


#Example with three genes (POLL, FANCM, POLQ) & I will use "H3K4me1" as example
1st step = Compare POLL, FANCM, POLQ and WT data points vs. H3K4me1 (in lin or log space)

step 1 = plot 

```{r}
# Data table and compute log2MMEJ
log2.ratio.follow.up.detail.all <- chromatin.followup %>% select(bio.rep, NHEJMMEJratio, gene,barcode,well,plate,gRNA) %>% mutate(log2ratio = log2(NHEJMMEJratio))

wt.set.follow.ratio <- filter(log2.ratio.follow.up.detail.all , gene == "tracr") %>% dplyr::group_by(barcode, bio.rep) %>% dplyr::summarise(wt.log2ratio = mean(log2ratio, na.rm = T))

#wt.set.ratio.all <- filter(log2.ratio.screen.detail.all , sample == "WT") %>% dplyr::select(barcode, t.rep,library, wt.log2ratio = log2ratio)

```


# 2nd step check compute distances and plot them
```{r}
# Each replicate has a different value (compute differences by replicate)
log2.distance.follow.up.ratio.all <- log2.ratio.follow.up.detail.all %>% filter(!gene %in% c("tracr","no")) %>% left_join(wt.set.follow.ratio) %>% mutate(log2.ratio = log2ratio - wt.log2ratio)

mean.log2.distance.follow.ratio.all <- log2.distance.follow.up.ratio.all %>% dplyr::group_by(barcode,gene,gRNA) %>% dplyr::summarise(m.log2.ratio.dist = mean(log2.ratio, na.rm = T)) %>% na.omit() %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

#Compute mean log2 fold change
mean.log2.fc.follow <- mean.log2.distance.follow.ratio.all %>% group_by(gene,gRNA) %>% dplyr::summarise(mean.fc = mean(m.log2.ratio.dist, na.rm = T))


```

# Extract slopes for all genes
```{r}
# Perform analysis across all features for three test genes
chromatin.features <- colnames(mean.log2.distance.follow.ratio.all)[7:31]

#
slope.protein.features.all.ratio.followup <- tibble(gene = NA, gRNA = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

for(k in unique(mean.log2.distance.follow.ratio.all$gRNA)) {
for (i in unique(mean.log2.distance.follow.ratio.all$gene)) {
  for (j in colnames(mean.log2.distance.follow.ratio.all)[7:31]) {
    model.dt <- mean.log2.distance.follow.ratio.all %>% filter(gene == i & gRNA == k)
    model.epistasis.log2 <- lm(formula = m.log2.ratio.dist ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
    slope.protein.features.all.ratio.followup <- slope.protein.features.all.ratio.followup %>% add_row(gene = i,gRNA = k, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
  }
}
}

# slope protein features inversee
#slope.protein.features.all.ratio.inverse <- tibble(gene = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

#for (i in unique(mean.log2.distance.ratio.all$gene)) {
#  for (j in colnames(mean.log2.distance.ratio.all)[7:31]) {
#    model.dt <- mean.log2.distance.ratio.all %>% filter(gene == i)
#    model.epistasis.log2 <- lm(formula = m.log2.ratio.inv ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
#    slope.protein.features.all.ratio.inverse <- slope.protein.features.all.ratio.inverse %>% add_row(gene = i, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
#  }
#}

#ratio.all.epistasis <- slope.protein.features.all.ratio %>% select(gene,feature,NM.slope = slope.log2, NM.p.value = p.value,term) %>% left_join(slope.protein.features.all.ratio.inverse %>% select(gene,feature,MN.slope = slope.log2, MN.pvalue = p.value, term))


#Calculate max value of all histone modification
max.values <- (mean.log2.distance.follow.ratio.all)[7:31] %>% distinct() %>% melt() %>% dplyr::group_by(variable) %>% slice_max(value, n= 3) %>% dplyr::group_by(variable) %>% dplyr::summarise(high = mean(value)) %>% select(feature = variable,high)

# Make all negative epistatic interactions 0
epistasis.ratio.sign.followup <- slope.protein.features.all.ratio.followup %>% left_join(mean.log2.fc.follow) %>% filter(complete.cases(.)) %>% dcast(gene+ gRNA + feature + mean.fc ~ term, value.var = "slope.log2") %>% select(gene,gRNA,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])', mean.fc) %>% left_join(max.values) %>% mutate(sign.presence = slope*0.5 + intercept,cross.x = -intercept/slope, epistasis.sign = slope*sign(sign.presence), epistasis.final = case_when(cross.x < 1 ~ slope, cross.x > 1 ~ 0)) 


epistasis.ratio.sign.followup.dcast <- epistasis.ratio.sign.followup %>% dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")
pheatmap(t(epistasis.ratio.sign.dcast))


```

# Test matrix
```{r}
# 
epistasis.ratio.sign.dcast <- epistasis.ratio.sign %>% dcast(gene ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "gene")

```
#Save every step
```{r}
# Step 1
saveRDS(data.for.plotting.all.ratio,file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")

#Step 2
saveRDS(mean.log2.distance.ratio.all, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

# Save all proteins correlation final
saveRDS(epistasis.ratio.sign.followup, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128_ratio_epistasis_followup.rds")

```

