---
title: "xv20211121_epistasis_model_bas"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(reshape2)
library(broom)
library(dendextend)
library(pls)
library(broom)
library(stats)
library(utils)
```

# Import data tables Inhibitors
```{r}
#Chromatin data
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#inhibitor table
inhibitor.table <- tibble(exp = as.character(c(15:21,36:42,57:63)), drug = rep(c("DMSO","DNAPKi","ATMi","PARPi","ATRi_VE821","ATRi_VE822","GFP"),3),replicate = rep(c("R1","R2","R3"), each = 7))

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)
PARPi.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210922_E1664_PARPi_follow_up/rmarkdown/xv20210922_E1664_indel_script_PARPi.rds")

```

#Data processing (E1627: DNA repair inhibitors)
```{r}
# Process kinase inhibitor data
inhibitor.data.processed <- inhibitor.data %>% 
                                  mutate(MMEJscore = pct_del_7 / (pct_del_7 + pct_ins_1),
                                                  NHEJscore = pct_ins_1 / (pct_del_7 + pct_ins_1),
                                                  freqMMEJ = pct_del_7,
                                                  freqNHEJ = pct_ins_1,
                                                  freqCut = 1 - pct_wt,
                                                  NHEJMMEJratio = pct_ins_1 / pct_del_7
                                                  ) %>%
                                select(-grep("pct", colnames(.)), -grep("ins",colnames(.)),-grep("del",colnames(.)), -c("wt","Inf")) %>% distinct()


sel.inhibitor.data <- inhibitor.data.processed %>% select(replicate, drug, barcode, MMEJscore,NHEJMMEJratio) %>% mutate(exp = "PIKK")
```


# Data processing (E1664: PARP inhibitors)
```{r}
# Process inhibitors data
# Remove unnecesary columns and add gene info - Filter out wells with less that 30 mutation reads
# Create a dt with all the descriptive variables
sample.table.PARPi <- read.table("/home/x.vergara/XV_P3_ChromDSBScreen/xv20210922_E1664_PARPi_follow_up/data/xv20210922_E1664_sample_file.txt", header = T)

PARPi.data.processed <- PARPi.data %>% mutate(MMEJscore = pct_del_7 / (pct_del_7 + pct_ins_1),
                                                  NHEJscore = pct_ins_1 / (pct_del_7 + pct_ins_1),
                                                  freqMMEJ = pct_del_7,
                                                  freqNHEJ = pct_ins_1,
                                                  freqCut = 1 - pct_wt,
                                                  NHEJMMEJratio = pct_ins_1 / pct_del_7
                                                  ) %>%
                                select(-grep("pct", colnames(.)), -grep("ins",colnames(.)),-grep("del",colnames(.)), -c("wt","Inf","ssODN")) %>% distinct()%>% left_join(sample.table.PARPi)
# Without vorinostat data
PARPi.data.processed.DMSO <- filter(PARPi.data.processed, is.na(Vorinostat)) %>% select(replicate = Replicate, drug = Sample, barcode, MMEJscore,NHEJMMEJratio) %>% mutate(exp = "E1664")
```

#Bind data from both experiments
```{r}
#Combine with log2.inhibitor 
combined.inhibitor.data <- bind_rows(PARPi.data.processed.DMSO,sel.inhibitor.data) %>% mutate(drug = case_when(drug == "PARPi" ~ "olaparib", T ~ drug))
```

#This is a analysis pipeline to study chromatin context dependencies (CCD)
#This is a proposal to get to a final version where we can infer statistical significance from the data

#I am currently working on these steps:
  1- log2 MMEJ:NHEJ ratio calculation
  2- Filter based on z-score of this data: Filter genes that have 1 significant IPR (z-score > 1.95) in at least 2 replicates.
  3- Compute log2 MMEJ:NHEJ fold changes based on WT values: Compare with WT samples (This shows that data is reproducible and R are better for KO than WT)
  4- Compute CCDs with WT wells (random CCDs) and KO samples
    A- Table with mean CCDs
    B- Table with all CCDs per replicate
  5- Test for non-random CCDs (KS test, I could even use the WT from the other screen, it won't change much). These are the interactions that matter and they correlate decently between replicates
    A- non-random genes based on mean CCDs
    B- non-random genes based on all CCDs per replicate
  

step 1 = plot 
```{r}
# Data table and compute log2MMEJ
log2.inhibitors <- combined.inhibitor.data  %>% select(replicate, NHEJMMEJratio, drug,barcode, exp) %>% mutate(log2ratio = log2(1/NHEJMMEJratio))

# Plot differences
data.for.plotting.inhibitors <- log2.inhibitors %>% dplyr::group_by(drug,barcode,exp) %>% dplyr::summarise(m.log2ratio = mean(log2ratio, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))
```


# 2nd step check compute distances and plot them
```{r}
#Calculate DMSO condition
DMSO.ratio <- log2.inhibitors %>% filter(drug == "DMSO") %>% ungroup() %>%  select(barcode,DMSO.log2ratio = log2ratio, exp, replicate) 

# Each replicate has a different value (compute differences by replicate)
log2.distance.ratio.inhibitors <- log2.inhibitors %>% left_join(DMSO.ratio, by = c("barcode","exp","replicate")) %>% dplyr::group_by(drug, barcode,exp, replicate) %>% mutate(log2.foldchange = log2ratio - DMSO.log2ratio) 

mean.log2.distance.ratio.inhibitors <- log2.distance.ratio.inhibitors %>% dplyr::group_by(barcode,drug) %>% dplyr::summarise(mean.log2.foldchange = mean(log2.foldchange, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID")) %>% filter(drug != "GFP")
```

# Do I have enough power to test single perturbations
```{r}
log2.distance.ratio.inhibitors.noGFP <- log2.distance.ratio.inhibitors %>% filter(drug != "GFP")

inh.wilc.test <- tibble(drug = NA, barcode = NA, p.value = NA, mean = NA)

# Test if we have enough power to test wilcox test per IPR
for( i in unique(log2.distance.ratio.inhibitors.noGFP$drug)) {
  for(j in unique(log2.distance.ratio.inhibitors.noGFP$barcode)) {
    IPR.drug <- log2.distance.ratio.inhibitors.noGFP %>% filter(drug == i & barcode == j) %>% pull(log2.foldchange)
   if (length(IPR.drug) < 3) {
next
}
    IPR.DMSO <- log2.distance.ratio.inhibitors.noGFP %>% filter(drug == "DMSO" & barcode == j) %>% pull(log2.foldchange)
    t_test <- t.test(IPR.drug,IPR.DMSO) %>% tidy()
    inh.wilc.test <- inh.wilc.test %>% add_row(drug = i, barcode = j, p.value = t_test %>% pull(p.value), mean = mean(IPR.drug, na.rm = T))
    }
  }


significant.inh.wilc.test <- inh.wilc.test %>% filter(drug != "DMSO") %>% mutate(p.adj = p.adjust(p.value, method = "BH")) %>% select(drug,barcode,p.value,p.adj, mean)

#what experiment have at least one IPR significant
significant.perturbations.IPR <- significant.inh.wilc.test %>% dplyr::group_by(drug) %>% dplyr::summarise(MMEJ_signif = sum(mean < 0 & p.adj < 0.05), NHEJ_signif = sum(mean > 0 & p.adj < 0.05)) %>% mutate(pathway_signif = case_when(MMEJ_signif == 0 & NHEJ_signif != 0 ~ "NHEJ", MMEJ_signif != 0 & NHEJ_signif == 0 ~ "MMEJ", MMEJ_signif != 0 & NHEJ_signif != 0 ~ "both", T ~ "none"))

```

# Run the model: For this I need average values => I'll include r.squared and adj.r.squared as well
```{r}
#DDR library
inhibitors.pvalue_c <- tibble(library = NA, drug = NA, num_comp = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)
for (i in unique(mean.log2.distance.ratio.inhibitors$drug)){
  for (j in 1:10){
gene.library.dt <- filter(mean.log2.distance.ratio.inhibitors, drug == i)
# Run a model per gene
set.seed(1)
PCR_model_DDR_test <- pcr(mean.log2.foldchange~CTCF+EZH2+H2AFZ+H3K27ac+H3K27me3+H3K36me3+H3K4me1+H3K4me2+H3K4me3+H3K79me2+H3K9me2+H3K9me3+H4K5acK8ac+HDAC1+HDAC2+HDAC3+POL2+POL2AS2+SMC3+LMNB1+Dam+DNAse+m5C+late_replicating+TTseq, data=gene.library.dt , validation="CV")

pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = j)
combined.dt <- tibble(measured = gene.library.dt$mean.log2.foldchange, predicted = as.numeric(pcr_pred))
pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% glance()
inhibitors.pvalue_c <- inhibitors.pvalue_c %>% add_row(drug = i, num_comp = j, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value))
}
}

#significant genes
significant.inhibitors_p.value <- inhibitors.pvalue_c %>% left_join(significant.perturbations.IPR) %>% filter(pathway_signif != "none") %>% filter(num_comp == 3) %>% mutate(p.adj = p.adjust(p.value, method = "BH")) %>% select(drug,p.value,p.adj)

```

# Calculate CCDs for drugs: 
# A) Calculate this after computing the mean
```{r}
# Compute slopes for all experiments
repair.inhibitor.CCD <- tibble(drug = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

for(h in unique(mean.log2.distance.ratio.inhibitors$drug)){
  for (j in colnames(mean.log2.distance.ratio.inhibitors)[6:30]) {
    model.dt <- mean.log2.distance.ratio.inhibitors %>% filter(drug == h)
       if (nrow(model.dt) == 0) {
next
}
    model.epistasis.log2 <- lm(formula = mean.log2.foldchange ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
   repair.inhibitor.CCD <- repair.inhibitor.CCD %>% add_row(drug = h, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
  }
}

#Calculate KO proteins (terms)
repair.inhibitor.CCD.corrected <- repair.inhibitor.CCD %>% 
  reshape2::dcast(drug + feature ~ term, value.var = "slope.log2") %>% 
  dplyr::select(drug,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% 
  left_join(significant.inhibitors_p.value %>% select(drug, model_p.value = p.value, model_p.adj = p.adj)) %>% 
  left_join(significant.perturbations.IPR %>% select(drug, pathway_signif) %>% 
              distinct()) %>% mutate(cross.x = -intercept/slope, slope_cor = case_when(cross.x < 1 ~ slope, cross.x > 1 ~ 0)) %>% 
  mutate(CCD_value = case_when(pathway_signif == "MMEJ" & slope_cor < 0 ~ slope_cor, pathway_signif == "MMEJ" & slope_cor > 0 ~ 0, pathway_signif == "NHEJ" & slope > 0 ~ slope_cor, pathway_signif == "NHEJ" & slope < 0 ~ 0, pathway_signif == "none" ~ 0, T ~ slope_cor))
```

#Save every step
```{r}
#Export CCD of significant genes based on model
saveRDS(repair.inhibitor.CCD.corrected, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_DDR_CCD_mean_significant.rds")
```

#Output files
```{r}
setwd(in.dir)
#Save log2 ratio values
saveRDS_proof(log2.inhibitors, file = "data/processed_data/CCD_analysis/xv20220624_log2_MMEJ_NHEJ_ratios_inhibitors.rds")

#Save all log2 fold change values
saveRDS_proof(mean.log2.distance.ratio.inhibitors, file = "data/processed_data/CCD_analysis/xv20220624_fold_change_ratio_inhibitors.rds") #Will be used for Figure 1B, 2B and 2C

#Save summary files
saveRDS_proof(repair.inhibitor.CCD.corrected, file = "data/processed_data/xv20220624_DR_inhibitors_CCD_summary.rds") #This is the data frame that will be mainly used to generate all the figures

```


