---
title: "xv20220814_SCC_mutations_analysis"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}

```

## Libraries:

```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
library(vcfR)
library(GenomicRanges)
```

## Import data

```{r pressure, echo=FALSE}
setwd(in.dir)

#Conversion data table
sample_conversion <- read_xlsx("import/xv20220727_PCAWG_tumor_data/xv20220815_conversion_table_samples_TCGA.xlsx", col_names = T)

#Import sporadic mutations
sporadic_mutation_dt <- read_xlsx("import/Webster_et_al_2021_FA_tumours/xv20220329_T5_FA_SCC_SV.xlsx", col_names = T, sheet = 2, skip = 1)
#New analysis
sporadic_mutation_dt_bis <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/HNSC/BRASS/summary_shortened.txt")

#Import sporadic mutations
BRCA2_mutation_dt <- read_xlsx("import/Webster_et_al_2021_FA_tumours/xv20220329_T5_FA_SCC_SV.xlsx", col_names = T, sheet = 3, skip = 1)
#New analysis
BRCA2_mutation_dt_bis <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/BRCA2/BRASS/summary_shortened.txt")



```

#Import LADs data
```{r}
setwd(in.dir)
# LAD domains
LAD_atlas <- as.data.frame(read.table("import/xv20220329_LAD_atlas_OSF.bed.gz",header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(LAD_atlas) <- c("chr","start","end","length","strand","LAD_class")

#LAD_length plot control
LAD_length_tibble <- LAD_atlas %>% mutate(LAD_length = end - start)

```

#Curate data tables: What are the features that I am interested (only select deletions)
```{r}
#Select only deletions (in all four files)
sporadic_mutation_del_dt <- sporadic_mutation_dt %>% filter(svclass == "deletion" & !`micro-homology` %in% c(".","_") & `Brass Notation` != "_")
sporadic_mutation_del_dt_bis <- sporadic_mutation_dt_bis %>% separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",") %>% left_join(sample_conversion, by = c('sample' ='TCGA ID')) %>% filter(svclass == "deletion" & !"micro-homology" %in% c(".","_") & `Brass Notation` != "_")

#Filter out data with FANC anemia mutations
sporadic_mutation_del_filtered <- sporadic_mutation_del_dt_bis %>% filter(!sample %in% c("de8ef60b-4dbe-4aa8-adef-6f58cdfada29","e52ffa79-557a-4024-81f3-f3826c227ec5"))

#Check by inner join
common_part_sporadic <- inner_join(sporadic_mutation_del_dt,sporadic_mutation_del_filtered, by = c("start1","end1","chr2","start2","end2"))
#Numeric brass score
sporadic_brass_score <- sporadic_mutation_del_dt$`Brass Notation` %>% str_extract("(?<=score ).*(?=\\))") %>% as.numeric()
sporadic_brass_score_bis <- sporadic_mutation_del_dt_bis$`Brass Notation` %>% str_extract("(?<=score ).*(?=\\))") %>% as.numeric()

#BRCA2 data ()
BRCA2_mutation_del_dt <- BRCA2_mutation_dt %>% filter(svclass == "deletion"  & !"micro-homology" %in% c(".","_") & `Brass Notation` != "_")
BRCA2_mutation_del_dt_bis <- BRCA2_mutation_dt_bis  %>% separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",") %>% filter(svclass == "deletion"  & !"micro-homology" %in% c(".","_") & `Brass Notation` != "_")
#Check by inner join
common_part_BRCA2 <- inner_join(BRCA2_mutation_del_dt,BRCA2_mutation_del_dt_bis, by = c("start1","end1","chr2","start2","end2"))
#numeric brass score
BRCA2_brass_score <- BRCA2_mutation_del_dt$`Brass Notation` %>% str_extract("(?<=score ).*(?=\\))") %>% as.numeric()
BRCA2_brass_score_bis <- BRCA2_mutation_del_dt_bis$`Brass Notation` %>% str_extract("(?<=score ).*(?=\\))") %>% as.numeric()
```

#Map SV for BRCA2 and sporadic tumors
```{r}
#Create function to call unique LADs
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T)
unique_LADs_call <- function(x) {
  mutations_GR <- makeGRangesFromDataFrame(x, keep.extra.columns = T)
  LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutations_GR)
  LAD_overlaps_dt <- tibble(x[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6])
  LAD_counts <- LAD_overlaps_dt %>% dplyr::group_by(sample, id) %>% dplyr::summarize(count = n())
  single_LADs_iLADs <-LAD_counts %>% filter(count == 1) %>% select(sample, id)
  LAD_overlaps_dt %>% right_join(single_LADs_iLADs)
}

#Prepare the data table for mapping
mutation_location_export_spo_SV <- sporadic_mutation_del_dt %>% mutate(mut_id = paste(sample, `id/name`, sep = "_"), chr = paste0("chr", `# chr1`)) %>% select(chr, start = "start1", end = "start2", mut_id, sample, id = `id/name`)
mutation_location_export_spo_SV_bis <- sporadic_mutation_del_filtered %>% mutate(mut_id = paste(sample, `id/name`, sep = "_"), chr = paste0("chr", `chr1`)) %>% select(chr, start = "start1", end = "start2", mut_id, sample = `Name Supp. Table`, id = `id/name`)
mutation_location_export_BRCA2_SV <- BRCA2_mutation_del_dt %>% mutate(mut_id = paste(sample, `id/name`, sep = "_"), chr = paste0("chr", `# chr1`)) %>% select(chr, start = "start1", end = "start2", mut_id, sample, id = `id/name`)
mutation_location_export_BRCA2_SV_bis <- BRCA2_mutation_del_dt_bis %>% mutate(mut_id = paste(sample, `id/name`, sep = "_"), chr = paste0("chr", `chr1`)) %>% select(chr, start = "start1", end = "start2", mut_id, sample, id = `id/name`)

#Run function
spo_SV_LAD_maps <- unique_LADs_call(mutation_location_export_spo_SV) #526 total MH del
spo_SV_LAD_maps_bis <- unique_LADs_call(mutation_location_export_spo_SV_bis) #899 total MH del
BRCA2_SV_LAD_maps <- unique_LADs_call(mutation_location_export_BRCA2_SV) #2022 total MH del
BRCA2_SV_LAD_maps_bis <- unique_LADs_call(mutation_location_export_BRCA2_SV_bis) #2281 total MH del

#Filter for HPVneg and constitutive LADs or iLADs
spo_HPVneg_c_maps <- spo_SV_LAD_maps %>% filter(grepl("c", LAD_status) & grepl("Neg", sample))  %>% mutate(dataset = "manuscript", type = "sporadic") #200 total 
spo_HPVneg_c_maps_bis <- spo_SV_LAD_maps_bis %>% filter(grepl("c", LAD_status) & grepl("Neg", sample)) %>% mutate(dataset = "update", type = "sporadic") #336 total
BRCA2_SV_c_maps <- BRCA2_SV_LAD_maps %>% filter(grepl("c", LAD_status)) %>% mutate(dataset = "manuscript", type = "BRCA2mut")#997 total
BRCA2_SV_c_maps_bis <- BRCA2_SV_LAD_maps_bis %>% filter(grepl("c", LAD_status)) %>% mutate(dataset = "update", type = "BRCA2mut")#1116 total

#Calculate mean per sample
spo_HPVneg_SV_cLAD_maps_tumor <- spo_HPVneg_c_maps %>% dplyr::group_by(sample, LAD_status) %>% dplyr::summarise(mut_n = n())
spo_HPVneg_SV_cLAD_maps_tumor_bis <- spo_HPVneg_c_maps_bis %>% dplyr::group_by(sample, LAD_status) %>% dplyr::summarise(mut_n = n()) 
BRCA2_SV_cLAD_maps_tumor <- BRCA2_SV_c_maps %>% dplyr::group_by(sample, LAD_status) %>% dplyr::summarise(mut_n = n()) 
BRCA2_SV_cLAD_maps_tumor_bis <- BRCA2_SV_c_maps_bis %>% dplyr::group_by(sample, LAD_status) %>% dplyr::summarise(mut_n = n()) 

#dcast these tibbles
spo_HPVneg_SV_cLAD_maps_tumor_dcast <- spo_HPVneg_SV_cLAD_maps_tumor %>% reshape2::dcast(sample ~ LAD_status, value.var = "mut_n", fill = 0) #mean ciLAD 2.83 & cLAD 5.87
spo_HPVneg_SV_cLAD_maps_tumor_bis_dcast <- spo_HPVneg_SV_cLAD_maps_tumor_bis %>% reshape2::dcast(sample ~ LAD_status, value.var = "mut_n", fill = 0) #mean ciLAD 4.13 & cLAD 10.48 (median values 4 vs. 9)
BRCA2_SV_cLAD_maps_tumor_dcast <- BRCA2_SV_cLAD_maps_tumor %>% reshape2::dcast(sample ~ LAD_status, value.var = "mut_n", fill = 0) #mean ciLAD 13.37 & cLAD 11.55
BRCA2_SV_cLAD_maps_tumor_bis_dcast <- BRCA2_SV_cLAD_maps_tumor_bis %>% reshape2::dcast(sample ~ LAD_status, value.var = "mut_n", fill = 0) #mean ciLAD 14.6 & cLAD 13.3 (median 12 vs. 10.5)

```

#Summary table with all the number in here
```{r}
#Bind all data
mutation_BRCA2_dataset <- bind_rows(spo_HPVneg_c_maps,spo_HPVneg_c_maps_bis,BRCA2_SV_c_maps,BRCA2_SV_c_maps_bis) %>% mutate(del_length = end - start)
sample_n <- mutation_BRCA2_dataset %>% select(dataset, type, sample) %>% distinct() %>% dplyr::group_by(dataset,type) %>% dplyr::summarise(samp_n = n())

length_summary <- mutation_BRCA2_dataset %>% dplyr::group_by(dataset,type,LAD_status) %>% dplyr::summarise(perc_5_length = quantile(del_length, c(0.05)), perc_95_lenght = quantile(del_length, c(0.95)))
mut_summary <-  mutation_BRCA2_dataset %>% dplyr::group_by(sample, LAD_status,dataset, type) %>% dplyr::summarise(mut_n = n()) %>% ungroup() %>% dplyr::group_by(dataset, type,LAD_status) %>% dplyr::summarise(sum_mut = sum(mut_n)) %>% left_join(sample_n) %>% mutate(mean_mut = sum_mut/samp_n) %>% dplyr::select(LAD_status,dataset,type,mean_mut)
# table summarizing mean for each run
summary_mutation_BRCA2_mut <- length_summary %>% left_join(mut_summary)

print(summary_mutation_BRCA2_mut)

```

#Conclusion: The main message does not change with the new data. Deletion size it's a bit smaller with the new dataset and the numbers higher (it detects more mutations than before).

