---
title: "xv20211121_epistasis_model_bas"
output: html_document
---

# Script to generate figure #5 in the paper. This figure describes the layout of the screen and technical issues. Figure outline:
A- Pathway balance perturbation inhibitors
B- Heatmap matrix
C- Bigger epistatic interactions than KO
D- Correlation with olaparib
E- Correlation with ATMi

Data in text:
A-...
B-...

Supplementary figure 6:
A- Veliparib vs. PARP1
B- Talazoparib vs. PARP1
C- VE821 vs. ATR
D- VE822 vs. ATR

This script generates the plots for figure 4 and S5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(rstatix)
```


# Import data tables Inhibitors
```{r libraries}
#Chromatin data
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#inhibitor table
inhibitor.table <- tibble(exp = as.character(c(15:21,36:42,57:63)), drug = rep(c("DMSO","DNAPKi","ATMi","PARPi","ATRi_VE821","ATRi_VE822","GFP"),3),replicate = rep(c("R1","R2","R3"), each = 7))

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)
PARPi.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210922_E1664_PARPi_follow_up/rmarkdown/xv20210922_E1664_indel_script_PARPi.rds")

```

# Data processing
```{r}
# Process inhibitors data
# Remove unnecesary columns and add gene info - Filter out wells with less that 30 mutation reads
# Create a dt with all the descriptive variables
sample.table.PARPi <- read.table("/home/x.vergara/XV_P3_ChromDSBScreen/xv20210922_E1664_PARPi_follow_up/data/xv20210922_E1664_sample_file.txt", header = T)

PARPi.data.processed <- PARPi.data %>% mutate(MMEJscore = pct_del_7 / (pct_del_7 + pct_ins_1),
                                                  NHEJscore = pct_ins_1 / (pct_del_7 + pct_ins_1),
                                                  freqMMEJ = pct_del_7,
                                                  freqNHEJ = pct_ins_1,
                                                  freqCut = 1 - pct_wt,
                                                  NHEJMMEJratio = pct_ins_1 / pct_del_7
                                                  ) %>%
                                select(-grep("pct", colnames(.)), -grep("ins",colnames(.)),-grep("del",colnames(.)), -c("wt","Inf","ssODN")) %>% distinct()%>% left_join(sample.table.PARPi)
# Without vorinostat data
PARPi.data.processed.DMSO <- filter(PARPi.data.processed, is.na(Vorinostat)) %>% select(replicate = Replicate, drug = Sample, barcode, MMEJscore,NHEJMMEJratio) %>% mutate(exp = "E1664")

# Process kinase inhibitor data
inhibitor.data.processed <- inhibitor.data %>% 
                                  mutate(MMEJscore = pct_del_7 / (pct_del_7 + pct_ins_1),
                                                  NHEJscore = pct_ins_1 / (pct_del_7 + pct_ins_1),
                                                  freqMMEJ = pct_del_7,
                                                  freqNHEJ = pct_ins_1,
                                                  freqCut = 1 - pct_wt,
                                                  NHEJMMEJratio = pct_ins_1 / pct_del_7
                                                  ) %>%
                                select(-grep("pct", colnames(.)), -grep("ins",colnames(.)),-grep("del",colnames(.)), -c("wt","Inf")) %>% distinct()


sel.inhibitor.data <- inhibitor.data.processed %>% select(replicate, drug, barcode, MMEJscore,NHEJMMEJratio) %>% mutate(exp = "PIKK")
#Combine with log2.inhibitor 
combined.inhibitor.data <- bind_rows(PARPi.data.processed.DMSO,sel.inhibitor.data) %>% mutate(drug = case_when(drug == "PARPi" ~ "olaparib", T ~ drug))
```

step 1 = plot 

```{r}
# Data table and compute log2MMEJ
log2.inhibitors <- combined.inhibitor.data  %>% select(replicate, NHEJMMEJratio, drug,barcode, exp) %>% mutate(log2ratio = log2(NHEJMMEJratio))

# Plot differences
data.for.plotting.inhibitors <- log2.inhibitors %>% dplyr::group_by(drug,barcode,exp) %>% dplyr::summarise(m.log2ratio = mean(log2ratio, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

# Plot in log2 space
ggplot(data.for.plotting.inhibitors %>% filter(drug != "GFP"), aes(H3K4me1,m.log2ratio, color = drug)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "spearman") +
  theme_bw() +
  facet_wrap(~ exp)

```

# 2nd step check compute distances and plot them
```{r}
#Calculate DMSO condition
DMSO.ratio <- data.for.plotting.inhibitors %>% filter(drug == "DMSO") %>% ungroup() %>%  select(barcode,DMSO.log2ratio = m.log2ratio, exp) 

# Each replicate has a different value (compute differences by replicate)
log2.distance.ratio.inhibitors <- data.for.plotting.inhibitors  %>% left_join(DMSO.ratio, by = c("barcode","exp")) %>% dplyr::group_by(drug, barcode,exp) %>% mutate(log2.ratio = m.log2ratio - DMSO.log2ratio) 
mean.log2.distance.ratio.inhibitors <- log2.distance.ratio.inhibitors %>% dplyr::group_by(barcode,drug,exp) %>% dplyr::summarise(m.log2.ratio = mean(log2.ratio, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID")) %>% filter(drug != "GFP")

#Plot slopes
ggplot(mean.log2.distance.ratio.inhibitors %>% 
         filter(drug != "GFP"),
       aes(H3K4me1,m.log2.ratio, color = drug)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw() +
  facet_wrap(~exp)

 #Compute mean log2 fold change
mean.log2.fc.drug <- mean.log2.distance.ratio.inhibitors %>% group_by(drug) %>% dplyr::summarise(mean.fc = mean(m.log2.ratio, na.rm = T))
```

# Extract slopes for all genes
```{r}
# Perform analysis across all features for three test genes
chromatin.features <- colnames(mean.log2.distance.ratio.inhibitors)[7:31]

slope.protein.features.inhibitors.ratio <- tibble(drug = NA, feature = NA, slope.log2 = NA, term = NA, exp = NA, p.value = NA)

for (i in unique(mean.log2.distance.ratio.inhibitors$drug)) {
  for (j in colnames(mean.log2.distance.ratio.inhibitors)[7:31]) {
    for (k in unique(mean.log2.distance.ratio.inhibitors$exp)) {
    model.dt <- mean.log2.distance.ratio.inhibitors %>% filter(drug == i & exp == k)
    if (nrow(model.dt) < 1){
      next
    }
    model.epistasis.log2 <- lm(formula = m.log2.ratio ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
    slope.protein.features.inhibitors.ratio <- slope.protein.features.inhibitors.ratio %>% add_row(drug = i, feature = j, exp = k, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term),p.value = model.epistasis.log2 %>% pull(p.value))
    }
  }
}

#Calculate max value of all histone modification
max.values.inh <- (mean.log2.distance.ratio.inhibitors)[7:31] %>% distinct() %>% melt() %>% dplyr::group_by(variable) %>% slice_max(value, n= 3) %>% dplyr::group_by(variable) %>% dplyr::summarise(high = mean(value)) %>% select(feature = variable,high)

# Make all negative epistatic interactions 0
epistasis.ratio.sign.inhibitors <- slope.protein.features.inhibitors.ratio %>% filter(complete.cases(.)) %>% dcast(drug + feature ~ term, value.var = "slope.log2",fun.aggregate = mean) %>% select(drug,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% left_join(max.values) %>% mutate(sign.presence = slope*0.5 + intercept,cross.x = -intercept/slope, epistasis.sign = slope*sign(sign.presence), epistasis.final = case_when(cross.x < 1 ~ slope, cross.x > 1 ~ 0)) 

```

#Export data for figures and plotting
```{r}
#Export inhibitor perturbation data
saveRDS(combined.inhibitor.data, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds")

#Export epistasis interactions
saveRDS(epistasis.ratio.sign.inhibitors, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128_ratio_kinase_inhib_epistasis.rds")

```