---
title: "xv20220623_CCD_preprocessing"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go over different calculations of CCDs with the validation RPE1 DSB TRIP cell line:
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  log_path <- paste0(in.dir,"log")
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    #print(mode(log_path))
    #base::write(c("test", as.character(sys_time),"Output", object), file = log_path, ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```

# Libraries
```{r libraries, warning = FALSE}
# libraries:
library(tidyverse)
library(reshape2)
library(broom)
library(dendextend)
library(pls)
library(broom)
library(stats)
library(MASS)
library(rstatix)
library(parallel)
library(data.table)
library(readxl)
```


# Import raw indel counts from snakemake pipeline from C. Leemans
```{r import}

# Set directory to the mutation output folder of the CRISPR-TRIP snakemake script
setwd("/DATA/projects/DSBrepair/data/xv20230424_E2222_RPE1_kinase_inh_b/indelPCR_counts/")
# Import files in list and make individual tables
# I use this if all the samples are good. Here however I do not use all the samples. 
file.list <- list.files("/DATA/projects/DSBrepair/data/xv20230714_E2286_pool_pathway_assay/indelPCR_counts",pattern = '*[.]co', full.names = T) 

#Filter only samples with data for the experiment
file.list.filtered <- file.list[c(grep("RPEPro_ATM", file.list),grep("RPEPro_LBR2",file.list),grep("RPEDef_ATM", file.list),grep("RPEDef_LBR2",file.list))]

# import the data
df.list.indel <- mclapply(file.list.filtered, read.table, col.names=c('barcode', 'call', 'indel', 'count'), mc.cores = 20, stringsAsFactors = FALSE, header=T)

# rename the lists
names(df.list.indel) <- str_extract(file.list.filtered,"RPE.*(?=.co)")

# count the sample number
n.samples.indel <- length(df.list.indel)

#Import other data
RPE1_data_chrom_tib <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen","chromatin_data_RPE1") %>%
  filter(cell_line %in% c("DEF","PRO") & binsize == 2000  & barcode != "CTTGCATTAACAACCA") #Chromatin
RPE1_data_chrom_tib$late_replicating <- as.numeric(RPE1_data_chrom_tib$late_replicating)
dna_quantification <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen","dna_quantification") #DNA viability
proteins_gRNA <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/validation_screen","gRNA_gene") #gRNA gene conversion
RPE1_genomewide_dynamic_range <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/revisions/pools_balance", "RPE1_genomewide_ranges")
#log2 chromatin
PRO_log2_chrom <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_validation_RPE1/","PRO_log2")
DEF_log2_chrom <- readRDS_proof("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/CCD_validation_RPE1/","DEF_log2")

```

### Process indel calling and filter for real data
Set everything in a dataframe that contains barcodes, indel ratios, and efficiencies.
```{r indeldataframe}
# Generate a datatable with the number of indelreads per mutations, sample and barcode (For this script I will modify this function, to sum replicates with the same ID)
mut.list = mclapply(names(df.list.indel), function(exp){
    dt = data.table(df.list.indel[[exp]])
    dt[, indel:=as.character(indel)]
    dt[call=='wt' & indel=='2', indel:='ssODN']
    sum_count = data.table(exp=exp,
                           dt[, list(count=sum(count)),by=c('barcode', 'indel')])
    count_cast = data.table::dcast(sum_count[!is.na(indel),], exp + barcode ~ indel,
                      value.var='count')
    return(count_cast)
}, mc.cores=10)

#Bind all data frames (one per sample) together
indels.dt = do.call(rbind, c(mut.list, fill=T))
indels.dt[is.na(indels.dt)] = 0

#Filter only mapped barcodes
indels.dt <- filter(indels.dt, barcode %in% RPE1_data_chrom_tib$barcode)

#Change colnames in indels.dt
indel_cols <- names(indels.dt)[grep("[0-9]|ssODN|Inf", names(indels.dt))]
indel_cols <- gsub("-", "del_", indel_cols)
indel_cols[grep("^[0-9].*", indel_cols)] <- gsub("^", "ins_", indel_cols[grep("^[0-9].*", indel_cols)])
indel_cols[grep("ins_0", indel_cols)] <- "wt"
names(indels.dt) <- c("exp", "barcode", indel_cols)

# Generate indel frequency data table
indels.frequencies.dt = data.frame(indels.dt[,c('exp', 'barcode')],
                                  indels.dt[, -c(1:2)] /
                                    rowSums(indels.dt[, -c(1:2)]))

# Rename indels.frequencies.dt
indel_pct_cols <- gsub("^", "pct_", indel_cols)
names(indels.frequencies.dt)[3:length(names(indels.frequencies.dt))] <- indel_pct_cols

# Dimensions check
dim(indels.dt)
dim(indels.frequencies.dt)

# Save all three data frames as a single tibble
indels.tib <- as_tibble(indels.dt)
indels.frequencies.tib <- as_tibble(indels.frequencies.dt)


# Join indels & descr.variables
indels.tib$exp <- gsub("indel_","", indels.tib$exp)
indels.frequencies.tib$exp <- gsub("indel_","", indels.frequencies.tib$exp)

screen.tib <- left_join(indels.tib,indels.frequencies.tib, by = c("exp","barcode")) %>%
    separate(exp, into = c("cell_line","inhibitor","replicate","tech_replicate"), remove = F)

grouped_screen.tib <- screen.tib %>% 
  dplyr::group_by(cell_line, inhibitor, replicate,barcode) %>%
  dplyr::summarise(across(colnames(screen.tib)[7:138], sum))



dim(grouped_screen.tib)
```


#Plot control (supplementary figures)
```{r}
#Data table for control plots
control_plots_dt <- grouped_screen.tib %>% 
  ungroup() %>%
  mutate(proccessed_read = rowSums(.[,c(-1:-4)]), indel_reads = rowSums(.[,c("ins_1","del_7")])) %>% 
  dplyr::select(cell_line,inhibitor,replicate, barcode, proccessed_read, indel_reads, del_7, ins_1, wt)

#Data table for control plots
reads_sample <- control_plots_dt %>% 
  dplyr::group_by(cell_line,inhibitor,replicate) %>%
  dplyr::summarise(total_reads_sample = sum(proccessed_read))

#Left join number of total reads
summary_table_control_plots <- control_plots_dt %>% 
  left_join(reads_sample) %>%
  mutate(IPR_frequency = proccessed_read/total_reads_sample) %>%
  dplyr::group_by(cell_line, inhibitor, barcode) %>%
  mutate(mean_IPR_freq = mean(IPR_frequency))
```

#Plot barcode distribution
```{r, fig.width=2.5,fig.height=3}
#Plot IPR_frequency across replicates
ggplot(summary_table_control_plots,aes(replicate, IPR_frequency)) + 
  geom_line(aes(group = barcode)) +
  geom_point() +
  theme_bw() + 
  facet_wrap(~ cell_line + inhibitor) +
  theme(axis.title.x = element_blank())
```

```{r, fig.width=4, fig.height=3}
#Plot IPR_frequency vs. balance
ggplot() + 
  geom_point(data = summary_table_control_plots %>% filter(replicate != "R4"),aes(IPR_frequency, log2((del_7) / (ins_1)), color = IPR_frequency > 0.005)) +
  theme_bw() +
  theme(axis.title.x = element_blank()) + 
  ylab("log2MMEJ:NHEJ") +
  theme(legend.position = "none") +
  scale_color_manual(values = c("grey20","#9f86bf")) +
  facet_wrap(~ cell_line + inhibitor)
```
#Filter for IPRs for further analysis
```{r}
#Check when I start getting reproducible data?
filtered_screen_data <- grouped_screen.tib %>%
  left_join(summary_table_control_plots) %>%
  filter(replicate != "R4" & ins_1 > 40 & del_7 > 40) %>%
  mutate(cell_line = case_when(cell_line == "RPEPro" ~ "PRO",
                               cell_line == "RPEDef" ~ "DEF"))

#Summary IPR numbers
summary_IPR <- filtered_screen_data %>%
  dplyr::group_by(cell_line,inhibitor, barcode) %>%
  dplyr::summarise(count = n())

```


#Step 3: Calculation of changes in MMEJ:NHEJ balance and annotations

### Repair metric calculations (Formulas #1 & #2), filtering and annotation of samples
```{r}
# Remove unnecesary columns and add gene info - Filter out wells with less that 30 mutation reads
screen.data.ddr <- filtered_screen_data %>% 
  left_join(summary_table_control_plots) %>%
  mutate(log2MMEJNHEJratio = log2((del_7) /( ins_1))) 


#Select variables that will be used downstream
screen.data.tib <- screen.data.ddr %>%
  dplyr::select(-grep("pct", colnames(.)), -grep("ins",colnames(.)),-grep("del",colnames(.)), -c("wt","Inf","ssODN")) %>% distinct()



#Add DR library info
screen.data.names.tib <- screen.data.tib %>%
  filter(is.finite(log2MMEJNHEJratio))

#Summary numbers
samples_per_replicate <- screen.data.names.tib %>% ungroup() %>% dplyr::select(inhibitor,replicate) %>% distinct() %>% dplyr::group_by(replicate) %>% dplyr::summarise(counts = n(), percentage = round(((n()/4)*100), digits = 2))
IPR_per_sample <- screen.data.names.tib %>% dplyr::select(barcode,inhibitor,replicate) %>% distinct() %>% dplyr::group_by(inhibitor,replicate) %>% dplyr::summarise(counts = n()) %>% pull(counts) %>% median() %>% round(digits = 2)

#Print
print(samples_per_replicate)
print(IPR_per_sample)

```



### Calculation of differentials of log2 MMEJ:NHEJ (Formula #3)
```{r}
#Extract non-targetting samples from this dataset
wt.set.ratio <- filter(screen.data.names.tib , inhibitor == "LBR2") %>% dplyr::group_by(barcode, replicate,cell_line) %>% dplyr::summarise(wt.log2ratio = mean(log2MMEJNHEJratio, na.rm = T))

# Compute log2 fold changes in different set ups for all the KO samples
log2.distance.ratio <- screen.data.names.tib %>% 
  na.omit() %>% #Remove NA
  left_join(wt.set.ratio, by = c("replicate", "barcode","cell_line")) %>% #Join mock KOs
  mutate(log2.foldchange = log2MMEJNHEJratio - wt.log2ratio) %>% #Calculate differential of log2MMEJ:NHEJ 
  na.omit() %>%
  dplyr::group_by(barcode,inhibitor, cell_line) %>% #Group by sample
  dplyr::summarise(mean.log2foldchange = mean(log2.foldchange), n_rep = n()) %>% #Compute differential of log2MMEJ:NHEJ
  ungroup() %>% filter(n_rep == 3)
```
  
  
#Step 4: Identification of DR proteins with global effects on MMEJ:NHEJ balance
```{r}
#Compute mean MMEJ:NHEJ differential over all IPRs
mean.diff.balance <- log2.distance.ratio %>% 
  filter(!inhibitor %in% c("LBR2")) %>%
  group_by(inhibitor,cell_line) %>% 
  dplyr::summarise(mean.fc = mean(mean.log2foldchange, na.rm = T))

#Run t.test with all the data
test.mean.diff.balance <- log2.distance.ratio %>% 
  filter(!inhibitor %in% c("LBR2")) %>%
  dplyr::group_by(inhibitor, cell_line) %>%
  rstatix::t_test(mean.log2foldchange ~ 1, mu = 0) %>% 
  mutate(p.adj = p.adjust(p, method = "BH")) %>% 
  left_join(mean.diff.balance, by = c("inhibitor","cell_line"))

#Prepare data frame for export
mean.diff.balance.export <- test.mean.diff.balance %>%
  dplyr::select(inhibitor,cell_line,p,p.adj,mean.balance.diff = mean.fc, n)

#Significant MMEJ
mean_significant_MMEJ <- filter(mean.diff.balance.export, p.adj < 0.05 & mean.balance.diff < 0) %>% nrow() #149 proteins favor MMEJ
mean_significant_NHEJ <- filter(mean.diff.balance.export, p.adj < 0.05 & mean.balance.diff > 0) %>% nrow() #16 proteins favor NHEJ

#Print
print(paste("A total of", mean_significant_MMEJ, "DR proteins favor MMEJ globally with an FDR < 0.05"))
print(paste("A total of", mean_significant_NHEJ, "DR proteins favor NHEJ globally with an FDR < 0.05"))
```



# Step5: Identification of repair proteisn with CCDs: three-step linear modeling

#Plot number of barcodes
```{r}
#Number of barcodes
ggplot(test.mean.diff.balance) + 
  geom_col(aes(cell_line,n)) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  geom_hline(yintercept = 8, linetype = 2)


#Positive controls

```

## Sub-step A: Selection of DR proteins with Effects on MMEJ:NHEJ balance
#This selection is not as easy as before, I can't calculate a z-score because I don't have enough controls. 
I will select based on two criteria (either the mean is different than 0 or has at least two IPRs on one side with an FDR of 0.25)

```{r}
#Compute t-test per IPR
IPR_test <- screen.data.names.tib %>% 
  filter(!inhibitor %in% c("LBR2")) %>%
  na.omit() %>% #Remove NA
  left_join(wt.set.ratio, by = c("replicate", "barcode","cell_line")) %>% 
  mutate(log2.foldchange = log2MMEJNHEJratio - wt.log2ratio) %>%
  ungroup() %>%
  na.omit() %>%
  dplyr::group_by(barcode, inhibitor,cell_line) %>%
  mutate(n_rep = n()) %>%
  filter(n_rep == 3) %>% 
  t_test(log2.foldchange ~ 1, mu = 0)

#Per gene get IPRs that have fdr < 0.25
IPR_test_summary <- IPR_test %>%
  dplyr::group_by(inhibitor,cell_line) %>%
  dplyr::mutate(fdr = p.adjust(p, method = "BH")) %>%
  ungroup() %>%
  dplyr::group_by(inhibitor,cell_line) %>%
  filter(fdr < 0.25) %>%
  dplyr::summarise(IPR_NHEJ = sum(statistic > 0),
                   IPR_MMEJ = sum(statistic < 0))

#Summary table with effect sizes
summary_table_mean <- test.mean.diff.balance %>%
  filter(p.adj < 0.05) %>%
  mutate(pathway_global = case_when(statistic < 0 ~ "MMEJ",
                             statistic > 0 ~ "NHEJ")) %>%
  select(inhibitor, pathway_global, cell_line)

summary_table_IPR <- IPR_test_summary %>%
  mutate(pathway_IPR = case_when(IPR_MMEJ > 1 ~ "MMEJ",
                                 IPR_NHEJ > 1 ~ "NHEJ",
                                 IPR_NHEJ > 1 & IPR_MMEJ > 1 ~ "both")) %>%
  select(inhibitor, pathway_IPR,cell_line)

#Filter based on this data
pathway_summary <- summary_table_mean %>%
  full_join(summary_table_IPR) %>%
  mutate(pathway_signif = case_when(pathway_global == pathway_IPR ~ pathway_global,
                             is.na(pathway_global) & is.na(pathway_IPR) ~ "none",
                             is.na(pathway_global) ~ pathway_IPR,
                             is.na(pathway_IPR) ~ pathway_global)) %>%
  select(inhibitor, pathway_signif, cell_line)
```

# Substep B: Principal component regression to identify repair proteins with CCDs
```{r}
#Filter genes that do not significantly perturb balance and add chromatin info
DDR.KO.CCD.dt <- log2.distance.ratio %>%
  left_join(pathway_summary) %>%
  filter(pathway_signif %in% c("NHEJ", "MMEJ"))  %>%
  left_join(RPE1_data_chrom_tib, by = c("barcode", "cell_line"))#Join chromatin info

#Create an empty dt with CCDs of DDR proteins
CCDs_dt_RPE1 <- tibble(cell_line = NA, inhibitor = NA, r.squared = NA, adj.r.squared = NA,p.value = NA)


  for (i in unique(DDR.KO.CCD.dt$cell_line)){
    gene.library.dt <- filter(DDR.KO.CCD.dt, inhibitor == "ATMi" & cell_line == i)
    if (nrow(gene.library.dt) < 9 ){
      next
    }
    set.seed(1)
    PCR_model_DDR_test <- pls::pcr(mean.log2foldchange~ dam_H3K27me3+chip_H3K4me1+chip_H3K36me3+dam_LMNB1+chip_H3K4me2+chip_H3K4me3+chip_H3K27ac+dam_H3K9me2+dam_H3K9me3+late_replicating, data=gene.library.dt) #Run principal component regression
    pcr_pred <- predict(PCR_model_DDR_test, gene.library.dt, ncomp = 3) #Run model with three PC
    combined.dt <- tibble(measured = gene.library.dt$mean.log2foldchange, predicted = as.numeric(pcr_pred)) #Create table with real and predicted differentials
    pred_vs_estim <- lm(formula = measured ~ predicted, data = combined.dt) %>% broom::glance() #Predicted vs. measured correlation plot
    CCDs_dt_RPE1 <- CCDs_dt_RPE1 %>% add_row(inhibitor = "ATMi", cell_line = i, r.squared = pred_vs_estim %>% pull(r.squared), adj.r.squared = pred_vs_estim %>% pull(adj.r.squared), p.value = pred_vs_estim %>% pull(p.value)) #Extract valuable metrics
  }

#Adjust per cell line
CCDs_dt_RPE1_adjs <- CCDs_dt_RPE1%>% mutate(fdr = p.adjust(p.value, method = "BH")) %>% right_join(test.mean.diff.balance %>% select(inhibitor,n,cell_line))

```

#Sub-step C: Linear modeling to identify individual DR proteins - chromatin feature links
```{r}
#Create empty dataframe to calculate synergy scores
RPE1_KO_CCD_slopes <- tibble(inhibitor = NA, cell_line = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

#Loop to run linear models on the values
for (h in unique(DDR.KO.CCD.dt$cell_line)) {
  for (j in colnames(DDR.KO.CCD.dt)[c(7:15,17)]) { #Run this function for each of the 25 high quality chromatin features
    model.dt <- DDR.KO.CCD.dt %>% filter(cell_line == h & inhibitor == "ATMi") # And For each gene
   if (nrow(model.dt) < 9) {
next
}
    model.epistasis.log2 <- lm(formula = mean.log2foldchange ~ unlist(model.dt[j]), data = model.dt) %>% tidy() #Correlation analysis
   RPE1_KO_CCD_slopes <- RPE1_KO_CCD_slopes %>% add_row(inhibitor = "ATMi", cell_line = h, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value)) #Select valuable parameters and save them in data frame
  }
}

#Retain slopes that 
DR_KO_CCD_all_values <- RPE1_KO_CCD_slopes %>% 
  reshape2::dcast(inhibitor + cell_line + feature ~ term, value.var = "slope.log2") %>%  #dcast table
  dplyr::select(inhibitor, cell_line ,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])') %>% #Extract information for slopes only
  left_join(CCDs_dt_RPE1_adjs %>% 
              dplyr::select(cell_line,inhibitor, model_p.value = p.value, model_p.adj = fdr), by = c("inhibitor","cell_line")) %>% #Join with CCDs p_value data table and change some names
  left_join(pathway_summary, by = c("inhibitor","cell_line")) %>% #Join with individual IPR significance calls (z-test)
  mutate(CCD_value = case_when(pathway_signif == "MMEJ" & slope < 0 ~ slope, pathway_signif == "MMEJ" & slope > 0 ~ 0, pathway_signif == "NHEJ" & slope > 0 ~ slope, pathway_signif == "NHEJ" & slope < 0 ~ 0, T ~ slope)) %>% #Call M-synergies, N-synergies or no synergies based on the slope and MMEJ:NHEJ differentials
  na.omit()

#How many M-, N- and no synergies
synergy_per_gene <- DR_KO_CCD_all_values %>% filter(model_p.adj < 0.25) %>% dplyr::group_by(inhibitor, cell_line) %>% dplyr::summarise(M_synergy = sum(CCD_value < 0), N_synergy = sum(CCD_value > 0)) %>% distinct() %>% mutate(CCD_effect = case_when(N_synergy == 0 ~ "M_synergy", M_synergy == 0 ~ "N_synergy", T ~ "both"))

synergy_summaries <- synergy_per_gene %>% dplyr::group_by(CCD_effect) %>% dplyr::summarise(c = n())
```

#Step 6: Compute estimated ∆log2MMEJ:NHEJ scores for CCD effects and compared to global ∆log2MMEJ:NHEJ
##Pre-processing chunk: load all ChIP track z-scores
```{r}
##Compute estimated CCD ∆log2MMEJ:NHEJ score
estimated_FC_CCD <- DR_KO_CCD_all_values %>% 
  left_join(RPE1_genomewide_dynamic_range, by = c("feature" ="file")) %>% #Join 99% CI for chromatin
  mutate(estim_fc_CCD = range*CCD_value)
```

#Step 7: Data visualization
This step will be performed with the output files in separate scripts 

#Create table S7: Data to export all major features
```{r}
export_screen_result_TS7 <- pathway_summary %>%
  full_join(estimated_FC_CCD) %>% 
  full_join(mean.diff.balance.export, by = c("inhibitor","cell_line")) %>%
  mutate(chrom_feature = str_extract(feature, "(H3|l|L).*")) %>%
  dplyr::select(inhibitor, cell_line, CCD_model_p_adj = "model_p.adj",chrom_feature, CCD_synergy_score = "CCD_value",CCD_estim_diff = "estim_fc_CCD", global_diff = "mean.balance.diff", global_p_adj = "p.adj", n, pathway_signif) %>%
  mutate(global_effect = case_when(global_p_adj < 0.05 & global_diff < 0 ~ "favorMMEJ",
                                   global_p_adj < 0.05 & global_diff > 0 ~ "favorNHEJ",
                                   T ~ "none"),
         cell_line = case_when(export_screen_result_TS7)) %>%
  left_join(synergy_per_gene %>% 
              select(inhibitor, CCD_effect))

export_screen_result_TS7[is.na(export_screen_result_TS7$CCD_effect), "CCD_effect"] <- "none"
```


#Output files: Export all files needed for making figures
```{r, eval = FALSE, echo = FALSE}
setwd(in.dir)

#Step 1: Raw freqCut & log2MMEJNHEJ values
saveRDS_proof(screen.data.names.tib, file = "data/processed_data/CCD_validation_RPE1/xv20231015_inhibitor_screening_raw_data_repair_metrics.rds")

#Step 2: dcast data table for 
saveRDS_proof(log2.distance.ratio, file = "data/processed_data/CCD_validation_RPE1/xv20231015_inhibitor_differentials_log2_MMEJ_NHEJ_balance.rds")

#Step 4: diff log2 MMEJ:NHEJ values for DR proteins with chromatin scores
saveRDS_proof(DDR.KO.CCD.dt, file = "data/processed_data/CCD_validation_RPE1/xv20231015_inhibitor_log2_MMEJNHEJ_differentials_chromatin_KO.rds")

#Step 5: save global changes for all proteins
saveRDS_proof(mean.diff.balance.export, file = "data/processed_data/CCD_validation_RPE1/xv20231015_inhibitor_global_diff_MMEJNHEJ.rds")

#Save final results as rds
saveRDS_proof(export_screen_result_TS7, file = "data/processed_data/xv20231015_inhibitor_CCD_summary.rds")

```

#Export supplementary tables in the paper
```{r, eval = FALSE, echo = FALSE}
setwd(in.dir)
#Export for supplementary files (.xlsx file)

#Table S4: ∆log2MMEJ:NHEJ scores for 519 proteins and 19 IPRs.
#write.xlsx(export_screen_result_TS4, file = "data/supplementary_tables/xv20220819_Table_S4a_delta_log2_MMEJ_NHEJ.xlsx")

#Table S6: Genomic coordinates, chromatin feature scores and barcodes of 19 IPRs in K562 clone 5
#write.xlsx(export_screen_result_TS6, file = "data/supplementary_tables/xv20220819_Table_S6_clone_5_chromatin_features.xlsx")

#Table S7: Chromatin context dependent effects of proteins in the screen.
write.xlsx(export_screen_result_TS7, file = "data/supplementary_tables/xv20231123_Table_S7e_RPE1_inhibitor_global_CCD_MMEJ_NHEJ_results.xlsx", keepNA = T)

```
