---
title: "xv20220814_SCC_mutations_analysis"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go through the mutation accumulation in BRCA2 mutant tumors

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}

```

## Libraries:

```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
```

## Import data

```{r pressure, echo=FALSE}
setwd(in.dir)

#Conversion data table to detect HPV negative data
sample_conversion <- read_xlsx("import/xv20220727_PCAWG_tumor_data/xv20220815_conversion_table_samples_TCGA.xlsx", col_names = T)

#Both data-sets
sporadic_mutation_dt <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/HNSC/BRASS/summary_shortened.txt") #BRCA2 +/+ HNSCC
BRCA2_mutation_dt <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/BRCA2/BRASS/summary_shortened.txt") #BRCA2 -/- HNSCC
```

#Import LADs data
```{r}
setwd(in.dir)
# LAD domains
LAD_atlas <- as.data.frame(read.table("import/xv20220329_LAD_atlas_OSF.bed.gz",header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(LAD_atlas) <- c("chr","start","end","length","strand","LAD_class")

#LAD_length plot control
LAD_length_tibble <- LAD_atlas %>% mutate(LAD_length = end - start)

```

#Curate data tables: What are the features that I am interested (only select deletions)
```{r, warning = F}
#Select only deletions (in all four files)
sporadic_mutation_del_dt <- sporadic_mutation_dt %>% 
  separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",") %>% #Remove from sample column the second TCGA id
  left_join(sample_conversion, by = c('sample' ='TCGA ID')) %>% #Bind HPV data
  filter(svclass == "deletion" & !"micro-homology" %in% c(".","_") & `Brass Notation` != "_" & grepl("Neg", `Name Supp. Table`)) #Filter MH mediated deletions

#Filter out data with FANC anemia mutations
sporadic_mutation_del_filtered <- sporadic_mutation_del_dt %>% 
  filter(!sample %in% c("de8ef60b-4dbe-4aa8-adef-6f58cdfada29","e52ffa79-557a-4024-81f3-f3826c227ec5")) #sample with mutation in putative FA proteins

#Numeric brass score (quality call)
sporadic_brass_score <- sporadic_mutation_del_filtered$`Brass Notation` %>% 
  str_extract("(?<=score ).*(?=\\))") %>% as.numeric()

#BRCA2 mutant data
BRCA2_mutation_del_dt <- BRCA2_mutation_dt %>% 
  separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",") %>% 
  filter(svclass == "deletion"  & !"micro-homology" %in% c(".","_") & `Brass Notation` != "_") #Filter MH mediated deletions

#numeric brass score (Quality call)
BRCA2_brass_score <- BRCA2_mutation_del_dt$`Brass Notation` %>% 
  str_extract("(?<=score ).*(?=\\))") %>% as.numeric()

#Summarise quality call
plot(density(sporadic_brass_score), col = "black", "BRASS score (Sporadinc HNSCC samples)")
plot(density(BRCA2_brass_score), col = "blue", "BRASS score (BRCA2 mutant HNSCC samples)")

print(paste("We detect a total of", length(sporadic_brass_score), "MH mediated deletions in 23 sporadic HNSCC and ", length(BRCA2_brass_score), "MH mediated deletions in 40 BRCA2 mutant tumor"))

```
#Conclusion : We detect a total of 830 MH mediated deletions in 23 sporadic HNSCC and 2673 MH mediated deletions in 40 BRCA2 mutant tumor

#Map SV for BRCA2 and sporadic tumors
```{r}
#Create function to call unique LADs
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T) #Make LADs into GR ranges

#Function to map mutations in unique LADs or iLADs
unique_LADs_call <- function(x) {
  mutations_GR <- makeGRangesFromDataFrame(x, keep.extra.columns = T) #Make a GR object with the coordinated of the deletion
  LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutations_GR) #Find overlaps between deletions and LAD states
  LAD_overlaps_dt <- tibble(x[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6]) #Extract the LAD_status of the mapped deletion
  LAD_counts <- LAD_overlaps_dt %>% dplyr::group_by(sample, id) %>% dplyr::summarize(count = n()) #How many different overlaps a deletions has. count = 1 -> single map, if bigger means goes over border
  single_LADs_iLADs <-LAD_counts %>% filter(count == 1) %>% select(sample, id) #Filter these that are called only once
  LAD_overlaps_dt %>% right_join(single_LADs_iLADs) #Keep only uniquely called
}

#Prepare the data table for mapping
mutation_location_export_spo_SV <- sporadic_mutation_del_filtered %>% #Sporadic HNSCC
  mutate(mut_id = paste(sample, `id/name`, sep = "_"), #Calculate mut_id (sample_mut)
         chr = paste0("chr", `chr1`)) %>% #change chr format
  select(chr, start = "start1", end = "start2", mut_id, sample, id = `id/name`) #select needed columns

mutation_location_export_BRCA2_SV <- BRCA2_mutation_del_dt %>%  #BRCA2 mutant HNSCC
  mutate(mut_id = paste(sample, `id/name`, sep = "_"), #Calculate mut_id (sample_mut)
         chr = paste0("chr", `chr1`)) %>% #change chr format
  select(chr, start = "start1", end = "start2", mut_id, sample, id = `id/name`) #select needed columns

#Run function
spo_SV_LAD_maps <- unique_LADs_call(mutation_location_export_spo_SV) #689 uniquely mapped MH del in HNSCC tumors
BRCA2_SV_LAD_maps <- unique_LADs_call(mutation_location_export_BRCA2_SV) #2281 uniquely mapped MH del in BRCA2mut tumors

#Filter for HPVneg and constitutive LADs or iLADs
spo_HPVneg_c_maps <- spo_SV_LAD_maps %>% #Sporadic HNSCC
  filter(grepl("c", LAD_status)) %>% #Constitutive LADs and iLADs only
  mutate(type = "sporadic_HPVneg") #336 total

BRCA2_SV_c_maps <- BRCA2_SV_LAD_maps %>% #BRCA2 mutant HNSCC
  filter(grepl("c", LAD_status)) %>% #Constitutive LADs and iLADs only
  mutate(type = "BRCA2mut") #1116 total

```
#Conclusion: After mapping, we had 336 MH deletions that uniquely mapped to constitutive LADs and iLADs in 23 sporadic HNSCC and 1116 MH deletion in 40 BRCA2 mutant HNSCC.

#Summary table with all the number in here
```{r, warnings = F}
#Bind all data
mutation_BRCA2_dataset <- bind_rows(spo_HPVneg_c_maps,BRCA2_SV_c_maps) %>% mutate(del_length = end - start)

#Summarise number of tumors sequenced per genetic background
sample_n <- mutation_BRCA2_dataset %>% 
  select(type, sample) %>% 
  distinct() %>% 
  dplyr::group_by(type) %>% dplyr::summarise(samp_n = n())
# Length distribution 95CI
length_summary <- mutation_BRCA2_dataset %>% 
  dplyr::group_by(type,LAD_status) %>% 
  dplyr::summarise(perc_5_length = quantile(del_length, c(0.05)), #5% percentile
                   perc_95_lenght = quantile(del_length, c(0.95))) #95% percentile

#mean number of mutations per tumor and chromatin context
mut_summary <-  mutation_BRCA2_dataset %>% 
  dplyr::group_by(sample, LAD_status,type) %>% dplyr::summarise(mut_n = n()) %>% #Number of mutations per tumor and chromatin context
  ungroup() %>% 
  dplyr::group_by(type,LAD_status) %>% dplyr::summarise(sum_mut = sum(mut_n)) %>% 
  left_join(sample_n) %>% mutate(mean_mut = sum_mut/samp_n) %>% #Calculate mean including 0 values
  dplyr::select(LAD_status,type,mean_mut) #Select important variables

# table summarizing mean for each run
summary_mutation_BRCA2_mut <- length_summary %>% left_join(mut_summary)

#Total 95CI 
total_length_summary_5 <- quantile(mutation_BRCA2_dataset$del_length, c(0.05,0.95))


print(summary_mutation_BRCA2_mut)
```

#Conclusion: BRCA2 mediated MH deletions accumulate in a CCD manner.
#Export data
```{r}
setwd(in.dir)

# Export file with selected features
export_HNSCC_mutations <- mutation_BRCA2_dataset %>% select(chr, start, end, tum_id = "sample", mut_id = "id", LAD_status, type, del_length)

saveRDS_proof(export_HNSCC_mutations, file = "data/processed_data/xv20220815_SCC_SV_MH_deletions.rds")

```
