---
title: "xv20211121_epistasis_model_bas"
output: html_document
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
```

# Aesthetic legend for the whole paper
```{r}
# Pathway color palette (MMEJ, wt/no_effect, NHEJ)
pathway.colors <- c("#EB2030","grey90","#2E358F")
# Pathway symbol palette ()
pathway.symbol <- c(19,17)

# Chromatin color palette
#chrom_colors = c( c("euchromatin" = "#F7941D", "other" = "#838687",  "H3K27me3" = "#D21F8A", "triple" = "#662D91"))
chromatin.colors <- c("#D21F8A","#838687","#F7941D","#662D91")
```

# Import data tables

```{r, include= FALSE}
#dcast table
#ddr
ddr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_dcast.rds")
#chromatin
chr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_dcast.rds")
  
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

#All variable table
ddr.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20220110_screen_ddr_all_computations.rds")
ddr.screen.all.sel <- ddr.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio, b.rep = bio.rep,t.rep = FinRep,freqCut, freqMMEJ, freqNHEJ) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#Chromatin all 
chromatin.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20220110_chromatin_library_all.rds")
chromatin.screen.all.sel <- chromatin.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio, b.rep,t.rep, freqCut, freqMMEJ, freqNHEJ) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "Chromatin")

# bind both dataset
both.screen.all.sel <- bind_rows(ddr.screen.all.sel,chromatin.screen.all.sel) %>% filter(!gene %in% c("DMSO","LBR_ctrl"))

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"

#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

#Chromatin data
clone5_chrom_tib <- readRDS("/DATA/projects/DSBrepair/data/R/rs20200519_clone5_newdoms_chromatin.RDS")
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)

```

# I will test NHEJ:MMEJ ratio reproducibility first
```{r}
# Select WT population
wt.ddr.screen.all.data <- both.screen.all.sel %>% filter(sample == "WT")
wt.ddr.screen.all.dcast <- wt.ddr.screen.all.data %>% reshape2::dcast(barcode + well + plate + library ~ t.rep, value.var = "NHEJMMEJratio")
ggplot(wt.ddr.screen.all.data) + geom_density(aes(NHEJMMEJratio, color = library, linetype = t.rep))

ggplot(wt.ddr.screen.all.dcast) + geom_point(aes(R1,R4))

#Reproducibility looks good overall
```
# Compute MMEJ over all 
```{r}
#
over.all.data <- both.screen.all.sel %>% mutate(MMEJall = freqMMEJ/freqCut, NHEJall = freqNHEJ/freqCut)
```

#Example with three genes (POLL, FANCM, POLQ) & I will use "H3K4me1" as example
1st step = Compare POLL, FANCM, POLQ and WT data points vs. H3K4me1 (in lin or log space)

step 1 = plot 

```{r}
# Data table and compute log2MMEJ
log2.ratio.screen.detail.all <- over.all.data  %>% select(t.rep, MMEJall,NHEJall,library, gene,barcode,sample) %>% filter(!(gene == "KAT5" & library == "DNA_repair"))

wt.set.ratio <- filter(log2.ratio.screen.detail.all , sample == "WT") %>% dplyr::group_by(barcode, t.rep,library) %>% dplyr::summarise(wt.MMEJall = mean(MMEJall, na.rm = T), wt.NHEJall = mean(NHEJall, na.rm = T))


# Plot differences
data.for.plotting.all.ratio <- log2.ratio.screen.detail.all %>% bind_rows(filter(log2.ratio.screen.detail.all , sample == "WT") %>% mutate(gene = "WT")) %>% dplyr::group_by(gene,barcode) %>% dplyr::summarise(m.MMEJall = mean(MMEJall, na.rm = T), m.NHEJall = mean(NHEJall, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

ggplot(data.for.plotting.all.ratio  %>% 
         filter(gene %in% c("POLL","WT")),
       aes(H3K4me1,m.MMEJall, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()

ggplot(data.for.plotting.all.ratio  %>% 
         filter(gene %in% c("POLL","WT")),
       aes(H3K4me1,m.NHEJall, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()
```


# 2nd step check compute distances and plot them
```{r}
# Each replicate has a different value (compute differences by replicate)
log2.distance.ratio.all <- log2.ratio.screen.detail.all %>% filter(sample == "KO") %>% left_join(wt.set.ratio) %>% mutate(log2.fc.MMEJ = log2(MMEJall/wt.MMEJall),log2.fc.NHEJ = log2(NHEJall/wt.NHEJall))
mean.log2.distance.ratio.all <- log2.distance.ratio.all %>% dplyr::group_by(barcode,gene) %>% dplyr::summarise(m.log2.fc.MMEJ= mean(log2.fc.MMEJ, na.rm = T), m.log2.fc.NHEJ = mean(log2.fc.NHEJ,na.rm =T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

ggplot(mean.log2.distance.ratio.all %>% 
         filter(gene == "POLL"),
       aes(H3K4me1,m.log2.fc.MMEJ, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()


ggplot(mean.log2.distance.ratio.all %>% 
         filter(gene == "POLL"),
       aes(H3K4me1,m.log2.fc.NHEJ, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()

#Compute mean log2 fold change
mean.log2.fc <- mean.log2.distance.ratio.all %>% group_by(gene) %>% dplyr::summarise(mean.fc = mean(m.log2.ratio.dist, na.rm = T))

```
# Extract slopes for all genes
```{r}
# Perform analysis across all features for three test genes
chromatin.features <- colnames(mean.log2.distance.ratio.all)[7:31]

#
slope.protein.features.all.MMEJ <- tibble(gene = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

for (i in unique(mean.log2.distance.ratio.all$gene)) {
  for (j in colnames(mean.log2.distance.ratio.all)[7:31]) {
    model.dt <- mean.log2.distance.ratio.all %>% filter(gene == i)
    model.epistasis.log2 <- lm(formula = m.log2.fc.MMEJ ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
    slope.protein.features.all.MMEJ <- slope.protein.features.all.MMEJ %>% add_row(gene = i, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
  }
}

slope.protein.features.all.NHEJ <- tibble(gene = NA, feature = NA, slope.log2 = NA, term = NA,p.value = NA)

for (i in unique(mean.log2.distance.ratio.all$gene)) {
  for (j in colnames(mean.log2.distance.ratio.all)[7:31]) {
    model.dt <- mean.log2.distance.ratio.all %>% filter(gene == i)
    model.epistasis.log2 <- lm(formula = m.log2.fc.NHEJ ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
    slope.protein.features.all.NHEJ <- slope.protein.features.all.NHEJ %>% add_row(gene = i, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term), p.value = model.epistasis.log2 %>% pull(p.value))
  }
}

#Both calculations
MMEJ.NHEJ.all.epistasis <- slope.protein.features.all.MMEJ %>% select(gene,feature,MMEJ.slope = slope.log2, MMEJ.p.value = p.value,term) %>% left_join(slope.protein.features.all.NHEJ %>% select(gene,feature,NHEJ.slope = slope.log2, NHEJ.pvalue = p.value, term))

# Make all negative epistatic interactions 0
epistasis.ratio.sign <- slope.protein.features.all.ratio %>% filter(complete.cases(.)) %>% left_join(mean.log2.fc) %>% dcast(gene + feature + mean.fc ~ term, value.var = "slope.log2") %>% select(gene,feature,intercept = '(Intercept)', slope = 'unlist(model.dt[j])', mean.fc) %>% mutate(epistasis = case_when(mean.fc > 0 & slope > 0 ~ slope, mean.fc > 0 & slope <= 0 ~ 0, mean.fc <= 0 & slope > 0 ~ 0, mean.fc <= 0 & slope < 0 ~ slope))


epistasis.ratio.sign.dcast <- epistasis.ratio.sign %>% dcast(gene ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "gene")
pheatmap(t(epistasis.ratio.sign.dcast))



```

# Test matrix
```{r}
# 
epistasis.ratio.sign.dcast <- epistasis.ratio.sign %>% dcast(gene ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "gene")

```
#Save every step
```{r}
# Step 1
saveRDS(data.for.plotting.all.ratio,file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")

#Step 2
saveRDS(mean.log2.distance.ratio.all, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

# Save all proteins correlation final
saveRDS(epistasis.ratio.sign, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_all_genes.rds")

```

