---
title: "xv20211121_epistasis_model_bas"
output: html_document
---

In this document, I will go over different calculations of epistasis.4 different workflows:
First decision is whether to use log2 MMEJscore or MMEJscore
Second decision is comparing slopes or statistic differences between WT vs. KO slopes

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
```

# Aesthetic legend for the whole paper
```{r}
# Pathway color palette (MMEJ, wt/no_effect, NHEJ)
pathway.colors <- c("#EB2030","grey90","#2E358F")
# Pathway symbol palette ()
pathway.symbol <- c(19,17)

# Chromatin color palette
#chrom_colors = c( c("euchromatin" = "#F7941D", "other" = "#838687",  "H3K27me3" = "#D21F8A", "triple" = "#662D91"))
chromatin.colors <- c("#D21F8A","#838687","#F7941D","#662D91")
```

# Import data tables

```{r, include= FALSE}
#dcast table
#ddr
ddr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_dcast.rds")
#chromatin
chr.screen.dcast <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_dcast.rds")
  
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"


# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"

#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

#Chromatin data
clone5_chrom_tib <- readRDS("/DATA/projects/DSBrepair/data/R/rs20200519_clone5_newdoms_chromatin.RDS")
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

# Inhibitor experiment
inhibitor.data <- readRDS("~/XV_P3_ChromDSBScreen/xv20210716_E1627_ATR_inhibitor/data/xv20210716_E1627_indel_script_ATRi.rds") %>% left_join(inhibitor.table)

```

# Put both screens together
```{r}
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

screen.detail.others <- both.screen.detail %>% mutate(NHEJscore = 1 - MMEJscore)
```

#Example with three genes (POLL, FANCM, POLQ) & I will use "H3K4me1" as example
1st step = Compare POLL, FANCM, POLQ and WT data points vs. H3K4me1 (in lin or log space)

step 1 = plot 

```{r}
# Data table and compute log2MMEJ
log2.NHEJ.screen.detail.all <- screen.detail.others %>% select(t.rep, NHEJscore,library, gene,barcode) %>% filter(!(gene == "KAT5" & library == "DNA_repair"))

# WT data table set
both.screen.controls.nhej <- both.screen.controls %>% mutate(NHEJscore = 1 - MMEJscore)

wt.set.nhej <- filter(both.screen.controls.nhej, gene == "WT") %>% select(t.rep, NHEJscore,library, gene,barcode) %>% dplyr::group_by(barcode, t.rep,library) %>% dplyr::summarise(wt.NHEJ = mean(NHEJscore, na.rm = T))

# Plot differences
data.for.plotting.all.nhej <- wt.set %>% mutate(gene = "WT") %>% select(t.rep, library, gene, barcode,NHEJscore = wt.NHEJ) %>% bind_rows(log2.NHEJ.screen.detail.all) %>% mutate(rep_lib = paste(t.rep,library, sep = "_")) %>% dplyr::group_by(gene,barcode) %>% dplyr::summarise(m.NHEJscore= mean(NHEJscore, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

ggplot(data.for.plotting.all.nhej  %>% 
         filter(gene %in% c("POLL", "WT")),
       aes(H3K4me1,m.NHEJscore, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()

ggplot(data.for.plotting.all.nhej  %>% 
         filter(gene %in% c("POLL", "WT","FANCM")),
       aes(H3K4me1,log2(m.NHEJscore), color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()
```


# 2nd step check compute distances and plot them
```{r}
# Each replicate has a different value (compute differences by replicate)
log2.distance.nhej.all <- log2.NHEJ.screen.detail.all %>% left_join(wt.set.nhej) %>% mutate(log2.dist = log2(log2NHEJ/wt.log2NHEJ)))
mean.log2.distance.nhej.all <- log2.distance.nhej.all %>% dplyr::group_by(barcode,gene) %>% dplyr::summarise(m.log2.dist = mean(log2.dist, na.rm = T)) %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

#POLL
ggplot(mean.log2.distance.nhej.all %>% 
         filter(gene == "POLL"),
       aes(H3K4me1,m.log2.dist, color = gene)) + 
  geom_point() + 
  stat_cor(method = "spearman") + 
  geom_smooth(method = "lm") + 
  theme_bw()

```
# Extract slopes for all genes
```{r}
# Perform analysis across all features for three test genes
chromatin.features <- colnames(mean.log2.distance.nhej.all)[6:30]

slope.protein.features.all.nhej <- tibble(gene = NA, feature = NA, slope.log2 = NA, term = NA)

for (i in unique(mean.log2.distance.nhej.all$gene)) {
  for (j in colnames(mean.log2.distance.nhej.all)[6:30]) {
    model.dt <- mean.log2.distance.nhej.all %>% filter(gene == i)
    model.epistasis.log2 <- lm(formula = m.log2.dist ~ unlist(model.dt[j]), data = model.dt) %>% tidy()
    slope.protein.features.all.nhej <- slope.protein.features.all.nhej %>% add_row(gene = i, feature = j, slope.log2 = model.epistasis.log2 %>% pull(estimate), term = model.epistasis.log2 %>% pull(term))
  }
}

```
#Save every step
```{r}
# Step 1
#saveRDS(data.for.plotting.all,  file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_1_step.rds")

#Step 2
#saveRDS(mean.log2.distance.mmej.all, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_2_step.rds")

# Save all proteins correlation final
#saveRDS(slope.protein.features.all, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_all_genes.rds")

```

