---
title: "xv20220814_SCC_mutations_analysis"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

In this document, I will go through the mutation accumulation in BRCA2 mutant tumors

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}

```

## Libraries:

```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
```

## Import data

```{r pressure, echo=FALSE}
setwd(in.dir)

#Conversion data table to detect HPV negative data
sample_conversion <- read_xlsx("import/xv20220727_PCAWG_tumor_data/xv20220815_conversion_table_samples_TCGA.xlsx", col_names = T)

#Both data-sets
sporadic_mutation_dt <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/HNSC/BRASS/summary_shortened.txt") #BRCA2 +/+ HNSCC
BRCA2_mutation_dt <- read_delim("import/xv20220727_PCAWG_tumor_data/SV_Indels_PCAWG/BRCA2/BRASS/summary_shortened.txt") #BRCA2 -/- HNSCC

#Import data for indels classified by state
#BRCA2 mut
BRCA2_indel_files <- list.files(path = "import/xv20220915_INDEL_data_mathijs/xv20220920_CLASS_mathijs_data/BRCA2",full.names = T)
BRCA2_indel_data <- map_dfr(BRCA2_indel_files, function(x) {
  read.delim(x, header = T) %>% mutate(sample = gsub(".class","",gsub("import/xv20220915_INDEL_data_mathijs/xv20220920_CLASS_mathijs_data/BRCA2/","",x)))
})
#HNSCC mut
HNSCC_indel_files <- list.files(path = "import/xv20220915_INDEL_data_mathijs/xv20220920_CLASS_mathijs_data/HNSC",full.names = T)
HNSCC_indel_data <- map_dfr(HNSCC_indel_files, function(x) {
  read.delim(x, header = T) %>% mutate(sample = gsub(".class","",gsub("import/xv20220915_INDEL_data_mathijs/xv20220920_CLASS_mathijs_data/HNSC/","",x)))
})

```

#Import LADs data
```{r}
setwd(in.dir)
# LAD domains
LAD_atlas <- as.data.frame(read.table("import/xv20220329_LAD_atlas_OSF.bed.gz",header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(LAD_atlas) <- c("chr","start","end","length","strand","LAD_class")

#LAD_length plot control
LAD_length_tibble <- LAD_atlas %>% mutate(LAD_length = end - start)

```
#Curate indel patterns: What are the features that I am interested (only select deletions)
```{r, warning = F}
#Select only deletions (in all four files)
sporadic_indel_dt <- HNSCC_indel_data %>% 
  left_join(sample_conversion, by = c('sample' ='TCGA ID')) %>% #Bind HPV data
  filter(!sample %in% c("de8ef60b-4dbe-4aa8-adef-6f58cdfada29","e52ffa79-557a-4024-81f3-f3826c227ec5","d25a4c65-9cb4-4611-909e-e68f93408d84") &grepl("Neg", `Name Supp. Table`)) #sample with mutation in putative FA proteins

#Total mutations per tumor
summary_n_tumour_spo_indel <- sporadic_indel_dt %>% dplyr::group_by(sample) %>% dplyr::summarise(mut_total = n())

#BRCA2 mutant data
BRCA2_indel_dt <- BRCA2_indel_data

#Number of total MH flanked mutation per tumor
summary_n_tumour_BRCA2_indel <- BRCA2_indel_dt %>% dplyr::group_by(sample) %>% dplyr::summarise(mut_total = n())

print(paste("We detect a total of", sum(summary_n_tumour_spo_indel$mut_total), "MH mediated deletions in", nrow(summary_n_tumour_spo_indel), "sporadic HNSCC and ", sum(summary_n_tumour_BRCA2_indel$mut_total), "MH mediated deletions in",nrow(summary_n_tumour_BRCA2_indel),"BRCA2 mutant tumor"))

```

#Map indel data for BRCA2 and sporadic tumors
```{r}
#Create function to call unique LADs
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T) #Make LADs into GR ranges

#Function to map mutations in unique LADs or iLADs
LADs_status_call_indel <- function(x) {
  print(dim(x))
  mutations_GR <- makeGRangesFromDataFrame(x, keep.extra.columns = T) #Make a GR object with the coordinated of the deletion
  LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutations_GR) #Find overlaps between deletions and LAD states
  LAD_overlaps_dt <- tibble(x[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6]) #Extract the LAD_status of the mapped deletion
  LAD_counts <- LAD_overlaps_dt %>% dplyr::group_by(tumor_id, mutation_id, LAD_status) %>% dplyr::summarize(count = n()) #How many different overlaps a deletions has. count = 1 -> single map, if bigger means goes over border
  single_LADs_iLADs <-LAD_counts %>% reshape2::dcast(tumor_id + mutation_id ~ LAD_status, fill = 0) #Filter these that are called only once
  #LAD_overlaps_dt %>% right_join(single_LADs_iLADs) #Keep only uniquely called
}

#Prepare the data table for mapping
mutation_location_export_spo_indel <- sporadic_indel_dt %>% #Sporadic HNSCC
    mutate(chr = CHR, start = POSITION, end = POSITION + 1,mutation_id = 1:nrow(.)) %>% #change chr format
  select(chr, start, end, tumor_id = sample, mutation_id, indelClass, indelSubClass, indelRepeats, CLASS) #select needed columns

mutation_location_export_BRCA2_indel<- BRCA2_indel_dt %>%  #BRCA2 mutant HNSCC
  mutate(chr = CHR, start = POSITION, end = POSITION + 1,mutation_id = 1:nrow(.)) %>% #change chr format
  select(chr, start, end, tumor_id = sample, mutation_id, indelClass, indelSubClass, indelRepeats, CLASS) #select needed columns

#Run function (table A)
spo_indel_LAD_maps <- LADs_status_call_indel(mutation_location_export_spo_indel) %>% mutate(tumor_type = "HHNSC_indel") #14560 uniquely mapped MH del in HNSCC tumors
BRCA2_indel_LAD_maps <- LADs_status_call_indel(mutation_location_export_BRCA2_indel) %>% mutate(tumor_type = "BRCA2mut_indel") #47460 uniquely mapped MH del in BRCA2mut tumors

#Build df for export
HPVneg_HNSCC_export_indel <- mutation_location_export_spo_indel %>% 
  left_join(spo_indel_LAD_maps, by = c("tumor_id", "mutation_id")) %>% 
  mutate(CLASS = case_when(indelClass == ">1bp deletion at repeats" & indelSubClass == "5+" & indelRepeats == "1" ~ "cNHEJ",
                           T ~ CLASS))

BRCA2_HNSCC_export_indel <- mutation_location_export_BRCA2_indel %>% left_join(BRCA2_indel_LAD_maps, by = c("tumor_id", "mutation_id"))  %>% 
  mutate(CLASS = case_when(indelClass == ">1bp deletion at repeats" & indelSubClass == "5+" & indelRepeats == "1" ~ "cNHEJ",
                           T ~ CLASS))

print(paste("CONCLUSION: We detect a total of", nrow(HPVneg_HNSCC_export_indel), "SV in 23 sporadic HNSCC and ", nrow(BRCA2_HNSCC_export_indel), "SV in 40 BRCA2 mutant tumor"))
```
#Plot total number of mutations
```{r}
total_mutations <- HPVneg_HNSCC_export_indel %>% na.omit() %>% dplyr::group_by(CLASS) %>% dplyr::summarise(count = n(), mean_counts = n()/22, tumor_type = "HNSC_HPVneg") %>% 
  bind_rows(BRCA2_HNSCC_export_indel %>% na.omit() %>% dplyr::group_by(CLASS) %>% dplyr::summarise(count = n(), mean_counts = n()/41, tumor_type = "BRCA2_mut"))

#Log2 fold change
fold_change_total_mut <- total_mutations %>% reshape2::dcast(CLASS ~ tumor_type, value.var = "mean_counts") %>% mutate(fold_change_mutation = log2(BRCA2_mut/HNSC_HPVneg))

#Plot fold change increase and decrease
ggplot(fold_change_total_mut) + 
  geom_col(aes(CLASS, fold_change_mutation)) + 
  theme_bw() + 
  geom_hline(yintercept = 0, linetype = 2)

```


#Quick check of mutation accumulation
```{r}
#Check indels and see how they look quickly (HNSC)
summary_table_HNSCC <- HPVneg_HNSCC_export_indel %>% dplyr::group_by(CLASS) %>% dplyr::summarise(tot_cLAD = sum(cLAD, na.rm = T),
                                                                                     tot_ciLAD = sum(ciLAD, na.rm = T),
                                                                                     tot_fLAD = sum(fLAD, na.rm = T),
                                                                                     tot_fiLAD = sum(fiLAD, na.rm = T), 
                                                                                     tumortype = "HNSC")

#Check indels and see how they look quickly (HNSC)
summary_table_BRCA2 <- BRCA2_HNSCC_export_indel %>% dplyr::group_by(CLASS) %>% dplyr::summarise(tot_cLAD = sum(cLAD, na.rm = T),
                                                                                     tot_ciLAD = sum(ciLAD, na.rm = T),
                                                                                     tot_fLAD = sum(fLAD, na.rm = T),
                                                                                     tot_fiLAD = sum(fiLAD, na.rm = T), 
                                                                                     tumortype = "BRCA2")

#Plot all mutation
plot_mutation_chromatin_state <- summary_table_HNSCC %>% na.omit() %>% select(CLASS, cLAD  = tot_cLAD, ciLAD = tot_ciLAD, tumor_type = tumortype) %>% reshape2::melt() %>% mutate(mean_counts = value/22) %>% bind_rows(summary_table_BRCA2 %>% na.omit() %>% select(CLASS, cLAD  = tot_cLAD, ciLAD = tot_ciLAD, tumor_type = tumortype) %>% reshape2::melt() %>% mutate(mean_counts = value/41))

summary_per_tumor <- HPVneg_HNSCC_export_indel %>% mutate(tumor_type = "HNSC") %>% bind_rows(BRCA2_HNSCC_export_indel %>% mutate(tumor_type = "BRCA2")) %>% dplyr::group_by(CLASS, tumor_type, tumor_id) %>% dplyr::summarise(tot_cLAD = sum(cLAD, na.rm = T),
                                                                                     tot_ciLAD = sum(ciLAD, na.rm = T),
                                                                                     tot_fLAD = sum(fLAD, na.rm = T),
                                                                                     tot_fiLAD = sum(fiLAD, na.rm = T))
plot_mutation_chromatin_state_tumor <- summary_per_tumor %>% na.omit() %>% select(CLASS,tumor_id, cLAD  = tot_cLAD, ciLAD = tot_ciLAD, tumor_type) %>% reshape2::melt()

## Merge
summary_table <- bind_rows(summary_table_BRCA2, summary_table_HNSCC) %>% 
  rowwise() %>% 
  mutate(total_indel = sum(tot_cLAD, tot_ciLAD, tot_fLAD, tot_fiLAD)) %>%
  mutate_at(c("tot_cLAD","tot_ciLAD","tot_fLAD","tot_fiLAD"), ~ . / total_indel) %>%
  reshape2::melt() %>% filter(variable != "total_indel") %>%
  reshape2::dcast(CLASS ~ tumortype + variable) %>% 
  select(CLASS, HNSC_tot_ciLAD, BRCA2_tot_ciLAD, HNSC_tot_cLAD, BRCA2_tot_cLAD)

colnames(summary_table) <- gsub("_tot","", colnames(summary_table))

#Print summary table 
print(summary_table)
```

#Prepare dataframe for export
```{r}
#
indel_dataset_BRCA2_HNSC_supplementary <- HPVneg_HNSCC_export_indel %>% bind_rows(BRCA2_HNSCC_export_indel) 

indel_dataset_BRCA2_HNSC <- indel_dataset_BRCA2_HNSC_supplementary %>%
  filter(CLASS %in% c("cNHEJ","MMEJ"))

total_mutation_class <- indel_dataset_BRCA2_HNSC %>% dplyr::group_by(tumor_id, CLASS, tumor_type) %>% dplyr::summarise(total_mutation = n())

indel_dataset_BRCA2_HNSC_figure <- left_join(indel_dataset_BRCA2_HNSC,total_mutation_class)

```


#Curate data tables: What are the features that I am interested (only select deletions)
```{r, warning = F}
#Select only deletions (in all four files)
sporadic_mutation_del_dt <- sporadic_mutation_dt %>% 
  separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",") %>% #Remove from sample column the second TCGA id
  left_join(sample_conversion, by = c('sample' ='TCGA ID')) %>% #Bind HPV data
  filter(!sample %in% c("de8ef60b-4dbe-4aa8-adef-6f58cdfada29","e52ffa79-557a-4024-81f3-f3826c227ec5","d25a4c65-9cb4-4611-909e-e68f93408d84") &grepl("Neg", `Name Supp. Table`)) #sample with mutation in putative FA proteins

#Total mutations per tumor
summary_n_tumour_spo <- sporadic_mutation_del_dt %>% dplyr::group_by(sample) %>% dplyr::summarise(mut_total = n())

#Numeric brass score (quality call)
sporadic_brass_score <- sporadic_mutation_del_dt$`Brass Notation` %>% 
  str_extract("(?<=score ).*(?=\\))") %>% as.numeric()

#BRCA2 mutant data
BRCA2_mutation_del_dt <- BRCA2_mutation_dt %>% 
  separate(`# sample`, into = c("sample","other_TCGA_sample"), sep = ",")


#Number of total MH flanked mutation per tumor
summary_n_tumour_BRCA2 <- BRCA2_mutation_del_dt %>% dplyr::group_by(sample) %>% dplyr::summarise(mut_total = n())

#numeric brass score (Quality call)
print(paste("CONCLUSION: We detect a total of", length(sporadic_brass_score), "SV in 23 sporadic HNSCC and ", length(BRCA2_brass_score), "SV in 40 BRCA2 mutant tumor"))

```

#Map SV for BRCA2 and sporadic tumors
```{r}
#Create function to call unique LADs
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T) #Make LADs into GR ranges

#Function to map mutations in unique LADs or iLADs
LADs_status_call <- function(x) {
  print(dim(x))
  same_chr_SV <- filter(x, as.numeric(chr_end) == as.numeric(chr_start))
  print(dim(same_chr_SV))
  mutations_GR <- makeGRangesFromDataFrame(same_chr_SV, keep.extra.columns = T) #Make a GR object with the coordinated of the deletion
  LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutations_GR) #Find overlaps between deletions and LAD states
  LAD_overlaps_dt <- tibble(same_chr_SV[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6]) #Extract the LAD_status of the mapped deletion
  LAD_counts <- LAD_overlaps_dt %>% dplyr::group_by(tumor_id, mutation_id, LAD_status) %>% dplyr::summarize(count = n()) #How many different overlaps a deletions has. count = 1 -> single map, if bigger means goes over border
  single_LADs_iLADs <-LAD_counts %>% reshape2::dcast(tumor_id + mutation_id ~ LAD_status, fill = 0) #Filter these that are called only once
  #LAD_overlaps_dt %>% right_join(single_LADs_iLADs) #Keep only uniquely called
}

#Prepare the data table for mapping
mutation_location_export_spo_SV <- sporadic_mutation_del_dt %>% #Sporadic HNSCC
    mutate(chr = paste0("chr", `chr1`)) %>% #change chr format
  select(chr, start = "start1", end = "start2", tumor_id = sample, mutation_id = `id/name`, svclass,chr_end = "chr2", chr_start = "chr1") #select needed columns

mutation_location_export_BRCA2_SV <- BRCA2_mutation_del_dt %>%  #BRCA2 mutant HNSCC
  mutate(chr = paste0("chr", `chr1`)) %>% #change chr format
  select(chr, start = "start1", end = "start2", tumor_id = sample, mutation_id = `id/name`, svclass,chr_end = "chr2", chr_start = "chr1") #select needed columns

#Run function (table A)
spo_SV_LAD_maps <- LADs_status_call(mutation_location_export_spo_SV) %>% mutate(tumor_type = "HPVneg_HNSCC") #689 uniquely mapped MH del in HNSCC tumors
BRCA2_SV_LAD_maps <- LADs_status_call(mutation_location_export_BRCA2_SV) %>% mutate(tumor_type = "BRCA2mut_HNSCC")#2281 uniquely mapped MH del in BRCA2mut tumors

#Table b
sv_type_spo <- sporadic_mutation_del_dt %>% select(tumor_id = sample, mutation_id= `id/name`, svclass,`micro-homology`,`non-template`,`Brass Notation`)
sv_type_BRCA2 <- BRCA2_mutation_del_dt %>% select(tumor_id = sample, mutation_id = `id/name`, svclass,`micro-homology`,`non-template`,`Brass Notation`)

#Table C
BRASS_else_spo <- sporadic_mutation_del_dt %>% select(-svclass,-`micro-homology`,-`non-template`, - `readpair names`, -`Brass Notation`)
BRASS_else_BRCA2 <- BRCA2_mutation_del_dt %>% select(-svclass,-`micro-homology`,-`non-template`, -`readpair names`, -`Brass Notation`)


#Build df for export
HPVneg_HNSCC_export <- spo_SV_LAD_maps %>% left_join(sv_type_spo) %>% left_join(BRASS_else_spo, by = c("tumor_id" = "sample","mutation_id" = "id/name")) %>% select(-`Overlap SV`, - `Name Supp. Table`)

BRCA2_HNSCC_export <- BRCA2_SV_LAD_maps %>% left_join(sv_type_BRCA2) %>% left_join(BRASS_else_BRCA2, by = c("tumor_id" = "sample","mutation_id" = "id/name"))

```

```{r}
#Filter for HPVneg and constitutive LADs or iLADs
spo_HPVneg_c_maps <- HPVneg_HNSCC_export %>% #Sporadic HNSCC
  filter(svclass == "deletion" & length(`micro-homology`) > 1 & `micro-homology` != "." & `non-template` == "." & `Brass Notation` != "_") %>% #Filter MH mediated deletions
  filter((ciLAD == 1 | cLAD == 1 | fiLAD == 1 | fLAD == 1) & ciLAD + cLAD + fiLAD + fLAD == 1) %>% 
  mutate(type = "sporadic_HPVneg") #420 total

mutation_per_tumor_spo <- spo_HPVneg_c_maps %>% dplyr::group_by(tumor_id) %>% dplyr::summarise(total_MH_del = n())

final_spo_HPVneg_c_mutations <- left_join(spo_HPVneg_c_maps,mutation_per_tumor_spo, by = "tumor_id")

BRCA2_SV_c_maps <- BRCA2_HNSCC_export  %>% #Sporadic HNSCC
  filter(svclass == "deletion" & length(`micro-homology`) > 1 & `micro-homology` != "." & `non-template` == "." & `Brass Notation` != "_") %>% #Filter MH mediated deletions
  filter((ciLAD == 1 | cLAD == 1 | fiLAD == 1 | fLAD == 1) & ciLAD + cLAD + fiLAD + fLAD == 1) %>% 
  mutate(type = "BRCA2mut") #747 total

mutation_per_tumor_BRCA <- BRCA2_SV_c_maps %>% dplyr::group_by(tumor_id) %>% dplyr::summarise(total_MH_del = n())

final_BRCA2_c_mutations <- left_join(BRCA2_SV_c_maps,mutation_per_tumor_BRCA, by = "tumor_id")

#numeric brass score (Quality call)
print(paste("CONCLUSION: We detect a total of",nrow(spo_HPVneg_c_maps), "MH uniquely mapped in 23 sporadic HNSCC and ", nrow(BRCA2_SV_c_maps), "MH uniquely mapped in 40 BRCA2 mutant tumor"))

```

#Summary table with all the number in here
```{r, warnings = F}
#Bind all data
mutation_BRCA2_dataset <- bind_rows(final_spo_HPVneg_c_mutations,final_BRCA2_c_mutations) %>% mutate(del_length = start2 - start1)

#Summarise number of tumors sequenced per genetic background
sample_n <- mutation_BRCA2_dataset %>% 
  select(type, tumor_id) %>% 
  distinct() %>% 
  dplyr::group_by(type) %>% dplyr::summarise(samp_n = n())
# Length distribution 95CI
length_summary <- mutation_BRCA2_dataset %>% 
  dplyr::group_by(type,cLAD, ciLAD, fiLAD, fLAD) %>% 
  dplyr::summarise(perc_5_length = quantile(del_length, c(0.05)), #5% percentile
                   perc_95_lenght = quantile(del_length, c(0.95))) #95% percentile

#mean number of mutations per tumor and chromatin context
mut_summary <-  mutation_BRCA2_dataset %>% 
  dplyr::group_by(tumor_id, cLAD, ciLAD,fiLAD, fLAD,type) %>% dplyr::summarise(mut_n = n()) %>% #Number of mutations per tumor and chromatin context
  ungroup() %>% 
  dplyr::group_by(type,cLAD, ciLAD, fiLAD, fLAD) %>% dplyr::summarise(sum_mut = sum(mut_n)) %>% 
  left_join(sample_n, by = "type") %>% mutate(mean_mut = sum_mut/samp_n) %>% #Calculate mean including 0 values
  dplyr::select(cLAD, ciLAD, fiLAD, fLAD,type,mean_mut) #Select important variables

# table summarizing mean for each run
summary_mutation_BRCA2_mut <- length_summary %>% left_join(mut_summary,by = c("type", "cLAD", "ciLAD", "fiLAD", "fLAD"))

#Total 95CI 
total_length_summary_5 <- quantile(mutation_BRCA2_dataset$del_length, c(0.05,0.95))


print(summary_mutation_BRCA2_mut)
```

#Conclusion: BRCA2 mediated MH deletions accumulate in a CCD manner.
#Export data
```{r}
setwd(in.dir)

#Export files for plotting
#SV 
export_HNSCC_mutations <- mutation_BRCA2_dataset %>% select(chr = "chr1", start = "start1", end = "start2", tumor_id,total_MH_del, mutation_id, ciLAD, cLAD, fiLAD, fLAD, type, del_length, `Brass Notation`)
#Indels
saveRDS_proof(indel_dataset_BRCA2_HNSC_figure, file = "data/processed_data/xv20220922_SCC_indel_MMEJ_NHEJ_deletions.rds")

#Export file for supplementary figures with the BRASS notation
write.xlsx(HPVneg_HNSCC_export, file = "data/supplementary_tables/xv20220922_Table_S9a_SV_mutations_tumors_BRASS.xlsx")
write.xlsx(BRCA2_HNSCC_export, file = "data/supplementary_tables/xv20220922_Table_S9b_SV_mutations_tumors_BRASS.xlsx")
#Export indel calls
write.xlsx(HPVneg_HNSCC_export_indel, file = "data/supplementary_tables/xv20220922_Table_S10a_indel_mutations_tumors_COSMIC.xlsx")
write.xlsx(BRCA2_HNSCC_export_indel, file = "data/supplementary_tables/xv20220922_Table_S10b_indel_mutations_tumors_COSMIC.xlsx")

```
