---
title: "xv20220329_FA_SCC_mutations_analysis"
output: html_document
date: '2022-03-29'
---

I will use "Fanconi Anemia Pathway Deficiency Drives Copy Number Variation in Squamous Cell Carcinomas" dataset to assess if FA SCC tumors accumulate mutations differently than sporadic tumors.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries:

```{r libraries}
library(readxl)
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
```

## Import data

```{r pressure, echo=FALSE}
#Import FA mutations
FA_mutation_dt <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_FA_SCC_webster_et_al_S5.xlsx", col_names = T, sheet = 1, skip = 1)

#Import sporadic mutations
sporadic_mutation_dt <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_FA_SCC_webster_et_al_S5.xlsx", col_names = T, sheet = 2, skip = 1)

#Import sporadic mutations
BRCA2_mutation_dt <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_FA_SCC_webster_et_al_S5.xlsx", col_names = T, sheet = 3, skip = 1)
BRCA1_mutation_dt <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_FA_SCC_webster_et_al_S5.xlsx", col_names = T, sheet = 4, skip = 1)

FA_tumour_type <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_tumor_features.xlsx", col_names = T, skip = 1)

#Select tumours that are FANCA mutants
FA_tumour_type <- read_xlsx("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_tumor_features.xlsx", col_names = T, skip = 1) #Import
FA_tumour_type_cg <- FA_tumour_type %>% select(sample = "Individual ID Family/Patient/Sample", CD = "Complementation group") %>% na.omit() #Select complementation group
FANCA_mut_tumours <- FA_tumour_type_cg %>% filter(CD == "A") %>% pull(sample) %>% unique() #Pull FANCA mutants only
```


#Curate data tables: What are the features that I am interested (only select deletions)
```{r}
#Select only deletions
FA_mutation_del_dt <- FA_mutation_dt %>% filter(svclass == "deletion")
sporadic_mutation_del_dt <- sporadic_mutation_dt %>% filter(svclass == "deletion" & grepl("Neg", sample))
BRCA2_mutation_del_dt <- BRCA2_mutation_dt %>% filter(svclass == "deletion")
BRCA1_mutation_del_dt <- BRCA1_mutation_dt %>% filter(svclass == "deletion")

#Deletion frequency
del_freq_dt <- c(nrow(sporadic_mutation_del_dt)/nrow(sporadic_mutation_dt), nrow(FA_mutation_del_dt)/nrow(FA_mutation_dt), nrow(BRCA2_mutation_del_dt)/nrow(BRCA2_mutation_dt), nrow(BRCA1_mutation_del_dt)/nrow(BRCA1_mutation_dt))
deletion_frequency_summary <- tibble(tumor_type = c("sporadic","FA","BRCA1mut","BRCA2mut"), del_freq = del_freq_dt)
```

#Data table select columns I will use in the analysis
```{r}
#select data table
FA_mutation_del_sel_dt <- FA_mutation_del_dt %>% select(chr1 = '# chr1', start1, end1, strand1 = strand1...9, chr2, start2, end2,strand2 = strand2...10, id = 'id/name', svclass,non_t = `non-template`,MH = `micro-homology`, sample)
sporadic_mutation_del_sel_dt <- sporadic_mutation_del_dt %>% select(chr1 = '# chr1', start1, end1, strand1 = strand1...9, chr2, start2, end2,strand2 = strand2...10, id = 'id/name', svclass,non_t = `non-template`,MH = `micro-homology`, sample)
BRCA2_mutation_del_sel_dt <- BRCA2_mutation_del_dt %>% select(chr1 = '# chr1', start1, end1, strand1 = strand1...9, chr2, start2, end2,strand2 = strand2...10, id = 'id/name', svclass,non_t = `non-template`,MH = `micro-homology`, sample)
BRCA1_mutation_del_sel_dt <- BRCA1_mutation_del_dt %>% select(chr1 = '# chr1', start1, end1, strand1 = strand1...9, chr2, start2, end2,strand2 = strand2...10, id = 'id/name', svclass,non_t = `non-template`,MH = `micro-homology`, sample)

#Put all dataframes together
all_tumour_deletions <- FA_mutation_del_sel_dt %>% filter(sample %in% FANCA_mut_tumours) %>% bind_rows(sporadic_mutation_del_sel_dt) %>% bind_rows(BRCA2_mutation_del_sel_dt) %>% bind_rows(BRCA1_mutation_del_sel_dt)
```

#Import LADs data
```{r}
# LAD domains
LAD_atlas <- as.data.frame(read.table("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/xv20220329_LAD_atlas_OSF.bed.gz",header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(LAD_atlas) <- c("chr","start","end","length","strand","LAD_class")

#LAD_length plot control
LAD_length_tibble <- LAD_atlas %>% mutate(LAD_length = end - start)

```

#Data create MH_length data_table where I extract only MH mutations
```{r}
all_tumour_deletions_clasified <- all_tumour_deletions %>% mutate(del_length = start2-start1, del_class = case_when(grepl("[A-Z]", MH) ~ "MH",grepl("[A-Z]", non_t) ~ "non_t", T ~ "none"), type = case_when(grepl("BRCA2", sample) ~ "BRCA2mut", grepl("BRCA1", sample) ~ "BRCA1mut",grepl("HNSCC", sample) ~ "spo", grepl("F",sample) ~ "FA"))

#summaryse data non T
mutations.summarized <- all_tumour_deletions_clasified %>% dplyr::group_by(del_class, type) %>% dplyr::summarise(FA_count = n())
```

#Check trend across all tumours: total amount of mutations
It fits with what the screen predicts
- BRCA2 (*) = favors NHEJ/ represses MMEJ in euchromatin
- BRCA1 (ns) = favors MMEJ in triple_het
- FANCA (ns = 0.07) = favors MMEJ in polycomb
Expected outcomes:
- BRCA2mut = Relatively more MH mutations in euchromatin (ciLADs) vs cLADs compared to sporadic tumours
- BRCA1mut = Relatively less MH mutation in LADs vs iLADs compared compared to sporadic tumours
- FANCAmut = Relatively less MH mutations in H3K27me3 vs iH3K27me3 compared to sporadic tumours
Measurement = log2 fc Domain/iDomain
- BRCA2mut = Down
- BRCA1mut = Down
- FANCAmut = Down


```{r}
#Prepare the data table for export
mutation_location_export <- all_tumour_deletions_clasified %>% mutate(mut_id = paste(sample, id, sep = "_"), chr = paste0("chr", chr1)) %>% select(chr, start = "start1", end = "start2", strand = "strand1" ,mut_id, del_class, del_length)
mutation_location_GR <- makeGRangesFromDataFrame(mutation_location_export, keep.extra.columns = T)

#Crate ranges with LADS
LAD_atlas_ranges <- makeGRangesFromDataFrame(LAD_atlas, keep.extra.columns = T)
LAD_overlaps <- findOverlaps(LAD_atlas_ranges,mutation_location_GR)
LAD_overlaps_dt <- tibble(mutation_location_export[subjectHits(LAD_overlaps),], LAD_status =  LAD_atlas[queryHits(LAD_overlaps),6])

#Put both together
mut_data_LADs <- LAD_overlaps_dt %>% mutate(type = case_when(grepl("BRCA2", mut_id) ~ "BRCA2mut", grepl("BRCA1", mut_id) ~ "BRCA1mut",grepl("HNSCC", mut_id) ~ "spo", grepl("F",mut_id) ~ "FA"))  %>% separate(mut_id, sep = "_(?=[1-9])", into = c("sample","id"))
mut_data_LADs$id  <- as.double(mut_data_LADs$id)

# Some deletions are very big and span over LADs and iLADs (filter only deletions that occur within the same iLAD or LAD)
LAD_counts <- mut_data_LADs %>% dplyr::group_by(sample, id) %>% dplyr::summarize(count = n())
single_LADs_iLADs <-LAD_counts %>% filter(count == 1) %>% select(sample, id)

#Filter only FANCA tumours
sim.chrom.status.cd <- mut_data_LADs %>% right_join(single_LADs_iLADs) %>% left_join(FA_tumour_type_cg %>% distinct()) %>% mutate(type_cd = paste(type, CD, sep = "_")) %>% filter(type != "FA" | type_cd == "FA_A")
```

```{r}
# Export data
saveRDS(sim.chrom.status.cd, file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/processed_data/xv20220530_SCC_SV_MH_deletions.rds")
```

#All mutations chromatin environment independent
```{r}
#Total amount of mutations
ggplot(sim.chrom.status.cd) + geom_bar(aes(fct_relevel(type, c("spo")), fill = del_class)) + theme_bw()

# Normalize mutations per sample
mutations_tumour_norm <- sim.chrom.status.cd %>% dplyr::group_by(sample,type,del_class) %>% dplyr::summarise(count = n()) %>% ungroup() %>% dplyr::group_by(type,del_class) %>% dplyr::summarise(mean_c = mean(count))

# 
ggplot(mutations_tumour_norm) + geom_col(aes(fct_relevel(type, c("spo")),mean_c, fill = del_class)) + theme_bw() + ylab("Mutations per tumour") + xlab("Tumor class")

ggplot(mutations_tumour_norm %>% filter(del_class == "MH")) + geom_col(aes(fct_relevel(type, c("spo")),mean_c)) + theme_bw() + ylab("Mutations per tumour") + xlab("Tumor class")

#Plot split by LAD and iLAD
LAD.data.mut <- sim.chrom.status.cd %>% dplyr::group_by(type,del_class, LAD_status) %>% dplyr::summarise(count = n()) %>% left_join(mut_per_tumour) %>% rowwise() %>% mutate(mean_c_LAD = count/s_count) %>% ungroup()
LAD.data.mut.filtered <- LAD.data.mut %>% filter(LAD_status %in% c("ciLAD","cLAD") & del_class == "MH" & HPV == FALSE)
LAD.data.mut.filtered_del_class <- LAD.data.mut %>% filter(del_class == "MH")

ggplot(LAD.data.mut.filtered) + geom_col(aes(LAD_status, mean_c_LAD, fill = LAD_status)) + theme_bw() + ylab("Mutations per tumour") + xlab("Tumor class") + facet_wrap(~ type)


```

```{r}
# Make two plots: one for LADs & one for iLADS (total amount of mutations)

# Same with only HPV negative
sim.chrom.status.cd.hpv <- sim.chrom.status.cd %>% rowwise() %>% mutate(HPV = grepl("HPVPos", sample))
tumour.n <- sim.chrom.status.cd.hpv %>% dplyr::group_by(type) %>% dplyr::summarise(sample_n = length(unique(sample)))
plot.tumour.LAD <-  sim.chrom.status.cd.hpv %>% dplyr::group_by(type, LAD_status) %>% dplyr::summarise(mut_n = n()) %>% left_join(tumour.n, by = "type") %>% mutate(mut_tumour = mut_n/sample_n)

ggplot(plot.tumour.LAD %>% filter(LAD_status %in% c("cLAD","ciLAD"))) + geom_col(aes(x = LAD_status, y = mut_tumour, fill = LAD_status)) + facet_wrap(~fct_relevel(type, c("spo","BRCA2mut","BRCA1mut","FA")), scales = "free_y") + theme_bw() 

ggplot(sim.chrom.status.cd.hpv %>% filter(del_class == "MH" & !grepl("f",LAD_status))) + geom_bar(aes(x = LAD_status, fill = LAD_status)) + facet_wrap(~type) + theme(axis.text.x = element_text(angle = 90)) + theme_bw() 


#LADs (with all mutations)
ggplot(sim.chrom.status.cd %>% filter(del_class == "MH"  & !grepl("f",LAD_status))) + geom_bar(aes(x = LAD_status, fill = del_class)) + facet_wrap(~type)

#fold change differences (K27_me3)
fc.LAD <- sim.chrom.status.cd %>% filter(del_class == "MH") %>% dplyr::group_by(type, LAD_status) %>% dplyr::summarise(count = n()) %>% dcast(type ~ LAD_status) %>% mutate(LAD_iLAD_ratio = log2(cLAD/ciLAD))

#K27me3 (There's a subtle trend in total amount of mutations)
ggplot(sim.chrom.status.cd %>% filter(del_class == "MH" & !grepl("f",H3K27me3_status))) + geom_bar(aes(x = H3K27me3_status, fill = del_class)) + facet_wrap(~type) + theme(axis.text.x = element_text(angle = 90)) + theme_bw() + theme_bw()

#fold change differences (K27_me3)
fc.K27 <- sim.chrom.status.cd %>% filter(del_class == "MH") %>% dplyr::group_by(type, H3K27me3_status) %>% dplyr::summarise(count = n()) %>% dcast(type ~ H3K27me3_status) %>% mutate(K27_iK27_ratio = log2(cH3K27me3/ciH3K27me3))

```
Results (log2 fold change D/iD): 
- BRCA2mut: -0.19 (mut) vs 1.03 (spo) => Rel. more MH mut in iLADs (big diff., ~ 1.2)
- BRCA1mut: 0.29 (mut) vs 1.03 (spo) => Rel. less MH mut in LADs (not so big, ~ 0.75)
- FANCAmut: -0.2 (mut) vs -0.34 (spo) => Rel. more MH mut in K27me3 (minimal diff, ~ 0.15)

# Do I have enough mutations to split by tumour
```{r}
# LAD data
mutations_per_K27 <- sim.chrom.status.cd %>% filter(del_class == "MH") %>% dplyr::group_by(type, H3K27me3_status, sample) %>% dplyr::summarise(mut_count = n())


```


# Check by tumour (is this trend consistent)
```{r}
# Count ratio per tumour sample (K27me3)
per_sample_muts_K27 <- sim.chrom.status.cd %>% filter(del_class == "MH") %>% dplyr::group_by(type, H3K27me3_status, sample) %>% dplyr::summarise(mut_count = n() + 1) %>% dcast(type + sample ~ H3K27me3_status)
per_sample_muts_K27[is.na(per_sample_muts_K27)] <- 1
per_sample_muts_K27_calc <- per_sample_muts_K27 %>% mutate(K27_iK27_ratio = log2(cH3K27me3/ciH3K27me3))


#How many mutations have 0 mutations on that domain (check amount of mutations)
mutation_number_tumour_K27 <- per_sample_muts_K27_calc %>% select(type, sample, cH3K27me3,ciH3K27me3) %>% melt(id.names = c("type","sample"))

#plot
ggplot(mutation_number_tumour_K27) + geom_quasirandom(aes(fct_relevel(variable,"ciH3K27me3"),log10(value))) + facet_wrap(~fct_relevel(type, "spo"), nrow = 1) + theme_bw() + ylab("log10(mutation counts + 1)")

#Plot H3K27me3
ggplot(per_sample_muts_K27_calc)  + geom_boxplot(aes(type, K27_iK27_ratio),outlier.shape = NA) + geom_quasirandom(aes(type, K27_iK27_ratio))

#Calculate stats (K27me3_K27me3)
mut_type_K27 <-per_sample_muts_K27 %>% filter(type != "spo") %>% pull(type) %>% unique()
stat_K27 <- map_dfr(mut_type, .f = function(x) {
  spo_mut <- per_sample_muts_K27_calc %>% filter(type == "spo") %>% pull(K27_iK27_ratio)
  mut_mut <- per_sample_muts_K27_calc %>% filter(type == x) %>% pull(K27_iK27_ratio)
  wilcox.test(mut_mut,spo_mut) %>% tidy() %>% mutate(type = x)
})
stat_K27_adj <- stat_K27 %>% mutate(p.adj = p.adjust(p.value))

# Count ratio per tumour sample (LADs)
per_sample_muts_LAD <- sim.chrom.status.cd %>% filter(del_class == "MH") %>% dplyr::group_by(type, LAD_status, sample) %>% dplyr::summarise(mut_count = n() + 1) %>% dcast(type + sample ~ LAD_status)
per_sample_muts_LAD[is.na(per_sample_muts_LAD)] <- 1
per_sample_muts_LAD_calc <- per_sample_muts_LAD %>% mutate(LAD_iLAD_ratio = log2(cLAD/ciLAD))

#How many mutations have 0 mutations on that domain (check amount of mutations)
mutation_number_tumour_LAD <- per_sample_muts_LAD_calc %>% select(type, sample, cLAD,ciLAD) %>% melt(id.names = c("type","sample"))

#plot
ggplot(mutation_number_tumour_LAD) + geom_quasirandom(aes(fct_relevel(variable,"ciLAD"),log10(value))) + facet_wrap(~fct_relevel(type, "spo"), nrow = 1) + theme_bw() + ylab("log10(mutation counts + 1)")



#Plot LAD vs iLAD
ggplot(per_sample_muts_LAD_calc)  + geom_boxplot(aes(fct_relevel(type, c("spo"), ), LAD_iLAD_ratio, fill = type), alpha = 0.5,outlier.shape = NA) + geom_quasirandom(aes(type, LAD_iLAD_ratio)) + theme_bw()

#Calculate stats (LAD_iLAD)
mut_type <-per_sample_muts_LAD %>% filter(type != "spo") %>% pull(type) %>% unique()
stat_LAD <- map_dfr(mut_type, .f = function(x) {
  spo_mut <- per_sample_muts_LAD_calc %>% filter(type == "spo") %>% pull(LAD_iLAD_ratio)
  mut_mut <- per_sample_muts_LAD_calc %>% filter(type == x) %>% pull(LAD_iLAD_ratio)
  wilcox.test(mut_mut,spo_mut) %>% tidy() %>% mutate(type = x)
})
stat_LAD_adj <- stat_LAD %>% mutate(p.adj = p.adjust(p.value))

```

#With this we can show that BRCA2 chromatin specific paterns are mantained