---
title: "xv20220712_S1_set_up_screening"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure S1 in the short version of the paper. This figure describes the layout of the screen and supplementaries to the screening set up. The main point of this figure is to build up confidence in our reporter cell line and describe how these pathways are build up. For the short version, I'll keep the script for supplementary figures separate. Figure outline:
A- Plate design
B- Correlation in log2 pathway balance
C- Positive control effect size


This script generates the plots for figure 1

```{r}
knitr::opts_chunk$set(root.dir = '/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  if (grepl(Date, file) == T) {
    write(c(document_name, as.character(sys_time),"Output", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(path, pat) {
  full.path <- paste0(in.dir, path)
  document_path <- rstudioapi::getSourceEditorContext()$path %>% str_split(pattern = "/") %>% unlist()
  document_name <- document_path[length(document_path)]
  sys_time <- Sys.time()
  
  if(file.exists(full.path) == T){
    if(paste0(in.dir,path) != getwd()){
      setwd(paste0(in.dir, path))
  }
    file.list <- list.files(pattern = pat)
    correct_file <- str_sort(file.list, decreasing = T)[1]
    print(correct_file)
    write(c(document_name,as.character(sys_time) ,"Input", correct_file), file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
    readRDS(correct_file)
    
  }
  else{
  setwd(path)
  file.list <- list.files(pattern = pat)
  correct_file <- str_sort(file.list, decreasing = T)[1]
  print(correct_file)
  write(c(document_name, as.character(sys_time),"Input", correct_file),file = paste0(in.dir,"log"), ncolumns = 4, sep = "\t", append = T)
  readRDS(correct_file)
  }
}
```


# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
```



# Import data tables
```{r}
setwd(in.dir)
#Clone 5 chromatin Version cl20201026
clone5_z.score_chrom_tib <- readRDS_proof("/DATA/projects/DSBrepair/data/R","ChIP_zscore_selection")

#Import MMEJ:NHEJ balance ratio measurements per replicate
ddr.screen.all <- readRDS_proof("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data","screen_ddr_all_computations")
ddr.screen.all.sel <- ddr.screen.all %>% dplyr::select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep = bio.rep,t.rep = FinRep, freqCut) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#log2 fold change
log2_fold_change_IPR_KO <- readRDS_proof( "data/processed_data","fold_change_ratio_KO")
log2_fold_change_IPR_WT <- readRDS_proof( "data/processed_data/CCD_analysis","fold_change_ratio_WT")
log2_fold_change_IPR_POLQ <- readRDS_proof( "data/processed_data/CCD_analysis","fold_change_ratio_POLQ")


# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

clone5_bc_cluster <-  c("CATCCACCACACTTCA","ATATCGTTGCTGGAGA","GTACCTCTCGATAGTG","TCTTTTGAGGAGCTGA","CATTTCTGATCAATAA","TGGCCAATATTTGTCT","ACTGTCGAGTTGTCCG","GAGCGCGTCACCGGGT","GCGCACCCTTTAATTG","ACCCCTAAAGGCGCTG","CGGCCTGAAGGTCAGG","AGAAAATAATATGACG","TTGAACGCGGGCTCGG","CCGGGGACGTATGCAC","GCTAACATCACGAATC","ATACTATATTTAACGG","TATGGCTGTCGGGTAG","AGGGCGTAAAATATTT","TGTCCCTTAGTACTTT")
```

#Supplementary Figures

#Figure S1A: Plate design with controls

# Figure S1B: Correlation MMEJ:NHEJ balance
```{r}
#Compute the log2 MMEJ:NHEJ balance: Initial dataset
log2.ratio.screen.detail.ddr <- ddr.screen.all.sel %>% dplyr::select(t.rep, MMEJNHEJratio,library, gene,barcode,sample,well,plate) %>% mutate(log2ratio = log2(MMEJNHEJratio))

#dcast data table for plotting reproducibility of the ratio
balance.ratio.list <- log2.ratio.screen.detail.ddr %>% reshape2::dcast(gene + library + barcode + sample + well + plate ~ t.rep, value.var = "log2ratio") %>% arrange(sample)

# Plot reproducibility DNA_repair library
#R1 vs. R2 (Fig. S1B)
ggplot(balance.ratio.list %>% filter(library == "DNA_repair")) + 
  geom_point(aes(R1,R4, color = fct_relevel(sample,c("WT","POLQ"))))+ 
  geom_smooth(aes(R1,R4), method = "lm")+ 
  xlab("R1 (log2 MMEJ:NHEJ balance))") + ylab("R2 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R1,R4), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))

#R1 vs. R5 (Fig. S1C)
ggplot(balance.ratio.list %>% filter(library == "DNA_repair")) + 
  geom_point(aes(R1,R5, color = fct_relevel(sample,c("WT","POLQ"))))+ 
  geom_smooth(aes(R1,R5), method = "lm")+ 
  xlab("R1 (log2 MMEJ:NHEJ balance))") + ylab("R3 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R1,R5), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))

#R4 vs. R5 (Fig. S1D)
ggplot(balance.ratio.list %>% filter(library == "DNA_repair")) + 
  geom_point(aes(R4,R5, color = fct_relevel(sample,c("WT","POLQ","KO"))))+ 
  geom_smooth(aes(R4,R5), method = "lm", color = "red")+ 
  xlab("R2 (log2 MMEJ:NHEJ balance))") + ylab("R3 (log2 MMEJ:NHEJ balance)") + 
  theme_bw() + stat_cor(aes(R4,R5), label.y = 1.5, label.x = -4.25) +
  scale_color_manual(values = c("#949DCF","#465584","#DADADA")) +
  theme(legend.position = "top") + coord_fixed(xlim = c(-4.5,2), ylim = c(-4.5,2))


```


#Figure S1E: log2 fold change effects of controls and KO (dynamic range)

```{r}
log2_fold_change_IPR_POLQ_mean <- log2_fold_change_IPR_POLQ %>% dplyr::group_by(library,gene,barcode,sample,well,plate) %>% dplyr::summarise(mean.log2foldchange = mean(log2.foldchange, na.rm = T))

log2_fold_change_IPR_WT_mean <- log2_fold_change_IPR_WT %>% dplyr::group_by(library,gene,barcode,sample,well,plate) %>% dplyr::summarise(mean.log2foldchange = mean(log2.foldchange, na.rm = T))

log2_fold_change_IPR_bind <- log2_fold_change_IPR_KO %>% bind_rows(log2_fold_change_IPR_POLQ_mean, log2_fold_change_IPR_WT_mean)

#Plot log2 fold changes
ggplot(log2_fold_change_IPR_bind %>% arrange(sample)) +
  geom_quasirandom(aes(fct_relevel(barcode,clone5_bc_cluster),mean.log2foldchange, color = sample), dodge.width = 0.5) + theme_bw() + theme(axis.text.x = element_blank(), legend.position = "top") + scale_color_manual(values = c("#DADADA","#465584","#949DCF")) 

```
