---
title: "xv20211230_4_epistasis_model"
output: html_document
---

# Script to generate figure #4 in the paper. This figure describes how epistatic interactions are present between chromatin features and protein KOs. Figure outline:
A- Step by Step for KAT5 and H3K9me2
B- Matrix for all hits
C- Hits per library
D- Hits per pathway
E- Strong hits examples

Data in text:
A-...
B-...

Supplementary figure 5:
A- All proteins matrix
B- Balance perturbation strong hits

This script generates the plots for figure 4 and S5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)

```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128b_ratio_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_kinase_inhib_epistasis.rds")
```
# Figure 2A: Mock examples
```{r}
mock.slopes <- tibble(z_score = seq(-3,3,0.25), C = z_score*0 + 0.6,A = z_score*0.2 + 0.6,B = z_score*0.1 +0.6, D = z_score*(-0.1)+0.6, E = z_score*-0.1 - 0.6) %>% reshape2::melt(id.vars = "z_score")

ggplot(mock.slopes, aes(z_score,value,color = variable)) + geom_smooth() + theme_bw() + geom_vline(xintercept = 0, linetype = 2) + geom_hline(yintercept = 0, linetype = 2) + ylab("log2(fold change NHEJ:MMEJ)") + xlab("Feature signal (z-score)") + theme(legend.position = "none") + coord_cartesian(ylim= c(-1.5,1.5)) + scale_color_manual(values = c("black","#01665e","#80cdc1","grey70","#8c510a"))


bar.mock <- tibble(name = c("A","B","C","D","E"), slope = c(0.2,0.1,0,0,-0.1))

ggplot(bar.mock) + geom_col(aes(name,slope, fill = as.character(slope)))  + scale_fill_manual(values = rev(c("#01665e","#80cdc1","black","#8c510a")))+ theme_bw() + geom_hline(yintercept = 0, linetype = 2) + theme(legend.position = "none")
```

# Figure 2B-D: Examples
  #2B: Favors MMEJ - RAD50 & LMNB1
```{r}
# Plot RAD50 with LMNB1
ggplot(step2.epistasis %>% 
         filter(gene %in% c("RAD50")),
       aes(LMNB1,m.log2.ratio.dist)) + 
  geom_point() +
  geom_smooth(aes(color = 0.114),method = "lm") +
  stat_cor(label.x = -1.5, label.y = 1.4) +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  theme_bw() + ylab("log2(RAD50 FC NHEJ:MMEJ balance)") + theme(legend.position = "none")


```

#2C: No interaction - MDC1 and H2AFZ
```{r}
# Plot RAD50 with LMNB1
ggplot(step2.epistasis %>% 
         filter(gene == "MDC1"),
       aes(H2AFZ,m.log2.ratio.dist)) + 
  geom_point(color = "black") +
  geom_smooth(method = "lm", color = "black") +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.15,0.15)) +
  stat_cor()+
  theme_bw() + ylab("log2(MDC1 NHEJ:MMEJ balance)")
```
# Protein that favors NHEJ - POLL and H3K4me1
```{r}
# Plot POLL with H3K36me3
ggplot(step2.epistasis %>% 
         filter(gene %in% c("POLL")),
       aes(H3K36me3,m.log2.ratio.dist, color = -0.07)) + 
  geom_point(color = "black") +
  geom_smooth(method = "lm") +
  stat_cor(color = "black") +
  scale_color_gradient2(low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.1,0.15)) +
  theme_bw() + ylab("log2(FC NHEJ:MMEJ balance)") + theme(legend.position = "none")
```





# Figure 4B: Matrix for hits
```{r}
#genes to select
IPR.significance <- filter(per_19IPR_data, abs(IPR.z.score) > 1.95) %>% dplyr::group_by(gene) %>% mutate(significant.pert = case_when(IPR.z.score < 0 ~ "MMEJ", IPR.z.score > 0 ~ "NHEJ")) %>% select(gene,significant.pert) %>% distinct() %>% reshape2::dcast(gene ~ significant.pert) %>% mutate(signif.pert = case_when(NHEJ == "NHEJ" & is.na(MMEJ) ~ "NHEJ", MMEJ == "MMEJ" & is.na(NHEJ) ~ "MMEJ", T ~ "both")) %>% select(gene,signif.pert)

# Filter slopes
selected.slopes <-  slope.protein.features.all %>% left_join(IPR.significance, by = "gene") %>% na.omit() %>% mutate(epistasis.final.b = case_when(signif.pert == "MMEJ" & epistasis.final < 0 ~ 0, signif.pert == "NHEJ" & epistasis.final > 0 ~ 0, T ~ epistasis.final))

tmp.gene.hits <- filter(test.log2.dist, p.adj < 0.0001) %>% pull(gene)

hit.genes <- filter(hits.both.screens, pathway != "NA") %>% separate(ID_gene, into = "gene") %>% pull(gene) %>% unique()

#Select 200 genes with highest differences
big.rearrangement.2 <- selected.slopes %>% 
  filter(epistasis.final != 0) %>%
  dplyr::group_by(gene) %>%
  dplyr::summarise(delta = max(slope) - min(slope),
            stdev = sd(slope),
            highest = max(abs(epistasis.final)),
            med.interaction = median(abs(epistasis.final), na.rm = T)) %>%
  slice_max(med.interaction, n = 200) %>%
  pull(gene)

# dcaast with selected
slope.prot.features.dcast.hits <- selected.slopes %>% filter(gene %in% big.rearrangement.2) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.hits <- pheatmap(slope.prot.features.dcast.hits, silent = F)
top200 <- dist(slope.prot.features.dcast.hits)
row_dend = hclust(top200)
row_dend.hits <- as.dendrogram(row_dend)
row_dend.hits <- seriate_dendrogram(row_dend.hits, top200, method="OLO")

heatmap.gene.order.slope.diff.hits <- rownames(slope.prot.features.dcast.hits[as.hclust(row_dend.hits)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])

# This are the effects

# heatmap
ggplot(selected.slopes %>% filter(gene %in% big.rearrangement.2)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.hits),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + coord_fixed(expand = F,ratio = 2)

ggplot(slope.protein.features.all %>% filter(gene %in% big.rearrangement.2)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.hits),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 2)

# Dendrogram
ggdendrogram(heatmap.slope.diff.hits$tree_row)


```
#split by positive and negative interactions
```{r}
# Filter slopes
selected.slopes <-  slope.protein.features.all %>% left_join(IPR.significance, by = "gene") %>% na.omit() %>% mutate(epistasis.final.b = case_when(signif.pert == "MMEJ" & epistasis.final < 0 ~ 0, signif.pert == "NHEJ" & epistasis.final > 0 ~ 0, T ~ epistasis.final))

# How many genes have interactions in NHEJ, MMEJ or both
interaction.summary <- selected.slopes %>% dplyr::group_by(gene) %>% dplyr::summarise(mmej.ep = sum(epistasis.final.b > 0), nhej.ep = sum(epistasis.final.b < 0)) %>% mutate(gene.class = case_when(mmej.ep == 0 & nhej.ep > 0 ~ "all.nhej",nhej.ep == 0 & mmej.ep > 0 ~ "all.mmej", nhej.ep > 0 & mmej.ep > 0 ~ "both.ep", T ~ "no.ep"))

#slope.with.gene.feature
gene.f.slope.protein <- selected.slopes %>% left_join(interaction.summary)

epistasis.summary <- selected.slopes %>% 
  left_join(interaction.summary) %>%
  filter(epistasis.final.b != 0) %>%
  dplyr::group_by(gene,gene.class) %>%
  dplyr::summarise(delta = max(epistasis.final.b) - min(epistasis.final.b),
            stdev = sd(epistasis.final.b),
            highest = max(abs(epistasis.final.b)),
            med.interaction = median(abs(epistasis.final.b), na.rm = T)) 


#Summary for all.nhej
all.nhej.matrix <- epistasis.summary %>% filter(gene.class == "all.nhej") %>% arrange(desc(med.interaction))
nhej.genes.highlight <- all.nhej.matrix[1:100,] %>% pull(gene)

#Summary for all.nhej
all.mmej.matrix <- epistasis.summary %>% filter(gene.class == "all.mmej") %>% arrange(desc(med.interaction))
mmej.genes.highlight <- all.mmej.matrix[1:100,] %>% pull(gene)

#Summary for all mixed
mixed.matrix <- epistasis.summary %>% filter(gene.class == "both.ep") %>% arrange(desc(med.interaction))
mix.genes.highlight <- mixed.matrix[1:43,] %>% pull(gene)

```

# Number of genes per pathway
Filter applied: Gene contains at least one synergy > 0.03
```{r}
# Gene clasification per library
gene_library <- both.screen.detail %>% select(gene,library) %>% distinct() %>% dplyr::group_by(gene) %>% dplyr::summarise(librar = case_when(n() == 1 ~ library, n() == 2 ~ "both")) %>% distinct()

# genes.all.NHEJ.MMEJ
all.same.summary <- bind_rows(all.mmej.matrix,all.nhej.matrix) %>% filter(abs(highest) > 0.05) %>% select(gene,gene.class) %>% left_join(gene_library) %>% ungroup() %>% dplyr::group_by(librar,gene.class) %>% dplyr::summarise(counts = n())


# Plot
ggplot(all.same.summary) + geom_col(aes(fct_relevel(gene.class, c("all.nhej","all.mmej")), counts, fill = fct_relevel(librar,c("DNA_repair","both","Chromatin"))))  + theme_bw() + coord_cartesian(expand = T) + scale_fill_manual(values = c("#56B3E6","grey40","#E69F03")) + theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + theme(legend.position = "top") + labs(fill = "CRISPR library")

```


```{r}
# dcaast with selected
slope.prot.features.dcast.mmej <- gene.f.slope.protein %>% filter(gene %in% mmej.genes.highlight) %>% reshape2::dcast(gene ~ fct_relevel(feature,heatmap.chromatin.order), value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")

#Heatmap
mmej.heatmap.proteins <- pheatmap(t(slope.prot.features.dcast.mmej), silent = F, clustering_method = "ward.D2", cluster_rows = F)


mmej.matrix.gene.order <- rownames(slope.prot.features.dcast.mmej[mmej.heatmap.proteins$tree_col[["order"]],])

# Both sides
ggplot(selected.slopes %>% left_join(interaction.summary, by = "gene") %>% filter(gene %in% mmej.genes.highlight)) + 
  geom_tile(aes(fct_relevel(gene,mmej.matrix.gene.order),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final.b)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(ratio = 1)

#Dendrogram
ggdendrogram(mmej.heatmap.proteins$tree_col, labels = F)

```

```{r}
# dcaast with selected
slope.prot.features.dcast.nhej <- gene.f.slope.protein %>% filter(gene %in% nhej.genes.highlight) %>% reshape2::dcast(gene ~ fct_relevel(feature,heatmap.chromatin.order), value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.nhej <- pheatmap(t(slope.prot.features.dcast.nhej), silent = F, clustering_method = "ward.D2", cluster_rows = F)

heatmap.gene.order.slope.diff.nhej <- rownames(slope.prot.features.dcast.nhej[heatmap.slope.diff.nhej$tree_col[["order"]],])

# Both sides
ggplot(gene.f.slope.protein %>% left_join(interaction.summary, by = "gene") %>% filter(gene %in% nhej.genes.highlight)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.nhej),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final.b)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top",axis.title = element_blank()) + coord_fixed(ratio = 1)

#Dendrogram
ggdendrogram(heatmap.slope.diff.nhej $tree_col, labels = F)

```

```{r}
# dcast with selected
slope.prot.features.dcast.both <- gene.f.slope.protein %>% filter(gene %in% mix.genes.highlight) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.both <- pheatmap(t(slope.prot.features.dcast.both), silent = F, clustering_method = "ward.D", cluster_rows = F)


heatmap.gene.order.slope.diff.both <- rownames(slope.prot.features.dcast.both[heatmap.slope.diff.both$tree_col[["order"]],])


# Both sides
ggplot(selected.slopes %>% left_join(interaction.summary, by = "gene") %>% filter(gene %in% mix.genes.highlight)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.both),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final.b)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 0.5), legend.position = "top")

#Dendrogram
ggdendrogram(heatmap.slope.diff.both $tree_col, labels = F)


```

# Number in text

#NOTE: This will change, but every number that I refer in the text will available in this chunk
```{r}
```

#Supplementary figures

#Figure S5A: slope with all proteins
```{r}
# dcast all genes
slope.prot.features.dcast.all <- slope.protein.features.all %>% reshape2::dcast(gene ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.all <- pheatmap(slope.prot.features.dcast.all, silent = T)

heatmap.gene.order.slope.diff.all <- rownames(slope.prot.features.dcast.all[heatmap.slope.diff.all$tree_row[["order"]],])
heatmap.feature.order.slope.diff.all <- colnames(slope.prot.features.dcast.all[,heatmap.slope.diff.all$tree_col[["order"]]])
# heatmap
ggplot(slope.protein.features.all) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.all),fct_relevel(feature, heatmap.feature.order.slope.diff.all), fill = epistasis)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + 
  theme_bw() + theme(axis.text.x = element_blank(), axis.title.y = element_blank(), legend.position = "top",) + 
  coord_fixed(expand = F,ratio = 20)

# Dendrogram
ggdendrogram(heatmap.slope.diff.all$tree_row, labels = F)
```

#Figure S5B: genes in plot
```{r}
# Filter proteins highlighted 
ggplot(per_19IPR_data %>% filter(gene %in% c("POLL","RAD50","POLQ"))) +
  geom_quasirandom(aes(gene,IPR.z.score)) +
  geom_hline(yintercept = 0, linetype =2) +
  theme_bw()
```

#Figure S5C: Slope is correlated with perturbation
```{r}
#
effect.size <- slope.protein.features.all %>% filter(complete.cases(.) & gene %in% hit.genes)
# Plot slope vs effect size
summ.slope.library <- effect.size %>% dplyr::group_by(gene) %>% dplyr::summarise(med.slope = mean(epistasis, na.rm = T), mean.intercept = mean(intercept))

ggplot(summ.slope.library, aes(med.slope, mean.intercept)) + geom_point() + stat_cor(method = "spearman",cor.coef.name = "rho") + geom_smooth(method = "lm") +
  theme_bw()

```
