---
title: "xv20220117_1_General_effects"
author: "x.vergara"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Script to generate figure #1 in the short version of the paper. This figure describes the layout of the screen and general effects. The main point of this figure is to build up confidence in our reporter cell line and describe how these pathways are build up. For the short version, I'll keep the script for supplementary figures separate. Figure outline:
A- Cartoon explaining screen
B- Chromatin landscape where 19IPRs are integrated
C- Volcano plot with all perturbations (Only for DNA repair proteins)


This script generates the plots for figure 1

```{r}
knitr::opts_chunk$set(root.dir = '/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/rmarkdown/xv20210920_paper_scripts/short_version/knit_html/')

StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

#Change in.dir accordingly
in.dir <- "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/"

#Check if output contain datetag
saveRDS_proof <- function(object, file) {
  if (grepl(Date, file) == T) {
    saveRDS(object, file)
  }
  else {
    print("Incorrect datetag!")
  }
}

#Read rds check through different versions
readRDS_proof <- function(pattern) {
  file_name <- paste0("\d{8}_",file)
  matching_files <- list_files(pattern = "\d{8}_" & file_name)
  if (lenght(matching_files) == 1) {
    readRDS
  }
}
```


# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(rstatix)
library(stats)
library(tibble)
library(reshape2)
```



# Import data tables
```{r}
setwd(in.dir)
#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#log2 fold change
log2_fold_change_IPR <- readRDS( "data/processed_data/xv20220624_fold_change_ratio_KO.rds")

# heatmap chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")
```

#Main figures

#Figure 1A: Cartoon

# Figure 1C: Epigenetic features
```{r}
#Extract barcodes in DR screen
clone5_bc <- log2_fold_change_IPR %>% pull(barcode) %>% unique()
# Prepare data frame
clone_5_IPR <- filter(clone5_z.score_chrom_tib, ID %in% clone5_bc) %>% select(-pool,-binsize) %>% column_to_rownames(var = "ID")

#pheatmap
clone_5_heatmap <- pheatmap(clone_5_IPR, silent = T)

#Cluster IPRs by features
clone5_bc_cluster <- rownames(clone_5_IPR[clone_5_heatmap$tree_row[["order"]],])

#dt for plotting
clone_5_IPR_melt <- clone_5_IPR %>% rownames_to_column(var = "barcode") %>% melt(value.name = "chrom.z.score", var.name = "feature")

#Plot
ggplot(clone_5_IPR_melt) + 
  geom_tile(aes(fct_relevel(barcode, clone5_bc_cluster),fct_relevel(variable,heatmap.chromatin.order), fill = chrom.z.score)) +
  scale_fill_gradient2( low = "#009B9E",mid = "#F1F1F1", high = "#C75DAB")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 0.75)

```


#Figure 1B: Volcano plot only for DNA_repair proteins

```{r}
#Average replicates
log2_fold_change_IPR_mean <- log2_fold_change_IPR %>% group_by(gene, library, barcode,sample, well, plate) %>% dplyr::summarise(mean.log2.foldchange = mean(log2.foldchange, na.rm = T))

#Compute mean log2 fold change
mean.log2.fc <- log2_fold_change_IPR_mean %>% group_by(gene) %>% dplyr::summarise(mean.fc = mean(mean.log2.foldchange, na.rm = T))

#Run t.test with all the data
test.log2.dist <- log2_fold_change_IPR_mean %>% dplyr::group_by(gene) %>% t_test(mean.log2.foldchange ~ 0) %>% mutate(p.adj = p.adjust(p, method = "fdr")) %>% left_join(mean.log2.fc, by = "gene")

# hits mmej
hits.nhej.fdr <- test.log2.dist %>% filter(p.adj < 0.001 & mean.fc > 0) %>% pull(gene)

# hits nhej
hits.mmej.fdr <- test.log2.dist %>% filter(p.adj < 0.001 & mean.fc < 0) %>% pull(gene)

# Highlight three complexes
FA_core_complex <- tibble(gene = c("FANCL","FANCA","FANCB","FANCC","FANCE","FANCF","FANCG","FANCM","FAAP24","FAAP100","FANCD2","FANCI"), complex = "FA_proteins")
MRN_complex <- tibble(gene = c("MRE11","RAD50","NBN"), complex = "MRN_complex")
DNA_ligase_complex <- tibble(gene = c("LIG4","XRCC4","NHEJ1"), complex = "DNA_ligase_complex")

#Complexes to be highlighted in text
highlight.complex <- bind_rows(FA_core_complex,DNA_ligase_complex,MRN_complex)

#Proteins highlighted in the text
prots.highlighted <- c("POLL","LIG4","NHEJ1","MRE11","NBN","RBBP8","POLQ","RAD50","FANCA","RBX1","FANCF","FANCM","SHLD1")

# Plot and highlight main findings
ggplot(test.log2.dist %>% 
         left_join(highlight.complex)) +
  geom_point(aes(mean.fc,-log10(p.adj)), color = "grey40") + 
  geom_point(data = test.log2.dist %>% 
               left_join(highlight.complex) %>% 
               na.omit(),
             aes(mean.fc,-log10(p.adj), color = complex), size = 2) + theme_bw() + 
  geom_text_repel(data = subset(test.log2.dist,gene %in% prots.highlighted),
                  aes(mean.fc,-log10(p.adj), label = gene),max.overlaps = Inf, box.padding = 0.5) + geom_vline(xintercept = 0, linetype = 2) + geom_hline(yintercept = -log10(0.001), linetype = 2) + ylab("-log10(FDR)") + xlab("log2(Average fold change MMEJ:NHEJ balance)") + theme(legend.position = "top") + scale_x_continuous(breaks = c(-2,-1,0,1)) + scale_color_manual( values = c("#ca0020","#92c5de","#0571b0"))

```
