---
title: "xv20211230_5_epistasis_inhibitors"
output: html_document
---
# Script to generate figure #5 in the paper. This figure shows that epistatic interactions can be bigger if drugs are used instead of KO. Figure outline:
A- Effect of drug treatment in MMEJ:NHEJ balance
B- Matrix for all treatments
C- Slopes are bigger in inhibition than KO
D- PARP1 and olaparib correlation
E- ATM and ATMi correlation

Data in text:
A-...
B-...

Supplementary figure 6:
A- PARP1 vs. veliparib
B- PARP1 vs. talazoparib
C- ATRi vs. VE-821
D- ATRi vs. VE-822

This script generates the plots for figure 5 and S6

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(ggdendro)
```


# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220119_ratio_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128_ratio_kinase_inhib_epistasis.rds")

# Chromatin follow up experiment
epistasis.ratio.followup <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128_ratio_epistasis_followup.rds")

```

# Figure 3A: Chromatin vs. DNA repair 
```{r}
# For this we select
module_I_tripl_het_features <- tibble(feature = c("late_replicating","H3K9me3","H3K9me2","LMNB1"), module = "triple_het")
module_II_polycomb_features <- tibble(feature = c("EZH2","H3K27me3"), module = "polycomb")
module_III_euchr_features <- tibble(feature = c("H3K4me1","H3K27ac","H3K4me2","H3K4me3","H4K5acK8ac","TTseq","POL2","POL2AS2"), module = "euchromatin")
# Module summary
module_summary <- bind_rows(module_I_tripl_het_features,module_II_polycomb_features, module_III_euchr_features)


# First plot: what library
gene_library <- both.screen.detail %>% select(gene,library) %>% distinct() %>% dplyr::group_by(gene) %>% dplyr::summarise(librar = case_when(n() == 1 ~ library, n() == 2 ~ "both")) %>% distinct()

#Left_join data
module.summary <- slope.protein.features.all %>% left_join(module_summary, by = "feature") %>% dplyr::group_by(module,gene) %>% dplyr::summarise(mean = mean(epistasis.final, na.rm = T)) %>% left_join(gene_library, by = "gene")
module_III_genes <- module.summary %>% filter(module == "euchromatin" & mean > 0.02)
module_II_genes <- module.summary %>% filter(module == "polycomb" & mean > 0.02)
module_I_genes <- module.summary %>% filter(module == "triple_het" & mean > 0.02)

#
ggplot(module.summary %>% filter(mean > 0.02) %>% na.omit() %>% ungroup() %>% select(librar,gene,module)  %>% distinct() %>% group_by(module) %>% summarise(counts = n())) + geom_col(aes(module, counts)) + theme_bw() + coord_cartesian()

# 
ggplot(module.summary %>% filter(mean > 0.02) %>% na.omit() %>% ungroup() %>% select(librar,gene,module)  %>% distinct() %>% group_by(librar,module) %>% summarise(counts = n())) + geom_col(aes(module, counts, fill = fct_relevel(librar,c("DNA_repair","both","Chromatin"))), position = "fill") + theme_bw() + coord_cartesian(expand = F) + theme(legend.position = "top")
```
# CORUM database
```{r}
#Triple heterochromatin
module_I_gene_set <- module_I_genes %>% pull(gene) %>% unique()
module_I_gost <- gost(module_I_gene_set, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_I_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_")) %>% filter(p_val > 2)) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")


```
# Figure 4A: Complexes enriched in module III
```{r}
res.complex.genes <- c("BARD1","BRCA1","RBBP8","RAD50","TOPBP1","MRE11","NBN","BACH1","FANCD2","ATM","UMC1")

# dcaast with selected
slope.prot.features.dcast.res <- gene.f.slope.protein %>% filter(gene %in% res.complex.genes & feature %in% mmej_module_III_tripl_het_features$feature) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.res <- pheatmap(slope.prot.features.dcast.res, silent = F)
res <- dist(slope.prot.features.dcast.res)
row_dend.res = hclust(res)
row_dend.hits.res <- as.dendrogram(row_dend.res)
row_dend.hits.res <- seriate_dendrogram(row_dend.hits.res, res, method="OLO")

heatmap.gene.order.slope.diff.res<- rownames(slope.prot.features.dcast.res[as.hclust(row_dend.hits.res)[["order"]],])


# Both sides
ggplot(selected.slopes %>% filter(gene %in% res.complex.genes)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.res),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

```

#MLL3/4 complex in screening
```{r}
mll3_4_complex <- c("KMT2C","KMT2D","KDM6A","RBBP5","ASH2L","WDR5","DPY30","PAXIP1","PAGR1","NCOA6")

#Cluster by genes
mll3_4_dcast  <- selected.slopes %>% filter(gene %in% mll3_4_complex) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")
mll.prots.heatmap <- pheatmap(mll3_4_dcast,method = "ward.D2", silent = T)
mll.prots.clustering <- rownames(mll3_4_dcast[mll.prots.heatmap$tree_row[["order"]],])

# Both sides
ggplot(selected.slopes %>% filter(gene %in% mll3_4_complex)) + 
  geom_tile(aes(fct_relevel(gene,mll.prots.clustering),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

```
# Followup experiment: WRAD complex/RBBP5


## ALL MLL3/4 proteins

# Follow up experiment: RBBP5 (FDR < 0.1 = H3K9me2,H3K9me3 and LMNB1)
```{r}

#Select RBBP5 only
RBBP5.followup <- epistasis.ratio.followup %>% filter(gene == "RBBP5")

ggplot(RBBP5.followup, aes(fct_relevel(gRNA, c("Screen","Combo")),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits))) +
  geom_tile(aes(fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

#The whole WRAD complex
ASCOM.followup <- epistasis.ratio.followup %>% filter(gene %in% c("ASH2L","DPY30","RBBP5","WDR5")) %>% dplyr::select(gene,feature, epistasis.final,gRNA) %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_"))

ASCOM.combined.dcast <- ASCOM.followup %>% dcast(gene_gRNA ~ feature, value.var = "epistasis.final") %>% column_to_rownames("gene_gRNA") 

ASCOM.heatmap <- pheatmap(t(ASCOM.combined.dcast))
ASCOM.dist <- dist(ASCOM.combined.dcast)
row_dend.ascom = hclust(ASCOM.dist)
row_dend.hits.ascom <- as.dendrogram(row_dend.ascom)
row_dend.hits.ascom <- seriate_dendrogram(row_dend.hits.ascom,ASCOM.dist, method="OLO")



ascom.gene.order <- rownames(ASCOM.combined.dcast[as.hclust(row_dend.hits.ascom)[["order"]],])

ggplot(ASCOM.followup) +
  geom_tile(aes(fct_relevel(gene_gRNA, ascom.gene.order), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final)) + scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

```

# Follow up experiment: PAXIP1 (FDR < 0.1 all four)
```{r}

PAXIP1.screen <- slope.protein.features.all %>% 
  filter( gene == "PAXIP1") %>% mutate(gRNA = "Screen")

PAXIP1.followup <- epistasis.ratio.followup %>% filter(gene == "PAXIP1")

PAXIP1.combined <- bind_rows(PAXIP1.screen, PAXIP1.followup)

ggplot(PAXIP1.followup, aes(fct_relevel(gRNA, c("Screen","Combo")),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits))) +
  geom_tile(aes(fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# PTIP & PA1



```

# Follow up experiment: PA1 (FDR < 0.1 all four)
```{r}

PAXIP1.screen <- slope.protein.features.all %>% 
  filter( gene == "PA1") %>% mutate(gRNA = "Screen")

PA1.followup <- epistasis.ratio.followup %>% filter(gene == "PA1")

PAXIP1.combined <- bind_rows(PAXIP1.screen, PAXIP1.followup)

ggplot(PAXIP1.followup, aes(fct_relevel(gRNA, c("Screen","Combo")),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits))) +
  geom_tile(aes(fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

PA1.test <- PA1.followup %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater") %>% mutate(FDR = p.adjust(p, method = "fdr"))

# FDR < 0.1 = All four features

```

# Test for the whole complex with follow up
#Figure 4B: LMNB1
```{r}
screen.ASCOM <- slope.protein.features.all %>% 
  filter(gene %in% c("ASH2L","DPY30","RBBP5","WDR5")) %>% select(gene,feature,epistasis.final) %>% mutate(gRNA = "Screen")

ASCOM.followup <- epistasis.ratio.followup %>% filter(gene %in% c("ASH2L","DPY30","RBBP5","WDR5")) %>% select(gene,feature, epistasis.final,gRNA) %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_"))

ASCOM.combined <- bind_rows(screen.ASCOM,ASCOM.followup) %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_"))

ASCOM.combined.dcast <- ASCOM.followup %>% dcast(gene_gRNA ~ feature, value.var = "epistasis.final") %>% column_to_rownames("gene_gRNA") 

ASCOM.heatmap <- pheatmap(t(ASCOM.combined.dcast))
ASCOM.dist <- dist(ASCOM.combined.dcast)
row_dend.ascom = hclust(ASCOM.dist)
row_dend.hits.ascom <- as.dendrogram(row_dend.ascom)
row_dend.hits.ascom <- seriate_dendrogram(row_dend.hits.ascom,ASCOM.dist, method="OLO")



ascom.gene.order <- rownames(ASCOM.combined.dcast[as.hclust(row_dend.hits.ascom)[["order"]],])

ggplot(ASCOM.followup) +
  geom_tile(aes(fct_relevel(gene_gRNA, ascom.gene.order), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final)) + scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.05,0.1), oob = squish)  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# Complex perturbation
ASCOM.comined.test <- ASCOM.followup %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater") %>% mutate(FDR = p.adjust(p, method = "fdr"))

```

#Figure 4D
```{r}

PA1.PTIP.followup <- epistasis.ratio.followup %>% filter(gene %in% c("PAXIP1","PA1")) %>% dplyr::select(gene,feature, epistasis.final,gRNA)

PA1.combined <- PA1.PTIP.followup %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_")) 

PA1.combined.dcast <- PA1.combined %>% dcast(gene_gRNA ~ feature, value.var = "epistasis.final") %>% column_to_rownames("gene_gRNA")

PA1.heatmap <- pheatmap(t(PA1.combined.dcast))
PA1.dist <- dist(PA1.combined.dcast)
row_dend.pa1 = hclust(PA1.dist)
row_dend.hits.pa1 <- as.dendrogram(row_dend.pa1)
row_dend.hits.pa1 <- seriate_dendrogram(row_dend.hits.pa1,PA1.dist, method="OLO")

pa1.gene.order <- rownames(PA1.combined.dcast[as.hclust(row_dend.hits.pa1)[["order"]],])
.feature.order<- colnames(mll3.4.combined.dcast[,mll3.4.heatmap$tree_row[["order"]]])

ggplot(PA1.combined) +
  geom_tile(aes(fct_relevel(gene_gRNA,pa1.gene.order), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final)) + scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.075,0.075), oob = squish)  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")


```


#Figure 4D: NCOA
```{r}
screen.NCOA6 <- slope.protein.features.all %>% 
  filter(gene %in% c("NCOA6")) %>% select(gene,feature,epistasis.final) %>% mutate(gRNA = "Screen")

NCOA6.followup <- epistasis.ratio.followup %>% filter(gene %in% c("NCOA6")) %>% select(gene,feature, epistasis.final,gRNA)

NCOA6.combined <- bind_rows(screen.NCOA6,NCOA6.followup) %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_"))

NCOA6.combined.dcast <- NCOA6.combined %>% dcast(gene_gRNA ~ feature, value.var = "epistasis.final") %>% column_to_rownames("gene_gRNA")

NCOA6.heatmap <- pheatmap(t(NCOA6.combined.dcast))
NCOA6.dist <- dist(NCOA6.combined.dcast)
row_dend.NCOA6 = hclust(NCOA6.dist)
row_dend.hits.NCOA6 <- as.dendrogram(row_dend.NCOA6)
row_dend.hits.NCOA6 <- seriate_dendrogram(row_dend.hits.NCOA6,NCOA6.dist, method="OLO")

NCOA6.gene.order <- rownames(NCOA6.combined.dcast[as.hclust(row_dend.hits.NCOA6)[["order"]],])
NCOA6.feature.order<- colnames(NCOA6.combined.dcast[,NCOA6.heatmap$tree_row[["order"]]])

ggplot(NCOA6.combined) +
  geom_tile(aes(fct_relevel(gene_gRNA,NCOA6.gene.order), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final)) + scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.05,0.1), oob = squish)  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# Complex perturbation
NCOA6.combined.test <- NCOA6.combined %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater") %>% mutate(FDR = p.adjust(p, method = "fdr"))

```

#Figure 4D: Enzymatic components
```{r}
KMT.followup <- epistasis.ratio.followup %>% filter(gene %in% c("KMT2C","KMT2D","KDM6A","dKMT2CD")) %>% dplyr::select(gene,feature, epistasis.final,gRNA)

KMT.combined <- KMT.followup %>% mutate(gene_gRNA = paste(gene,gRNA, sep = "_"))

KMT.combined.dcast <- KMT.combined %>% dcast(gene_gRNA ~ feature, value.var = "epistasis.final") %>% column_to_rownames("gene_gRNA")

KMT.heatmap <- pheatmap(t(KMT.combined.dcast))
KMT.dist <- dist(KMT.combined.dcast)
row_dend.KMT = hclust(KMT.dist)
row_dend.hits.KMT <- as.dendrogram(row_dend.KMT)
row_dend.hits.KMT <- seriate_dendrogram(row_dend.hits.KMT,KMT.dist, method="OLO")

KMT.gene.order <- rownames(KMT.combined.dcast[as.hclust(row_dend.hits.KMT)[["order"]],])

ggplot(KMT.combined %>% filter(gene %in% c("KMT2C","KMT2D"))) +
  geom_tile(aes(fct_relevel(gene_gRNA,KMT.gene.order), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final)) + scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", limits = c(-0.05,0.1), oob = squish)  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# Complex perturbation
KMT.combined.test <- KMT.combined %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater") %>% mutate(FDR = p.adjust(p, method = "fdr"))

```
