---
title: "xv20211230_4_epistasis_model"
output: html_document
---

# Script to generate figure #4 in the paper. This figure describes how epistatic interactions are present between chromatin features and protein KOs. Figure outline:
A- Step by Step for KAT5 and H3K9me2
B- Matrix for all hits
C- Hits per library
D- Hits per pathway
E- Strong hits examples

Data in text:
A-...
B-...

Supplementary figure 5:
A- All proteins matrix
B- Balance perturbation strong hits

This script generates the plots for figure 4 and S5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)

```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_kinase_inhib_epistasis.rds")
```

#Figure 4A: Step by step explanation KAT5
```{r}
#Step 1: IPR WT vs. KAT5
ggplot(step2.epistasis %>% filter(gene %in% c("WT","KAT5"))) + geom_col(aes(fct_reorder(barcode,m.log2.ratio.dist), m.log2.ratio.dist, fill = gene)) + scale_color_manual(values = c("#EF6817","#21201F")) + theme_bw() +  ylab("log2(MMEJ:NHEJ ratio)") + theme(axis.text.x = element_blank(), legend.position = "top") + ylim(c(0,0.8))

# Difference vs. mean effect
step2.epistasis.effect <- step2.epistasis %>%
  group_by(gene) %>%
  dplyr::summarise(diff = max(m.log2.ratio.dist) - min(m.log2.ratio.dist), mean = abs(mean(m.log2.ratio.dist))) %>% melt()



ggplot(step2.epistasis.effect %>% filter(gene == "KAT5")) +
  geom_quasirandom(aes(variable,value)) + theme_bw()

hits.fdr <- test.log2.dist %>% filter(p.adj < 0.001) %>% pull(gene)

#Step2: Distance vs H3K9me2
ggplot(step2.epistasis %>% 
         filter(gene %in% c("KAT5")),
       aes(H3K9me2,m.log2.ratio.dist, color = gene)) + 
  geom_point() +
  geom_smooth(method = "lm") +
  scale_color_manual(values = "#EF6817") +
  theme_bw() + ylab("Distance WT vs. KO")

# Gent intercept and slope
slope.protein.features.all %>% filter(gene == "KAT5" & feature == "H3K9me2")

#Plot KAT5 interactions
ggplot(slope.protein.features.all %>% filter(gene == "KAT5")) +
  geom_col(aes(fct_relevel(feature,clone5_chromatin_cluster),slope, fill = gene)) + geom_vline(aes(xintercept = 0)) + theme_bw() + coord_flip() + scale_fill_manual(values = "#EF6817")

```

# Figure 4B: Matrix for hits
```{r}
#genes to select
tmp.gene.hits <- filter(test.log2.dist, p.adj < 0.0001) %>% pull(gene)

hit.genes <- filter(hits.both.screens, pathway != "NA") %>% separate(ID_gene, into = "gene") %>% pull(gene) %>% unique()

#Select 200 genes with highest differences
big.rearrangement <- slope.protein.features.all %>% 
  group_by(gene) %>%
  summarise(delta = max(slope) - min(slope),
            stdev = sd(slope)) %>%
  slice_max(delta, n = 200) %>%
  pull(gene)

# dcaast with selected
slope.prot.features.dcast.hits <- slope.protein.features.all %>% filter(gene %in% big.rearrangement) %>% reshape2::dcast(gene ~ feature, value.var = "slope") %>% column_to_rownames(var = "gene")



#Heatmap
heatmap.slope.diff.hits <- pheatmap(slope.prot.features.dcast.hits, silent = F)
top200 <- dist(slope.prot.features.dcast.hits)
row_dend = hclust(top200)
row_dend.hits <- as.dendrogram(row_dend)
row_dend.hits <- seriate_dendrogram(row_dend.hits, top200, method="OLO")

heatmap.gene.order.slope.diff.hits <- rownames(slope.prot.features.dcast.hits[as.hclust(row_dend.hits)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])

# This are the effects


# heatmap
ggplot(slope.protein.features.all %>% filter(gene %in% big.rearrangement)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.hits),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits), fill = slope)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 1)

# Dendrogram
ggdendrogram(heatmap.slope.diff.hits$tree_row)

# Relative slope
tmp.rel.slope <- slope.protein.features.all %>% filter(gene %in% tmp.gene.hits) %>% mutate(rel.epistasis = slope*6/mean.fc)
```

# Figure 4C: Epistasis per library
```{r}
slope.library <- slope.protein.features.all %>% left_join(hits.both.screens %>% separate(ID_gene, into = "gene", remove = T) %>% select(gene,library,pathway) %>% distinct()) %>% filter(library %in% c("Chromatin","DNA_repair"))

# Plot per library(all)
ggplot(slope.library) + geom_quasirandom(aes(library,abs(epistasis), color = library)) + theme_bw() + scale_color_manual(values = levels(as_factor(library.colors$color))) + theme(legend.position = "top")

#test
chromatin.genes.slope <- slope.library %>% filter(library == "Chromatin") %>% pull(epistasis) %>% abs()
dna_repair.genes.slope <- slope.library %>% filter(library == "DNA_repair") %>% pull(epistasis) %>% abs()

ks.test.library <- ks.test(chromatin.genes.slope,dna_repair.genes.slope)
```

#Figure 4D: Epistasis per pathway
```{r}
# Plot per library 
ggplot(slope.library %>% filter(pathway != "NA")) + geom_quasirandom(aes(pathway, abs(epistasis), color = pathway)) + theme_bw() + scale_color_manual(values = c("#EB2030","#2E358F")) + theme(legend.position = "top")

#test
NHEJ.genes.slope <- slope.library %>% filter(pathway == "NHEJ") %>% pull(epistasis) %>% abs()
MMEJ.genes.slope <- slope.library %>% filter(pathway == "MMEJ") %>% pull(epistasis) %>% abs()

ks.test.pathway <- ks.test(NHEJ.genes.slope,MMEJ.genes.slope)
```


#Figure 4E: Strong genes
```{r}
#Plot this for POLL, POLQ and RAD50
# Plot per library 
POLQ.POLL.RAD50 <- slope.library %>% filter(gene %in% c("POLL","POLQ","RAD50")) %>% select(gene,epistasis, library) %>% distinct()

ggplot(POLQ.POLL.RAD50, aes(gene,epistasis)) + geom_quasirandom() + theme_bw() + scale_color_manual(values = library.colors)

```

# Number in text

#NOTE: This will change, but every number that I refer in the text will available in this chunk
```{r}
```

#Supplementary figures

#Figure S5A: slope with all proteins
```{r}
# dcast all genes
slope.prot.features.dcast.all <- slope.protein.features.all %>% reshape2::dcast(gene ~ feature, value.var = "epistasis") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.all <- pheatmap(slope.prot.features.dcast.all, silent = T)

heatmap.gene.order.slope.diff.all <- rownames(slope.prot.features.dcast.all[heatmap.slope.diff.all$tree_row[["order"]],])
heatmap.feature.order.slope.diff.all <- colnames(slope.prot.features.dcast.all[,heatmap.slope.diff.all$tree_col[["order"]]])
# heatmap
ggplot(slope.protein.features.all) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.all),fct_relevel(feature, heatmap.feature.order.slope.diff.all), fill = epistasis)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e")  + 
  theme_bw() + theme(axis.text.x = element_blank(), axis.title.y = element_blank(), legend.position = "top",) + 
  coord_fixed(expand = F,ratio = 20)

# Dendrogram
ggdendrogram(heatmap.slope.diff.all$tree_row, labels = F)
```

#Figure S5B: genes in plot
```{r}
# Filter proteins highlighted 
ggplot(per_19IPR_data %>% filter(gene %in% c("POLL","RAD50","POLQ"))) +
  geom_quasirandom(aes(gene,IPR.z.score)) +
  geom_hline(yintercept = 0, linetype =2) +
  theme_bw()
```

#Figure S5C: Slope is correlated with perturbation
```{r}
#
effect.size <- slope.protein.features.all %>% filter(complete.cases(.) & gene %in% hit.genes)
# Plot slope vs effect size
summ.slope.library <- effect.size %>% dplyr::group_by(gene) %>% dplyr::summarise(med.slope = mean(epistasis, na.rm = T), mean.intercept = mean(intercept))

ggplot(summ.slope.library, aes(med.slope, mean.intercept)) + geom_point() + stat_cor(method = "spearman",cor.coef.name = "rho") + geom_smooth(method = "lm") +
  theme_bw()

```
