---
title: "xv20211230_4_epistasis_model"
output: html_document
---

# Script to generate figure #4 in the paper. This figure describes how epistatic interactions are present between chromatin features and protein KOs. Figure outline:
A- Step by Step for KAT5 and H3K9me2
B- Matrix for all hits
C- Hits per library
D- Hits per pathway
E- Strong hits examples

Data in text:
A-...
B-...

Supplementary figure 5:
A- All proteins matrix
B- Balance perturbation strong hits

This script generates the plots for figure 4 and S5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(STRINGdb)

```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% dplyr::select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
  <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220317_DDR_CCD_mean_significance.rds")

non.targetting.CCD_data <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220317_non_targetting_DDR_CCD_mean_significance.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS( "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_CCD_analysis_preprocessing.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
CCD_inhibitors_data <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_DDR_CCD_mean_significant.rds")

#Import chromatin data significant
chromatin_significant_CCD <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220328_chromatin_CCD_mean_significance.rds")

#Import chromatin data significant
chromatin_all_CCD <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220412_chromatin_CCD_mean_significance_all.rds")
```


# Figure 3A: Matrix for all hits
```{r}
#Filter significant hits
significant_chromatin_proteins <- chromatin_all_CCD %>% filter(model_p.adj < 0.05) %>% na.omit()

# dcaast with selected
slope.prot.features.dcast.hits_chr <- significant_chromatin_proteins %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap_chr <- pheatmap(t(slope.prot.features.dcast.hits_chr), silent = T, clustering_method = "ward.D", cluster_rows = F, cutree_cols = 4)
signif_prots.gene.order_chr <- rownames(slope.prot.features.dcast.hits_chr[signif_prots_heatmap_chr$tree_col[["order"]],])

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_chromatin_proteins) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order_chr),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

# Dendrogram
ggdendrogram(signif_prots_heatmap_chr$tree_col)

```

#Test if proteins in screening have higher interaction with proteins that interact
#Import string database (not all genes are picked by this analysis)
```{r}
#Load string database
#Download string_db and save it as string_db
string_import <- read.table(file = "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/9606.protein.links.v11.5.txt.gz", header = T)
string_import_high_conf <- filter(string_import, combined_score > 750)
string_import_norm_conf <- filter(string_import, combined_score > 400)

#Import gene name
string_alias <- read.delim(file = "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import/9606.protein.aliases.v11.5.txt.gz")
string_alias_name <- string_alias %>% select(string_ID = X.string_protein_id, gene = alias) %>% distinct()
#Load all gene names (I can olny find 362 genes out of 591)
chromatin_genes <- unique(chromatin_all_CCD$gene)

#Check which proteins are missing
string_corr <- significant_chromatin_proteins %>% left_join(string_alias_name)

string_in_chr <- filter(string_alias_name, gene %in% chromatin_genes) %>% mutate(CCD_s = case_when(gene %in% unique(significant_chromatin_proteins$gene) ~ "s", T ~ "non_s"))
string_names_1 <- string_in_chr %>% select(protein1 = string_ID, gene1 =  gene, CCD_s1 = CCD_s)
string_names_2 <- string_in_chr %>% select(protein2 = string_ID, gene2 =  gene, CCD_s2 = CCD_s)

#filter interactions data (with 750 combined_score filtering) => 1744/65341 ~ 2.7% of all interaction comb. spece
interactions_chr_chr<- string_import_high_conf %>% left_join(string_names_1) %>% left_join(string_names_2) %>% na.omit() %>% mutate(s_CCD = paste(CCD_s1, CCD_s2, sep = "."))

```

#
```{r}
# function to get correlation
correlation.features <- function(x,y) {
  A_gene_features <- filter(chromatin_all_CCD, gene == x) %>% select(feature,CCD_value,gene) %>% pull(CCD_value)
  B_gene_features <- filter(chromatin_all_CCD, gene == y) %>% select(feature,CCD_value, gene) %>% pull(CCD_value)
  lm(B_gene_features ~ A_gene_features) %>% glance() %>% mutate(gene1 = x, gene2 = y)
}

# Run function
corr_list <- map2_dfr(interactions_chr_chr$gene1, interactions_chr_chr$gene2, correlation.features)

#Join with interaction dt
int_corr_dt <- interactions_chr_chr %>% left_join(corr_list)
int_corr_signif <- int_corr_dt %>% filter(s_CCD == "s.s")


```

#Calculate for all hits in chromatin libraries
```{r}
#
signif.protein.vector <- significant_chromatin_proteins %>% pull(gene) %>% unique()
comb_signif <- combn(signif.protein.vector, 2, simplify = F) 
comb_signif_dt <- do.call("rbind", comb_signif) %>% as.data.frame()
colnames(comb_signif_dt) <- c("gene1","gene2")

# All possible significant combinations
corr_significant <-  map2_dfr(comb_signif_dt$gene1, comb_signif_dt$gene2, correlation.features)

# Proteins that interact tend to have higher significance level
corr_s_value <- corr_significant %>% left_join(interactions_chr_chr %>% filter(s_CCD == "s.s") %>% select(gene1, gene2, s_CCD)) %>% mutate(interaction = case_when(s_CCD == "s.s" ~ TRUE, T ~ FALSE)) %>% mutate(p.adj = p.adjust(p.value))

#Density plot comparing all of them
ggplot(corr_s_value) + 
  geom_density(aes(r.squared, color = interaction)) + 
  theme_bw()

#Get bootstrap approach (select n = 130 extra ones)

sample_correlations <- tibble()

for (i in c(1:100)){
  set.seed(i)
  t <- corr_s_value %>% filter(interaction == F) %>% sample_n(239) %>% mutate(seed = i)
  sample_correlations <- sample_correlations %>% bind_rows(t)
}

# Proteins that interact tend to get higher interactions (Small difference but higher)
ggplot(sample_correlations) + 
    geom_density(aes(r.squared, group = seed), color = "grey80") + 
    geom_density(data = corr_s_value %>% filter(interaction == T), aes(r.squared, color = interaction)) +
    theme_bw()

# Median interaction r.square
median_interaction <- sample_correlations %>% group_by(seed) %>% summarise(mean = mean(r.squared))
interactor.mean <- corr_s_value %>% filter(interaction == T) %>% pull(r.squared) %>% mean()

stat_test <- t.test(median_interaction$mean, mu = interactor.mean)
```

# Calculate this for DDR proteins
```{r}
correlation.features.DDR <- function(x,y) {
  A_gene_features <- filter(slope.protein.features.all, gene == x) %>% select(feature,CCD_value,gene) %>% pull(CCD_value)
  B_gene_features <- filter(slope.protein.features.all, gene == y) %>% select(feature,CCD_value, gene) %>% pull(CCD_value)
  lm(B_gene_features ~ A_gene_features) %>% glance() %>% mutate(gene1 = x, gene2 = y)
}

#Import data from interaction network
#Load all gene names (I can olny find 362 genes out of 591)
ddr_genes <- unique(slope.protein.features.all$gene)

#Check which proteins are missing
string_corr.ddr <- significant_proteins %>% left_join(string_alias_name)

string_in_chr.DDR <- filter(string_alias_name, gene %in% ddr_genes) %>% mutate(CCD_s = case_when(gene %in% unique(significant_proteins$gene) ~ "s", T ~ "non_s"))
string_names_1.DDR <- string_in_chr.DDR %>% select(protein1 = string_ID, gene1 =  gene, CCD_s1 = CCD_s)
string_names_2.DDR <- string_in_chr.DDR %>% select(protein2 = string_ID, gene2 =  gene, CCD_s2 = CCD_s)

#filter interactions data (with 750 combined_score filtering) => 1744/65341 ~ 2.7% of all interaction comb. spece
interactions_chr_chr.DDR<- string_import_high_conf %>% left_join(string_names_1.DDR) %>% left_join(string_names_2.DDR) %>% na.omit() %>% mutate(s_CCD = paste(CCD_s1, CCD_s2, sep = "."))

#
signif.protein.vector.DDR <- significant_proteins %>% pull(gene) %>% unique()
comb_signif.DDR <- combn(signif.protein.vector.DDR, 2, simplify = F) 
comb_signif_dt.DDR <- do.call("rbind", comb_signif.DDR) %>% as.data.frame()
colnames(comb_signif_dt.DDR) <- c("gene1","gene2")

# All possible significant combinations
corr_significant.DDR <-  map2_dfr(comb_signif_dt.DDR$gene1, comb_signif_dt.DDR$gene2, correlation.features.DDR)

# Proteins that interact tend to have higher significance level
corr_s_value.DDR <- corr_significant.DDR %>% left_join(interactions_chr_chr.DDR %>% filter(s_CCD == "s.s") %>% select(gene1, gene2, s_CCD) %>% distinct()) %>% mutate(interaction = case_when(s_CCD == "s.s" ~ TRUE, T ~ FALSE)) %>% mutate(p.adj = p.adjust(p.value))

#Density plot comparing all of them
ggplot(corr_s_value.DDR) + 
  geom_density(aes(r.squared, color = interaction)) + 
  theme_bw()

#Get bootstrap approach (select n = 130 extra ones)
sample_correlations.ddr <- tibble()

for (i in c(1:100)){
  set.seed(i)
  t <- corr_s_value.DDR %>% filter(interaction == F) %>% sample_n(430) %>% mutate(seed = i)
  sample_correlations.ddr <- sample_correlations.ddr %>% bind_rows(t)
}

# Proteins that interact tend to get higher interactions (Small difference but higher)
ggplot(sample_correlations.ddr) + 
    geom_density(aes(adj.r.squared, group = seed), color = "grey80") + 
    geom_density(data = corr_s_value.DDR %>% filter(interaction == T), aes(adj.r.squared, color = interaction)) +
    theme_bw()

# Median interaction r.square
median_interaction.ddr <- sample_correlations.ddr %>% group_by(seed) %>% summarise(mean = mean(r.squared))
interactor.mean.ddr <- corr_s_value.DDR %>% filter(interaction == T) %>% pull(r.squared) %>% mean()

stat_test.ddr <- t.test(median_interaction.ddr$mean, mu = interactor.mean.ddr)
```

# Correlate both matrices at the same time
```{r}
#Bind both libraries together
sign.both.libraries <- significant_proteins %>% bind_rows(significant_chromatin_proteins)
sign.b.libraries.dt <- sign.both.libraries %>% left_join(string_alias_name)

#All possible combinations
signif.protein.vector.both <- sign.both.libraries %>% pull(gene) %>% unique()
comb_signif.both <- combn(signif.protein.vector.both, 2, simplify = F) 
comb_signif_dt.both <- do.call("rbind", comb_signif.both) %>% as.data.frame()
colnames(comb_signif_dt.both) <- c("gene1","gene2")


correlation.features.both <- function(x,y) {
  A_gene_features <- filter(sign.both.libraries, gene == x) %>% dplyr::group_by(gene,feature) %>% dplyr::summarise(CCD_value = mean(CCD_value, na.rm = T), .groups = 'drop') %>% pull(CCD_value)
  B_gene_features <- filter(sign.both.libraries, gene == y)%>% dplyr::group_by(gene,feature) %>% dplyr::summarise(CCD_value = mean(CCD_value, na.rm = T), .groups = 'drop') %>% pull(CCD_value)
  lm(B_gene_features ~ A_gene_features) %>% glance() %>% mutate(gene1 = x, gene2 = y)
}

# All possible significant combinations
corr_significant.both <-  map2_dfr(comb_signif_dt.both$gene1, comb_signif_dt.both$gene2, correlation.features.both)

#Interactions between these genes
cor_signif_alias <- corr_significant.both %>% left_join(string_alias_name %>% select(protein1 = string_ID, gene1 = gene), ) %>% left_join(string_alias_name %>% select(protein2 = string_ID, gene2 = gene)) %>% mutate(p.adj = p.adjust(p.value)) 

#What library each gene belongs
library.genes <- sign.b.libraries.dt  %>% select(library, gene) %>% distinct()

#Interactions between these genes
cor_signif_alias_lig <- cor_signif_alias %>% left_join(sign.b.libraries.dt %>% select(library1 = library, gene1 = gene) %>% distinct()) %>% left_join(sign.b.libraries.dt %>% select(library2 = library, gene2 = gene) %>% distinct())

cor_s_alias_f <- cor_signif_alias_lig %>% left_join(string_import_high_conf) %>% mutate(interaction = case_when(is.na(combined_score) ~ FALSE, T ~ TRUE), class = paste0(library1, "_", library2)) %>% mutate(class = case_when(class %in% c("Chromatin_DNA_repair", "DNA_repair_Chromatin") ~ "DNA_repair_Chromatin", T ~ class))

cor_s_alias_final <- cor_s_alias_f %>% select(r.squared, p.value,gene1, gene2, p.adj, interaction, class) %>% dplyr::group_by(r.squared, p.value,gene1, gene2, p.adj, class) %>% dplyr::summarise(interaction = case_when(sum(interaction) == 0 ~ F, T ~ T))

#Density plot comparing all of them
ggplot(cor_s_alias_final) + 
  geom_density(aes(r.squared, color = interaction)) + facet_wrap(~ class) + 
  theme_bw()

#Get sampling approach

sample_correlations.both <- tibble()

for (j in unique(cor_s_alias_final$class)) {
  n <- cor_s_alias_final %>% dplyr::group_by(class, interaction) %>% dplyr::summarise(c = n()) %>% filter(interaction == T & class == j) %>% pull(c)
   for (i in c(1:1000)){
  set.seed(i)
  t <- cor_s_alias_final %>% ungroup() %>% filter(class == j) %>% sample_n(n) %>% mutate(seed =i)
  sample_correlations.both <- sample_correlations.both %>% bind_rows(t)
  }
}

sample_correlation_mean <- sample_correlations.both %>% dplyr::group_by(class, seed) %>% dplyr::summarise(m = mean(r.squared))
real_correlation_mean <- cor_s_alias_final %>% dplyr::group_by(interaction,class) %>% dplyr::summarise(m = mean(r.squared))
#
ggplot(sample_correlation_mean) + 
  geom_histogram(aes(m)) + 
  geom_linerange(data = real_correlation_mean,aes(m,ymin = 0, ymax = 200, color = interaction)) + 
  facet_wrap(~ class) +
  coord_cartesian(expand = F, xlim = c(0.2, 0.4)) +
  theme_bw()


# Proteins that interact tend to get higher interactions (Small difference but higher)
ggplot(sample_correlations.both) + 
    geom_density(aes(r.squared, group = seed), color = "grey80") + facet_wrap(~ class) + 
    geom_density(data = cor_s_alias_final %>% filter(interaction == T), aes(r.squared, color = interaction)) +
    theme_bw()

```

#Call concordant pairs more carefully
```{r}
#conc table


concordant.features.both <- function(x,y) {
  A_gene_features <- filter(sign.both.libraries, gene == x) %>% dplyr::group_by(gene,feature) %>% dplyr::summarise(CCD_sign = sign(mean(CCD_value, na.rm = T)), .groups = 'drop') %>% dcast(feature ~ gene, value.var = "CCD_sign")
  B_gene_features <- filter(sign.both.libraries, gene == y)%>% dplyr::group_by(gene,feature) %>% dplyr::summarise(CCD_sign = sign(mean(CCD_value, na.rm = T)), .groups = 'drop') %>% dcast(feature ~ gene, value.var = "CCD_sign")
  cont_table <- A_gene_features %>% left_join(B_gene_features, by = "feature") %>% select(-feature) %>% table() %>% as_tibble()
  conc_pairs <- cont_table[which(cont_table[1] == cont_table[2] & cont_table[1] != 0),] %>% pull(n) %>% sum()
  conc_pairs_0 <- cont_table[which(cont_table[1] == cont_table[2] & cont_table[1] == 0),] %>% pull(n) %>% sum()
  n_conc_pairs <- cont_table[which(cont_table[1] != cont_table[2] & cont_table[1] != 0 & cont_table[2] != 0),] %>% pull(n) %>% sum()
  n_conc_pairs_0 <- cont_table[which(cont_table[1] != cont_table[2] & (cont_table[1] == 0 | cont_table[2] == 0)),] %>% pull(n) %>% sum()
  tibble(gene1 = x, gene2 = y,library1 =  conc = conc_pairs, conc_0 = conc_pairs_0 , non_conc = n_conc_pairs, non_conc_0 = n_conc_pairs_0)
}

#Map all the concordant and discordant features
conc_significant.both <- map2_dfr(comb_signif_dt.both$gene1, comb_signif_dt.both$gene2, concordant.features.both)

#Conc hits
conc_significant.hits <- conc_significant.both %>%  mutate(conc_pair = case_when(non_conc != 0 ~ FALSE, conc <= (non_conc + non_conc_0) ~ FALSE, T ~ TRUE)) %>% left_join(string_alias_name %>% select(protein1 = string_ID, gene1 = gene), ) %>% left_join(string_alias_name %>% select(protein2 = string_ID, gene2 = gene))

conc_significant.hits.f <- conc_significant.hits %>% left_join(sign.b.libraries.dt %>% select(library1 = library, gene1 = gene) %>% distinct()) %>% left_join(sign.b.libraries.dt %>% select(library2 = library, gene2 = gene) %>% distinct()) %>% left_join(string_import_high_conf) %>% mutate(interaction = case_when(is.na(combined_score) ~ FALSE, T ~ TRUE), class = paste0(library1, "_", library2)) %>% mutate(class = case_when(class %in% c("Chromatin_DNA_repair", "DNA_repair_Chromatin") ~ "DNA_repair_Chromatin", T ~ class))

conc_significant.hits.finale <- conc_significant.hits.f %>% select(conc_pair,gene1, gene2, interaction, class) %>% dplyr::group_by(conc_pair,gene1, gene2, class) %>% dplyr::summarise(interaction = case_when(sum(interaction) == 0 ~ F, T ~ T))

# Draw n = interaction TRUE
sample_conc_interaction.both <- tibble()
for (j in unique(conc_significant.hits.finale$class)) {
  n <- conc_significant.hits.finale %>% dplyr::group_by(class, interaction) %>% dplyr::summarise(c = n()) %>% filter(interaction == T & class == j) %>% pull(c)
   for (i in c(1:1000)){
  set.seed(i)
  t <- conc_significant.hits.finale %>% ungroup() %>% filter(class == j) %>% sample_n(n) %>% mutate(seed =i)
  sample_conc_interaction.both <- sample_conc_interaction.both %>% bind_rows(t)
  }
}

# Draw n = conc_pairs TRUE
sample_conc_pairs.both <- tibble()
for (j in unique(conc_significant.hits.finale$class)) {
  n <- conc_significant.hits.finale %>% dplyr::group_by(class, conc_pair) %>% dplyr::summarise(c = n()) %>% filter(conc_pair == T & class == j) %>% pull(c)
   for (i in c(1:1000)){
  set.seed(i)
  t <- conc_significant.hits.finale %>% ungroup() %>% filter(class == j) %>% sample_n(n) %>% mutate(seed =i)
  sample_conc_pairs.both <- sample_conc_pairs.both %>% bind_rows(t)
  }
}

#Plot histogram
sample_c_p_both <- sample_conc_interaction.both %>% dplyr::group_by(seed, class,conc_pair) %>% dplyr::summarise(c = n())
real_c_p_both <- conc_significant.hits.finale %>% filter(interaction == T) %>% dplyr::group_by(conc_pair, class) %>% dplyr::summarise(c = n())

# Plot histogram for sampled data
ggplot() +  
  geom_histogram(data = sample_c_p_both %>% filter(conc_pair == T), aes(c)) +
  geom_linerange(data = real_c_p_both %>% filter(conc_pair == T), aes(c, ymin = 0, ymax = 200), color = "blue") + facet_wrap(~ class, scales = "free_x") + theme_bw()

#Plot histogram
sample_c_i_both <- sample_conc_pairs.both %>% dplyr::group_by(seed, interaction, class) %>% dplyr::summarise(c = n())
real_c_i_both <- conc_significant.hits.finale %>% filter(conc_pair == T) %>% dplyr::group_by(interaction, class) %>% dplyr::summarise(c = n())

# Plot histogram for sampled data
ggplot() +  
  geom_histogram(data = sample_c_i_both %>% filter(interaction == T), aes(c)) +
  geom_linerange(data = real_c_i_both %>% filter(interaction == T), aes(c, ymin = 0, ymax = 200), color = "blue") + facet_wrap(~ class, scales = "free_x") + theme_bw()

```

# Are pairs more conserved across similar proteins?
# Correlate both matrices at the same time (I think this is not a good strategy to do)
```{r}
#Call clusters (M vs. N synergies)
m_n_synergies <- sign.b.libraries.dt %>% dplyr::group_by(library, gene) %>% dplyr::summarise(cluster = case_when(sum(CCD_value > 0) == 0 ~ "M_synergy", sum(CCD_value < 0) == 0 ~ "N_synergy", T ~ "both_synergy")) %>% left_join(string_alias_name)  %>% ungroup()

#All possible combinations with hits
signif.protein.vector_string <- m_n_synergies %>% pull(string_ID) %>% unique()
comb_signif.both_string <- combn(signif.protein.vector_string, 2, simplify = F) 
comb_signif_dt.both_string <- do.call("rbind", comb_signif.both_string) %>% as.data.frame()
colnames(comb_signif_dt.both_string) <- c("protein1","protein2")

# Pairs that are in the string database
hit_hit_string_pairs <- string_import_high_conf %>% inner_join(comb_signif_dt.both_string)


#Check if these occur within same synergies
conc.synergies <- hit_hit_string_pairs %>% left_join(m_n_synergies %>% select(library1 = library, gene1 = gene, cluster1 = cluster, protein1 = string_ID) %>% distinct()) %>% left_join(m_n_synergies %>% select(library2 = library, gene2 = gene, cluster2 = cluster, protein2 = string_ID) %>% distinct()) %>% mutate(conc_inter = cluster1 == cluster2)

conc.synergies.class <- conc.synergies %>% mutate(class = paste0(library1, "_", library2)) %>% mutate(class = case_when(class %in% c("Chromatin_DNA_repair", "DNA_repair_Chromatin") ~ "DNA_repair_Chromatin", T ~ class))

# All possible pairs
all_hit_hit_pairs <- comb_signif_dt.both_string %>% left_join(m_n_synergies %>% select(library1 = library, gene1 = gene, cluster1 = cluster, protein1 = string_ID) %>% distinct()) %>% left_join(m_n_synergies %>% select(library2 = library, gene2 = gene, cluster2 = cluster, protein2 = string_ID) %>% distinct()) %>% mutate(conc_inter = cluster1 == cluster2)

all_h_h_pairs_class <- all_hit_hit_pairs %>% mutate(class = paste0(library1, "_", library2)) %>% mutate(class = case_when(class %in% c("Chromatin_DNA_repair", "DNA_repair_Chromatin") ~ "DNA_repair_Chromatin", T ~ class))

# Pure chance 
sample_concordant.both <- tibble()

for (j in unique(conc.synergies.class$class)) {
  n <- conc.synergies.class %>% filter(class == j) %>% nrow()
   for (i in c(1:1000)){
  set.seed(i)
  t <- all_h_h_pairs_class %>% ungroup() %>% filter(class == j) %>% sample_n(n) %>% mutate(seed =i)
  sample_concordant.both <- sample_concordant.both %>% bind_rows(t)
  }
}

#Plot histogram
sample_c_both <- sample_concordant.both %>% dplyr::group_by(seed, conc_inter, class) %>% dplyr::summarise(c = n())
real_c_both <- conc.synergies.class %>% dplyr::group_by(conc_inter, class) %>% dplyr::summarise(c = n())

ggplot() +  
  geom_histogram(data = sample_c_both %>% filter(conc_inter == T), aes(c)) +
  geom_linerange(data = real_c_both %>% filter(conc_inter == T), aes(c, ymin = 0, ymax = 200), color = "red") + facet_wrap(~ class, scales = "free_x") + theme_bw()

#Plot histogram
sample_i_both <- sample_concordant.both %>% dplyr::group_by(seed, interaction, class) %>% dplyr::summarise(c = n())
real_i_both <- conc.synergies.class %>% dplyr::group_by(interaction, class) %>% dplyr::summarise(c = n())

ggplot() +  
  geom_histogram(data = sample_c_both %>% filter(conc_inter == T), aes(c)) +
  geom_linerange(data = real_c_both %>% filter(conc_inter == T), aes(c, ymin = 0, ymax = 200), color = "red") + facet_wrap(~ class, scales = "free_x") + theme_bw()


ggplot + geom_quasirandom(data = sample_concordant.both %>% dplyr::group_by(seed, conc_inter, class) %>% dplyr::summarise(c = n()), aes(conc_inter, c)) + facet_wrap(~ class)

```

#Do DNA_repair Chromatin pairs create a nice network
```{r}
#Chromatin proteins with conc. role with a DDR protein
chromatin.proteins.ddr <- conc_significant.hits.finale %>% filter(interaction == T & conc_pair == T & class == "DNA_repair_Chromatin") %>% select(gene1,gene2) %>% distinct()

chr_chr_genes_filter_a <- chromatin.proteins.ddr %>% pull(gene2) %>% unique()
#This does not add any new name
chr_chr_genes_filter_b <- chromatin.proteins.ddr %>% pull(gene1) %>% unique()

chr_ddr_genes <- c(chr_chr_genes_filter_a,chr_chr_genes_filter_b) %>% unique()

# Make a heatmap with all these proteins
significant_ddr_chr_proteins <- bind_rows(significant_chromatin_proteins, significant_proteins) %>% filter(gene %in% chr_ddr_genes)

# dcaast with selected
sign_ddr_chr_dcast <- significant_ddr_chr_proteins %>% reshape2::dcast(gene + library ~ feature, value.var = "CCD_value")
rownames(sign_ddr_chr_dcast) <- make.names(sign_ddr_chr_dcast[,1], unique = TRUE)
sign_ddr_chr_dcast_clean <- sign_ddr_chr_dcast %>% select(-gene, -library)


#Heatmap
signif_heatmap_ddr_chr <- pheatmap(t(sign_ddr_chr_dcast_clean), silent = T, clustering_method = "ward.D2", cluster_rows = F, cutree_cols = T)
signif_prots_ddr_chr.gene.order <- rownames(sign_ddr_chr_dcast[signif_heatmap_ddr_chr$tree_col[["order"]],])

# heatmap: All heatmaps together (DDR only)
ggplot(significant_ddr_chr_proteins %>% filter(gene %in% chr_ddr_genes & library == "Chromatin")) + 
  geom_tile(aes(fct_relevel(gene,signif_prots_ddr_chr.gene.order),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

#Chromatin network
network.chromatin.prots <- conc_significant.hits.finale %>% filter(interaction == T & conc_pair == T & (class == "DNA_repair_Chromatin" | class == "Chromatin_Chromatin" & (gene2 %in% chr_chr_genes_filter_a | gene1 %in% chr_chr_genes_filter_a)))

#Export network
write.csv(network.chromatin.prots, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220426_chromatin_protein_interaction.csv")

write.csv(conc_significant.hits.finale %>% filter(interaction == T & conc_pair == T & class == "DNA_repair_Chromatin"), file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220426_chr_DDR_interaction.csv")


```

#Do genes that interact have networks
```{r}
#ENetwork with proteins that have interactions
interacting.prots <- cor_s_alias_final %>% filter(interaction == T)
correlating.prots <- cor_s_alias_final %>% filter(p.adj < 0.05)
corr.int.prots <- cor_s_alias_final %>% filter(p.adj < 0.05 & interaction == T)

#Export these networks
write.xlsx(interacting.prots, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220422_prots_ccd_interacting.csv")

write.csv(corr.int.prots, file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/export/xv20220422_prots_ccd_interacting_corr.csv")

```


# What features are driving in each cluster
```{r}
#Test wilcox test if
significant_proteins_clust <- significant_proteins %>% left_join(cluster_tibble)

#Data frame for p_value
cluster.feature.test <- tibble(cluster_name = NA, feature = NA, p.value = NA, mean = NA, sd = NA)

for(i in unique(significant_proteins_clust$cluster_name)) {
  for(j in unique(significant_proteins_clust$feature)) {
  cluster_data <- significant_proteins_clust %>% filter(cluster_name == i, feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(cluster_data,control_data) %>% tidy()
  cluster.feature.test <- cluster.feature.test %>% add_row(cluster_name = i, feature = j, p.value = wil_t %>% pull(p.value), mean = mean(cluster_data, na.rm = T), sd = sd(cluster_data, na.rm = T))
  }
}

#Apply fdr correction
cluster.feature.test.adj <- cluster.feature.test %>% mutate(p.adj = p.adjust(p.value))

# Features that drive clustering
ggplot(cluster.feature.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature, heatmap.chromatin.order),mean, fill = p.adj < 0.01)) + facet_wrap(~ cluster_name, nrow = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + scale_fill_manual(values = c("grey60","orange"))

```

# Highlight FA proteins
```{r}
#dcast epistatic interactions (FANC proteins)
FA.proteins.CCD <- slope.protein.features.all %>% filter(gene %in% c("FANCL","FAAP100","FAAP24","FANCA","FANCB","FANCC","FANCE","FANCF","FANCG","FANCM","FANCD2","FANCI")) 

FA.proteins.dcast <- FA.proteins.CCD %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

ggplot(FA.proteins.CCD) + 
  geom_tile(aes(fct_relevel(gene,c("FAAP24","FANCM","FANCA","FANCG","FANCF","FANCC","FAAP100","FANCE","FANCL","FANCB","FANCI","FANCD2")), fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.075,-0.025))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

#Data frame for p_value
FA.prots.feautres.test <- tibble(feature = NA, p.value = NA, mean = NA, sd = NA)

  for(j in unique(FA.proteins.CCD$feature)) {
  fa_data <- FA.proteins.CCD %>% filter(feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(fa_data,control_data) %>% tidy()
  FA.prots.feautres.test <- FA.prots.feautres.test %>% add_row(feature = j, p.value = wil_t %>% pull(p.value), mean = mean(fa_data, na.rm = T), sd = sd(fa_data, na.rm = T))
  }

#Apply fdr correction
FA.prots.feautres.test.adj <- FA.prots.feautres.test %>% mutate(p.adj = p.adjust(p.value))

#Plot significant features (test summary plot)
ggplot(FA.prots.feautres.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature,heatmap.chromatin.order),mean, fill = p.adj < 0.05)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + geom_quasirandom(data = FA.proteins.CCD, aes(feature,CCD_value))

```

# Highlight FA proteins
```{r}
#dcast epistatic interactions (PARP proteins)
PARP.proteins.CCD <- slope.protein.features.all %>% filter(grepl("PAR", gene) & !gene %in% c("PARK7","PARPBP"))

PARP.proteins.dcast <- PARP.proteins.CCD %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

ggplot(PARP.proteins.CCD) + 
  geom_tile(aes(gene, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.2, -0.1,0), limits = c(-0.2,0))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

#dcast epistatic interactions (PARP inhibitors)
PARP.inhibitors <- CCD_inhibitors_data %>% filter(drug %in% c("olaparib","talazoparib"))

PARP.inhibitors.dcast <- PARP.inhibitors %>% reshape2::dcast(drug ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "drug")

ggplot(PARP.inhibitors) + 
  geom_tile(aes(drug, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.2, -0.1,0), limits = c(-0.2,0))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# bind inhibitors and KO and test for driver features
combined.PARP.test <- PARP.proteins.CCD %>% select(feature, CCD_value) %>% bind_rows(PARP.inhibitors %>% select(feature, CCD_value))


#Data frame for p_value
PARP.prots.feautres.test <- tibble(feature = NA, p.value = NA, mean = NA, sd = NA)

  for(j in unique(combined.PARP.test$feature)) {
  parp_data <- combined.PARP.test %>% filter(feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(parp_data,control_data) %>% tidy()
  PARP.prots.feautres.test <- PARP.prots.feautres.test %>% add_row(feature = j, p.value = wil_t %>% pull(p.value), mean = mean(parp_data, na.rm = T), sd = sd(parp_data, na.rm = T))
  }

#Apply fdr correction
PARP.prots.feautres.test.adj <- PARP.prots.feautres.test %>% mutate(p.adj = p.adjust(p.value))

#Plot significant features (test summary plot)
ggplot(PARP.prots.feautres.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature,heatmap.chromatin.order),mean, fill = p.adj < 0.05)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + geom_quasirandom(data = combined.PARP.test, aes(feature,CCD_value))

```
# DNAPK inhibitor vs. DNAPK ko
```{r}
#dcast epistatic interactions (PARP proteins)
DNAPK.proteins.CCD <- slope.protein.features.all %>% filter(gene == "PRKDC") %>% select(gene, feature, CCD_value)
DNAPK.inhibitors <- CCD_inhibitors_data %>% filter(drug == "DNAPKi") %>% select(gene = drug, feature, CCD_value)

# Bind rows
DNAPK.perturbations <- bind_rows(DNAPK.proteins.CCD, DNAPK.inhibitors)

# DNAPK inhibition
ggplot(DNAPK.perturbations) + 
    geom_tile(aes(gene, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
    scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(0, 0.05, 0.1, 0.15))  + 
    coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + geom_point(data = subset(DNAPK.perturbations, CCD_value != 0), aes(gene, feature), shape = 4)

# Try to show a correlation plot
DNAPK.perturbations.dcast <- DNAPK.perturbations %>% reshape2::dcast(feature ~ gene, value.var = "CCD_value")
ggplot(DNAPK.perturbations.dcast,aes(PRKDC,DNAPKi)) + geom_point() + stat_cor() + geom_smooth(method = "lm") + theme_bw() + geom_abline(linetype = 2)

```

# Figure 4B: Matrix for hits
```{r}
# Separate data based on the significant IPRs that they have (based on z-score)
significant_proteins_chromatin <- chromatin_significant_CCD %>% filter(model_p.adj < 0.05)

# dcaast with selected
slope.prot.features.dcast.hits_chr <- significant_proteins_chromatin %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap_chr <- pheatmap(t(slope.prot.features.dcast.hits_chr), silent = T, clustering_method = "ward.D", cluster_rows = F, cutree_cols = 4)
signif_prots.gene.order_chr <- rownames(slope.prot.features.dcast.hits_chr[signif_prots_heatmap_chr$tree_col[["order"]],])

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_proteins_chromatin) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order_chr),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + xlab("Chromatin proteins with significant chromatin dependencies") + ylab("Chromatin features") + coord_fixed(ratio = 1)

# Dendrogram
ggdendrogram(signif_prots_heatmap_chr$tree_col)

```

