---
title: "xv20211230_4_epistasis_model"
output: html_document
---

# Script to generate figure #4 in the paper. This figure describes how epistatic interactions are present between chromatin features and protein KOs. Figure outline:
A- Step by Step for KAT5 and H3K9me2
B- Matrix for all hits
C- Hits per library
D- Hits per pathway
E- Strong hits examples

Data in text:
A-...
B-...

Supplementary figure 5:
A- All proteins matrix
B- Balance perturbation strong hits

This script generates the plots for figure 4 and S5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(STRINGdb)

```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% dplyr::select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
  <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220317_DDR_CCD_mean_significance.rds")

non.targetting.CCD_data <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220317_non_targetting_DDR_CCD_mean_significance.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS( "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_CCD_analysis_preprocessing.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
CCD_inhibitors_data <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220318_DDR_CCD_mean_significant.rds")

#Import chromatin data significant
chromatin_significant_CCD <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220328_chromatin_CCD_mean_significance.rds")

#Import chromatin data significant
chromatin_all_CCD <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220412_chromatin_CCD_mean_significance_all.rds")
```


# Figure 3A: Matrix for all hits
```{r}
#Filter significant hits
significant_chromatin_proteins <- chromatin_all_CCD %>% filter(model_p.adj < 0.05) %>% na.omit()

# dcaast with selected
slope.prot.features.dcast.hits_chr <- significant_chromatin_proteins %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap_chr <- pheatmap(t(slope.prot.features.dcast.hits_chr), silent = T, clustering_method = "ward.D", cluster_rows = F, cutree_cols = 4)
signif_prots.gene.order_chr <- rownames(slope.prot.features.dcast.hits_chr[signif_prots_heatmap_chr$tree_col[["order"]],])

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_chromatin_proteins) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order_chr),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

# Dendrogram
ggdendrogram(signif_prots_heatmap_chr$tree_col)

```

#Test this with significant proteins
```{r}
#Load string database
#Download string_db and save it as string_db
string_db <- STRINGdb$new(version="11.5", species = 9606, input_directory = "/DATA/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/import")

#Load significant gene names
signif.hits <- significant_chromatin_proteins %>% select(gene, model_p.adj) %>% distinct()
chromatin_sign_hits <- string_db$mp(signif.hits$gene)



```


# Figure 4B: Matrix for hits
```{r}
# Separate data based on the significant IPRs that they have (based on z-score)
significant_proteins <- slope.protein.features.all %>% filter(model_p.adj < 0.05)

# dcaast with selected
slope.prot.features.dcast.hits <- significant_proteins %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap <- pheatmap(t(slope.prot.features.dcast.hits), silent = T, clustering_method = "ward.D", cluster_rows = F, cutree_cols = 4)
signif_prots.gene.order <- rownames(slope.prot.features.dcast.hits[signif_prots_heatmap$tree_col[["order"]],])

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_proteins) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + coord_fixed(expand = F,ratio = 1) + xlab("DNA repair proteins with significant chromatin dependencies") + ylab("Chromatin features")

# Dendrogram
ggdendrogram(signif_prots_heatmap$tree_col)

#Extract 4 clusters
cluster_data <- cutree(signif_prots_heatmap$tree_col, k=4) 
cluster_tibble <- tibble(gene = names(cluster_data), cluster = cluster_data) %>% mutate(cluster_name = case_when(cluster == "1" ~ "MMEJ_3", cluster == "2" ~ "MMEJ_2", cluster == "3" ~ "MMEJ_1", cluster == "4" ~"NHEJ"))

#Number of genes per cluster
ggplot(cluster_tibble) + geom_bar(aes(cluster_name)) + theme_bw()

```

# What features are driving in each cluster
```{r}
#Test wilcox test if
significant_proteins_clust <- significant_proteins %>% left_join(cluster_tibble)

#Data frame for p_value
cluster.feature.test <- tibble(cluster_name = NA, feature = NA, p.value = NA, mean = NA, sd = NA)

for(i in unique(significant_proteins_clust$cluster_name)) {
  for(j in unique(significant_proteins_clust$feature)) {
  cluster_data <- significant_proteins_clust %>% filter(cluster_name == i, feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(cluster_data,control_data) %>% tidy()
  cluster.feature.test <- cluster.feature.test %>% add_row(cluster_name = i, feature = j, p.value = wil_t %>% pull(p.value), mean = mean(cluster_data, na.rm = T), sd = sd(cluster_data, na.rm = T))
  }
}

#Apply fdr correction
cluster.feature.test.adj <- cluster.feature.test %>% mutate(p.adj = p.adjust(p.value))

# Features that drive clustering
ggplot(cluster.feature.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature, heatmap.chromatin.order),mean, fill = p.adj < 0.01)) + facet_wrap(~ cluster_name, nrow = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + scale_fill_manual(values = c("grey60","orange"))

```

# Gost per cluster
```{r}
# Complexes in H3K27me3 (Fanconi anemia)
complex_mmej_1 <- gost(cluster_tibble %>% filter(cluster == 1) %>% pull(gene), source = "CORUM")

#Complexes with triple heterochromatin (Resection machinery)
complex_mmej_2 <- gost(cluster_tibble %>% filter(cluster == 2) %>% pull(gene), source = "CORUM")

#Complexes with euchromatn (no significant complexes) => PARP1 and PARG?
complex_mmej_3 <- gost(cluster_tibble %>% filter(cluster == 3) %>% pull(gene), source = "CORUM")

#Complexes with euchromatn (BRCC complex and COP9 signalosome)
complex_mmej_4 <- gost(cluster_tibble %>% filter(cluster == 4) %>% pull(gene), source = "CORUM")

```

# Highlight FA proteins
```{r}
#dcast epistatic interactions (FANC proteins)
FA.proteins.CCD <- slope.protein.features.all %>% filter(gene %in% c("FANCL","FAAP100","FAAP24","FANCA","FANCB","FANCC","FANCE","FANCF","FANCG","FANCM","FANCD2","FANCI")) 

FA.proteins.dcast <- FA.proteins.CCD %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

ggplot(FA.proteins.CCD) + 
  geom_tile(aes(fct_relevel(gene,c("FAAP24","FANCM","FANCA","FANCG","FANCF","FANCC","FAAP100","FANCE","FANCL","FANCB","FANCI","FANCD2")), fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.075,-0.025))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

#Data frame for p_value
FA.prots.feautres.test <- tibble(feature = NA, p.value = NA, mean = NA, sd = NA)

  for(j in unique(FA.proteins.CCD$feature)) {
  fa_data <- FA.proteins.CCD %>% filter(feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(fa_data,control_data) %>% tidy()
  FA.prots.feautres.test <- FA.prots.feautres.test %>% add_row(feature = j, p.value = wil_t %>% pull(p.value), mean = mean(fa_data, na.rm = T), sd = sd(fa_data, na.rm = T))
  }

#Apply fdr correction
FA.prots.feautres.test.adj <- FA.prots.feautres.test %>% mutate(p.adj = p.adjust(p.value))

#Plot significant features (test summary plot)
ggplot(FA.prots.feautres.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature,heatmap.chromatin.order),mean, fill = p.adj < 0.05)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + geom_quasirandom(data = FA.proteins.CCD, aes(feature,CCD_value))

```

# Highlight FA proteins
```{r}
#dcast epistatic interactions (PARP proteins)
PARP.proteins.CCD <- slope.protein.features.all %>% filter(grepl("PAR", gene) & !gene %in% c("PARK7","PARPBP"))

PARP.proteins.dcast <- PARP.proteins.CCD %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

ggplot(PARP.proteins.CCD) + 
  geom_tile(aes(gene, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.2, -0.1,0), limits = c(-0.2,0))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

#dcast epistatic interactions (PARP inhibitors)
PARP.inhibitors <- CCD_inhibitors_data %>% filter(drug %in% c("olaparib","talazoparib"))

PARP.inhibitors.dcast <- PARP.inhibitors %>% reshape2::dcast(drug ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "drug")

ggplot(PARP.inhibitors) + 
  geom_tile(aes(drug, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(-0.2, -0.1,0), limits = c(-0.2,0))  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# bind inhibitors and KO and test for driver features
combined.PARP.test <- PARP.proteins.CCD %>% select(feature, CCD_value) %>% bind_rows(PARP.inhibitors %>% select(feature, CCD_value))


#Data frame for p_value
PARP.prots.feautres.test <- tibble(feature = NA, p.value = NA, mean = NA, sd = NA)

  for(j in unique(combined.PARP.test$feature)) {
  parp_data <- combined.PARP.test %>% filter(feature == j) %>% pull(CCD_value)
  control_data <- non.targetting.CCD_data %>% filter(feature == j) %>% pull(CCD_value)
  wil_t <- wilcox.test(parp_data,control_data) %>% tidy()
  PARP.prots.feautres.test <- PARP.prots.feautres.test %>% add_row(feature = j, p.value = wil_t %>% pull(p.value), mean = mean(parp_data, na.rm = T), sd = sd(parp_data, na.rm = T))
  }

#Apply fdr correction
PARP.prots.feautres.test.adj <- PARP.prots.feautres.test %>% mutate(p.adj = p.adjust(p.value))

#Plot significant features (test summary plot)
ggplot(PARP.prots.feautres.test.adj %>% na.omit()) + geom_col(aes(fct_relevel(feature,heatmap.chromatin.order),mean, fill = p.adj < 0.05)) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + geom_quasirandom(data = combined.PARP.test, aes(feature,CCD_value))

```
# DNAPK inhibitor vs. DNAPK ko
```{r}
#dcast epistatic interactions (PARP proteins)
DNAPK.proteins.CCD <- slope.protein.features.all %>% filter(gene == "PRKDC") %>% select(gene, feature, CCD_value)
DNAPK.inhibitors <- CCD_inhibitors_data %>% filter(drug == "DNAPKi") %>% select(gene = drug, feature, CCD_value)

# Bind rows
DNAPK.perturbations <- bind_rows(DNAPK.proteins.CCD, DNAPK.inhibitors)

# DNAPK inhibition
ggplot(DNAPK.perturbations) + 
    geom_tile(aes(gene, fct_relevel(feature,heatmap.chromatin.order), fill = CCD_value)) +
    scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e", breaks = c(0, 0.05, 0.1, 0.15))  + 
    coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + geom_point(data = subset(DNAPK.perturbations, CCD_value != 0), aes(gene, feature), shape = 4)

# Try to show a correlation plot
DNAPK.perturbations.dcast <- DNAPK.perturbations %>% reshape2::dcast(feature ~ gene, value.var = "CCD_value")
ggplot(DNAPK.perturbations.dcast,aes(PRKDC,DNAPKi)) + geom_point() + stat_cor() + geom_smooth(method = "lm") + theme_bw() + geom_abline(linetype = 2)

```

# Figure 4B: Matrix for hits
```{r}
# Separate data based on the significant IPRs that they have (based on z-score)
significant_proteins_chromatin <- chromatin_significant_CCD %>% filter(model_p.adj < 0.05)

# dcaast with selected
slope.prot.features.dcast.hits_chr <- significant_proteins_chromatin %>% reshape2::dcast(gene ~ feature, value.var = "CCD_value") %>% column_to_rownames(var = "gene")

#Heatmap
signif_prots_heatmap_chr <- pheatmap(t(slope.prot.features.dcast.hits_chr), silent = T, clustering_method = "ward.D", cluster_rows = F, cutree_cols = 4)
signif_prots.gene.order_chr <- rownames(slope.prot.features.dcast.hits_chr[signif_prots_heatmap_chr$tree_col[["order"]],])

# This are the effects

# heatmap: All heatmaps together
ggplot(significant_proteins_chromatin) + 
  geom_tile(aes(fct_relevel(gene,signif_prots.gene.order_chr),fct_relevel(feature, heatmap.chromatin.order), fill = CCD_value)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.15,0.15), oob = squish, breaks = c(-0.1,0,0.1)) + theme_bw() +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), legend.position = "top") + xlab("Chromatin proteins with significant chromatin dependencies") + ylab("Chromatin features") + coord_fixed(ratio = 1)

# Dendrogram
ggdendrogram(signif_prots_heatmap_chr$tree_col)

```

