---
title: "xv20211230_5_epistasis_inhibitors"
output: html_document
---
# Script to generate figure #5 in the paper. This figure shows that epistatic interactions can be bigger if drugs are used instead of KO. Figure outline:
A- Effect of drug treatment in MMEJ:NHEJ balance
B- Matrix for all treatments
C- Slopes are bigger in inhibition than KO
D- PARP1 and olaparib correlation
E- ATM and ATMi correlation

Data in text:
A-...
B-...

Supplementary figure 6:
A- PARP1 vs. veliparib
B- PARP1 vs. talazoparib
C- ATRi vs. VE-821
D- ATRi vs. VE-822

This script generates the plots for figure 5 and S6

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(dendextend)
library(Hmisc)
library(ggpubr)
library(rstatix)
library(ggdendro)
```


# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6","grey80"), label = c("Chromatin","DNA_repair","both"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128b_ratio_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220110_ratio_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20220128_ratio_kinase_inhib_epistasis.rds")

#Check values
chip.values.DDR <- readRDS("/DATA/projects/DSBrepair/data/R/rs20220120_chip_residuals_clone5.RDS")

```
# Number of proteins in each module
```{r}
module_III_trip_het <- gene.f.slope.protein %>% filter(epistasis.final.b > 0.03 & feature %in% c("H3K9me2","H3K9me3","late_replicating","LMNB1")) %>% pull(gene) %>% unique()
module_II_polycomb <- gene.f.slope.protein %>% filter(epistasis.final.b > 0.03 & feature %in% c("EZH2","H3K27me3")) %>% pull(gene) %>% unique()
module_I_euchr <- gene.f.slope.protein %>% filter(epistasis.final.b > 0.03 & feature %in% c("H3K4me1","H3K27ac","H3K4me2","H3K4me3","H4K5acK8ac")) %>% pull(gene) %>% unique()

```

# Figure 3A: Chromatin vs. DNA repair
```{r}
module.summary <- selected.slopes %>% left_join(interaction.summary, by = "gene") %>% left_join(module_summary, by = c("feature")) %>% dplyr::group_by(module,gene) %>% dplyr::summarise(mean = mean(epistasis.final, na.rm = T)) %>% left_join(gene_library, by = "gene")

#Number of proteins in module I (n =8 of 60, 13.3%)
I <- module.summary %>% filter(gene %in% c(module_I_euchr)) %>% na.omit() %>% ungroup() %>% dplyr::select(librar,gene) %>% distinct() %>% dplyr::group_by(librar) %>% dplyr::summarise(counts = n()) %>% mutate(module = "I")

#Number of proteins in module II (n =20 of 80, 25%)
II <- module.summary %>% filter(gene %in% c(module_II_polycomb)) %>% na.omit() %>% ungroup() %>% dplyr::select(librar,gene)  %>% distinct() %>% dplyr::group_by(librar) %>% dplyr::summarise(counts = n()) %>% mutate(module = "II")

#Number of proteins in module III (n =81 of 177, 46%)
III <- module.summary %>% filter(gene %in% c(module_III_trip_het)) %>% na.omit() %>% ungroup() %>% dplyr::select(librar,gene)  %>% distinct() %>% dplyr::group_by(librar) %>% dplyr::summarise(counts = n()) %>% mutate(module = "III")

#Bind cols for plot
all.modules <- bind_rows(I,II,III)

# 
ggplot(all.modules) + geom_col(aes(module, counts, fill = fct_relevel(librar,c("DNA_repair","both","Chromatin"))))  + theme_bw() + coord_cartesian(expand = T) + scale_fill_manual(values = c("#56B3E6","grey40","#E69F03")) + theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + theme(legend.position = "top") + labs(fill = "CRISPR library")

  
```

# Figure 3B: GOST analysis for all each module
```{r}
module_I_gost <- gost(module_I_euchr, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_I_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")

module_II_gost <- gost(module_II_polycomb, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_II_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")

module_III_gost <- gost(module_III_trip_het, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_III_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")
```
#Plot RFC & missmatch proteins
```{r}
# Select RFC genes and missmatch proteins
RCF_proteins <- c("RFC2","RFC3","RFC5")
missmatch <- c("MSH2","MSH6","PMS1","PMS2","RPA1","LIG1","POLD","PCNA","MLH1")

#Cluster by genes
RFC.prots.dcast <- selected.slopes %>% filter(gene %in% RCF_proteins) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")
RFC.prots.heatmap <- pheatmap(RFC.prots.dcast,method = "ward.D2", silent = F)
RFC.prots.clustering <- rownames(RFC.prots.dcast[RFC.prots.heatmap$tree_row[["order"]],])

#Cluster by genes
missmatch.prots.dcast <- selected.slopes %>% filter(gene %in% missmatch) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")
missmatch.prots.heatmap <- pheatmap(missmatch.prots.dcast,method = "ward.D2", silent = F)
missmatch.prots.clustering <- rownames(missmatch.prots.dcast[missmatch.prots.heatmap$tree_row[["order"]],])

# Both sides (RFC)
ggplot(selected.slopes %>% filter(gene %in% RCF_proteins)) + 
  geom_tile(aes(fct_relevel(gene,RFC.prots.clustering),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

# Both sides (RFC)
ggplot(selected.slopes %>% filter(gene %in% missmatch)) + 
  geom_tile(aes(fct_relevel(gene,missmatch.prots.clustering),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()


```
# Figure 3C: PARP inhibitors
```{r}
# Select PARP proteins
PAR.proteins <- selected.slopes %>% filter(gene %in% c("PARP3","PARP1","PARG","PARP2","PARP4")) %>% dplyr::select(drug = gene, feature,intercept,slope,mean.fc,high,sign.presence,epistasis.sign,epistasis.final = epistasis.final.b)

# Select PARP inhibitors
PARP.inhibitors <- slope.protein.features.inhibitors.filt %>% filter(drug %in% c("olaparib","talazoparib","veliparib"))


# PAR genes dcast (cluster genes)
PAR.prots.dcast <- PAR.proteins %>% reshape2::dcast(drug ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "drug")
PAR.prots.heatmap <- pheatmap(PAR.prots.dcast, method = "ward.D2", silent = F)
PAR.prots.clustering <- rownames(PAR.prots.dcast[PAR.prots.heatmap$tree_row[["order"]],])

# PAR genes dcast (cluster genes)
PARP.inh.dcast <- PARP.inhibitors %>% reshape2::dcast(drug ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "drug")
PARP.inh.heatmap <- pheatmap(PARP.inh.dcast, method = "ward.D2", silent = F)
PARP.inh.clustering <- rownames(PARP.inh.dcast[PARP.inh.heatmap$tree_row[["order"]],])


# Heatmap with proteins only
ggplot(PAR.proteins) + 
  geom_tile(aes(fct_relevel(drug, PAR.prots.clustering),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.11,0.11), oob = squish, breaks = c(-0.1,0,0.1))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

# Heatmap with PARP inhibitors
ggplot(PARP.inhibitors) + 
  geom_tile(aes(fct_relevel(drug,PARP.inh.clustering),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.11,0.11), oob = squish, breaks = c(-0.1,0,0.1))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()


```


# CORUM database
```{r}
#Triple heterochromatin
module_I_gene_set <- module_I_genes %>% pull(gene) %>% unique()
module_I_gost <- gost(module_I_gene_set, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_I_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")

#Polycomb
module_II_gene_set <- module_II_genes %>% pull(gene) %>% unique()
module_II_gost <- gost(module_II_gene_set, organism = "hsapiens",sources = "CORUM")
# FANC anemia complex
ggplot(module_II_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")

#Euchromatin
module_III_gene_set <- module_III_genes %>% pull(gene) %>% unique()
module_III_gost <- gost(module_III_gene_set, organism = "hsapiens",sources = "CORUM")
# BASC complex
ggplot(module_III_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")


#nhej 
module_I_nhej_gene_set <- module_nhej %>% pull(gene) %>% unique()
module_I_nhej_gost <- gost(module_I_nhej_gene_set, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_I_nhej_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_minimal() + xlab("-log10(p-adj)") + ylab("")


#nhej 
module_nhej_III_gene_set <- module_nhej_III %>% pull(gene) %>% unique()
module_III_nhej_gost <- gost(module_nhej_III_gene_set, organism = "hsapiens",sources = "CORUM")
# Signalling factors
# MLL4 complex
ggplot(module_III_nhej_gost$result %>% dplyr::group_by(term_name, term_id) %>% dplyr::summarise(p_val = -log10(mean(p_value)), CORUM_name = paste0(term_name, term_id, sep = "_"))) + geom_col(aes(p_val,reorder(CORUM_name,p_val)), fill = "#EB2030") + coord_cartesian(expand = F) + theme_bw() + xlab("-log10(p-adj)") + ylab("")

```
# Figure 3B: Complexes enriched in module I
```{r}
BASC.complex.genes <- c("BLM","BRCA1","MLH1","MSH2","MSH6","RFC1","RFC2","RFC4","RAD50","MRE11","NBN","ATM")

# dcaast with selected
slope.prot.features.dcast.BASC <- gene.f.slope.protein %>% filter(gene %in% BASC.complex.genes & feature %in% mmej_module_I_euchr_features$feature) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.BASC <- pheatmap(slope.prot.features.dcast.BASC, silent = F)
BASC <- dist(slope.prot.features.dcast.BASC)
row_dend.BASC = hclust(BASC)
row_dend.hits.BASC <- as.dendrogram(row_dend.BASC)
row_dend.hits.BASC <- seriate_dendrogram(row_dend.hits.BASC, BASC, method="OLO")

heatmap.gene.order.slope.diff.BASC <- rownames(slope.prot.features.dcast.BASC[as.hclust(row_dend.hits.BASC)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])

#chromatin order
heatmap.chromatin.order <- c("HDAC3","LMNB1","late_replicating","H3K9me2","H3K9me3","HDAC1","EZH2","H3K27me3","CTCF","SMC3","m5C","HDAC2","POL2","POL2AS2","H3K36me3","TTseq","Dam","H3K79me2","H2AFZ","DNAse","H3K4me3","H4K5acK8ac","H3K4me1","H3K27ac","H3K4me2")

# Both sides
ggplot(selected.slopes %>% filter(gene %in% BASC.complex.genes)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.BASC),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

```
# Figure 3B: Complexes enriched in module I
```{r}
RFC.complex.genes <- c("CCNA2","CDK2","LIG1","POLA1","POLD1","POLE","TOP1","PARP1","RFC1","RFC2")

# dcaast with selected
slope.prot.features.dcast.RFC <- gene.f.slope.protein %>% filter(gene %in% RFC.complex.genes & feature %in% mmej_module_I_euchr_features$feature) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.RFC <- pheatmap(slope.prot.features.dcast.RFC, silent = F)
RFC <- dist(slope.prot.features.dcast.RFC)
row_dend.RFC = hclust(RFC)
row_dend.hits.RFC <- as.dendrogram(row_dend.RFC)
row_dend.hits.RFC <- seriate_dendrogram(row_dend.hits.RFC, RFC, method="OLO")

heatmap.gene.order.slope.diff.RFC<- rownames(slope.prot.features.dcast.RFC[as.hclust(row_dend.hits.RFC)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])

# Both sides
ggplot(selected.slopes %>% filter(gene %in% RFC.complex.genes)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.RFC),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e", limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()


```


# Figure 3B: Complexes enriched in module I
```{r}
MSH.complex.genes <- c("MLH1","MSH2","MSH6","PMS2","PCNA")

# dcaast with selected
slope.prot.features.dcast.MSH <- gene.f.slope.protein %>% filter(gene %in% MSH.complex.genes & feature %in% mmej_module_I_euchr_features$feature) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")

#Heatmap
heatmap.slope.diff.MSH <- pheatmap(slope.prot.features.dcast.MSH, silent = F)
MSH <- dist(slope.prot.features.dcast.MSH)
row_dend.MSH = hclust(MSH)
row_dend.hits.MSH <- as.dendrogram(row_dend.MSH)
row_dend.hits.MSH <- seriate_dendrogram(row_dend.hits.MSH, MSH, method="OLO")

heatmap.gene.order.slope.diff.MSH<- rownames(slope.prot.features.dcast.MSH[as.hclust(row_dend.hits.MSH)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])

# Both sides
ggplot(selected.slopes %>% filter(gene %in% MSH.complex.genes)) + 
  geom_tile(aes(fct_relevel(gene,heatmap.gene.order.slope.diff.MSH),fct_relevel(feature, heatmap.chromatin.order.slope.diff.hits), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.075,0.075), oob = squish, breaks = c(-0.05,0,0.05))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

```
# Figure 3C: PARP inhibitors
```{r}
# Plot data with both of them
proteins.PIKK <- slope.protein.features.all %>% filter(gene %in% c("PARP3","PARP1","PARG","PARP2","PARP4")) %>% select(drug = gene, feature,intercept,slope,mean.fc,high,sign.presence,epistasis.sign,epistasis.final)

PAR.prots <- slope.protein.features.inhibitors.filt %>% filter(drug %in% c("olaparib","talazoparib","veliparib")) %>% bind_rows(proteins.PIKK) 


# dcaast with selected
slope.prot.features.dcast.PAR <- PAR.prots %>% reshape2::dcast(drug ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "drug")

#Heatmap
heatmap.slope.diff.PAR <- pheatmap(slope.prot.features.dcast.PAR, method = "ward.D2", silent = F)
PAR<- dist(slope.prot.features.dcast.PAR)
row_dend.PAR= hclust(PAR)
row_dend.hits.PAR <- as.dendrogram(row_dend.PAR)
row_dend.hits.PAR <- seriate_dendrogram(row_dend.hits.PAR, PAR, method="OLO")

heatmap.drug.order.slope.diff.PAR<- rownames(slope.prot.features.dcast.PAR[as.hclust(row_dend.hits.PAR)[["order"]],])

# Both sides
ggplot(PAR.prots) + 
  geom_tile(aes(fct_relevel(drug,c("PARP1","PARP2","PARP4","PARP3","PARG","olaparib","talazoparib","veliparib")),fct_relevel(feature, heatmap.chromatin.order), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a",mid = "#f5f5f5", high = "#01665e",limits = c(-0.11,0.11), oob = squish, breaks = c(-0.1,0,0.1))  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position = "top") + coord_fixed()

```

# FANCONI anemia
#Figure 3H: FANC proteins
```{r}
#dcast epistatic interactions (FANC proteins)
slope.prot.features.scale.dcast.FA <- selected.slopes %>% filter(gene %in% c("FANCL","FAAP100","FAAP24","FANCA","FANCB","FANCC","FANCE","FANCF","FANCG","FANCM","FANCD2","FANCI") & feature %in% module_II_polycomb_features$feature) %>% reshape2::dcast(gene ~ feature, value.var = "epistasis.final.b") %>% column_to_rownames(var = "gene")


ggplot(selected.slopes %>% filter(gene %in% c("FAAP24","FANCM","FANCA","FANCG","FANCF","FANCC","FAAP100","FANCE","FANCL","FANCB","FANCI","FANCD2"))) + 
  geom_tile(aes(fct_relevel(gene,c("FAAP24","FANCM","FANCA","FANCG","FANCF","FANCC","FAAP100","FANCE","FANCL","FANCB","FANCI","FANCD2")), fct_relevel(feature,heatmap.chromatin.order), fill = epistasis.final.b)) +
  scale_fill_gradient2(low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

FA.test <- gene.f.slope.protein %>% filter(gene %in% c("FAAP24","FANCM","FANCA","FANCG","FANCF","FANCC","FAAP100","FANCE","FANCL","FANCB","FANCI","FANCD2")) %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater") %>% mutate(fdr = p.adjust(p, method = "fdr"))

```

#Figure 5A: Effect in pathway balance
```{r}
# Mean effect over replicates per experiment
mean.combined.inhibitor.data <- combined.inhibitor.data.filt %>% dplyr::group_by(drug,barcode,exp) %>% dplyr::summarise(m.ratio = mean(NHEJMMEJratio), na.rm = T)

#DMSO control data
DMSO.control <- mean.combined.inhibitor.data %>% filter(drug == "DMSO") %>% ungroup() %>%  select(barcode,DMSO.ratio = m.ratio, exp)

# Calculate log2 differences
log2.fc.drugs <- mean.combined.inhibitor.data %>% left_join(DMSO.control, by = c("barcode", "exp")) %>% mutate(log2.fc = log2(m.ratio/DMSO.ratio))

# Data plot
plot.dt.log2.fc.drugs <- log2.fc.drugs %>% filter(!drug %in% c("DMSO","GFP")) %>% dplyr::group_by(barcode,drug) %>% dplyr::summarise(log2.fc = mean(log2.fc, na.rm = T))

# Plot
ggplot(plot.dt.log2.fc.drugs, aes(drug,log2.fc)) + 
geom_quasirandom() + 
stat_mean(geom = "point", color = "red") +
geom_hline(yintercept = 0, linetype = 2) + 
theme_bw() +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

test.perturbation.drug <- plot.dt.log2.fc.drugs %>%
  dplyr::group_by(drug) %>%
  wilcox_test(log2.fc ~ 0)

```


# Figure 6B: Heatmap
```{r}
# Plot data with both of them
proteins.PIKK <- slope.protein.features.all %>% filter(gene %in% c("ATM","ATR","PARP1","PRKDC","POLL")) %>% select(drug = gene, feature,intercept,slope,mean.fc,high,sign.presence,epistasis.sign,epistasis.final)

#dcast epistatic interactions
slope.prot.features.scale.dcast.inhibitors <- slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% reshape2::dcast(drug ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "drug")

#Heatmap for slope differences
heatmap.slope.diff.inhibitors <- pheatmap(slope.prot.features.scale.dcast.inhibitors, silent = T)
heatmap.gene.order.slope.diff.inhibitors <- rownames(slope.prot.features.scale.dcast.inhibitors[heatmap.slope.diff.inhibitors$tree_row[["order"]],])

#heatmap.gene.order.slope.diff.inhibitors <- rownames(slope.prot.features.scale.dcast.inhibitors[dendextend::rotate(heatmap.slope.diff.inhibitors$tree_row, order = c("DNAPKi","ATMi","ATRi_VE821","ATRi_VE822","veliparib","talazoparib","olaparib"))[["order"]],])
heatmap.chromatin.order.slope.diff.inhibitors <- colnames(slope.prot.features.scale.dcast.inhibitors[heatmap.slope.diff.inhibitors$tree_col[["order"]]])


#Heatmap

inh.arrange <- dist(slope.prot.features.scale.dcast.inhibitors)
row_dend.inh = hclust(inh.arrange)
row_dend.inhi <- as.dendrogram(row_dend.inh)
row_dend.inhibitors <- seriate_dendrogram(row_dend.inhi, inh.arrange, method="OLO")

heatmap.gene.order.slope.diff.hits <- rownames(slope.prot.features.scale.dcast.inhibitors[as.hclust(row_dend.inhibitors)[["order"]],])
heatmap.chromatin.order.slope.diff.hits <- colnames(slope.prot.features.dcast.hits[,heatmap.slope.diff.hits$tree_col[["order"]]])




#Plot correlations (As a heatmap with corrected pvalues)
ggplot(slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK)) + 
  geom_tile(aes(fct_relevel(feature, heatmap.chromatin.order.slope.diff.inhibitors),fct_relevel(drug,heatmap.gene.order.slope.diff.inhibitors), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# dendogram
ggdendrogram(dendextend::rotate(heatmap.slope.diff.inhibitors$tree_row, order = c("DNAPKi","ATMi","ATRi_VE821","ATRi_VE822","veliparib","talazoparib","olaparib")), rotate =T)

# Alternative plot: PRKDC
ggplot(slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% filter(drug %in% c("PRKDC","DNAPKi"))) + 
  geom_tile(aes(fct_relevel(feature, heatmap.chromatin.order.slope.diff.inhibitors),fct_relevel(drug,heatmap.gene.order.slope.diff.inhibitors), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# Alternative plot: olaparib, PARP1, talazoparib, veliparib
ggplot(slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% filter(drug %in% c("PARP1","PARP3","olaparib","talazoparib","veliparib"))) + 
  geom_tile(aes(fct_relevel(feature, heatmap.chromatin.order.slope.diff.inhibitors),fct_relevel(drug,heatmap.gene.order.slope.diff.inhibitors), fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")


# 
ggplot(slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% filter(drug %in% c("PARP1","PARP3","olaparib","talazoparib","veliparib"))) +
  geom_tile(aes(fct_relevel(drug,rev(heatmap.gene.order.slope.diff.inhibitors)),fct_relevel(feature, heatmap.chromatin.order.slope.diff.inhibitors),fill = epistasis.final)) +
  scale_fill_gradient2( low = "#8c510a" ,mid = "#f5f5f5", high = "#01665e")  + 
  coord_fixed(expand = F, ratio = 1) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top")

# PARP test
PARP.test <- slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% filter(drug %in% c("PARP1","PARP3","olaparib","talazoparib","veliparib")) %>% group_by(feature) %>% t_test(epistasis.final ~ 0, alternative = "greater")

```
# Figure 3B: Correlation between fold.change and mean(interaction)
```{r}
epistasis.drug <- slope.protein.features.inhibitors.filt %>% bind_rows(proteins.PIKK) %>% filter(epistasis.final != 0) %>% dplyr::group_by(drug,mean.fc) %>% dplyr::summarise(mean.epistasis = mean(epistasis.final, na.rm = T), max.epistasis = sign()*(max(abs(epistasis.final), na.rm = T))) %>% mutate(class = case_when(grepl("i",drug) ~ "inhibitor", T ~ "KO"))

ggplot(epistasis.drug %>% select(mean.fc,max.epistasis,class) %>% distinct(), aes(mean.fc, max.epistasis, color = class, group = "1")) + geom_point() + geom_smooth(method = "loess") + stat_cor(method = "spearman") + geom_hline(yintercept = 0, linetype = 2) + geom_vline(xintercept = 0, linetype = 2) + theme_bw() + xlab("log2(fold change NHEJ:MMEJ balance)") + coord_fixed(ratio = 12) + theme(legend.position = "top")


```

# Figure 3C: Correlation within chromatin features
#Figure 7A: Correlations between features

```{r}
slope.prot.features.dcast.all <- slope.protein.features.all %>% filter(epistasis.final != 0) %>% dcast(gene ~ feature, value.var = "epistasis.final") %>% column_to_rownames(var = "gene")


#Feature order by correlation
cor.features.epistasis <- cor(slope.prot.features.dcast.all, method = "spearman", use = "pairwise.complete.obs")
heatmap.features.epistasis <- pheatmap(cor.features.epistasis, silent = T)
heatmap.features.epistasis.rows <- rownames(cor.features.epistasis[heatmap.features.epistasis$tree_row[["order"]],])

# Calculate p-values
#rcorr.features.z.score <- rcorr(as.matrix(clone5_z.score_chrom_tib %>% filter(barcode %in% clone5bcs) %>% select(-ID, -barcode,-binsize, -pool,-binsize)), type = "spearman")
#p.value.adj <- as.data.frame(rcorr.features.z.score$P) %>% rownames_to_column(var = "feature_1") %>% reshape2::melt() %>% filter(complete.cases(.)) %>% mutate(p.adj = p.adjust(value, method = "BH"))

# corr.features.z.score
features.z.score <- clone5_z.score_chrom_tib %>% filter(ID %in%  unique(both.screen.detail$barcode)) %>% select(-ID,-binsize, -pool,-binsize) 

corr.features.z.score <- cor(features.z.score, method = "spearman") %>% as.data.frame()%>% rownames_to_column(var = "feature_1") %>% reshape2::melt() %>% filter(complete.cases(.)) %>%select(feature_1, feature_2 = variable, cor_chrom = value)
melt.cor.features.epistasis <- melt(cor.features.epistasis) %>% select(feature_1 = Var1, feature_2 = Var2, cor_ep = value)


# get a value column with values that are Both, only in one or the other
corr_names <- corr.features.z.score %>% left_join(melt.cor.features.epistasis ) %>% mutate(classes = case_when((abs(cor_chrom) > 0.6 & abs(cor_ep) < 0.6) ~ "Chrom", (abs(cor_chrom) < 0.6 & abs(cor_ep) > 0.6) ~ "Epistasis", (abs(cor_chrom) >0.6 & abs(cor_ep) >0.6) ~ "Both", T ~ "NA")) %>% mutate(diff = cor_ep - cor_chrom)


# Make correlation plots for features
ggplot(melt.cor.features.epistasis) + 
geom_tile(aes(fct_relevel(feature_1, heatmap.features.epistasis.rows), fct_relevel(feature_2,heatmap.features.epistasis.rows), fill = cor_ep)) +
  scale_fill_gradient2( low = "#009B9E" ,mid = "#F1F1F1", high = "#C75DAB", limits = c(-1,1), na.value = "#2E358F")  + 
  coord_fixed(expand = F) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), axis.title = element_blank()) + 
  geom_point(data = subset(corr_names, classes != "NA"), aes(fct_relevel(feature_1, heatmap.features.epistasis.rows), fct_relevel(feature_2,heatmap.features.epistasis.rows), shape = classes)) + scale_shape_manual(values = c(4,17,16)) + scale_x_discrete(position = "top")

#Dendrogram
ggdendrogram(heatmap.features.epistasis$tree_row, rotate = T)


# 
ggplot(slope.prot.features.dcast.all, aes(CTCF,SMC3)) + geom_point(size = 0.5) + stat_cor(method = "spearman", cor.coef.name = "rho") + geom_smooth(method = "lm") + theme_bw()  + coord_fixed(ratio = 1)

ggplot(features.z.score, aes(CTCF,SMC3)) + geom_point(size = 0.5) + stat_cor(method = "spearman", cor.coef.name = "rho") + geom_smooth(method = "lm") + theme_bw()  + coord_fixed(ratio = 1)


ggplot(slope.prot.features.dcast.all, aes(EZH2,H3K27me3)) + geom_point(size = 0.5) + stat_cor(method = "spearman", cor.coef.name = "rho") + geom_smooth(method = "lm") + theme_bw()  + coord_fixed(ratio = 1)

ggplot(features.z.score, aes(EZH2,H3K27me3)) + geom_point(size = 0.5) + stat_cor(method = "spearman", cor.coef.name = "rho") + geom_smooth(method = "lm") + theme_bw()  + coord_fixed(ratio = 1)
```