---
title: "xv20220117_1_General_effects"
output: html_document
---
# Script to generate figure #1 in the short version of the paper. This figure describes the layout of the screen and general effects. The main point of this figure is to build up confidence in our reporter cell line and describe how these pathways are build up. For the short version, I'll keep the script for supplementary figures separate. Figure outline:
A- Cartoon explaining screen
B- Chromatin landscape where 19IPRs are integrated
C- Volcano plot with all perturbations (Only for DNA repair proteins)


This script generates the plots for figure 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r libraries}
# libraries:
library(tidyverse)
library(data.table)
library(parallel)
library(gtools)
library(pheatmap)
library(ggbeeswarm)
library(reshape2)
library(gridExtra)
library(broom)
library(ggrepel)
library(gprofiler2)
library(ggdendro)
```

# Aesthetic legend for the whole paper
This are the colors that I will use for the whole paper. This chunk will be copied in every file.
```{r}
# Pathway color palette
pathway.colors <- tibble(color = c("#EB2030","grey90","#2E358F"), label = c("MMEJ","no_effect","NHEJ"), feature = "pathway_balance")

#Library colors 
library.colors <- tibble(color = c("#E69F03","#56B3E6"), label = c("Chromatin","DNA_repair"),  feature = "library")

#KAT5 example color
KAT5.example <- tibble(color = "#EF6817", label = "KAT5", feature = "example")

#Epistatic interaction colors
slope.colors <- tibble(color = c("#01665e","#f5f5f5","#8c510a"),label = c("negative","none","positive"), feature = "epistasis")

#Inhibitor and protein KO colors
inh.colors <- tibble(color = c("#2A52BE","#2B3035"), label = c("drug","protein"), feature = "inh_KO")

#Four complex selection
complex.colors <- tibble(color = c("#EA3442","#EA5C68","#2E358F","#EA717A"), label = c("NHEJ","FANC_core_complex","MRN_complex","RNF8_RNF168"), feature = "protein_complexes")

#RNF8_RNF168 slope examples
ubq.response.colors <-  tibble(color = c("#1732ef","#179eef","#ef6817","#efd417"), label = c("RNF8_RNF168","H2AK15ub","H2AK15ac","H1ub"), feature = "RNF8_RNF168_example")

# Chromatin correlations
chr.correlation.colors <- tibble(color = c("#009B9E","#F1F1F1","#C75DAB"), label = c("negative","none","positive"), feature = "chromatin_correlation")

#Bind all and plot
paper.palette <- bind_rows(pathway.colors, library.colors,KAT5.example,slope.colors,inh.colors, complex.colors,ubq.response.colors,chr.correlation.colors) 
ggplot(paper.palette) +
  geom_tile(aes(label,"1", fill = color, width = 1)) +
  geom_text(aes(label,"1", label = color)) +
  scale_fill_manual(values = levels(as.factor(paper.palette$color))) +
  facet_wrap(~ feature, scales = "free", nrow = 4, strip.position = "top") +
  theme_bw() +
  theme(legend.position = "none")
```

# Import data tables
```{r libraries}
#detailed table
#ddr
ddr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_scores.rds")
colnames(ddr.screen.detail)[9] <- "b.rep"
colnames(ddr.screen.detail)[10] <- "t.rep"
ddr.screen.detail$library <- "DNA_repair"
#chromatin
chr.screen.detail <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_mmej.rds")
chr.screen.detail$library <- "Chromatin"

# I think I don't need these ones for this figure
#Control table (WT & POLQ)
#ddr
ddr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210609_screen_ddr_controls.rds")
colnames(ddr.screen.controls)[9] <- "b.rep"
colnames(ddr.screen.controls)[10] <- "t.rep"
ddr.screen.controls$library <- "DNA_repair"

#chromatin
chr.screen.controls <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20201014_frozen_chz_scores_controls.rds")
chr.screen.controls$library <- "Chromatin"

#All variable table
ddr.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20220110_screen_ddr_all_computations.rds")
ddr.screen.all.sel <- ddr.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep = bio.rep,t.rep = FinRep) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "DNA_repair")

#Chromatin all 
chromatin.screen.all <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv20200421_ChromDSBscreen/data/xv20220110_chromatin_library_all.rds")
chromatin.screen.all.sel <- chromatin.screen.all %>% select(well,plate,sample,gene,barcode,NHEJMMEJratio,MMEJscore, b.rep,t.rep) %>% mutate(MMEJNHEJratio = 1/NHEJMMEJratio, library = "Chromatin")

# bind both dataset
both.screen.all.sel <- bind_rows(ddr.screen.all.sel,chromatin.screen.all.sel) %>% filter(!gene %in% c("DMSO","LBR_ctrl"))


#indel data
indel.data.ddr <- readRDS(file = "/home/x.vergara/XV_P3_ChromDSBScreen/xv20201204_agm_ddr_screen/data/xv20210526_screen_DDR_indel_data.rds")

# Chromatin follow up data
chromatin.followup <- readRDS("/home/x.vergara/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211005_chromatin_follow_up.rds")

# Chromatin followup mutate IPRscore
chromatin.fup.IPR <- chromatin.followup %>% filter(plate == "P1") %>% dplyr::group_by(well,gene,gRNA,barcode,plate) %>% dplyr::summarise(IPR.z.score = sum(z.score,na.rm = T)/sqrt(n()), count = n())

# Put both screens together
#detailed table with all the variables
both.screen.detail <- bind_rows(ddr.screen.detail, chr.screen.detail) %>% mutate(ID_gene = paste(gene, library, sep = "_")) %>% filter(sample == "KO" & !ID_gene %in% c("DMSO_DNA_repair","LBR_ctrl_DNA_repair"))

both.screen.gene.value.pre <- both.screen.detail %>% select(ID_gene, b.rep, t.rep,barcode, mmej.z.score) %>% 
  mutate(tech.rep = case_when(t.rep == "R5" ~ "R3", t.rep == "R4" ~ "R2", T ~ t.rep)) %>%
  dplyr::group_by(ID_gene,tech.rep) %>% 
  dplyr::summarise(replicate.score = mean(mmej.z.score, na.rm = T)) %>% 
  dplyr::group_by(ID_gene) %>%
  dplyr::mutate(IPR.z.score = sum(replicate.score, na.rm = T)/sqrt(n()))

#Data for plotting reproducibility
both.screen.gene.value <- both.screen.gene.value.pre %>%
  reshape2::dcast(ID_gene ~ tech.rep, value.var = "IPR.z.score") %>% 
  filter(complete.cases(.)) %>% 
  mutate(comb.data = (R1+R2+R3)/sqrt(3))

# Data to call hits
hits.both.screens <- both.screen.gene.value.pre %>% mutate(pathway = case_when(IPR.z.score < -1.95 ~ "MMEJ", IPR.z.score > 1.95 ~ "NHEJ", T ~ "NA"), library = case_when(grepl("DNA", ID_gene) ~ "DNA_repair", grepl("Chromatin", ID_gene) ~ "Chromatin"))

# Filter
per_19IPR_data <- both.screen.detail %>% 
  dplyr::group_by(gene,barcode,library) %>% 
  dplyr::summarise(IPR.z.score = sum(mmej.z.score, na.rm = T)/sqrt(n()), count = n()) %>% ungroup()

#Clone 5 chromatin
clone5_z.score_chrom_tib <- readRDS('/DATA/projects/DSBrepair/data/R/cl20201026_ChIP_zscore_selection.RDS')

#Epistasis analysis
slope.protein.features.all <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_all_genes.rds")

step1.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_1_step.rds")
step2.epistasis <- readRDS("~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_epistasis_2_step.rds")

#Inhibitor perturbation data
combined.inhibitor.data.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_effect.rds") 

#Export epistasis interactions
slope.protein.features.inhibitors.filt <- readRDS(file = "~/XV_P3_ChromDSBScreen/xv_CRISPR_screen_figures/data/xv20211230_kinase_inhib_epistasis.rds")
```

#Main figures

#Figure 1A: Cartoon

# Figure 1B: Epigenetic features (CHECK!)
```{r}
clone5_bc <- both.screen.detail %>% pull(barcode) %>% unique()
# Prepare data frame
clone_5_IPR <- filter(clone5_z.score_chrom_tib, ID %in% clone5_bc) %>% select(-pool,-binsize) %>% column_to_rownames(var = "ID")

#pheatmap
clone_5_heatmap <- pheatmap(clone_5_IPR, silent = T, clustering_method = "ward.D")

#Arrange col & row
clone5_bc_cluster <- rownames(clone_5_IPR[clone_5_heatmap$tree_row[["order"]],])
clone5_chromatin_cluster <- colnames(clone_5_IPR[,clone_5_heatmap$tree_col[["order"]]])

#dt for plotting
clone_5_IPR_melt <- clone_5_IPR %>% rownames_to_column(var = "barcode") %>% melt(value.name = "chrom.z.score", var.name = "feature")

#Plot
ggplot(clone_5_IPR_melt) + 
  geom_tile(aes(fct_relevel(barcode, clone5_bc_cluster),fct_relevel(variable,heatmap.chromatin.order), fill = chrom.z.score)) +
  scale_fill_gradient2( low = "#009B9E",mid = "#F1F1F1", high = "#C75DAB")  + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title = element_blank(), legend.position = "top") + coord_fixed(expand = F,ratio = 0.75)

#dendrogram barcodes
ggdendrogram(clone_5_heatmap$tree_row, rotate = T)
#denrogram chromatin
ggdendrogram(clone_5_heatmap$tree_col)

```


#Figure 1D: Volcano plot only for DNA_repair proteins
NOTE:
If you use ggbreak in published research, please cite the following paper:

S Xu, M Chen, T Feng, L Zhan, L Zhou, G Yu. Use ggbreak to effectively utilize plotting space
to deal with large datasets and outliers. Frontiers in Genetics. 2021, 12:774846. doi:
10.3389/fgene.2021.774846 

```{r}
# Each replicate has a different value (compute differences by replicate)
log2.distance.ratio.all <- log2.ratio.screen.detail.all %>% filter(sample == "KO" & library == "DNA_repair") %>% left_join(wt.set.ratio) %>% mutate(log2.ratio = log2ratio - wt.log2ratio)
mean.log2.distance.ratio.all <- log2.distance.ratio.all %>% dplyr::group_by(barcode,gene) %>% dplyr::summarise(m.log2.ratio.dist = mean(log2.ratio, na.rm = T)) %>% na.omit() %>% left_join(clone5_z.score_chrom_tib, by = c("barcode" = "ID"))

#Compute mean log2 fold change
mean.log2.fc <- mean.log2.distance.ratio.all %>% group_by(gene) %>% dplyr::summarise(mean.fc = mean(m.log2.ratio.dist, na.rm = T))

test.log2.dist <- mean.log2.distance.ratio.all %>% dplyr::group_by(gene) %>% t_test(m.log2.ratio.dist ~ 0) %>% mutate(p.adj = p.adjust(p, method = "fdr")) %>% left_join(mean.log2.fc, by = "gene")

# hits mmej
hits.nhej.fdr <- test.log2.dist %>% filter(p.adj < 0.001 & mean.fc > 0) %>% pull(gene)

hits.nhej.fdr.5 <- test.log2.dist %>% filter(p.adj < 0.05 & mean.fc > 0) %>% pull(gene)

#Run GO profiler (this gives the whole list that will go to supplementary)
gotest.nhej.fdr <- gost(hits.mmej.fdr, organism = "hsapiens",sources = "CORUM")

# hits nhej
hits.mmej.fdr <- test.log2.dist %>% filter(p.adj < 0.001 & mean.fc < 0) %>% pull(gene)

hits.mmej.fdr.5 <- test.log2.dist %>% filter(p.adj < 0.05 & mean.fc < 0) %>% pull(gene)

#Run GO profiler (this gives the whole list that will go to supplementary) => NO significant hits
gotest.mmej.fdr <- gost(hits.nhej.fdr, organism = "hsapiens",sources = "CORUM")

# Highlight three complexes
FA_complex_ID_2378 <- tibble(gene = c("BLM","TOP3A","FANCL","FANCA","FANCB","FANCC","FANCE","FANCF","FANCG","RMI","RPA1","FANCM"), complex = "FA_comlplex_ID:2378")
MRN_complex <- tibble(gene = c("MRE11","RAD50","NBN","SHLD1"), complex = "MRN_complex")
DNA_ligase_complex_ID_365 <- tibble(gene = c("LIG4","XRCC4","NHEJ1","FANCD2"), complex = "DNA_ligase_complex")

highlight.complex <- bind_rows(FA_complex_ID_2378,DNA_ligase_complex_ID_365,MRN_complex)

# Plot and highlight main findings (Plot A: highlight MMEJ foctors)

ggplot(test.log2.dist %>% 
         left_join(highlight.complex)) +
  geom_point(aes(mean.fc,-log10(p.adj)), color = "grey40") + 
  geom_point(data = test.log2.dist %>% 
               left_join(highlight.complex) %>% 
               na.omit(),
             aes(mean.fc,-log10(p.adj), color = complex), size = 2) + theme_bw() + 
  geom_text_repel(data = subset(test.log2.dist,(-log10(p.adj) > 5 & mean.fc < -0.3) | (-log10(p.adj) > 5 & mean.fc > -0)),
                  aes(mean.fc,-log10(p.adj), label = gene),max.overlaps = Inf, box.padding = 0.5) + geom_vline(xintercept = 0, linetype = 2) + geom_hline(yintercept = -log10(0.001), linetype = 2) + ylab("-log10(FDR)") + xlab("log2(Average fold change MMEJ:NHEJ balance)") + theme(legend.position = "top") + scale_x_continuous(breaks = c(-2,-1,0,1)) + scale_color_manual( values = c("#ca0020","#92c5de","#0571b0"))

#Export instructions

```
# All figures can be found on: "
# This is the end of this script for main figure 1 => Supplementaries will be performed in a separate rmarkdown file